### 2017-02-11 Draft-xuezhou.Rmd
github test

---
title: "Draft-assinment1"
author: "Kun Shen, Xiang Li, Xue Zhou"
date: "February 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction
# 2. Objective
# 3. Dataset Preparation
Our analysis focuses on the data of year 2012 -2013 of 23 variables. We chose the median income 6 years after graduation as our predictor variable, and eliminated many fields with bunch of missing values as well as some fields not pertainig enougn to post-graduation income. We then fosued primarily on a set of 23 variables, including gender, age, ethnic, acedemic costs, family income, completetion rates and replayment rate.

The following libraries have been used in the analysis: 
```{r, message=FALSE}

library(readr)
library(dplyr)

```

We loaded ScoreCard from the US Dapartment of Education. The main dataset spans nearly 20 years of data and covering multiple sources including IPEDS, NSLDS, and Department of Treasury. To ensure the maximum data accurary and availablity, we choose the subset of year 2012 - 2013 as our ..... 
```{r, message=FALSE}
fulldata = read_csv("/Users/xuezhou/Desktop/MERGED2012_13_PP.csv")

### 2017-02-12 Draft-xuezhou.Rmd
combine the member's parts

@@
 library(dplyr)
+library(ggplot2)
+library(tidyr)
 
@@
 ```{r, message=FALSE}
-fulldata = read_csv("/Users/xuezhou/Desktop/MERGED2012_13_PP.csv")
+
+data = read.csv("MERGED2012_13_PP.csv",header = T, na.strings = "NULL")
+
+```
+
+The full dataset contains 1743 variables. 36 variables were intially selected to analyze.
+```{r, message=FALSE}
+
+variable = select(data, STABBR, PREDDEG, LOCALE2, MENONLY, WOMENONLY, CONTROL, ADM_RATE, SATMTMID, SATVRMID, SATWRMID, SAT_AVG,
+                  NPT4_PUB,
+                  NPT4_PRIV,
+                  UG,
+                  UG_NRA,
+                  UG_UNKN,
+                  UG_WHITENH,
+                  UG_BLACKNH,
+                  UG_API,
+                  UG_AIANOLD,
+                  UG_HISPOLD,
+                  DEP_STAT_PCT_IND,
+                  PAR_ED_PCT_1STGEN,
+                  MD_FAMINC,
+                  MEDIAN_HH_INC,
+                  POVERTY_RATE,
+                  PCTPELL,
+                  PCTFLOAN,
+                  GRAD_DEBT_MDN,
+                  WDRAW_DEBT_MDN,
+                  APPL_SCH_PCT_GE5,
+                  C150_4,
+                  MD_EARN_WNE_P6,
+                  MN_EARN_WNE_P10,
+                  RPY_3YR_RT_SUPP,
+                  CDR3)
+head(variable)
+```
+
+After loading the dataset, we change variables to correct data types. 
+```{r, message=FALSE}
+
+class(as.data.frame(apply(variable[,-c(1,2,6)],2,as.numeric)))
+
+#numeric
+variable[,-c(1,2,6)] %>%
+    apply(2, as.numeric) %>%
+    as.data.frame() %>%
+    {.} -> variable_num
+
+#categorical
+variable$preF=factor(variable$PREDDEG,levels=c(0,1,2,3,4),labels = c("NotClassified","Certificate-degree","Associate's-degree","Bachelor's-degree","Graduate-degree"))
+variable$ctrF=factor(variable$CONTROL,levels = c(1,2,3),labels = c("Public","Private Nonprofit","Private For-profit"))
+
+#combine
+data_new <- cbind(variable_num,variable$preF, variable$ctrF,variable$STABBR)
+mean(is.na(data_new$NPT4_PUB))
+```
+
+Deal with missing values
+
+```{r, message=FALSE}
+
+for (i in 1:length(data_new$NPT4_PRIV)) {
+   if (is.na(data_new$NPT4_PRIV[i])) {
+       data_new$NPT4_PRIV[i] = data_new$NPT4_PUB[i]
+   }
+}
+mean(is.na(data_new$NPT4_PRIV))
+data_new = data_new[,-9]
+colnames(data_new)[9] <- "NPT4_COMBINE"
+```
+
+Deal with missing values
+```{r, message=FALSE}
+
+data_clean = data_new[,colMeans(is.na(data_new)) < 0.25]
+write.csv(data_clean,"data_clean.csv")
+```
+
+#Variable summary
+For numeric variabls, we calculate the basic summary statistics such as median, min, max, mode and quantile.
+We also use visualizations such hisgroms and boxplots to understand the distribution of the values.
+For categorical variables, we use frequency table to understand the frequcies of each categories.Box chart is used as a visualization tool.
+
+| Tables        | Are           | Cool  |
+| ------------- |:-------------:| -----:|
+|col 3 is       | right-aligned | $1600 |
+|col 2 is       | centered      |   $12 |
+|zebra stripes  | are neat      |    $1 |
+
+| Variable Name    | Data Type         | Categorical or Continuous         | hi       |
+| --------------  | ----------------- | --------------------------------- | -------- |
+| NPT4_COMBINE    | Numeric           |  Continuous                       |  Cost    |
+| DEP_STAT_PCT_IND | Numeric           |  Continuous                       |     n    |
+| PAR_ED_PCT_1STGEN| Numeric           |  Continuous                       |      n   |
+| MD_FAMINC        | Numeric           |  Continuous                       |       n  |
+| PCTPELL          | Numeric           |  Continuous                       |n         |
+| PCTFLOAN         | Numeric           |  Continuous                       | n        |
+| GRAD_DEBT_MDN    | Numeric           |  Continuous                       |  nn      |
+| WDRAW_DEBT_MDN   | Numeric           |  Continuous                       |    n     |
+| MD_EARN_WNE_P6   | Numeric           |  Continuous                       |     n    |
+| MN_EARN_WNE_P10  | Numeric           |  Continuous                       |       n  |
+| RPY_3YR_RT_SUPP  | Numeric           |  Continuous                       |      n   |
+| CDR3             | Numeric           |  Continuous                       | n        |
+| variable$qreF    | Factor            |  Categorical                      |  n       |
+| variable$ctrF    | Factor            |  Categorical                      |   n      |
+| variable$STABBR  | Factor            |  Categorical                      |   n      |

### 2017-02-12 Draft-xuezhou.Rmd
added
added summary

@@
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE,warning=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
@@
 The following libraries have been used in the analysis: 
-```{r, message=FALSE}
+```{r, message=FALSE,warning=FALSE}
 
@@
 The main dataset spans nearly 20 years of data and covering multiple sources including IPEDS, NSLDS, and Department of Treasury.
-To ensure the maximum data accurary and availablity, we choose the subset of year 2012 - 2013 as our .....
-```{r, message=FALSE}
+To ensure the maximum data accurary and availablity, we choose the subset of year 2012 - 2013 as our analysis sample.
+```{r, message=FALSE,warning=FALSE}
 
@@
 The full dataset contains 1743 variables. 36 variables were intially selected to analyze.
-```{r, message=FALSE}
+```{r, message=FALSE,warning=FALSE}
 
@@
                   CDR3)
-head(variable)
+
 ```
 
-After loading the dataset, we change variables to correct data types. 
-```{r, message=FALSE}
+After loading the dataset, we change variables to the correct data types. 
+```{r, message=FALSE,warning=FALSE}
 
@@
 
-#combine
+#combine the three datafram to generate a new dataset. 
 data_new <- cbind(variable_num,variable$preF, variable$ctrF,variable$STABBR)
+dim(data_new)
+
 mean(is.na(data_new$NPT4_PUB))
@@
 
-Deal with missing values
-
-```{r, message=FALSE}
+```{r, message=FALSE,warning=FALSE}
 
@@
 
-Deal with missing values
-```{r, message=FALSE}
+In order to choose the variables that have enough valid data related to the topic, we filtered and considered variables with percent missing value less than 25%.
+The new data set was named as “data_clean” for further analysis.
+```{r, message=FALSE,warning=FALSE}
 
@@
 For numeric variabls, we calculate the basic summary statistics such as median, min, max, mode and quantile.
-We also use visualizations such hisgroms and boxplots to understand the distribution of the values.
+We also use visualizations such as hisgroms and boxplots to understand the distribution of the values.
 For categorical variables, we use frequency table to understand the frequcies of each categories.Box chart is used as a visualization tool.
 
-| Tables        | Are           | Cool  |
-| ------------- |:-------------:| -----:|
-|col 3 is       | right-aligned | $1600 |
-|col 2 is       | centered      |   $12 |
-|zebra stripes  | are neat      |    $1 |
+The table below lists the variables that will be used for the further analysis: 12 of them are continuous variables while 3 of them are categorical variables.
+Our data set now contains 120 rows and 15 columns.
+120 rows consist of 20 years’ of data for each of the chosen six countries.
+The 15 columns consist of country name, country code, year, and the twelve indicators.
 
@@
 | variable$STABBR  | Factor            |  Categorical                      |   n      |
+
+we will summarize the data set using dplyr functions in the variables summary section.
+
+#4 Goals
+#5 Summary of variables
+##5.1 Individual variable summary
+
+###5.1.1 Average net price for Title IV institutions(both public and private institutions)
+We are using the code below to summarize the continuous data
+
+```{R, message=FALSE, warinig = FALSE}
+
+summarize(group_by(data_clean, variable$STABBR),
+          minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),
+          maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),
+          meanNPT4_COMBINE=mean(NPT4_COMBINE,na.rm=TRUE),
+          medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE),
+          sdNPT4_COMBINE = sd(NPT4_COMBINE,na.rm=TRUE))
+
+data_clean %>%
+  group_by(variable$STABBR) %>%
+    summarise(minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),    maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE))  %>%
+    arrange(medianNPT4_COMBINE)   %>%
+{.}  ->test
+
+test
+
+#Here is the quartile statistics :
+summary(data_clean$NPT4_COMBINE,na.rm=TRUE)
+quantile(data_clean$NPT4_COMBINE,na.rm=TRUE)
+
+  #histogram for NPT_COM
+  ggplot(data=data_clean , aes(NPT4_COMBINE)) + 
+    geom_histogram(breaks=seq(0, 90000, by =2000), 
+                   col="red", 
+                   aes(fill=..count..),na.rm=TRUE)+
+    labs(title="Histogram for NPT_COM") +
+    labs(x="NPT_COM", y="Count")
+```
+
+##5.2 Relationship between variables
+##5.3 Comparison between variables
+
+#6. Findings and Conclusion

### 2017-02-12 Draft-xuezhou.Rmd
xiang

--- 2017-02-11 Draft-xuezhou.Rmd
@@
 ```{r, message=FALSE}
-fulldata = read_csv("/Users/xuezhou/Desktop/MERGED2012_13_PP.csv")
+
+data = read.csv("MERGED2012_13_PP.csv",header = T, na.strings = "NULL")
+data = read.csv("MERGED2012_13_PP.csv",header = T, na.strings = c("NULL","PrivacySuppressed"))
+
+```
+
+The full dataset contains 1743 variables. 36 variables were intially selected to analyze.
+```{r, message=FALSE}
+
+variable = select(data, STABBR, PREDDEG, LOCALE2, MENONLY, WOMENONLY, CONTROL, ADM_RATE, SATMTMID, SATVRMID, SATWRMID, SAT_AVG,
+                  NPT4_PUB,
+                  NPT4_PRIV,
+                  UG,
+                  UG_NRA,
+                  UG_UNKN,
+                  UG_WHITENH,
+                  UG_BLACKNH,
+                  UG_API,
+                  UG_AIANOLD,
+                  UG_HISPOLD,
+                  DEP_STAT_PCT_IND,
+                  PAR_ED_PCT_1STGEN,
+                  MD_FAMINC,
+                  MEDIAN_HH_INC,
+                  POVERTY_RATE,
+                  PCTPELL,
+                  PCTFLOAN,
+                  GRAD_DEBT_MDN,
+                  WDRAW_DEBT_MDN,
+                  C150_4,
+                  MD_EARN_WNE_P6,
+                  RPY_3YR_RT_SUPP,
+                  CDR3)
+```
+
+After loading the dataset, we change variables to correct data types. 
+```{r, message=FALSE}
+
+#numeric
+variable[,-c(1,2,6)] %>%
+    apply(2, as.numeric) %>%
+    as.data.frame() %>%
+    {.} -> variable_num
+
+#categorical
+variable$preF=factor(variable$PREDDEG,levels=c(0,1,2,3,4),labels = c("NotClassified","Certificate-degree","Associate's-degree","Bachelor's-degree","Graduate-degree"))
+variable$ctrF=factor(variable$CONTROL,levels = c(1,2,3),labels = c("Public","Private Nonprofit","Private For-profit"))
+
+#combine
+data_new <- cbind(variable_num,variable$preF, variable$ctrF,variable$STABBR)
+```
+
+Deal with missing values
+
+```{r, message=FALSE}
+
+for (i in 1:length(data_new$NPT4_PRIV)) {
+   if (is.na(data_new$NPT4_PRIV[i])) {
+       data_new$NPT4_PRIV[i] = data_new$NPT4_PUB[i]
+   }
+}
+mean(is.na(data_new$NPT4_PRIV))
+data_new = data_new[,-9]
+colnames(data_new)[9] <- "NPT4_COMBINE"
+```
+
+Deal with missing values
+```{r, message=FALSE}
+
+data_clean = data_new[,colMeans(is.na(data_new)) < 0.25]
+write.csv(data_clean,"data_clean.csv")
+```
+
+#Variable summary
+For numeric variabls, we calculate the basic summary statistics such as median, min, max, mode and quantile.
+We also use visualizations such hisgroms and boxplots to understand the distribution of the values.
+For categorical variables, we use frequency table to understand the frequcies of each categories.Box chart is used as a visualization tool.
+
+| --------------  | ----------------- | --------------------------------- | -------- |
+| NPT4_COMBINE    | Numeric           |  Continuous                       |  Cost    |
+| CDR3             | Numeric           |  Continuous                       | n        |
+| variable$qreF    | Factor            |  Categorical                      |  n       |
+| variable$ctrF    | Factor            |  Categorical                      |   n      |
+| variable$STABBR  | Factor            |  Categorical                      |   n      |

### 2017-02-12 Draft-xuezhou.Rmd
xiang

@@
 library(tidyr)
+library(plyr)
 
@@
                   WDRAW_DEBT_MDN,
+                  APPL_SCH_PCT_GE2,
                   C150_4,
                   MD_EARN_WNE_P6,
+                  MD_EARN_WNE_P10,
                   RPY_3YR_RT_SUPP,
                   CDR3)
+
 ```
@@
 data_new <- cbind(variable_num,variable$preF, variable$ctrF,variable$STABBR)
+
 ```
@@
 
+| Variable Name    | Data Type         | Categorical or Continuous         | Group       |
 | --------------  | ----------------- | --------------------------------- | -------- |
@@
 | variable$STABBR  | Factor            |  Categorical                      |   n      |
+| DEP_STAT_PCT_IND | Numeric           |  Continuous                       |     Student    |
+| PAR_ED_PCT_1STGEN| Numeric           |  Continuous                       |      Student   |
+| MD_FAMINC        | Numeric           |  Continuous                       |       Student  |
+| PCTPELL          | Numeric           |  Continuous                       |Aid         |
+| PCTFLOAN         | Numeric           |  Continuous                       | Aid        |
+| GRAD_DEBT_MDN    | Numeric           |  Continuous                       |  Aid      |
+| APPL_SCH_PCT_GE2 | Numeric           |  Continuous                       | Student        |
+| WDRAW_DEBT_MDN   | Numeric           |  Continuous                       |    Aid     |
+| MD_EARN_WNE_P6   | Numeric           |  Continuous                       |     Earning    |
+| MD_EARN_WNE_P10  | Numeric           |  Continuous                       |       Earning  |
+| RPY_3YR_RT_SUPP  | Numeric           |  Continuous                       |      Repayment   |
+| CDR3             | Numeric           |  Continuous                       | Repayment        |
+| variable$qreF    | Factor            |  Categorical                      |  School       |
+| variable$ctrF    | Factor            |  Categorical                      |   School      |
+| variable$STABBR  | Factor            |  Categorical                      |   School      |
+
+##School
+
+###1. State
+```{r, message=FALSE}
+state_freq = count(data_clean,"variable$STABBR")[order(-count(data_clean,'variable$STABBR')$freq),]
+state_freq[1:15,]
+data_clean$STABBR1 <- with(data_clean, reorder(variable$STABBR, variable$STABBR, function(x) -length(x)))
+data_clean %>% ggplot(aes(x= STABBR1)) +
+  geom_histogram(stat = "count") 
+```
+
+###2. Predominant Degree
+```{r, message=FALSE}
+prop.table(table(data_clean$`variable$preF`)) * 100
+data_clean %>% ggplot(aes(x=variable$`preF`, fill = 'blue')) + 
+  geom_histogram(stat = "count") +
+  xlab("Degree") 
+```
+
+###3. Control of Institution
+
+```{r, message=FALSE}
+prop.table(table(data_clean$`variable$ctrF`)) * 100
+data_clean %>% ggplot(aes(x=variable$`ctrF`, fill = 'blue')) + 
+  geom_histogram(stat = "count") +
+  xlab("Degree")
+```
+
+##Student
+###3. How many application do students submit?
+```{r, message=FALSE}
+college_app_2_clean = data_clean$APPL_SCH_PCT_GE2[!is.na(data_clean$APPL_SCH_PCT_GE2)]*100
+summary(college_app_2_clean)
+
+college_app_2_clean %>% as.data.frame() %>%
+    mutate(college_app_1 = 100-college_app_2_clean)%>%
+    {.}  -> college_app
+colnames(college_app)[1] <- "two_or_greater"
+
+college_app %>% gather(key = num_app, value = perc_student) %>%
+{.} -> college_app
+
+college_app %>%
+    ggplot(aes(factor(num_app), perc_student)) +
+    geom_boxplot() 
+```
+
+##Cost
+###1.Net Price
+
+##Aid
+
+###1. The median debt for students who have completed
+
+###2. The median debt for students who have not completed
+```{r, message=FALSE}
+median_debt_non_null = data_clean$WDRAW_DEBT_MDN[!is.na(data_clean$WDRAW_DEBT_MDN)]
+summary(median_debt_non_null)
+sd(median_debt_non_null)
+
+#plot a ggplot to better illustrate
+data_clean %>% ggplot(aes(x = WDRAW_DEBT_MDN)) +
+                geom_histogram(binwidth = 450) #warning?
+
+ggplot(data=data_clean,
+       aes(x=WDRAW_DEBT_MDN,
+           fill = I('red'))) +
+    geom_density(alpha=0.5) +
+    scale_x_continuous("Median Debt for Students Who Have Completed")
+```
+
+##Earning
+
+###. Earning after graduation
+```{r,message=FALSE}
+earning = select(data_clean, MD_EARN_WNE_P6, MD_EARN_WNE_P10 )
+earning_clean = na.omit(earning)
+earning_clean %>%
+    mutate(salary_increase = MD_EARN_WNE_P10-MD_EARN_WNE_P6) %>%
+    {.} -> earning_clean
+summary(earning_clean)
+apply(earning_clean, 2, sd) #dispersion increases
+
+earning_clean %>% gather(key = years_from_entry, value = salary) %>%
+    {.} -> earning_clean1
+
+earning_clean1[earning_clean1$years_from_entry != 'salary_increase',] %>% 
+    ggplot(aes(x = salary, fill =years_from_entry )) +
+    geom_density(alpha = 0.5) + 
+    xlim(0, 150000)
+```
+
+##Repayment
+
+###1. 3-year Repayment rate
+```{r,message=FALSE}
+repayment_clean = na.omit(data_clean$RPY_3YR_RT_SUPP)
+summary(repayment_clean)
+qplot(repayment_clean, geom = "histogram")
+```
+
+###2. 3-year Default rate
+```{r,message=FALSE}
+default_clean = na.omit(data_clean$CDR3)
+qplot(default_clean, geom = "histogram")
+
+summary(data_clean$RPY_3YR_RT_SUPP + data_clean$CDR3)
+```

### 2017-02-12 Draft-xuezhou.Rmd
Variable summary
integrate summary part (7 variables)

--- 2017-02-12 Draft-xuezhou.Rmd
@@
 
-```{r setup, include=FALSE,warning=FALSE}
+```{r setup, include=FALSE,warning=FALSE, message = FALSE}
 knitr::opts_chunk$set(echo = TRUE)
@@
 #combine the three datafram to generate a new dataset. 
-data_new <- cbind(variable_num,variable$preF, variable$ctrF,variable$STABBR)
+data_new <- cbind(variable_num, variable$preF, variable$ctrF,variable$STABBR)
 dim(data_new)
@@
 
+Considering the variable NPT4_PUB and NPT4_PRIV are mutually exclusive, to maximize the data integrity, we integrated the information of the two coloumns into a single cololumn, named with "NPT4_COMBINE".
+The new coloumn records the average net price for TITLE IV institutions of both public and private.
 ```{r, message=FALSE,warning=FALSE}
@@
 
-| Variable Name    | Data Type         | Categorical or Continuous         | hi       |
-| --------------  | ----------------- | --------------------------------- | -------- |
-| NPT4_COMBINE    | Numeric           |  Continuous                       |  Cost    |
-| DEP_STAT_PCT_IND | Numeric           |  Continuous                       |     n    |
-| PAR_ED_PCT_1STGEN| Numeric           |  Continuous                       |      n   |
-| MD_FAMINC        | Numeric           |  Continuous                       |       n  |
-| PCTPELL          | Numeric           |  Continuous                       |n         |
-| PCTFLOAN         | Numeric           |  Continuous                       | n        |
-| GRAD_DEBT_MDN    | Numeric           |  Continuous                       |  nn      |
+| Variable Name    | Data Type         | Description                                                         | Categoriy in Documentation       |
+| --------------  | ------------------ | ---------------------------------------------------------------------| --------------------------------|
+| NPT4_COMBINE    | Numeric            |average net price for TITLE IV institutions of both public and private.|  Cost                           |
+| DEP_STAT_PCT_IND | Numeric           |Percentage of students who are financially independent                |     Cost                        |
+| PAR_ED_PCT_1STGEN| Numeric           |Percentage first-generation students                                  |      Sudent Body                |
+| MD_FAMINC        | Numeric           |Median family income in real 2015 dollars                             |       Earning                   |
+| PCTPELL          | Numeric           |Percentage of undergraduates who receive a Pell Grant                 |        Financial Aid            |
+| PCTFLOAN         | Numeric           |Percent of all federal undergraduate students receiving a federal student loan      |  Financial Aid    |
+| GRAD_DEBT_MDN    | Numeric           |The median debt for students who have completed                       |   Financial Aid                 |
 | WDRAW_DEBT_MDN   | Numeric           |  Continuous                       |    n     |
@@
 ###5.1.1 Average net price for Title IV institutions(both public and private institutions)
-We are using the code below to summarize the continuous data
+We are using the code below to summarize the variable.
 
@@
 
-summarize(group_by(data_clean, variable$STABBR),
+summarize(data_clean,
           minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),
@@
 
-data_clean %>%
-  group_by(variable$STABBR) %>%
-    summarise(minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),    maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE))  %>%
-    arrange(medianNPT4_COMBINE)   %>%
-{.}  ->test
+```
+From the summary results, we learned that the maximum value of the variable is \$87,414 while the minimum value is about \$-19,360.
+Considering the elements are collected from Title IV institutions, we reasoned that the negative values may be due to the fact that some intituions's finainal aid supoort can cover the students' all the costs, meanwhile overbalancing the cost amount.
+There are not much differerences between mean value and median value, which is around \$15,000.
+We can conclude that the average net price costs for most institutuins are falling into the this range.
+The standard deviation value is 7604.884 which is relatively large, indicating a great spread in the data.
 
-test
-
-#Here is the quartile statistics :
-summary(data_clean$NPT4_COMBINE,na.rm=TRUE)
-quantile(data_clean$NPT4_COMBINE,na.rm=TRUE)
+Below is the histogram for this variable which is able to provide us with a visual distrubution presentation. 
+```{R}
 
@@
     labs(title="Histogram for NPT_COM") +
-    labs(x="NPT_COM", y="Count")
+    labs(x="NPT_COM", y="Count") +
+    theme(plot.title = element_text(hjust = 0.5))
+```
+From the histogram, we notice that the average net cost for IV institutions typically is 16000 -18000.
+The values between 16000 - 20000 are quite frequent.
+There were an unsually lagrge number of costs.
+The institutions with cost of more than 50000 were atypical.
+Due to the atypical large value, the histogram is abviously right skewed.
+Without those extreme values, the histogram would be reasonably symmetric.
+
+Then we grouped by the data into the state level to take a further analysis.
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(variable$STABBR) %>%
+    summarise(minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),    
+    maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE))  %>%
+       arrange(desc(medianNPT4_COMBINE))   %>%
+{.}  ->NPT4_STATE
+
+NPT4_STATE
+
+```
+Baesd on the output, we can see the highest median value for avreage net price exists in DC while the lowest median value is in AS(American Samoa).Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are all states conceiving amount of well-known educational institutions.
+
+###5.1.2 Percentage of students who are financially independent
+
+Here is the code to generate the summary statistics: 
+```{R}
+
+summarize(data_clean,
+          minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND*100 ,na.rm=TRUE),
+          maxDEP_STAT_PCT_IND=max(DEP_STAT_PCT_IND*100,na.rm=TRUE),
+          meanDEP_STAT_PCT_IND=mean(DEP_STAT_PCT_IND*100,na.rm=TRUE),
+          medianDEP_STAT_PCT_IND=median(DEP_STAT_PCT_IND*100,na.rm=TRUE),
+          sdDEP_STAT_PCT_IND = sd(DEP_STAT_PCT_IND*100,na.rm=TRUE))
+
+```
+From the summary results, we learned that the maximum value of the variable is 98.56% while the minimum value is about 0，73%.
+From the mean value, we can tell that about in most colleges, around 56% of the students are financially independent.
+The standatd deviation value of 23.3% indicates a relative high different among different colleges, which can also be concluded from the difference between maximum value and minimum value.
+Below is the code to generate a visual representation in the histgram.
+```{R}
+
+  #histogram for NPT_COM
+  ggplot(data=data_clean , aes(DEP_STAT_PCT_IND)) + 
+    geom_histogram(breaks=seq(0.1, 1, by =0.1), 
+                   col="red", 
+                   aes(fill=..count..),na.rm=TRUE)+
+    labs(title="Histogram for DEP_STAT_PCT_IND") +
+    labs(x="DEP_STAT_PCT_IND", y="Count") +
+    theme(plot.title = element_text(hjust = 0.5))
+```
+From the histogram, we notice that the value between 0.7 to 0.8 are quite frequent. There was no obvious extrme value in the sample, thus the histogram is normally distributed.
+
+Then we grouped by the data into the control type(Private For-profit, Public, Private Nonprofit) to take a further analysis.
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(variable$ctrF) %>%
+    summarise(minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND ,na.rm=TRUE),    
+    maxDEP_STAT_PCT_IND=max(DEP_STAT_PCT_IND,na.rm=TRUE),medianDEP_STAT_PCT_IND=median(DEP_STAT_PCT_IND,na.rm=TRUE))  %>%
+       arrange(desc(medianDEP_STAT_PCT_IND))   %>%
+{.}  ->DEP_STAT_PCT_IND_CONTROL
+
+DEP_STAT_PCT_IND_CONTROL
+
+```
+From the result, we can tell the highest median value is in Privately For-profit colleges, which means that in most private for-profit colleges, the students are more likely to be financial independent.
+
+###5.1.3 Percentage first-generation students
+This element identifies the share of federally aided students at the institution who report that they are first generaion on the FAFSA (Free Application for Federal Student Aid).
+First generation students are defined as sudents in which neither parents completed college.
+
+We are using the code below to summarize this variable.
+
+```{R, message=FALSE, warinig = FALSE}
+
+summarize(data_clean,
+          minPAR_ED_PCT_1STGEN=min(PAR_ED_PCT_1STGEN*100 ,na.rm=TRUE),
+          maxPAR_ED_PCT_1STGEN=max(PAR_ED_PCT_1STGEN*100,na.rm=TRUE),
+          meanPAR_ED_PCT_1STGEN=mean(PAR_ED_PCT_1STGEN*100,na.rm=TRUE),
+          medianPAR_ED_PCT_1STGEN=median(PAR_ED_PCT_1STGEN*100,na.rm=TRUE),
+          sdPAR_ED_PCT_1STGEN = sd(PAR_ED_PCT_1STGEN*100,na.rm=TRUE))
+
+```
+From the output, we can tell that the highest value is 95.25% while the lowest value is 6.87%.
+From the mean value of 48.08%, we can tell that about for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college.
+The standard deviation value of 12.2% indicated a relative high difference of this element among different colleges.
+Below is the code to generate a visual representation in the histgram.
+```{R}
+
+  ggplot(data=data_clean , aes(PAR_ED_PCT_1STGEN)) + 
+    geom_histogram(breaks=seq(0.1, 1, by =0.1), 
+                   col="red", 
+                   aes(fill=..count..),na.rm=TRUE)+
+    labs(title="Histogram for PAR_ED_PCT_1STGEN") +
+    labs(x="PAR_ED_PCT_1STGEN", y="Count") +
+    theme(plot.title = element_text(hjust = 0.5))
+```
+From the histogram, we can tell the share of first generarion students is most frequently falling into the range between 0.5 and 0.6.
+The value with the value more than 0.8 are rare.
+The colleges with the share of first generarion students value less than 0.5 are much more than the ones with the value larger than 0.5.
+Thus the histogram is seriously left skewed.
+
+Then we grouped by the data into the state level to take a further analysis.
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(variable$STABBR ) %>%
+    summarise(minPAR_ED_PCT_1STGEN=min(PAR_ED_PCT_1STGEN ,na.rm=TRUE),    
+    maxPAR_ED_PCT_1STGEN=max(PAR_ED_PCT_1STGEN,na.rm=TRUE),medianPAR_ED_PCT_1STGEN=median(PAR_ED_PCT_1STGEN,na.rm=TRUE))  %>%
+       arrange(medianPAR_ED_PCT_1STGEN)   %>%
+{.}  ->PAR_ED_PCT_1STGEN_STATE
+
+PAR_ED_PCT_1STGEN_STATE
+
+data_clean %>%
+  group_by(variable$ctrF ) %>%
+    summarise(minPAR_ED_PCT_1STGEN=min(PAR_ED_PCT_1STGEN ,na.rm=TRUE),    
+    maxPAR_ED_PCT_1STGEN=max(PAR_ED_PCT_1STGEN,na.rm=TRUE),medianPAR_ED_PCT_1STGEN=median(PAR_ED_PCT_1STGEN,na.rm=TRUE))  %>%
+       arrange(desc(medianPAR_ED_PCT_1STGEN))   %>%
+{.}  ->PAR_ED_PCT_1STGEN_STATE
+
+PAR_ED_PCT_1STGEN_STATE
+
 ```
 
+The results showed that the universities located in ND, SD, VT, MN and MT had the lowest value of Percentage first-generation students, which indicated that the students in thoses five states would have a better family education background.
+The five states with the highest share of parents first-generation are LA, NJ, MP, MV, MH.
+
+###5.1.4 Median family income in real 2015 dollars
+This elemnt is recording the median family income of students, one of the undergraduate student demographics for earnings cohorts.
+
+We are using the code below to summarize this variable.
+
+```{R, message=FALSE, warinig = FALSE}
+
+summarize(data_clean,
+          minMD_FAMINC=min(MD_FAMINC ,na.rm=TRUE),
+          maxMD_FAMINC=max(MD_FAMINC,na.rm=TRUE),
+          meanMD_FAMINC=mean(MD_FAMINC,na.rm=TRUE),
+          medianMD_FAMINC=median(MD_FAMINC,na.rm=TRUE),
+          sdMD_FAMINC = sd(MD_FAMINC,na.rm=TRUE))
+
+```
+From the output, we can tell that the highest value is 155,606.6	 while the lowest value is 0.
+From the median value of 19,262.26, we can tell that among the colleges in the sample, the students in half of them had the family income below 19,262.26	.
+The standard deviation value is 18,824.31,  indicating a relative high difference of this element among different colleges.
+Below is the code to generate a visual representation in the histgram.
+
+```{R}
+
+  ggplot(data=data_clean , aes(MD_FAMINC)) + 
+    geom_histogram(breaks=seq(0, 156000, by =3000), 
+                   col="red", 
+                   aes(fill=..count..),na.rm=TRUE)+
+    labs(title="Histogram for MD_FAMINC") +
+    labs(x="MD_FAMINC", y="Count") +
+    theme(plot.title = element_text(hjust = 0.5))
+```
+From the histogram, we are able to conclude that in most colleges, students' family income are falling into the range of 12,000 - 18,000.
+Due to the atypical large values, the histogram is obviouslu skewed to right.
+The value larger than 18,000 are tend to be extreme and the cases larger than 75000 are rare and atypical, which means rare colleges have the students with family income more than 75000.
+
+###5.1.5 Percentage of undergraduates who receive a Pell Grant
+
+This elemnt, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received Pell Grants in a given year, which is an important measure of the access an insiitution provides to low-income students.
+We are using the code below to summarize this variable.
+
+```{R, message=FALSE, warinig = FALSE}
+
+summarize(data_clean,
+          minPCTPELL=min(PCTPELL*100 ,na.rm=TRUE),
+          maxPCTPELL=max(PCTPELL*100,na.rm=TRUE),
+          meanPCTPELL=mean(PCTPELL*100,na.rm=TRUE),
+          medianPCTPELL=median(PCTPELL*100,na.rm=TRUE),
+          sdPCTPELL = sd(PCTPELL*100,na.rm=TRUE))
+
+```
+From the output, we can tell that the range is between 0 to 100%.	 Among the colleges in the sample, half colleges have this rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs.
+The standard deviation value is 22.57%,  indicating a relative high difference of this element among different colleges, which can also be noticed from the largest range (100%).
+Below is the code to generate a visual representation in the histgram.
+
+```{R}
+
+  ggplot(data=data_clean , aes(PCTPELL)) + 
+    geom_histogram(breaks=seq(0, 1, by =0.1), 
+                   col="red", 
+                   aes(fill=..count..),na.rm=TRUE)+
+    labs(title="Histogram for PCTPELL") +
+    labs(x="PCTPELL", y="Count") +
+    theme(plot.title = element_text(hjust = 0.5))
+```
+From the histogram, we are able to conclude that in most colleges, 40% to 50% of the undergraduate students are able to receive the Pell Grant.
+With little skewness, the histogram is resonalbe symmetric.
+
+Then we grouped by the data into the college type level to take a further analysis.
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(variable$ctrF ) %>%
+    summarise(minPCTPELL=min(PCTPELL ,na.rm=TRUE),    
+    maxPCTPELL=max(PCTPELL,na.rm=TRUE),medianPCTPELL=median(PCTPELL,na.rm=TRUE))  %>%
+       arrange(medianPCTPELL)   %>%
+{.}  ->PCTPELL_CONTROL
+
+PCTPELL_CONTROL
+
+```
+From the result, the Private Nonprofit institutions and Public institutions have a similar low median values while the provate For- profit institutions have a hgihest value.
+This is resonable because generally speanking, the private for-profit colleges may have much higher costs than the private nonprofit and public institutions.
+The low-income students were more likely to seek a Pell Grant for assistant.
+Additionally, in some states, state financial aid may be sufficient to cover costs at community colleges; thus the students in thoes schools may not need to seek for addtional financial aid like Pell Grant.
+
+###5.1.6 Percent of all federal undergraduate students receiving a federal student loan
+
+This elemnt, as reported in the IPEDs Student Finaicial Aid component, show the share of undergraduate students who received federal loans in a given year.
+It can provide important context to figures related to debt, repayment and non-repayment.
+We are using the code below to summarize this variable.
+
+```{R, message=FALSE, warinig = FALSE}
+
+summarize(data_clean,
+          minPCTFLOAN=min(PCTFLOAN*100 ,na.rm=TRUE),
+          maxPCTFLOAN=max(PCTFLOAN*100,na.rm=TRUE),
+          meanPCTFLOAN=mean(PCTFLOAN*100,na.rm=TRUE),
+          medianPCTFLOAN=median(PCTFLOAN*100,na.rm=TRUE),
+          sdPCTFLOAN = sd(PCTFLOAN*100,na.rm=TRUE))
+
+```
+From the output, we can tell that the range is between 0 to 100%.	 Among the colleges in the sample, half colleges have this rate as high as 60.08% , which means around 60% of undergraduare students are able to receive the fedaral student loans.
+The standard deviation value is 29.26%,  indicating a relative high difference of this element among different colleges, which can also be noticed from the largest range (100%).
+Below is the code to generate a visual representation in the histgram.
+
+```{R}
+
+  ggplot(data=data_clean , aes(PCTFLOAN)) + 
+    geom_histogram(breaks=seq(0, 1, by =0.1), 
+                   col="red", 
+                   aes(fill=..count..),na.rm=TRUE)+
+    labs(title="Histogram for PCTFLOAN") +
+    labs(x="PCTFLOAN", y="Count") +
+    theme(plot.title = element_text(hjust = 0.5))
+```
+From the histogram, it's obvious that there are two peaks in the graph. The rates between 0- 10% and the rates between 60% to 80% are similarly frequent compated to the other ranges.
+
+Then we grouped by the data into the college type level to take a further analysis.
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(variable$STABBR ) %>%
+    summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
+    maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
+       arrange(medianPCTFLOAN)   %>%
+{.}  ->PCTFLOAN_STATE
+
+PCTFLOAN_STATE
+
+data_clean %>%
+  group_by(variable$ctrF ) %>%
+    summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
+    maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
+       arrange(medianPCTFLOAN)   %>%
+{.}  ->PCTFLOAN_CONTROL
+
+PCTFLOAN_CONTROL
+
+```
+From the result, the Private institutions have a much larger value than the public institutions.
+With a much higher costs burden, it makes sence that the low-income sutdents in private institutions are much more likely to seek financial aid than the public institution students.
+
+###5.1.7 The median debt for students who have completed
+GRAD_DEBT_MDN
+
+This element is the median loan debt acculated at the institution by all student borrowers of federal lonans who have habe completed the education.We are using the code below to summarize this variable.
+```{R, message=FALSE, warinig = FALSE}
+
+summarize(data_clean,
+          minGRAD_DEBT_MDN=min(GRAD_DEBT_MDN ,na.rm=TRUE),
+          maxGRAD_DEBT_MDN=max(GRAD_DEBT_MDN,na.rm=TRUE),
+          meanGRAD_DEBT_MDN=mean(GRAD_DEBT_MDN,na.rm=TRUE),
+          medianGRAD_DEBT_MDN=median(GRAD_DEBT_MDN,na.rm=TRUE),
+          sdGRAD_DEBT_MDN = sd(GRAD_DEBT_MDN,na.rm=TRUE))
+
+```
+From the output, we can tell that the highest value is 48,250	 while the lowest value is 1448.
+From the median value of 14000, we can tell that among the colleges in the sample, the students who have completed the deducaiton from half of them had the median loan debt of 14000.
+The standard deviation value is 8020.958,  indicating a relative high difference of this element among different colleges.
+Below is the code to generate a visual representation in the histgram.
+
+```{r}
+
+ ggplot(data=data_clean , aes(GRAD_DEBT_MDN)) + 
+    geom_histogram(breaks=seq(0, 50000, by =2000), 
+                   col="red", 
+                   aes(fill=..count..),na.rm=TRUE)+
+    labs(title="Histogram for GRAD_DEBT_MDN") +
+    labs(x="GRAD_DEBT_MDN", y="Count") +
+    theme(plot.title = element_text(hjust = 0.5))
+
+```
+From the histogram, it is obvious that most students in the sampled colleges still have 8000- 10000 debt amount after they completed the education.
+This amount is related to the federal loan amount provided by each institution.
+Additionally, it is asscociated with students' ability to repay the loan during the acdemic period.
+For example, for the students who were capabiblity of working through the acedemic years or securing jobs before graduation may have a less value for this variable.
+
 ##5.2 Relationship between variables
+PCTPELL vs family income 
+PCTFLOAN vs family income
+PCTPELL vs NPT4
+PCTPELL  VS family independent
+family independent   vs family income 
+first generation vs family income
+pell vs control
+pell vs degree type
+pell vs states
+loan vs debt, repayment, non-repay
+loan vs cost
+
+loan vs control
+
 ##5.3 Comparison between variables
+PCTPELL vs PCTFLOAN
 

### 2017-02-13 Draft-xuezhou-sjb.Rmd
integreted version

--- 2017-02-12 Draft-xuezhou.Rmd
@@
 
-# 1. Introduction
-# 2. Objective
-# 3. Dataset Preparation
+#1. Introduction
+#2. Objective
+#3. Dataset Preparation
 Our analysis focuses on the data of year 2012 -2013 of 23 variables.
@@
 library(tidyr)
+library(ggvis)
 
@@
 
-#Variable summary
+#4. Goals
+#5. Summary of variables
+
 For numeric variabls, we calculate the basic summary statistics such as median, min, max, mode and quantile.
@@
 | MD_EARN_WNE_P6   | Numeric           |  Continuous                       |     n    |
-| MN_EARN_WNE_P10  | Numeric           |  Continuous                       |       n  |
+| MD_EARN_WNE_P10  | Numeric           |  Continuous                       |       n  |
 | RPY_3YR_RT_SUPP  | Numeric           |  Continuous                       |      n   |
@@
 | variable$STABBR  | Factor            |  Categorical                      |   n      |
+| APPL_SCH_PCT_GE2 | Numeric           |  Continuous                       | Student        |
 
@@
 
-#4 Goals
-#5 Summary of variables
 ##5.1 Individual variable summary
 
-###5.1.1 Average net price for Title IV institutions(both public and private institutions)
-We are using the code below to summarize the variable.
+###5.1.1 School
+####5.1.1.1. State
+State variable has 59 levels, which include 55 states in the United States, as well as.........
+From the following frequncy table, we see that California has the most University-equivalent institutions across the nation, followed by Texas and Ner York.
+It has about 800 universities!
+A bar chart is plotted below to show the top states with the most universities.
+Then summary of frequency table also shows that the state having the least university only has one, and 50% of state has more than 132 universities in the state.
+```{r, message=FALSE}
 
-```{R, message=FALSE, warinig = FALSE}
-
-summarize(data_clean,
-          minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),
-          maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),
-          meanNPT4_COMBINE=mean(NPT4_COMBINE,na.rm=TRUE),
-          medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE),
-          sdNPT4_COMBINE = sd(NPT4_COMBINE,na.rm=TRUE))
+state_freq = count(data_clean,"variable$STABBR")[order(-count(data_clean,'variable$STABBR')$freq),]
+state_freq[1:15,]
+summary(state_freq)
+state_freq[1:15,] %>% ggplot(aes(x=factor(variable.STABBR), y=freq)) +
+  geom_bar(stat = 'identity')
 
 ```
-From the summary results, we learned that the maximum value of the variable is \$87,414 while the minimum value is about \$-19,360.
-Considering the elements are collected from Title IV institutions, we reasoned that the negative values may be due to the fact that some intituions's finainal aid supoort can cover the students' all the costs, meanwhile overbalancing the cost amount.
-There are not much differerences between mean value and median value, which is around \$15,000.
-We can conclude that the average net price costs for most institutuins are falling into the this range.
-The standard deviation value is 7604.884 which is relatively large, indicating a great spread in the data.
 
-Below is the histogram for this variable which is able to provide us with a visual distrubution presentation. 
-```{R}
+####5.1.1.2. Predominant Degree
+Then we analyze the type of degree universities mainly offer.
+There are four types of degree taken into accout: Certificate-degree, Associate's degree, Bachelor's degree, and Graduate Degree.
+About 5% of the population are classified in terms of the predomiant degree they offer.
+43.2% offer Certificate degree as their predominant degree, 27.3% predominantly offer Bachelor degree and only 4.23% are classified as Graduate degree predominant.
 
-  #histogram for NPT_COM
-  ggplot(data=data_clean , aes(NPT4_COMBINE)) + 
-    geom_histogram(breaks=seq(0, 90000, by =2000), 
-                   col="red", 
-                   aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for NPT_COM") +
-    labs(x="NPT_COM", y="Count") +
-    theme(plot.title = element_text(hjust = 0.5))
+```{r, message=FALSE}
+prop.table(table(data_clean$`variable$preF`)) * 100
+data_clean %>% ggplot(aes(x=variable$`preF`, fill = 'blue')) + 
+  geom_histogram(stat = "count") +
+  xlab("Degree") 
 ```
-From the histogram, we notice that the average net cost for IV institutions typically is 16000 -18000.
-The values between 16000 - 20000 are quite frequent.
-There were an unsually lagrge number of costs.
-The institutions with cost of more than 50000 were atypical.
-Due to the atypical large value, the histogram is abviously right skewed.
-Without those extreme values, the histogram would be reasonably symmetric.
 
-Then we grouped by the data into the state level to take a further analysis.
-```{R message=FALSE, warinig = FALSE}
+####5.1.1.3. Control of Institution
+In addition to State and predominant degree, control of institution is another categorical variable in the dataset.
+The label of Control is to indicate whether or not a institution is institution is privately owned and whether or not it is a for-profit organization.
+There are three levels of this variable: Public, Private Nonprofit, and Private-profit.
+Public school is assumed to be non-profit so it was not further divided into two subgroups.
+Below we show the percentage summary of each category and a bar char is plotted to facilitate understanding.
+About 50% out of all institution are private for-profit university, which is the *most dominant type*.
+Public and Private Nonprofit are pretty tie in terms of number of observation in each category.
 
-data_clean %>%
-  group_by(variable$STABBR) %>%
-    summarise(minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),    
-    maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE))  %>%
-       arrange(desc(medianNPT4_COMBINE))   %>%
-{.}  ->NPT4_STATE
+```{r, message=FALSE}
+prop.table(table(data_clean$`variable$ctrF`)) * 100
+data_clean %>% ggplot(aes(x=variable$`ctrF`, fill = 'blue')) + 
+  geom_histogram(stat = "count") +
+  xlab("Degree")
+```
 
-NPT4_STATE
-
-```
-Baesd on the output, we can see the highest median value for avreage net price exists in DC while the lowest median value is in AS(American Samoa).Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are all states conceiving amount of well-known educational institutions.
-
-###5.1.2 Percentage of students who are financially independent
-
-Here is the code to generate the summary statistics: 
-```{R}
-
-summarize(data_clean,
-          minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND*100 ,na.rm=TRUE),
-          maxDEP_STAT_PCT_IND=max(DEP_STAT_PCT_IND*100,na.rm=TRUE),
-          meanDEP_STAT_PCT_IND=mean(DEP_STAT_PCT_IND*100,na.rm=TRUE),
-          medianDEP_STAT_PCT_IND=median(DEP_STAT_PCT_IND*100,na.rm=TRUE),
-          sdDEP_STAT_PCT_IND = sd(DEP_STAT_PCT_IND*100,na.rm=TRUE))
-
-```
-From the summary results, we learned that the maximum value of the variable is 98.56% while the minimum value is about 0，73%.
-From the mean value, we can tell that about in most colleges, around 56% of the students are financially independent.
-The standatd deviation value of 23.3% indicates a relative high different among different colleges, which can also be concluded from the difference between maximum value and minimum value.
-Below is the code to generate a visual representation in the histgram.
-```{R}
-
-  #histogram for NPT_COM
-  ggplot(data=data_clean , aes(DEP_STAT_PCT_IND)) + 
-    geom_histogram(breaks=seq(0.1, 1, by =0.1), 
-                   col="red", 
-                   aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for DEP_STAT_PCT_IND") +
-    labs(x="DEP_STAT_PCT_IND", y="Count") +
-    theme(plot.title = element_text(hjust = 0.5))
-```
-From the histogram, we notice that the value between 0.7 to 0.8 are quite frequent. There was no obvious extrme value in the sample, thus the histogram is normally distributed.
-
-Then we grouped by the data into the control type(Private For-profit, Public, Private Nonprofit) to take a further analysis.
-```{R message=FALSE, warinig = FALSE}
-
-data_clean %>%
-  group_by(variable$ctrF) %>%
-    summarise(minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND ,na.rm=TRUE),    
-    maxDEP_STAT_PCT_IND=max(DEP_STAT_PCT_IND,na.rm=TRUE),medianDEP_STAT_PCT_IND=median(DEP_STAT_PCT_IND,na.rm=TRUE))  %>%
-       arrange(desc(medianDEP_STAT_PCT_IND))   %>%
-{.}  ->DEP_STAT_PCT_IND_CONTROL
-
-DEP_STAT_PCT_IND_CONTROL
-
-```
-From the result, we can tell the highest median value is in Privately For-profit colleges, which means that in most private for-profit colleges, the students are more likely to be financial independent.
-
-###5.1.3 Percentage first-generation students
+###5.1.2 Student Body
+####5.1.2.1 Percentage first-generation students
 This element identifies the share of federally aided students at the institution who report that they are first generaion on the FAFSA (Free Application for Federal Student Aid).
@@
 
-###5.1.4 Median family income in real 2015 dollars
-This elemnt is recording the median family income of students, one of the undergraduate student demographics for earnings cohorts.
+####5.1.2.2 Number of applications
+The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who list two and more than two institutions when he apply for financial aid when apply for universities.
+This information is only available to students who submit their applications for federal aid, and not cover the whole student body.
+Nevertheless, this variable still can provides some useful insights about how many universities a student usually apply to.
 
-We are using the code below to summarize this variable.
+The summry statistics show that median percentage of student submitting two or more than two application is 41.67 across all univerisities.
+The institution having the least percentage only has 4% of student applying to more than one university, while the school with the highest percent has as high as 90% of student applying to more than one university.
+The boxplot compare the the percent of student applying to only one university and that of student applying to more than 1.
+We can clearly see that a larger portion of students only list one instituiton when submitting their application for financial aid.
 
+```{r, message=FALSE}
+college_app_2_clean = data_clean$APPL_SCH_PCT_GE2[!is.na(data_clean$APPL_SCH_PCT_GE2)]*100
+summary(college_app_2_clean)
+
+college_app_2_clean %>% as.data.frame() %>%
+    mutate(college_app_1 = 100-college_app_2_clean)%>%
+    {.}  -> college_app
+colnames(college_app)[1] <- "two_or_greater"
+
+college_app %>% gather(key = num_app, value = perc_student) %>%
+{.} -> college_app
+
+college_app %>%
+    ggplot(aes(factor(num_app), perc_student)) +
+    geom_boxplot() 
+```
+
+###5.1.3 Cost
+####5.1.3.1 Average net price for Title IV institutions(both public and private institutions)
+
+We are using the code below to summarize the variable. 
 ```{R, message=FALSE, warinig = FALSE}
@@
 summarize(data_clean,
-          minMD_FAMINC=min(MD_FAMINC ,na.rm=TRUE),
-          maxMD_FAMINC=max(MD_FAMINC,na.rm=TRUE),
-          meanMD_FAMINC=mean(MD_FAMINC,na.rm=TRUE),
-          medianMD_FAMINC=median(MD_FAMINC,na.rm=TRUE),
-          sdMD_FAMINC = sd(MD_FAMINC,na.rm=TRUE))
+          minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),
+          maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),
+          meanNPT4_COMBINE=mean(NPT4_COMBINE,na.rm=TRUE),
+          medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE),
+          sdNPT4_COMBINE = sd(NPT4_COMBINE,na.rm=TRUE))
 
 ```
-From the output, we can tell that the highest value is 155,606.6	 while the lowest value is 0.
-From the median value of 19,262.26, we can tell that among the colleges in the sample, the students in half of them had the family income below 19,262.26	.
-The standard deviation value is 18,824.31,  indicating a relative high difference of this element among different colleges.
-Below is the code to generate a visual representation in the histgram.
+From the summary results, we learned that the maximum value of the variable is \$87,414 while the minimum value is about \$-19,360.
+Considering the elements are collected from Title IV institutions, we reasoned that the negative values may be due to the fact that some intituions's finainal aid supoort can cover the students' all the costs, meanwhile overbalancing the cost amount.
+There are not much differerences between mean value and median value, which is around \$15,000.
+We can conclude that the average net price costs for most institutuins are falling into the this range.
+The standard deviation value is 7604.884 which is relatively large, indicating a great spread in the data.
 
+Below is the histogram for this variable which is able to provide us with a visual distrubution presentation. 
 ```{R}
 
-  ggplot(data=data_clean , aes(MD_FAMINC)) + 
-    geom_histogram(breaks=seq(0, 156000, by =3000), 
+  #histogram for NPT_COM
+  ggplot(data=data_clean , aes(NPT4_COMBINE)) + 
+    geom_histogram(breaks=seq(0, 90000, by =2000), 
                    col="red", 
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for MD_FAMINC") +
-    labs(x="MD_FAMINC", y="Count") +
+    labs(title="Histogram for NPT_COM") +
+    labs(x="NPT_COM", y="Count") +
     theme(plot.title = element_text(hjust = 0.5))
 ```
-From the histogram, we are able to conclude that in most colleges, students' family income are falling into the range of 12,000 - 18,000.
-Due to the atypical large values, the histogram is obviouslu skewed to right.
-The value larger than 18,000 are tend to be extreme and the cases larger than 75000 are rare and atypical, which means rare colleges have the students with family income more than 75000.
+From the histogram, we notice that the average net cost for IV institutions typically is 16000 -18000.
+The values between 16000 - 20000 are quite frequent.
+There were an unsually lagrge number of costs.
+The institutions with cost of more than 50000 were atypical.
+Due to the atypical large value, the histogram is abviously right skewed.
+Without those extreme values, the histogram would be reasonably symmetric.
 
-###5.1.5 Percentage of undergraduates who receive a Pell Grant
+Then we grouped by the data into the state level to take a further analysis.
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(variable$STABBR) %>%
+    summarise(minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),    
+    maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE))  %>%
+       arrange(desc(medianNPT4_COMBINE))   %>%
+{.}  ->NPT4_STATE
+
+NPT4_STATE
+
+```
+
+below is the code to generate stacked hisogram
+``` {r}
+data_clean %>%
+  group_by(variable$STABBR, na.rm=TRUE) %>%
+  ggvis(x=~NPT4_COMBINE,na.rm=TRUE) %>%
+  layer_histograms(width=5, stack=TRUE, fill=~variable$STABBR,na.rm=TRUE)
+
+data_clean %>%
+  ggplot(aes(x=NPT4_COMBINE)) +
+  geom_bar(binwidth=50,aes(fill=variable$ctrF,na.rm=TRUE))
+
+data_clean %>%
+  ggplot(aes(x=NPT4_COMBINE,na.rm=TRUE)) +
+  geom_bar(binwidth=3000,aes(fill=variable$ctrF,na.rm=TRUE),breaks=seq(-20000, 90000, by =3000))
+
+```
+Baesd on the output, we can see the highest median value for avreage net price exists in DC while the lowest median value is in AS(American Samoa).Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are all states conceiving amount of well-known educational institutions.
+
+data_clean %>%
+  ggplot(aes(x=NPT4_COMBINE,na.rm=TRUE)) +
+  geom_bar(binwidth=50,aes(fill=variable$ctrF,na.rm=TRUE),breaks=seq(-20000, 90000, by =50))
+
+####5.1.3.2 Percentage of students who are financially independent
+
+Here is the code to generate the summary statistics: 
+```{R}
+
+summarize(data_clean,
+          minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND*100 ,na.rm=TRUE),
+          maxDEP_STAT_PCT_IND=max(DEP_STAT_PCT_IND*100,na.rm=TRUE),
+          meanDEP_STAT_PCT_IND=mean(DEP_STAT_PCT_IND*100,na.rm=TRUE),
+          medianDEP_STAT_PCT_IND=median(DEP_STAT_PCT_IND*100,na.rm=TRUE),
+          sdDEP_STAT_PCT_IND = sd(DEP_STAT_PCT_IND*100,na.rm=TRUE))
+
+```
+From the summary results, we learned that the maximum value of the variable is 98.56% while the minimum value is about 0，73%.
+From the mean value, we can tell that about in most colleges, around 56% of the students are financially independent.
+The standatd deviation value of 23.3% indicates a relative high different among different colleges, which can also be concluded from the difference between maximum value and minimum value.
+Below is the code to generate a visual representation in the histgram.
+```{R}
+
+  #histogram for NPT_COM
+  ggplot(data=data_clean , aes(DEP_STAT_PCT_IND)) + 
+    geom_histogram(breaks=seq(0.1, 1, by =0.1), 
+                   col="red", 
+                   aes(fill=..count..),na.rm=TRUE)+
+    labs(title="Histogram for DEP_STAT_PCT_IND") +
+    labs(x="DEP_STAT_PCT_IND", y="Count") +
+    theme(plot.title = element_text(hjust = 0.5))
+```
+From the histogram, we notice that the value between 0.7 to 0.8 are quite frequent. There was no obvious extrme value in the sample, thus the histogram is normally distributed.
+
+Then we grouped by the data into the control type(Private For-profit, Public, Private Nonprofit) to take a further analysis.
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(variable$ctrF) %>%
+    summarise(minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND ,na.rm=TRUE),    
+    maxDEP_STAT_PCT_IND=max(DEP_STAT_PCT_IND,na.rm=TRUE),medianDEP_STAT_PCT_IND=median(DEP_STAT_PCT_IND,na.rm=TRUE))  %>%
+       arrange(desc(medianDEP_STAT_PCT_IND))   %>%
+{.}  ->DEP_STAT_PCT_IND_CONTROL
+
+DEP_STAT_PCT_IND_CONTROL
+
+```
+From the result, we can tell the highest median value is in Privately For-profit colleges, which means that in most private for-profit colleges, the students are more likely to be financial independent.
+
+###5.1.4 Financial Aid 
+####5.1.4.1 Percentage of undergraduates who receive a Pell Grant
 
@@
 
-###5.1.6 Percent of all federal undergraduate students receiving a federal student loan
+####5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan
 
@@
 
+#stacked histogram
+data_clean %>%
+  ggplot(aes(x=PCTFLOAN)) +
+  geom_bar(binwidth=5,aes(fill=variable$ctrF ))
+
 ```
@@
 
-###5.1.7 The median debt for students who have completed
+####5.1.4.3 The median debt for students who have completed
 GRAD_DEBT_MDN
@@
 
-##5.2 Relationship between variables
+####5.1.4.4 The median debt for students who have not completed
+WDRAW_DEBT_MDN describe the median number of debt fot students who ahve not completed their study.
+In other words, they are still attending the university.
+The summary statistics show that student in 50% of institution have a median debt of 5500.The median debt for attending students can be as much as 27620 for some school and as low as 600 for other.
+So we also calculate standard deviation to measure the dispenser of the distribution.
+The standard deviation is 2415.623 about 95% of institution have a median debt ranging from 1167.754  to 10830.246.
+The histogram that the distribution of the median roughly follows a standard distribution with a sharp peak and slightly heavy tail.
+```{r, message=FALSE}
+median_debt_non_null = data_clean$WDRAW_DEBT_MDN[!is.na(data_clean$WDRAW_DEBT_MDN)]
+summary(median_debt_non_null)
+sd(median_debt_non_null)
+
+#plot a ggplot to better illustrate
+data_clean %>% ggplot(aes(x = WDRAW_DEBT_MDN)) +
+                geom_histogram(binwidth = 450) #warning?
+
+ggplot(data=data_clean,
+       aes(x=WDRAW_DEBT_MDN,
+           fill = I('red'))) +
+    geom_density(alpha=0.5) +
+    scale_x_continuous("Median Debt for Students Who Have Completed")
+
+```
+
+###5.1.5 Earning
+####5.1.5.1 Median family income in real 2015 dollars
+This elemnt is recording the median family income of students, one of the undergraduate student demographics for earnings cohorts.
+
+We are using the code below to summarize this variable.
+
+```{R, message=FALSE, warinig = FALSE}
+
+summarize(data_clean,
+          minMD_FAMINC=min(MD_FAMINC ,na.rm=TRUE),
+          maxMD_FAMINC=max(MD_FAMINC,na.rm=TRUE),
+          meanMD_FAMINC=mean(MD_FAMINC,na.rm=TRUE),
+          medianMD_FAMINC=median(MD_FAMINC,na.rm=TRUE),
+          sdMD_FAMINC = sd(MD_FAMINC,na.rm=TRUE))
+
+```
+From the output, we can tell that the highest value is 155,606.6	 while the lowest value is 0.
+From the median value of 19,262.26, we can tell that among the colleges in the sample, the students in half of them had the family income below 19,262.26	.
+The standard deviation value is 18,824.31,  indicating a relative high difference of this element among different colleges.
+Below is the code to generate a visual representation in the histgram.
+
+```{R}
+
+  ggplot(data=data_clean , aes(MD_FAMINC)) + 
+    geom_histogram(breaks=seq(0, 156000, by =3000), 
+                   col="red", 
+                   aes(fill=..count..),na.rm=TRUE)+
+    labs(title="Histogram for MD_FAMINC") +
+    labs(x="MD_FAMINC", y="Count") +
+    theme(plot.title = element_text(hjust = 0.5))
+```
+From the histogram, we are able to conclude that in most colleges, students' family income are falling into the range of 12,000 - 18,000.
+Due to the atypical large values, the histogram is obviouslu skewed to right.
+The value larger than 18,000 are tend to be extreme and the cases larger than 75000 are rare and atypical, which means rare colleges have the students with family income more than 75000.
+
+####5.1.5.2 Earning after graduation
+We selected two variables MD_EARN_WNE_P6, MD_EARN_WNE_P10 in the earning group to study students' post-graduation earning.
+MD_EARN_WNE_P6 is the median earning for students six years after their entry, which corresponds to two years after gradutaion.
+Likewise, MD_EARN_WNE_P10 is the median earning ten years after their entry, which corresponds to four years after graduation.
+These two metrics give us an idea of how much money students are making after their graduation, in both short and medium terms,
+
+In addition to these two variables, a derived variable is also calculated as the difference between the salary 2 years post-graduation and another two years later.
+The median earning of 50% institutions is below the 25300 below, while the maximum median earining is 120200.
+After two years, the median earning increase by about $5500.
+The histogram gives a more clear idea of the earning distribution.
+Most of institution with students having a median earining between 10000 to 50000 two years after graduation.
+The distribution of earining 4 years after graduation falls at the right of the distribution of 2 years post-graduation, which means that the earning increases as students' experience increase.
+Beside, the fout-year post-graduation earning has a wider spread than that of two years.
+That indicate that the there is a greater gap between who are earning more and who are earning less, which can be confirmed by the increase in standard deviation.
+
+```{r,message=FALSE}
+earning = select(data_clean, MD_EARN_WNE_P6, MD_EARN_WNE_P10 )
+earning_clean = na.omit(earning)
+earning_clean %>%
+    mutate(salary_increase = MD_EARN_WNE_P10-MD_EARN_WNE_P6) %>%
+    {.} -> earning_clean
+summary(earning_clean)
+apply(earning_clean, 2, sd) #dispersion increases
+
+earning_clean %>% gather(key = years_from_entry, value = salary) %>%
+    {.} -> earning_clean1
+
+earning_clean1[earning_clean1$years_from_entry != 'salary_increase',] %>% 
+    ggplot(aes(x = salary, fill =years_from_entry )) +
+    geom_density(alpha = 0.5) + 
+    xlim(0, 150000)
+
+```
+
+###5.1.6 Repayment 
+####5.1.6.1 3-year Repayment rate
+The variable RPY_3YR_RT_SUPP measure the percentage of student who entered repayemtn from October 1, 2010 to September 30, 2011, and made progress in the following three years.
+Repayment rates are generally considered most sensitive than default rates(we study later), which measure only the worse-case scenario for repayment outcomes.
+From the summary statistics, we see that 50% institutions have 40% of students who are making progress in their repayment.
+A histogram is also plotted to show the distribution, clearly with a heavy tail.
+
+```{r,message=FALSE}
+repayment_clean = na.omit(data_clean$RPY_3YR_RT_SUPP)
+summary(repayment_clean)
+qplot(repayment_clean, geom = "histogram")
+```
+
+####5.1.6.2 3-year Default rate
+The three-year cohort default rate describe the repayment in the opposite of repayment rate.
+In this dataset, it was calculated using the cohort of borrowers who entered repayment on their federal student loans between Octorber 1, 2010 to September 30, 2011, and who defaulted before September 2012.
+The median default rate aross all universities is 13%, which means that 50% of universities have 13 out of 100 students defaulted their loan!
+
+```{r,message=FALSE}
+default_clean = na.omit(data_clean$CDR3)
+summary(default_clean)
+qplot(default_clean, geom = "histogram")
+
+summary(data_clean$RPY_3YR_RT_SUPP + data_clean$CDR3)
+```
+
+#6. Relationship between variables
 PCTPELL vs family income 
@@
 
-##5.3 Comparison between variables
+Hello world
+
+#7. Comparison between variables
 PCTPELL vs PCTFLOAN
 
-#6. Findings and Conclusion
+#8. Findings and Conclusion

### 2017-02-13 Draft-xuezhou-sjb.Rmd
added relationship

@@
 ```{r setup, include=FALSE,warning=FALSE, message = FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
 ```
@@
 library(ggvis)
+library(plyr)
 
@@
 
-data = read.csv("MERGED2012_13_PP.csv",header = T, na.strings = "NULL")
+data = read.csv("MERGED2012_13_PP.csv",header = T, na.strings = c("NULL","PrivacySuppressed"))
 
@@
                   WDRAW_DEBT_MDN,
-                  APPL_SCH_PCT_GE5,
+                  APPL_SCH_PCT_GE2,
                   C150_4,
                   MD_EARN_WNE_P6,
-                  MN_EARN_WNE_P10,
+                  MD_EARN_WNE_P10,
                   RPY_3YR_RT_SUPP,
@@
 The standard deviation value of 12.2% indicated a relative high difference of this element among different colleges.
+
 Below is the code to generate a visual representation in the histgram. 
@@
     geom_histogram(breaks=seq(0.1, 1, by =0.1), 
-                   col="red", 
+                   col="white", 
                    aes(fill=..count..),na.rm=TRUE)+
@@
 Thus the histogram is seriously left skewed.
-
-Then we grouped by the data into the state level to take a further analysis.
-```{R message=FALSE, warinig = FALSE}
-
-data_clean %>%
-  group_by(variable$STABBR ) %>%
-    summarise(minPAR_ED_PCT_1STGEN=min(PAR_ED_PCT_1STGEN ,na.rm=TRUE),    
-    maxPAR_ED_PCT_1STGEN=max(PAR_ED_PCT_1STGEN,na.rm=TRUE),medianPAR_ED_PCT_1STGEN=median(PAR_ED_PCT_1STGEN,na.rm=TRUE))  %>%
-       arrange(medianPAR_ED_PCT_1STGEN)   %>%
-{.}  ->PAR_ED_PCT_1STGEN_STATE
-
-PAR_ED_PCT_1STGEN_STATE
-
-data_clean %>%
-  group_by(variable$ctrF ) %>%
-    summarise(minPAR_ED_PCT_1STGEN=min(PAR_ED_PCT_1STGEN ,na.rm=TRUE),    
-    maxPAR_ED_PCT_1STGEN=max(PAR_ED_PCT_1STGEN,na.rm=TRUE),medianPAR_ED_PCT_1STGEN=median(PAR_ED_PCT_1STGEN,na.rm=TRUE))  %>%
-       arrange(desc(medianPAR_ED_PCT_1STGEN))   %>%
-{.}  ->PAR_ED_PCT_1STGEN_STATE
-
-PAR_ED_PCT_1STGEN_STATE
-
-```
-
-The results showed that the universities located in ND, SD, VT, MN and MT had the lowest value of Percentage first-generation students, which indicated that the students in thoses five states would have a better family education background.
-The five states with the highest share of parents first-generation are LA, NJ, MP, MV, MH.
 
@@
 
-Then we grouped by the data into the state level to take a further analysis.
-```{R message=FALSE, warinig = FALSE}
-
-data_clean %>%
-  group_by(variable$STABBR) %>%
-    summarise(minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),    
-    maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE))  %>%
-       arrange(desc(medianNPT4_COMBINE))   %>%
-{.}  ->NPT4_STATE
-
-NPT4_STATE
-
-```
-
-below is the code to generate stacked hisogram
-``` {r}
-data_clean %>%
-  group_by(variable$STABBR, na.rm=TRUE) %>%
-  ggvis(x=~NPT4_COMBINE,na.rm=TRUE) %>%
-  layer_histograms(width=5, stack=TRUE, fill=~variable$STABBR,na.rm=TRUE)
-
-data_clean %>%
-  ggplot(aes(x=NPT4_COMBINE)) +
-  geom_bar(binwidth=50,aes(fill=variable$ctrF,na.rm=TRUE))
-
-data_clean %>%
-  ggplot(aes(x=NPT4_COMBINE,na.rm=TRUE)) +
-  geom_bar(binwidth=3000,aes(fill=variable$ctrF,na.rm=TRUE),breaks=seq(-20000, 90000, by =3000))
-
-```
-Baesd on the output, we can see the highest median value for avreage net price exists in DC while the lowest median value is in AS(American Samoa).Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are all states conceiving amount of well-known educational institutions.
-
-data_clean %>%
-  ggplot(aes(x=NPT4_COMBINE,na.rm=TRUE)) +
-  geom_bar(binwidth=50,aes(fill=variable$ctrF,na.rm=TRUE),breaks=seq(-20000, 90000, by =50))
-
 ####5.1.3.2 Percentage of students who are financially independent
@@
     geom_histogram(breaks=seq(0, 1, by =0.1), 
-                   col="red", 
+                   col="white", 
                    aes(fill=..count..),na.rm=TRUE)+
@@
 With little skewness, the histogram is resonalbe symmetric.
-
-Then we grouped by the data into the college type level to take a further analysis.
-```{R message=FALSE, warinig = FALSE}
-
-data_clean %>%
-  group_by(variable$ctrF ) %>%
-    summarise(minPCTPELL=min(PCTPELL ,na.rm=TRUE),    
-    maxPCTPELL=max(PCTPELL,na.rm=TRUE),medianPCTPELL=median(PCTPELL,na.rm=TRUE))  %>%
-       arrange(medianPCTPELL)   %>%
-{.}  ->PCTPELL_CONTROL
-
-PCTPELL_CONTROL
-
-```
-From the result, the Private Nonprofit institutions and Public institutions have a similar low median values while the provate For- profit institutions have a hgihest value.
-This is resonable because generally speanking, the private for-profit colleges may have much higher costs than the private nonprofit and public institutions.
-The low-income students were more likely to seek a Pell Grant for assistant.
-Additionally, in some states, state financial aid may be sufficient to cover costs at community colleges; thus the students in thoes schools may not need to seek for addtional financial aid like Pell Grant.
 
@@
   group_by(variable$ctrF ) %>%
-    summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
+    dplsummarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
     maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
@@
 ```
+
 From the result, the Private institutions have a much larger value than the public institutions.
@@
 ####5.1.4.3 The median debt for students who have completed
-GRAD_DEBT_MDN
 
@@
 #6. Relationship between variables
-PCTPELL vs family income 
-PCTFLOAN vs family income
-PCTPELL vs NPT4
-PCTPELL  VS family independent
-family independent   vs family income 
-first generation vs family income
-pell vs control
-pell vs degree type
-pell vs states
-loan vs debt, repayment, non-repay
-loan vs cost
+#6.1 NPT4_COMBINE VS State
 
-loan vs control
+To analyze the relationship between average net cost for Title IV institutions and State, we grouped the data into the state level to take a further look.
+```{R message=FALSE, warinig = FALSE}
 
-Hello world
+data_clean %>%
+  group_by(variable$STABBR) %>%
+    dplyr::summarise(minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),    
+    maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE))  %>%
+       arrange(desc(medianNPT4_COMBINE))   %>%
+{.}  ->NPT4_STATE
+
+NPT4_STATE
+
+```
+Baesd on the output, we can see the highest median value for avreage net price exists in DC while the lowest median value is in AS(American Samoa).Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are all states conceiving amount of well-known educational institutions.
+
+#6.2 Debt VS Cost
+We assume there would be a strong linear relationship between the average net price for Title IV institutions and the median debt for students who have completed.
+We hypothese that the students who completed education in colleges with high costs will have a highe value of median debt.
+We use the following code to generate the scatterplot between these two variables to verify our assumptions.
+
+```{R}
+
+ggplot(data_clean, aes(x =NPT4_COMBINE, y = GRAD_DEBT_MDN)) + geom_point(color="red")+  geom_smooth(method = "lm", se = TRUE)+
+  labs(title="Scatterplot between Cost and Debt ") +
+    theme(plot.title = element_text(hjust = 0.5))
+
+```
+
+From the scatter plot, it is obvious that the NPT4_COMBINE is in positive linear relationship with GRAD_DEBT_MDN: the data show an uphill pattern from left to right; As the average net price for Title IV institutions increase, the value of median debt for students who have completed increases as well.
+
+#6.3 Percentage of Pell Grant VS college type(CONTROL)
+
+To compare the percentages of Pell Grant in different control type(), we grouped the Percentage of Pell Grant data into the college type level to take a further analysis.
+
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(variable$ctrF ) %>%
+    dplyr::summarise(minPCTPELL=min(PCTPELL ,na.rm=TRUE),    
+    maxPCTPELL=max(PCTPELL,na.rm=TRUE),medianPCTPELL=median(PCTPELL,na.rm=TRUE))  %>%
+       arrange(medianPCTPELL)   %>%
+{.}  ->PCTPELL_CONTROL
+
+PCTPELL_CONTROL
+
+```
+From the result, the Private Nonprofit institutions and Public institutions have a similar low median values while the provate For- profit institutions have a hgihest value.
+This is resonable because generally speanking, the private for-profit colleges may have much higher costs than the private nonprofit and public institutions.
+The low-income students were more likely to seek a Pell Grant for assistant.
+Additionally, in some states, state financial aid may be sufficient to cover costs at community colleges; thus the students in thoes schools may not need to seek for addtional financial aid like Pell Grant.
+
+```{r}
+
+ggplot(data_clean, aes(x=PCTPELL, fill=variable$ctrF)) + geom_histogram(alpha=0.9, position="identity") +
+  labs(title="Overlaying histogram between PCTPELL and Control ") +
+    theme(plot.title = element_text(hjust = 0.5))
+
+```
+From the overlaying histograms, the percentage of students who receive Pell Grant are quite frequent for Private For-profit institutions between 0.7 and 0.75 ; the percentage of students who receive Pell Grant are quite frequent around 0.4 for Public institutions and Private Nonprofit institutions.
+The histogram for private for profit is slightly left skewed due to atypical small value near 0.
+The histogram for public and private nonprofit institions are reasonably symmetric
+
+#6.4  Percentage first-generation students VS Status
+To compare the Percentage first-generation students with the different states, we grouped data into the state level to take a further analysis.
+
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(variable$STABBR ) %>%
+    dplyr::summarise(minPAR_ED_PCT_1STGEN=min(PAR_ED_PCT_1STGEN ,na.rm=TRUE),    
+    maxPAR_ED_PCT_1STGEN=max(PAR_ED_PCT_1STGEN,na.rm=TRUE),medianPAR_ED_PCT_1STGEN=median(PAR_ED_PCT_1STGEN,na.rm=TRUE))  %>%
+       arrange(medianPAR_ED_PCT_1STGEN)   %>%
+{.}  ->PAR_ED_PCT_1STGEN_STATE
+
+PAR_ED_PCT_1STGEN_STATE
+
+```
+
+The results showed that the universities located in ND, SD, VT, MN and MT had the lowest value of Percentage first-generation students, which indicated that the students in thoses five states would have a better family education background.
+The five states with the highest share of parents first-generation are LA, NJ, MP, MV, MH.
+
+#6.5 family income with Pell Grant
+We assume there would be a strong linear relationship between family income and the percentage of students who can receive the Pell Grant.
+We hypothese that colleges with small value of family income would tend to have high percentages of Pell Grant.
+The following codes are used to generate the scatter pkot between these two varaibles.
+
+```{r}
+
+ggplot(data_clean, aes(x =MD_FAMINC, y = PCTPELL)) + geom_point(color="red")+  geom_smooth(method = "lm", se = TRUE)+
+  labs(title="Scatterplot between MD_FAMINC and PCTPELL ") +
+    theme(plot.title = element_text(hjust = 0.5))
+
+```
+From the scatter plot, it is obvious that the MD_FAMINC is in a negative linear relationship with PCTPELL: the data show a downhill pattern from left to right; As the MD_FAMINC increase, the value of PCTPELL decreases.
 
 #7. Comparison between variables
-PCTPELL vs PCTFLOAN
+
+#7.1 PCTPELL VS PCTFLOAN
+
+```{r}
+
+library(ggplot2)
+ggplot(data = data_clean) +
+  geom_histogram(breaks=seq(0, 1, by =0.05),aes(x = PCTPELL, y=(..count..)), 
+                       alpha=0.9, fill ="red",position="dodge") +
+  geom_histogram(breaks=seq(0, 1, by =0.05),aes(x = PCTFLOAN, y=(..count..)), 
+                       alpha=0.9,fill ="blue",position="dodge") +
+  labs(title="Overlaying Histogram of PCTPELL and PCTFLOAN ") +
+    theme(plot.title = element_text(hjust = 0.5)) + 
+  scale_colour_manual("", 
+                      breaks = c("PCTPELL", "PCTFLOAN"),
+                      values = c("PCTPELL"="red", "PCTFLOAN"="blue"))
+
+```
+
+aid_compare -> data_clean[c(5:6)]
+class(aid_compare)
 

### 2017-02-13 Draft-xuezhou-sjb.Rmd
changed summary panckage

@@
   group_by(variable$ctrF) %>%
-    summarise(minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND ,na.rm=TRUE),    
+    dplyr::summarise(minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND ,na.rm=TRUE),    
     maxDEP_STAT_PCT_IND=max(DEP_STAT_PCT_IND,na.rm=TRUE),medianDEP_STAT_PCT_IND=median(DEP_STAT_PCT_IND,na.rm=TRUE))  %>%
@@
   group_by(variable$STABBR ) %>%
-    summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
+    dplyr::summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
     maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
@@
   group_by(variable$ctrF ) %>%
-    dplsummarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
+    dplyr::summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
     maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
@@
   ggplot(aes(x=PCTFLOAN)) +
-  geom_bar(binwidth=5,aes(fill=variable$ctrF ))
+  geom_bar(binwidth=0.1,aes(fill=variable$ctrF ))
 

### 2017-02-13 Draft-xuezhou-sjb.Rmd
update

@@
 
+#6.6 cost and median debts with earning
+
+Next, we analyze the relationship between the amount the debt and post-graduation earning.
+There are variable in the dataset to summarize the median amount for student body, one for students who already graduated and one for students who have not completed their study.
+We decide debts for students who have not completed to study in this problem, since it may can be used as a predictor in the future modeling.
+First, we use scatter plot to visualize the relationship between earning and the amount of debts, the plot clearly shows a positive relationship.
+Then we further calculate pearson correlation to confirm our hypothesis.
+The resulting coefficient 0.44 indicatesa moderate linear relationship between earning and debts.
+
+```{r, message=FALSE}
+
+data_clean %>% ggplot(aes(x = WDRAW_DEBT_MDN, y = MD_EARN_WNE_P6)) + 
+                geom_jitter() +
+                geom_smooth(method = "lm", se = FALSE, col = 'red')
+
+data_clean %>% select(WDRAW_DEBT_MDN, MD_EARN_WNE_P6) %>%
+      na.omit() %>%
+       {.} -> earn_debt
+
+cor(earn_debt$WDRAW_DEBT_MDN,earn_debt$MD_EARN_WNE_P6)
+
+```
+
+##6.6 default with earning
+
+Understanding the relationship between the default rate of load repayment and post-graduation earning on one hand can help federal financial commitee to improve their decision about load granting.
+If the the default rate increases as the earning increase, the commitee may prefer to grant more financial aid to institution whose students have higher earning.
+One the other hand, the default rate of repayment for a univeristy can also help predict the post-graduation earning for its students.
+A scatter plot is created to visualized the relationship between these two variables.
+From the chart, we see the earning seem to show a decreasing trend as the default rate increases, which shows a negative relationship between earning and default rate.
+Then we create the correlation coefficient of these two variables.
+The resulting r = -0.33 shows that correlation between earning and default rate is negative but the relationship is not very strong, this may be due the influences of some outliers.
+To conclude, the negative relationship between earning and default rate of loan repayment needs larger population or detail data processing to confirm.
+
+```{r, message=FALSE}
+data_clean %>% ggplot(aes(x = CDR3, y = MD_EARN_WNE_P6)) + 
+                geom_jitter() +
+                geom_smooth(method = "lm", se = FALSE, col = 'red')
+
+data_clean %>% select(CDR3, MD_EARN_WNE_P6) %>%
+      na.omit() %>%
+       {.} -> earn_default
+
+cor(earn_default$CDR3,earn_default$MD_EARN_WNE_P6)
+
+```
+
+##6.7 degree type vs. earning
+
+The higher education one receives, the higher the earning is after graduation.
+This is an assumption we are making when investing higher education.
+But is it really the case.
+In this question, we would like to study if it's worthwhile to invest in higher degree.
+From the boxplot below, we see that the higher the education, the higher the median earning 2 years after graduation.
+Especially, a graduate-degree give a hugh rise to one's earning.
+
+```{r, message=FALSE}
+
+data_clean %>%
+  ggplot(aes(x=`variable$preF`,
+             y=MD_EARN_WNE_P6)) +
+  geom_boxplot(aes(fill = `variable$preF`))
+
+```
+
 #7. Comparison between variables

### 2017-02-13 Draft-xuezhou-sjb.Rmd
fill the table

@@
 | GRAD_DEBT_MDN    | Numeric           |The median debt for students who have completed                       |   Financial Aid                 |
-| WDRAW_DEBT_MDN   | Numeric           |  Continuous                       |    n     |
-| MD_EARN_WNE_P6   | Numeric           |  Continuous                       |     n    |
-| MD_EARN_WNE_P10  | Numeric           |  Continuous                       |       n  |
-| RPY_3YR_RT_SUPP  | Numeric           |  Continuous                       |      n   |
-| CDR3             | Numeric           |  Continuous                       | n        |
-| variable$qreF    | Factor            |  Categorical                      |  n       |
-| variable$ctrF    | Factor            |  Categorical                      |   n      |
-| variable$STABBR  | Factor            |  Categorical                      |   n      |
-| APPL_SCH_PCT_GE2 | Numeric           |  Continuous                       | Student        |
+| WDRAW_DEBT_MDN   | Numeric           |The median debt for sutdents who have not completed                      |Financial Aid |
+| MD_EARN_WNE_P6   | Numeric           |Median income six years after entry                       |     earning   |
+| MD_EARN_WNE_P10  | Numeric           |Median income ten years after entry                       |     earning  |
+| RPY_3YR_RT_SUPP  | Numeric           |Repayment rate 3 years after entering repayment                       |Repayment   |
+| CDR3             | Numeric           |Loan default rate 3 years afer entering repayment                       |Repayment        |
+| variable$qreF    | Factor            |Predominant degree offered                      |School       |
+| variable$ctrF    | Factor            |Control of institution                      |School     |
+| variable$STABBR  | Factor            |State                      |School     |
+| APPL_SCH_PCT_GE2 | Numeric           |Percentage of students submitting 2 or more than 2 applications                       | Student        |
 

### 2017-02-14 Draft-xuezhou-sjb.Rmd
local
local changes
added kun's part

@@
 #1. Introduction
+The College Scorecard project is designed to increase the transparency of the college data, working as a navigator for students to make informed decisions about their choice for higher education.
+The information gathered in the dataset also gives students and families more power to differentiate postsecondary institutions and get an idea about how well colleges are preparing their students to be successful.
+The dataset is created to address the lack of information about college costs, earnings, federal student aid, admission rate etc.
+The data provides insights into the performance of institutions that receive federal aid dollars, and the outcomes of the students of those institutions.
+It also gives extra support for students and families to weight the tradeoffs on choosing the right college according to their own needs and educational goals.
+
+These data are collected by United States Department of Education with the goal of helping students and families make more educated decisions about college and compare vital information easily.
+
 #2. Objective
+
 #3. Dataset Preparation
@@
 
-#4. Goals
-#5. Summary of variables
-
 For numeric variabls, we calculate the basic summary statistics such as median, min, max, mode and quantile.
@@
 
+#4. Goals
+#5. Summary of variables
 ##5.1 Individual variable summary
@@
 
+#6.7 predominant degree, control  and earning after 6 year of entry
+The following plot shows two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry.
+Our team uses the function geom_tile()[ggplot2 package] to visualize the correlation matrix.
+
+```{r}
+ data_clean %>%
+ group_by(preF, ctrF) %>%
+summarize(avg.earn_p6=mean(MD_EARN_WNE_P6)) %>%
+ ggplot(aes(x=ctrF, y=preF)) +
+  geom_tile(aes(fill=avg.earn_p6))
+
+```
+Mean value of the earning is more representative for the two categories, preF and ctrF.
+For the range of average earning, the darkest color shows the lowest earning, while the lightest color shows the highest value.
+According to the plot, we realize that ‘Private For-profit college’ whose predominant degree award is ‘Graduate degree’ has the lowest average earning with the minimum average value at \$3100, while ‘Public college’ whose predominant degree award is ‘Graduate degree’ has the highest average earning with the maximum average value at $45850.
+
+#6.8 predominant degree + control (count)
+The following plot shows the relationship between two factor variables (predominant degree award variable and control variable).
+Our team uses the function geom_bar() to map the counting values to y, ctrF variable to x and the preF variable to the fill of the bar.
+
+```{R}
+data_clean %>%
++ ggplot(aes(x=ctrF))+
++ geom_bar(aes(fill=preF), position = 'dodge')
+```
+According to the histogram comparison of predominant degree award among three types of colleges, we realize that for public college, Associate’s-degree is most frequent awarded in the public colleges.
+Bachelor’s-degree is most frequent awarded in the private nonprofit colleges.
+Certificate-degree is most frequent awarded in private for-profit colleges.
+
+Although the ‘Graduate-degree’ awarded by public colleges shows the least frequency, it gains the highest average earnings of students working and not enrolled 6 years after entry compared with other degrees awarded by different colleges.
+
+Our team further uses the function xtabs() to calculate the exact number for each predominant degree awarded by different types of colleges.
+```{R}
+> data_clean %>%
++ xtabs(formula = ~preF+ctrF)
+```
+
+#6.9 control + earning
+The following plot shows one factor variable (control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry).
+We use the
+geom_violin() using the default scale = “area” argument and geom_boxplot() function to explore the relationship.
+The plot superimposed with boxplots shows side-by-side violin plots of median earnings of students working and not enrolled 6 years after entry according to different college types.
+
+```{r}
+ data_clean %>%
++ ggplot(aes(x=ctrF, y=MD_EARN_WNE_P6)) +
++ geom_violin(aes(fill=ctrF), scale = 'area') +
++ geom_boxplot(aes(fill=ctrF), width = 0.25, outlier.size = 1.25)
+```
+Obvious differences contain in these three boxplots.
+The boxplot of ‘Public college’ is comparatively short, while the boxplot of ‘Private Nonprofit’ is comparatively tall.
+The ‘Private For-profit college’ has the lowest median earning.
+
 #7. Comparison between variables
-
 #7.1 PCTPELL VS PCTFLOAN

### 2017-02-15 Draft-xuezhou-sjb.Rmd
add TOC

@@
 
+#Table of Contents
+*1. Introduction
+*2. Goal
+*3. Dataset Preparation
+*4. Objective
+*5. Summary of variables
+    *5.1 Individual variable summary
+       *5.1.1 School
+          *5.1.1.1 State
+          *5.1.1.2 Predominant Degree
+          *5.1.1.3 Control of Institution
+       *5.1.2 Student Body
+          *5.1.2.1 Percentage first-generation students
+          *5.1.2.2 Number of applications
+       *5.1.3 Cost 
+          *5.1.3.1 Average net price for Title IV institutions(both public and private institutions)
+          *5.1.3.2 Percentage of students who are financially independent
+       *5.1.4 Financial Aid
+          *5.1.4.1 Percentage of undergraduates who receive a Pell Grant
+          *5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan
+          *5.1.4.3 The median debt for students who have completed
+          *5.1.4.4 The median debt for students who have not completed
+       *5.1.5 Earning
+          *5.1.5.1 Median family income in real 2015 dollars
+          *5.1.5.2 Earning after graduation
+       *5.1.6 Repayment
+          *5.1.6.1 3-year Repayment rate
+          *5.1.6.2 3-year Default rate
+*6. Relationship between variables
+    *6.1 NPT4_COMBINE VS State
+    *6.2 Debt VS Cost
+    *6.3 Percentage of Pell Grant VS college type(CONTROL)
+    *6.4 Percentage first-generation students VS Statu
+    *6.5 family income with Pell Grant
+    *6.6 cost and median debts with earning
+    *6.7 default with earning
+    *6.8 degree type vs. earning
+    *6.9 predominant degree, control  and earning after 6 year of entry
+    *6.10 predominant degree + control (count)
+    *6.11 control + earning
+*7. Comparison between variables
+*8. Findings and Conclusion
+
 #1. Introduction
@@
 
-#2. Objective
-
+#2. Goal
 #3. Dataset Preparation
@@
 
-| Variable Name    | Data Type         | Description                                                         | Categoriy in Documentation       |
+| Variable Name    | Data Type         | Description                                                          |    Category in Documentation       |
 | --------------  | ------------------ | ---------------------------------------------------------------------| --------------------------------|
@@
 
-#4. Goals
+#4. Objective
 #5. Summary of variables
@@
 #6. Relationship between variables
-#6.1 NPT4_COMBINE VS State
+##6.1. NPT4_COMBINE VS State
 
@@
 
-#6.2 Debt VS Cost
+##6.2. Debt VS Cost
 We assume there would be a strong linear relationship between the average net price for Title IV institutions and the median debt for students who have completed.
@@
 
-#6.3 Percentage of Pell Grant VS college type(CONTROL)
+##6.3. Percentage of Pell Grant VS college type(CONTROL)
 
@@
 
-#6.4  Percentage first-generation students VS Status
+##6.4  Percentage first-generation students VS Status
 To compare the Percentage first-generation students with the different states, we grouped data into the state level to take a further analysis.
@@
 
-#6.5 family income with Pell Grant
+##6.5 family income with Pell Grant
 We assume there would be a strong linear relationship between family income and the percentage of students who can receive the Pell Grant.
@@
 
-#6.6 cost and median debts with earning
+##6.6 cost and median debts with earning
 
@@
 
-##6.6 default with earning
+##6.7 default with earning
 
@@
 
-##6.7 degree type vs. earning
+##6.8 degree type vs. earning
 
@@
 
-#6.7 predominant degree, control  and earning after 6 year of entry
+##6.9 predominant degree, control  and earning after 6 year of entry
 The following plot shows two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry.
@@
 
-#6.8 predominant degree + control (count)
+##6.10 predominant degree + control (count)
 The following plot shows the relationship between two factor variables (predominant degree award variable and control variable).
@@
 ```{R}
-> data_clean %>%
-+ xtabs(formula = ~preF+ctrF)
+ data_clean %>%
+ xtabs(formula = ~preF+ctrF)
 ```
 
-#6.9 control + earning
+ #6.11 control + earning
 The following plot shows one factor variable (control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry).

### 2017-02-15 Draft-xuezhou-sjb.Rmd
Update Draft-xuezhou-sjb.Rmd

@@
 #Table of Contents
-*1. Introduction
-*2. Goal
-*3. Dataset Preparation
-*4. Objective
-*5. Summary of variables
+*1.Introduction
+*2.Goal
+*3.Dataset Preparation
+*4.Objective
+*5.Summary of variables
  *5.1 Individual variable summary
@@
     *5.1.6.2 3-year Default rate
-*6. Relationship between variables
+*6.Relationship between variables
  *6.1 NPT4_COMBINE VS State

### 2017-02-15 Draft-xuezhou-sjb.Rmd
Update Draft-xuezhou-sjb.Rmd

@@
 #Table of Contents
-*1.Introduction
-*2.Goal
-*3.Dataset Preparation
-*4.Objective
-*5.Summary of variables
- *5.1 Individual variable summary
-  *5.1.1 School
-   *5.1.1.1 State
-   *5.1.1.2 Predominant Degree
-   *5.1.1.3 Control of Institution
-  *5.1.2 Student Body
-    *5.1.2.1 Percentage first-generation students
-    *5.1.2.2 Number of applications
-  *5.1.3 Cost 
-    *5.1.3.1 Average net price for Title IV institutions(both public and private institutions)
-    *5.1.3.2 Percentage of students who are financially independent
-  *5.1.4 Financial Aid
-    *5.1.4.1 Percentage of undergraduates who receive a Pell Grant
-    *5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan
-    *5.1.4.3 The median debt for students who have completed
-    *5.1.4.4 The median debt for students who have not completed
-  *5.1.5 Earning
-    *5.1.5.1 Median family income in real 2015 dollars
-    *5.1.5.2 Earning after graduation
-  *5.1.6 Repayment
-    *5.1.6.1 3-year Repayment rate
-    *5.1.6.2 3-year Default rate
-*6.Relationship between variables
- *6.1 NPT4_COMBINE VS State
- *6.2 Debt VS Cost
- *6.3 Percentage of Pell Grant VS college type(CONTROL)
- *6.4 Percentage first-generation students VS Statu
- *6.5 family income with Pell Grant
- *6.6 cost and median debts with earning
- *6.7 default with earning
- *6.8 degree type vs. earning
- *6.9 predominant degree, control  and earning after 6 year of entry
- *6.10 predominant degree + control (count)
- *6.11 control + earning
-*7. Comparison between variables
-*8. Findings and Conclusion
+- 1.Introduction
+* 2.Goal
+* 3.Dataset Preparation
+* 4.Objective
+* 5.Summary of variables
+ * 5.1 Individual variable summary
+  * 5.1.1 School
+   * 5.1.1.1 State
+   * 5.1.1.2 Predominant Degree
+   * 5.1.1.3 Control of Institution
+  * 5.1.2 Student Body
+    * 5.1.2.1 Percentage first-generation students
+    * 5.1.2.2 Number of applications
+  * 5.1.3 Cost 
+    * 5.1.3.1 Average net price for Title IV institutions(both public and private institutions)
+    * 5.1.3.2 Percentage of students who are financially independent
+  * 5.1.4 Financial Aid
+    * 5.1.4.1 Percentage of undergraduates who receive a Pell Grant
+    * 5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan
+    * 5.1.4.3 The median debt for students who have completed
+    * 5.1.4.4 The median debt for students who have not completed
+  * 5.1.5 Earning
+    * 5.1.5.1 Median family income in real 2015 dollars
+    * 5.1.5.2 Earning after graduation
+  * 5.1.6 Repayment
+    * 5.1.6.1 3-year Repayment rate
+    * 5.1.6.2 3-year Default rate
+* 6.Relationship between variables
+ * 6.1 NPT4_COMBINE VS State
+ * 6.2 Debt VS Cost
+ * 6.3 Percentage of Pell Grant VS college type(CONTROL)
+ * 6.4 Percentage first-generation students VS Statu
+ * 6.5 family income with Pell Grant
+ * 6.6 cost and median debts with earning
+ * 6.7 default with earning
+ * 6.8 degree type vs. earning
+ * 6.9 predominant degree, control  and earning after 6 year of entry
+ * 6.10 predominant degree + control (count)
+ * 6.11 control + earning
+* 7. Comparison between variables
+* 8. Findings and Conclusion
 

### 2017-02-15 Draft-xuezhou-sjb.Rmd
update

@@
 #Table of Contents
-- 1.Introduction
+* 1.Introduction
 * 2.Goal

### 2017-02-15 Draft-xuezhou-sjb.Rmd
update
hi

@@
 #Table of Contents
-* 1.Introduction
+* 1.[Introduction](#1. Introduction)
 * 2.Goal

### 2017-02-15 Draft-xuezhou-sjb.Rmd
Update Draft-xuezhou-sjb.Rmd

--- 2017-02-15 Draft-xuezhou-sjb.Rmd
@@
 #Table of Contents
-* 1.Introduction
+* 1.[Introduction](#Introduction)
 * 2.Goal
@@
 
-#1. Introduction
+#1. Introduction<a id="Introduction"></a>
 The College Scorecard project is designed to increase the transparency of the college data, working as a navigator for students to make informed decisions about their choice for higher education.

### 2017-02-19 Draft-xuezhou-sjb.Rmd
update

@@
 * 1.[Introduction](#Introduction)
-* 2.Goal
-* 3.Dataset Preparation
-* 4.Objective
-* 5.Summary of variables
+* 2.[Goal](#Goal)
+* 3.[Dataset Preparation](#dataprep)
+* 4.[Objective](#objective)
+* 5.[Summary of variables](#variablesummary)
  * 5.1 Individual variable summary
@@
   * 5.1.2 Student Body
-    * 5.1.2.1 Percentage first-generation students
+    * 5.1.2.1 [Percentage first-generation students](#5.1.2.1)
     * 5.1.2.2 Number of applications
   * 5.1.3 Cost 
-    * 5.1.3.1 Average net price for Title IV institutions(both public and private institutions)
-    * 5.1.3.2 Percentage of students who are financially independent
+    * 5.1.3.1 [Average net price for Title IV institutions(both public and private institutions)](#5.1.3.1)
+    * 5.1.3.2 [Percentage of students who are financially independent](#5.1.3.2)
   * 5.1.4 Financial Aid
-    * 5.1.4.1 Percentage of undergraduates who receive a Pell Grant
-    * 5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan
-    * 5.1.4.3 The median debt for students who have completed
+    * 5.1.4.1 [Percentage of undergraduates who receive a Pell Grant](#5.1.4.1)
+    * 5.1.4.2 [Percent of all federal undergraduate students receiving a federal student loan](#5.1.4.2l)
+    * 5.1.4.3 [The median debt for students who have completed](#5.1.4.3)
     * 5.1.4.4 The median debt for students who have not completed
   * 5.1.5 Earning
-    * 5.1.5.1 Median family income in real 2015 dollars
+    * 5.1.5.1 [Median family income in real 2015 dollars](#5.1.5.1)
     * 5.1.5.2 Earning after graduation
@@
 * 6.Relationship between variables
- * 6.1 NPT4_COMBINE VS State
- * 6.2 Debt VS Cost
- * 6.3 Percentage of Pell Grant VS college type(CONTROL)
- * 6.4 Percentage first-generation students VS Statu
- * 6.5 family income with Pell Grant
- * 6.6 cost and median debts with earning
- * 6.7 default with earning
- * 6.8 degree type vs. earning
- * 6.9 predominant degree, control  and earning after 6 year of entry
- * 6.10 predominant degree + control (count)
- * 6.11 control + earning
-* 7. Comparison between variables
-* 8. Findings and Conclusion
+ * 6.1 [The relationship between NPT4_COMBINE and State](#6.1)
+ * 6.2 [The relationship between Debt and Cost](#6.2)
+ * 6.3 [The relationship between Percentage of Pell Grant and college type(CONTROL)](#6.3)
+ * 6.4 [The relationship between Percentage first-generation students and States](#6.4)
+ * 6.5 [The relationship between family income and percentage of Pell Grant](#6.5)
+ * 6.6 [The relationship between DEP_STAT_PCT_IND and Control](#6.6)
+ * 6.7 [The relationship between PCTLOAN and Control](#6.7)
+ * 6.8 cost and median debts with earning
+ * 6.9 default with earning
+ * 6.10 degree type vs. earning
+ * 6.11 predominant degree, control  and earning after 6 year of entry
+ * 6.12 predominant degree + control (count)
+ * 6.13 control + earning
+ * 7. Findings and Conclusion
 
@@
 
-#2. Goal
-#3. Dataset Preparation
+#2. Goal<a id="Goal"></a>
+#3. Dataset Preparation<a id="dataprep"></a>
 Our analysis focuses on the data of year 2012 -2013 of 23 variables.
@@
 
+Considering there is some special characters in the ID coloumn, we first rename the first column's name to be selectable. 
+```{R}
+colnames(data)[1] = 'UNITID'
+```
+
 The full dataset contains 1743 variables. 36 variables were intially selected to analyze.
@@
 
-variable = select(data, STABBR, PREDDEG, LOCALE2, MENONLY, WOMENONLY, CONTROL, ADM_RATE, SATMTMID, SATVRMID, SATWRMID, SAT_AVG,
+variable = select(data, UNITID, INSTNM, STABBR, PREDDEG, LOCALE2, MENONLY, WOMENONLY, CONTROL, ADM_RATE, SATMTMID, SATVRMID, SATWRMID, SAT_AVG,
                   NPT4_PUB,
@@
 
-class(as.data.frame(apply(variable[,-c(1,2,6)],2,as.numeric)))
+#need to be commented out 
+class(as.data.frame(apply(variable[,-c(2,3,4,8)],2,as.numeric)))
 
-#numeric
-variable[,-c(1,2,6)] %>%
+#get rid of factor variables: INSTNM, STABBR, PREDDEG, CONTROL and change the data types to numeric 
+variable[,-c(2,3,4,8)] %>%
     apply(2, as.numeric) %>%
@@
 
-#categorical
+#change the categorical level of the categorical variables:PREDDEG, CONTROL
 variable$preF=factor(variable$PREDDEG,levels=c(0,1,2,3,4),labels = c("NotClassified","Certificate-degree","Associate's-degree","Bachelor's-degree","Graduate-degree"))
+
 variable$ctrF=factor(variable$CONTROL,levels = c(1,2,3),labels = c("Public","Private Nonprofit","Private For-profit"))
@@
 #combine the three datafram to generate a new dataset. 
-data_new <- cbind(variable_num, variable$preF, variable$ctrF,variable$STABBR)
-dim(data_new)
+data_new <- cbind(variable_num, variable$preF, variable$ctrF,variable$INSTNM, variable$STABBR)
 
+#rename the categorical varaibles to approriate readable names
+colnames(data_new)[35] = 'PREDDEG_factor'
+colnames(data_new)[36] = 'Control_factor'
+colnames(data_new)[37] = 'INSTNM_factor'
+colnames(data_new)[38] = 'STABBR_factor'
+
+#need to be commented out
 mean(is.na(data_new$NPT4_PUB))
@@
 mean(is.na(data_new$NPT4_PRIV))
-data_new = data_new[,-9] 
-colnames(data_new)[9] <- "NPT4_COMBINE"
+data_new = data_new[,-10] 
+colnames(data_new)[10] <- "NPT4_COMBINE"
+
+mean(is.na(data_new$NPT4_COMBINE))
 ```
@@
 In order to choose the variables that have enough valid data related to the topic, we filtered and considered variables with percent missing value less than 25%.
-The new data set was named as “data_clean” for further analysis.
+The new data set was named as ???data_clean???
+for further analysis.
 ```{r, message=FALSE,warning=FALSE}
@@
 data_clean = data_new[,colMeans(is.na(data_new)) < 0.25]
+
+#rearrange the coloumn names
+data_clean = data_clean[c(1,17,18,15,16,2:14)]
+
 write.csv(data_clean,"data_clean.csv")
 ```
+The data_clean will be the cleaned dataset we used in the further variable summaries and comparisons
 
@@
 Our data set now contains 120 rows and 15 columns.
-120 rows consist of 20 years’ of data for each of the chosen six countries.
+120 rows consist of 20 years???
+of data for each of the chosen six countries.
 The 15 columns consist of country name, country code, year, and the twelve indicators.
@@
 
-we will summarize the data set using dplyr functions in the variables summary section.
+we will summarize the data set using dplyr and ggplot funtions in the variables summary section.
 
-#4. Objective
-#5. Summary of variables
+#4. Objective<a id="objective"></a>
+Over years, Many students see college as an investment to help them earn more and live better lives after graduation.
+Amount of findints have been conducted to analyze the influence of college quality on students' post graudation earning based on *colleges characterestis, student's demographic information, as well as family charateristics as well as federal finanical support to low-income students.
+It has become more and more important to understand the fac- tors that contribute to post-graduation earnings and the ability to repay student loans.
+
+Our objective in this stage is to conduct an exploration analysis on the CollegeScore dataset, investigating the individual variable's summary statistics and the relationship between these variabls, so as to better prepare for further analaysis on the factors affecting students' post college earnings with some advanced data mining techniques.
+
+#5. Summary of variables<a id="variablesummary"></a>
 ##5.1 Individual variable summary
@@
 
-We are using the code below to summarize this variable.
+With the followin codes, we extracted the institions with the highest and lowest percentage of first-generation students.
+Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
+```{R, message=FALSE, warinig = FALSE}
 
-```{R, message=FALSE, warinig = FALSE}
+#the institions with the highest and lowest percentage of first-generation students
+data_clean %>%
+ slice(c(which.min(PAR_ED_PCT_1STGEN),which.max(PAR_ED_PCT_1STGEN))) %>%
+  select (c(1,2,3,8))
 
@@
 ```
-From the output, we can tell that the highest value is 95.25% while the lowest value is 6.87%.
+From the output, we can tell that the highest value existed in  La Belle Beauty Academy with the value of 95.25%.
+The lowest value is 6.87%, which existed in Wheaton College.
 From the mean value of 48.08%, we can tell that about for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college.
@@
 
-We are using the code below to summarize the variable. 
+With the followin codes, we extracted the IV institutions with the highest and lowest Average net price.
+Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
+
 ```{R, message=FALSE, warinig = FALSE}
+
+data_clean %>%
+ slice(c(which.min(NPT4_COMBINE),which.max(NPT4_COMBINE))) %>%
+  select (c(1,2,3,6))
 
@@
 ```
-From the summary results, we learned that the maximum value of the variable is \$87,414 while the minimum value is about \$-19,360.
+From the summary results, we learned that the maximum value of the variable is $87,414, which is for Aerosim Flight Academy located in FL.
+while the minimum value is about $-19,360 - the value for Berea College located in KY.
 Considering the elements are collected from Title IV institutions, we reasoned that the negative values may be due to the fact that some intituions's finainal aid supoort can cover the students' all the costs, meanwhile overbalancing the cost amount.
-There are not much differerences between mean value and median value, which is around \$15,000.
+There are not much differerences between mean value and median value, which is around $15,000.
 We can conclude that the average net price costs for most institutuins are falling into the this range.
@@
 
-Here is the code to generate the summary statistics: 
+With the followin codes, we extracted the institions which have the highest and lowest percentage of students who are financially independent.
+Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
 ```{R}
+
+data_clean %>%
+ slice(c(which.min(DEP_STAT_PCT_IND),which.max(DEP_STAT_PCT_IND))) %>%
+  select (c(1,2,3,7))
 
@@
 ```
-From the summary results, we learned that the maximum value of the variable is 98.56% while the minimum value is about 0，73%.
+From the summary results, we learned that the maximum value of the variable is 98.56%(Capella University) while the minimum value is about 0.73%(University of Notre Dame).
 From the mean value, we can tell that about in most colleges, around 56% of the students are financially independent.
@@
 Below is the code to generate a visual representation in the histgram.
+
 ```{R}
@@
 
-Then we grouped by the data into the control type(Private For-profit, Public, Private Nonprofit) to take a further analysis.
-```{R message=FALSE, warinig = FALSE}
-
-data_clean %>%
-  group_by(variable$ctrF) %>%
-    dplyr::summarise(minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND ,na.rm=TRUE),    
-    maxDEP_STAT_PCT_IND=max(DEP_STAT_PCT_IND,na.rm=TRUE),medianDEP_STAT_PCT_IND=median(DEP_STAT_PCT_IND,na.rm=TRUE))  %>%
-       arrange(desc(medianDEP_STAT_PCT_IND))   %>%
-{.}  ->DEP_STAT_PCT_IND_CONTROL
-
-DEP_STAT_PCT_IND_CONTROL
-
-```
-From the result, we can tell the highest median value is in Privately For-profit colleges, which means that in most private for-profit colleges, the students are more likely to be financial independent.
-
 ###5.1.4 Financial Aid 
@@
 This elemnt, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received Pell Grants in a given year, which is an important measure of the access an insiitution provides to low-income students.
-We are using the code below to summarize this variable.
+
+We are using the code below to summarize this variable and extraced the instition with the highest and lowest value.
 
 ```{R, message=FALSE, warinig = FALSE}
+
+data_clean %>%
+ slice(c(which.min(PCTPELL),which.max(PCTPELL))) %>%
+  select (c(1,2,3,10))
 
@@
 ```
-From the output, we can tell that the range is between 0 to 100%.	 Among the colleges in the sample, half colleges have this rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs.
+From the output, we can tell that the range is between 0 to 100%.	No students in Thunderbird School of Global Management has received the Pell grant while all of the students in American Conservatory Theater have receiced the Pell Grant.
+Among the colleges in the sample, half colleges have this rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs.
 The standard deviation value is 22.57%,  indicating a relative high difference of this element among different colleges, which can also be noticed from the largest range (100%).
@@
 It can provide important context to figures related to debt, repayment and non-repayment.
-We are using the code below to summarize this variable.
+We are using the code below to summarize this variable and extracted the institutions with highest and lowest value.
 
 ```{R, message=FALSE, warinig = FALSE}
+
+data_clean %>%
+ slice(c(which.min(PCTFLOAN),which.max(PCTFLOAN))) %>%
+  select (c(1,2,3,11))
 
@@
 ```
-From the output, we can tell that the range is between 0 to 100%.	 Among the colleges in the sample, half colleges have this rate as high as 60.08% , which means around 60% of undergraduare students are able to receive the fedaral student loans.
+From the output, we can tell that the range is between 0 to 100%.	As we can recall from the previous variable(PCTPELL), American Conservatory Theater has both the PCTLOAN and PCTPELL the value of 100%.
+Gadsden State Community College has no students reveiced federal loans in the given year.
+Among the colleges in the sample, half colleges have this rate as high as 60.08% , which means around 60% of undergraduare students are able to receive the fedaral student loans.
 The standard deviation value is 29.26%,  indicating a relative high difference of this element among different colleges, which can also be noticed from the largest range (100%).
@@
 
-Then we grouped by the data into the college type level to take a further analysis.
-```{R message=FALSE, warinig = FALSE}
+####5.1.4.3 The median debt for students who have completed
+
+This element is the median loan debt acculated at the institution by all student borrowers of federal lonans who have habe completed the education.
+We are using the code below to summarize this variable.
+```{R, message=FALSE, warinig = FALSE}
 
 data_clean %>%
-  group_by(variable$STABBR ) %>%
-    dplyr::summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
-    maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
-       arrange(medianPCTFLOAN)   %>%
-{.}  ->PCTFLOAN_STATE
-
-PCTFLOAN_STATE
-
-data_clean %>%
-  group_by(variable$ctrF ) %>%
-    dplyr::summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
-    maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
-       arrange(medianPCTFLOAN)   %>%
-{.}  ->PCTFLOAN_CONTROL
-
-PCTFLOAN_CONTROL
-
-#stacked histogram
-data_clean %>%
-  ggplot(aes(x=PCTFLOAN)) +
-  geom_bar(binwidth=0.1,aes(fill=variable$ctrF ))
-
-```
-
-From the result, the Private institutions have a much larger value than the public institutions.
-With a much higher costs burden, it makes sence that the low-income sutdents in private institutions are much more likely to seek financial aid than the public institution students.
-
-####5.1.4.3 The median debt for students who have completed
-
-This element is the median loan debt acculated at the institution by all student borrowers of federal lonans who have habe completed the education.We are using the code below to summarize this variable.
-```{R, message=FALSE, warinig = FALSE}
+ slice(c(which.min(GRAD_DEBT_MDN),which.max(GRAD_DEBT_MDN))) %>%
+  select (c(1,2,3,12))
 
@@
 ```
-From the output, we can tell that the highest value is 48,250	 while the lowest value is 1448.
+From the output, we can tell that the highest value is 48,250 - Southwest University of Visual Arts-Tucson, while the lowest value is 1448 - South Florida Institute of Technology.
 From the median value of 14000, we can tell that among the colleges in the sample, the students who have completed the deducaiton from half of them had the median loan debt of 14000.
@@
 
-####5.1.4.4 The median debt for students who have not completed
+####5.1.4.4 The median debt for students who have not completed<a id="5.1.4.4"></a>
 WDRAW_DEBT_MDN describe the median number of debt fot students who ahve not completed their study.
@@
 ###5.1.5 Earning
-####5.1.5.1 Median family income in real 2015 dollars
+####5.1.5.1 Median family income in real 2015 dollars<a id="5.1.5.1"></a>
 This elemnt is recording the median family income of students, one of the undergraduate student demographics for earnings cohorts.
@@
 ```{R, message=FALSE, warinig = FALSE}
+
+data_clean %>%
+ slice(c(which.min(MD_FAMINC),which.max(MD_FAMINC))) %>%
+  select (c(1,2,3,9))
 
@@
 ```
-From the output, we can tell that the highest value is 155,606.6	 while the lowest value is 0.
+From the output, we can tell that the highest value is 155,606.6 -Jewish Theological Seminary of America, while the lowest value is 0 -Associated Technical College-Los Angeles.
 From the median value of 19,262.26, we can tell that among the colleges in the sample, the students in half of them had the family income below 19,262.26	.
@@
 
-#6. Relationship between variables
-##6.1. NPT4_COMBINE VS State
+#6. Relationship between variables<a id="6"></a>
+
+##6.1. The relationship between NPT4_COMBINE and State<a id="6.1"></a>
 
 To analyze the relationship between average net cost for Title IV institutions and State, we grouped the data into the state level to take a further look.
+We would like to investigate the state distribution of the Title IV instituions based on the average net costs.
 ```{R message=FALSE, warinig = FALSE}
@@
 
-##6.2. Debt VS Cost
+##6.2. The relationship between Debt and Cost<a id="6.2"></a>
 We assume there would be a strong linear relationship between the average net price for Title IV institutions and the median debt for students who have completed.
@@
 
-##6.3. Percentage of Pell Grant VS college type(CONTROL)
+##6.3. The relationship between Percentage of Pell Grant and college type(CONTROL)<a id="6.3"></a>
 
-To compare the percentages of Pell Grant in different control type(), we grouped the Percentage of Pell Grant data into the college type level to take a further analysis.
+To compare the percentages of Pell Grant in different control type, we grouped the Percentage of Pell Grant data into the college type level to take a further analysis.
+We would like to investigate the percentage of Pell Grant based on the  college type.
 
@@
 {.}  ->PCTPELL_CONTROL
-
 PCTPELL_CONTROL
@@
 
-##6.4  Percentage first-generation students VS Status
-To compare the Percentage first-generation students with the different states, we grouped data into the state level to take a further analysis.
+##6.4. The relationship between Percentage first-generation students and States<a id="6.4"></a>
+To compare the Percentage first-generation students with the different states, we grouped data into the state level to take a further analysis.We would like to investigate the percentage of first-generation students based on the status distribution.
 
@@
 
-##6.5 family income with Pell Grant
+##6.5. The relationship between family income and percentage of Pell Grant<a id="6.5"></a>
  We assume there would be a strong linear relationship between family income and the percentage of students who can receive the Pell Grant.
@@
 From the scatter plot, it is obvious that the MD_FAMINC is in a negative linear relationship with PCTPELL: the data show a downhill pattern from left to right; As the MD_FAMINC increase, the value of PCTPELL decreases.
+It makes sense in the real world applcation: for institutions with low percentage of students who receive the Pell Grant, the financial aids like Pell Grant may not be as necessary as they are in the instituions with high value of PCTPELL, thus these collges students will tend to have a high median family income after they graduate.
 
-##6.6 cost and median debts with earning
+##6.6. The relationship between DEP_STAT_PCT_IND and Control<a id="6.6"></a>
 
+To compare the percentage of financially-independent students in institutions with different college type, we grouped by the data into the control type(Private For-profit, Public, Private Nonprofit) to take a further analysis.
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(variable$ctrF) %>%
+    dplyr::summarise(minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND ,na.rm=TRUE),    
+    maxDEP_STAT_PCT_IND=max(DEP_STAT_PCT_IND,na.rm=TRUE),medianDEP_STAT_PCT_IND=median(DEP_STAT_PCT_IND,na.rm=TRUE))  %>%
+       arrange(desc(medianDEP_STAT_PCT_IND))   %>%
+{.}  ->DEP_STAT_PCT_IND_CONTROL
+
+DEP_STAT_PCT_IND_CONTROL
+
+```
+From the result, we can tell the highest median value is in Privately For-profit colleges, which means that in most private for-profit colleges, the students are more likely to be financial independent.
+
+##6.7. The relationship between PCTLOAN and Control.<a id="6.7"></a>
+
+To compare the percentage of students who receive the federal loans in the given year based on the schools control type,we grouped by the data into the control type(Private For-profit, Public, Private Nonprofit) to take a further analysis.
+A stacked histogram is also requested to generate a more vidual presentation.
+
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(variable$ctrF ) %>%
+    dplyr::summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
+    maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
+       arrange(medianPCTFLOAN)   %>%
+{.}  ->PCTFLOAN_CONTROL
+
+PCTFLOAN_CONTROL
+
+#stacked histogram
+data_clean %>%
+  ggplot(aes(x=PCTFLOAN)) +
+  geom_bar(binwidth=0.1,aes(fill=Control_factor ))
+
+```
+
+From the result, the Private institutions have a much larger median value than the public institutions.
+With a much higher costs burden, it makes sence that the low-income sutdents in private institutions are much more likely to seek financial aid than the public institution students.
+From the stacked histogram, we can tell that for all the three college type institutions, the most frequent ranges appear between 0.6 and 0.8, which is consistent with the findings we observe in the individual variable summary section.
+
+##6.8 cost and median debts with earning
 Next, we analyze the relationship between the amount the debt and post-graduation earning.
@@
 
-##6.7 default with earning
+##6.9 default with earning
 
@@
 
-##6.8 degree type vs. earning
+##6.10 degree type vs. earning
 
@@
 
-##6.9 predominant degree, control  and earning after 6 year of entry
+##6.11 predominant degree, control  and earning after 6 year of entry
 The following plot shows two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry.
@@
 For the range of average earning, the darkest color shows the lowest earning, while the lightest color shows the highest value.
-According to the plot, we realize that ‘Private For-profit college’ whose predominant degree award is ‘Graduate degree’ has the lowest average earning with the minimum average value at \$3100, while ‘Public college’ whose predominant degree award is ‘Graduate degree’ has the highest average earning with the maximum average value at $45850.
+According to the plot, we realize that ???Private For-profit college???
+whose predominant degree award is ???Graduate degree???
+has the lowest average earning with the minimum average value at \$3100, while ???Public college???
+whose predominant degree award is ???Graduate degree???
+has the highest average earning with the maximum average value at $45850.
 
-##6.10 predominant degree + control (count)
+##6.12 predominant degree + control (count)
 The following plot shows the relationship between two factor variables (predominant degree award variable and control variable).
@@
 ```
-According to the histogram comparison of predominant degree award among three types of colleges, we realize that for public college, Associate’s-degree is most frequent awarded in the public colleges.
-Bachelor’s-degree is most frequent awarded in the private nonprofit colleges.
+According to the histogram comparison of predominant degree award among three types of colleges, we realize that for public college, Associate???s-degree is most frequent awarded in the public colleges.
+Bachelor???s-degree is most frequent awarded in the private nonprofit colleges.
 Certificate-degree is most frequent awarded in private for-profit colleges.
 
-Although the ‘Graduate-degree’ awarded by public colleges shows the least frequency, it gains the highest average earnings of students working and not enrolled 6 years after entry compared with other degrees awarded by different colleges.
+Although the ???Graduate-degree???
+awarded by public colleges shows the least frequency, it gains the highest average earnings of students working and not enrolled 6 years after entry compared with other degrees awarded by different colleges.
 
@@
 We use the
-geom_violin() using the default scale = “area” argument and geom_boxplot() function to explore the relationship.
+geom_violin() using the default scale = ???area???
+argument and geom_boxplot() function to explore the relationship.
 The plot superimposed with boxplots shows side-by-side violin plots of median earnings of students working and not enrolled 6 years after entry according to different college types.
@@
 Obvious differences contain in these three boxplots.
-The boxplot of ‘Public college’ is comparatively short, while the boxplot of ‘Private Nonprofit’ is comparatively tall.
-The ‘Private For-profit college’ has the lowest median earning.
+The boxplot of ???Public college???
+is comparatively short, while the boxplot of ???Private Nonprofit???
+is comparatively tall.
+The ???Private For-profit college???
+has the lowest median earning.
 
-#7. Comparison between variables
- #7.1 PCTPELL VS PCTFLOAN
-
-```{r}
-
-library(ggplot2)
-ggplot(data = data_clean) +
-  geom_histogram(breaks=seq(0, 1, by =0.05),aes(x = PCTPELL, y=(..count..)), 
-                       alpha=0.9, fill ="red",position="dodge") +
-  geom_histogram(breaks=seq(0, 1, by =0.05),aes(x = PCTFLOAN, y=(..count..)), 
-                       alpha=0.9,fill ="blue",position="dodge") +
-  labs(title="Overlaying Histogram of PCTPELL and PCTFLOAN ") +
-    theme(plot.title = element_text(hjust = 0.5)) + 
-  scale_colour_manual("", 
-                      breaks = c("PCTPELL", "PCTFLOAN"),
-                      values = c("PCTPELL"="red", "PCTFLOAN"="blue"))
-
-```
-
-aid_compare -> data_clean[c(5:6)]
-class(aid_compare)
-
-#8. Findings and Conclusion
+#7. Findings and Conclusion

### 2017-02-19 Draft-xuezhou-sjb.Rmd
variale table

@@
 
-| Variable Name    | Data Type         | Description                                                          |    Category in Documentation       |
-| --------------  | ------------------ | ---------------------------------------------------------------------| --------------------------------|
-| NPT4_COMBINE    | Numeric            |average net price for TITLE IV institutions of both public and private.|  Cost                       |
-| DEP_STAT_PCT_IND | Numeric           |Percentage of students who are financially independent                |     Cost                        |
-| PAR_ED_PCT_1STGEN| Numeric           |Percentage first-generation students                                  |      Sudent Body                |
-| MD_FAMINC        | Numeric           |Median family income in real 2015 dollars                             |       Earning                   |
-| PCTPELL          | Numeric           |Percentage of undergraduates who receive a Pell Grant                 |        Financial Aid            |
-| PCTFLOAN         | Numeric           |Percent of all federal undergraduate students receiving a federal student loan      |  Financial Aid    |
-| GRAD_DEBT_MDN    | Numeric           |The median debt for students who have completed                       |   Financial Aid                 |
-| WDRAW_DEBT_MDN   | Numeric           |The median debt for sutdents who have not completed                   |Financial Aid |
-| MD_EARN_WNE_P6   | Numeric           |Median income six years after entry                                   |     earning   |
-| MD_EARN_WNE_P10  | Numeric           |Median income ten years after entry                                   |     earning  |
-| RPY_3YR_RT_SUPP  | Numeric           |Repayment rate 3 years after entering repayment                       |Repayment   |
-| CDR3             | Numeric           |Loan default rate 3 years afer entering repayment                     |Repayment        |
-| variable$qreF    | Factor            |Predominant degree offered                                            |School       |
-| variable$ctrF    | Factor            |Control of institution                                                |School     |
-| variable$STABBR  | Factor            |State                                                                 |School     |
-| APPL_SCH_PCT_GE2 | Numeric           |Percentage of students submitting 2 or more than 2 applications       | Student        |
+| Variable Name    | Data Type |  Unit  |  Description                                                          |    Category in Documentation       |
+| --------------  | -----------|- ----|-----------------------------------------------------------------| --------------------------------|
+| NPT4_COMBINE    | Numeric   | $         |average net price for TITLE IV institutions of both public and private.|  Cost                       |
+| DEP_STAT_PCT_IND | Numeric  | $         |Percentage of students who are financially independent                |     Cost                        |
+| PAR_ED_PCT_1STGEN| Numeric  | %         |Percentage first-generation students                                  |      Sudent Body                |
+| MD_FAMINC        | Numeric  | $         |Median family income in real 2015 dollars                             |       Earning                   |
+| PCTPELL          | Numeric  | %         |Percentage of undergraduates who receive a Pell Grant                 |        Financial Aid            |
+| PCTFLOAN         | Numeric  | %         |Percent of all federal undergraduate students receiving a federal student loan      |  Financial Aid    |
+| GRAD_DEBT_MDN    | Numeric  | $         |The median debt for students who have completed                       |   Financial Aid                 |
+| WDRAW_DEBT_MDN   | Numeric  | $         |The median debt for sutdents who have not completed                   |Financial Aid |
+| MD_EARN_WNE_P6   | Numeric  | $         |Median income six years after entry                                   |     earning   |
+| MD_EARN_WNE_P10  | Numeric  | $         |Median income ten years after entry                                   |     earning  |
+| RPY_3YR_RT_SUPP  | Numeric  | %         |Repayment rate 3 years after entering repayment                       |Repayment   |
+| CDR3             | Numeric  | %         |Loan default rate 3 years afer entering repayment                     |Repayment        |
+| PREDDEG_factor    | Factor  | N/A          |Predominant degree offered                                            |School       |
+| Control_factor    | Factor  | N/A          |Control of institution                                                |School     |
+| STABBR_factor  | Factor     | N/A       |State                                                                 |School     |
+| APPL_SCH_PCT_GE2 | Numeric  | %         |Percentage of students submitting 2 or more than 2 applications       | Student        |
+| UNITID | Numeric | N/A| Unique ID for the institutions | unknown
+| INSTNM | Factor |  N/A|name for the institutions | unknown
 

### 2017-02-19 Draft-xuezhou-sjb.Rmd
update

--- 2017-02-19 Draft-xuezhou-sjb.Rmd
@@
 
-| Variable Name    | Data Type         | Description                                                          |    Category in Documentation       |
-| --------------  | ------------------ | ---------------------------------------------------------------------| --------------------------------|
-| NPT4_COMBINE    | Numeric            |average net price for TITLE IV institutions of both public and private.|  Cost                       |
-| DEP_STAT_PCT_IND | Numeric           |Percentage of students who are financially independent                |     Cost                        |
-| PAR_ED_PCT_1STGEN| Numeric           |Percentage first-generation students                                  |      Sudent Body                |
-| MD_FAMINC        | Numeric           |Median family income in real 2015 dollars                             |       Earning                   |
-| PCTPELL          | Numeric           |Percentage of undergraduates who receive a Pell Grant                 |        Financial Aid            |
-| PCTFLOAN         | Numeric           |Percent of all federal undergraduate students receiving a federal student loan      |  Financial Aid    |
-| GRAD_DEBT_MDN    | Numeric           |The median debt for students who have completed                       |   Financial Aid                 |
-| WDRAW_DEBT_MDN   | Numeric           |The median debt for sutdents who have not completed                   |Financial Aid |
-| MD_EARN_WNE_P6   | Numeric           |Median income six years after entry                                   |     earning   |
-| MD_EARN_WNE_P10  | Numeric           |Median income ten years after entry                                   |     earning  |
-| RPY_3YR_RT_SUPP  | Numeric           |Repayment rate 3 years after entering repayment                       |Repayment   |
-| CDR3             | Numeric           |Loan default rate 3 years afer entering repayment                     |Repayment        |
-| variable$qreF    | Factor            |Predominant degree offered                                            |School       |
-| variable$ctrF    | Factor            |Control of institution                                                |School     |
-| variable$STABBR  | Factor            |State                                                                 |School     |
-| APPL_SCH_PCT_GE2 | Numeric           |Percentage of students submitting 2 or more than 2 applications       | Student        |
+| Variable Name    | Data Type |  Unit  |  Description                                                          |    Category in Documentation       |
+| --------------   | ----------|--------|-----------------------------------------------------------------------| --------------------------------|
+| NPT4_COMBINE    | Numeric   | $         |average net price for TITLE IV institutions of both public and private.|  Cost                       |
+| DEP_STAT_PCT_IND | Numeric  | $         |Percentage of students who are financially independent                |     Cost                        |
+| PAR_ED_PCT_1STGEN| Numeric  | %         |Percentage first-generation students                                  |      Sudent Body                |
+| MD_FAMINC        | Numeric  | $         |Median family income in real 2015 dollars                             |       Earning                   |
+| PCTPELL          | Numeric  | %         |Percentage of undergraduates who receive a Pell Grant                 |        Financial Aid            |
+| PCTFLOAN         | Numeric  | %         |Percent of all federal undergraduate students receiving a federal student loan      |  Financial Aid    |
+| GRAD_DEBT_MDN    | Numeric  | $         |The median debt for students who have completed                       |   Financial Aid                 |
+| WDRAW_DEBT_MDN   | Numeric  | $         |The median debt for sutdents who have not completed                   |Financial Aid |
+| MD_EARN_WNE_P6   | Numeric  | $         |Median income six years after entry                                   |     earning   |
+| MD_EARN_WNE_P10  | Numeric  | $         |Median income ten years after entry                                   |     earning  |
+| RPY_3YR_RT_SUPP  | Numeric  | %         |Repayment rate 3 years after entering repayment                       |Repayment   |
+| CDR3             | Numeric  | %         |Loan default rate 3 years afer entering repayment                     |Repayment        |
+| PREDDEG_factor    | Factor  | N/A          |Predominant degree offered                                            |School       |
+| Control_factor    | Factor  | N/A          |Control of institution                                                |School     |
+| STABBR_factor  | Factor     | N/A       |State                                                                 |School     |
+| APPL_SCH_PCT_GE2 | Numeric  | %         |Percentage of students submitting 2 or more than 2 applications       | Student        |
+| UNITID | Numeric | N/A      | Unique ID for the institutions | unknown |
+| INSTNM | Factor |  N/A      |name for the institutions       | unknown |
 

### 2017-02-19 Draft-xuezhou-sjb.Rmd
Update Draft-xuezhou-sjb.Rmd
update toc

@@
 * 5.[Summary of variables](#variablesummary)
- * 5.1 Individual variable summary
-  * 5.1.1 School
-    * 5.1.1.1 State
-    * 5.1.1.2 Predominant Degree
-    * 5.1.1.3 Control of Institution
+ * 5.1 [Individual variable summary](#5.1)
+  * 5.1.1 [School](#5.1.1)
+    * 5.1.1.1 [State](#5.1.1.1)
+    * 5.1.1.2 [Predominant Degree](#5.1.1.2)
+    * 5.1.1.3 [Control of Institution](#5.1.1.3)
   * 5.1.2 Student Body
     * 5.1.2.1 [Percentage first-generation students](#5.1.2.1)
-    * 5.1.2.2 Number of applications
+    * 5.1.2.2 [Number of applications](#5.1.2.2)
   * 5.1.3 Cost 
@@
     * 5.1.4.3 [The median debt for students who have completed](#5.1.4.3)
-    * 5.1.4.4 The median debt for students who have not completed
+    * 5.1.4.4 [The median debt for students who have not completed](#5.1.4.4)
   * 5.1.5 Earning
     * 5.1.5.1 [Median family income in real 2015 dollars](#5.1.5.1)
-    * 5.1.5.2 Earning after graduation
+    * 5.1.5.2 [Earning after graduation](#5.1.5.2)
   * 5.1.6 Repayment
-    * 5.1.6.1 3-year Repayment rate
-    * 5.1.6.2 3-year Default rate
+    * 5.1.6.1 [3-year Repayment rate](#5.1.6.1)
+    * 5.1.6.2 [3-year Default rate](#5.1.6.2)
 * 6.Relationship between variables
@@
  * 6.7 [The relationship between PCTLOAN and Control](#6.7)
- * 6.8 cost and median debts with earning
- * 6.9 default with earning
- * 6.10 degree type vs. earning
- * 6.11 predominant degree, control  and earning after 6 year of entry
- * 6.12 predominant degree + control (count)
- * 6.13 control + earning
- * 7. Findings and Conclusion
+ * 6.8 [cost and median debts with earning](#6.8)
+ * 6.9 [default with earning](#6.9)
+ * 6.10 [degree type vs. earning](#6.10)
+ * 6.11 [predominant degree, control  and earning after 6 year of entry](#6.11)
+ * 6.12 [predominant degree + control (count)](#6.12)
+ * 6.13 [control + earning](#6.13)
+* 7. [Findings and Conclusion](#7)
 
@@
 ###5.1.2 Student Body
-####5.1.2.1 Percentage first-generation students
+####5.1.2.1 Percentage first-generation students<a id="5.1.2.1"></a>
 This element identifies the share of federally aided students at the institution who report that they are first generaion on the FAFSA (Free Application for Federal Student Aid).
@@
 
-####5.1.2.2 Number of applications
+####5.1.2.2 Number of applications<a id="5.1.2.2"></a>
 The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who list two and more than two institutions when he apply for financial aid when apply for universities.
@@
 
-###5.1.3 Cost
-####5.1.3.1 Average net price for Title IV institutions(both public and private institutions)
+###5.1.3 Cost<a id="5.1.3"></a>
+####5.1.3.1 Average net price for Title IV institutions(both public and private institutions)<a id="5.1.3.1"></a>
 
@@
 
-####5.1.3.2 Percentage of students who are financially independent
+####5.1.3.2 Percentage of students who are financially independent<a id="5.1.3.2"></a>
 
@@
 ###5.1.4 Financial Aid 
-####5.1.4.1 Percentage of undergraduates who receive a Pell Grant
+####5.1.4.1 Percentage of undergraduates who receive a Pell Grant<a id="5.1.4.1"></a>
 
@@
 
-####5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan
+####5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan<a id="5.1.4.2"></a>
 
@@
 
-####5.1.4.3 The median debt for students who have completed
+####5.1.4.3 The median debt for students who have completed<a id="5.1.4.3"></a>
 
@@
 
-####5.1.5.2 Earning after graduation
+####5.1.5.2 Earning after graduation<a id="5.1.5.2"></a>
 We selected two variables MD_EARN_WNE_P6, MD_EARN_WNE_P10 in the earning group to study students' post-graduation earning.
@@
 ###5.1.6 Repayment 
-####5.1.6.1 3-year Repayment rate
+####5.1.6.1 3-year Repayment rate<a id="5.1.6.1"></a>
 The variable RPY_3YR_RT_SUPP measure the percentage of student who entered repayemtn from October 1, 2010 to September 30, 2011, and made progress in the following three years.
@@
 
-####5.1.6.2 3-year Default rate
+####5.1.6.2 3-year Default rate<a id="5.1.6.2"></a>
 The three-year cohort default rate describe the repayment in the opposite of repayment rate.

### 2017-02-19 Draft-xuezhou-sjb.Rmd

@@
 * 5.[Summary of variables](#variablesummary)
+
  * 5.1 [Individual variable summary](#5.1)

### 2017-02-19 Draft-xuezhou-sjb.Rmd
update

@@
 
-#5. Summary of variables<a id="variablesummary"></a>
-##5.1 Individual variable summary
+#5. Summary of variables<a id="5"></a>
+##5.1 Individual variable summary<a id="5.1"></a>
 
-###5.1.1 School
-####5.1.1.1. State
+###5.1.1 School<a id="5.1.1"></a> 
+####5.1.1.1. State<a id="5.1.1.1"></a> 
 State variable has 59 levels, which include 55 states in the United States, as well as.........
@@
 
-####5.1.1.2. Predominant Degree
+####5.1.1.2. Predominant Degree<a id="5.1.1.2"></a> 
 Then we analyze the type of degree universities mainly offer.
@@
 
-####5.1.1.3. Control of Institution
+####5.1.1.3. Control of Institution<a id="5.1.1.3"></a> 
 In addition to State and predominant degree, control of institution is another categorical variable in the dataset.
@@
 
-###5.1.2 Student Body
+###5.1.2 Student Body<a id="5.1.2"></a> 
 ####5.1.2.1 Percentage first-generation students<a id="5.1.2.1"></a>
@@
 
-###5.1.4 Financial Aid 
+###5.1.4 Financial Aid<a id="5.1.4"></a>  
 ####5.1.4.1 Percentage of undergraduates who receive a Pell Grant<a id="5.1.4.1"></a>
@@
 
-###5.1.5 Earning
+###5.1.5 Earning<a id="5.1.5"></a> 
 ####5.1.5.1 Median family income in real 2015 dollars<a id="5.1.5.1"></a>
@@
 
-##6.8 cost and median debts with earning
+##6.8 cost and median debts with earning<a id="6.8"></a> 
 Next, we analyze the relationship between the amount the debt and post-graduation earning.
@@
 
-##6.9 default with earning
+##6.9 default with earning<a id="6.9"></a>
 
@@
 
-##6.10 degree type vs. earning
+##6.10 degree type vs. earning<a id="6.10"></a>
 
@@
 
-##6.11 predominant degree, control  and earning after 6 year of entry
+##6.11 predominant degree, control  and earning after 6 year of entry<a id="6.11"></a> 
 The following plot shows two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry.
@@
 
-##6.12 predominant degree + control (count)
+##6.12 predominant degree + control (count)<a id="6.12"></a> 
 The following plot shows the relationship between two factor variables (predominant degree award variable and control variable).
@@
 
- #6.11 control + earning
+ #6.13 control + earning<a id="6.13"></a> 
 The following plot shows one factor variable (control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry).
@@
 
-#7. Findings and Conclusion
+#7. Findings and Conclusion<a id="7"></a>

### 2017-02-19 Draft-xuezhou-sjb.Rmd

@@
 * 5.[Summary of variables](#variablesummary)
-
 * 5.1 [Individual variable summary](#5.1)

### 2017-02-20 Draft-xuezhou-sjb.Rmd
update

--- 2017-02-19 Draft-xuezhou-sjb.Rmd
@@
 * 1.[Introduction](#Introduction)
-* 2.[Goal](#Goal)
-* 3.[Dataset Preparation](#dataprep)
-* 4.[Objective](#objective)
-* 5.[Summary of variables](#variablesummary)
+* 2.Goal
 
- * 5.1 [Individual variable summary](#5.1)
-  * 5.1.1 [School](#5.1.1)
-    * 5.1.1.1 [State](#5.1.1.1)
-    * 5.1.1.2 [Predominant Degree](#5.1.1.2)
-    * 5.1.1.3 [Control of Institution](#5.1.1.3)
+College is an investment.
+It's well-known that a major motivation for students to attend a university is higher post-graduation earning.
+So will your school make sure you get a return?
+With collge costs and student-debt levels reaching an all-time high, it's important to know which schools produce the highest earning alumni and what factors can help us predict post-graduate earnings for students in a university.
+
+The goal of our study is to understand what factors are predictive of students' post-graduate earning through individual variable analysis and relationship detection.
+To this end, we first explore a set of variables that are carefully selected to represent a school's profile, getting to know the distribution of each variable.
+Then we further analyze how these variables are related, particually the relationship between earning and other variables.
+
+* 3.Dataset Preparation
+* 4.Objective
+* 5.Summary of variables
+ * 5.1 Individual variable summary
+  * 5.1.1 School
+    * 5.1.1.1 State
+    * 5.1.1.2 Predominant Degree
+    * 5.1.1.3 Control of Institution
   * 5.1.2 Student Body
-    * 5.1.2.1 [Percentage first-generation students](#5.1.2.1)
-    * 5.1.2.2 [Number of applications](#5.1.2.2)
+    * 5.1.2.1 Percentage first-generation students
+    * 5.1.2.2 Number of applications
   * 5.1.3 Cost 
-    * 5.1.3.1 [Average net price for Title IV institutions(both public and private institutions)](#5.1.3.1)
-    * 5.1.3.2 [Percentage of students who are financially independent](#5.1.3.2)
+    * 5.1.3.1 Average net price for Title IV institutions(both public and private institutions)
+    * 5.1.3.2 Percentage of students who are financially independent
   * 5.1.4 Financial Aid
-    * 5.1.4.1 [Percentage of undergraduates who receive a Pell Grant](#5.1.4.1)
-    * 5.1.4.2 [Percent of all federal undergraduate students receiving a federal student loan](#5.1.4.2l)
-    * 5.1.4.3 [The median debt for students who have completed](#5.1.4.3)
-    * 5.1.4.4 [The median debt for students who have not completed](#5.1.4.4)
+    * 5.1.4.1 Percentage of undergraduates who receive a Pell Grant
+    * 5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan
+    * 5.1.4.3 The median debt for students who have completed
+    * 5.1.4.4 The median debt for students who have not completed
   * 5.1.5 Earning
-    * 5.1.5.1 [Median family income in real 2015 dollars](#5.1.5.1)
-    * 5.1.5.2 [Earning after graduation](#5.1.5.2)
+    * 5.1.5.1 Median family income in real 2015 dollars
+    * 5.1.5.2 Earning after graduation
   * 5.1.6 Repayment
-    * 5.1.6.1 [3-year Repayment rate](#5.1.6.1)
-    * 5.1.6.2 [3-year Default rate](#5.1.6.2)
+    * 5.1.6.1 3-year Repayment rate
+    * 5.1.6.2 3-year Default rate
 * 6.Relationship between variables
- * 6.1 [The relationship between NPT4_COMBINE and State](#6.1)
- * 6.2 [The relationship between Debt and Cost](#6.2)
- * 6.3 [The relationship between Percentage of Pell Grant and college type(CONTROL)](#6.3)
- * 6.4 [The relationship between Percentage first-generation students and States](#6.4)
- * 6.5 [The relationship between family income and percentage of Pell Grant](#6.5)
- * 6.6 [The relationship between DEP_STAT_PCT_IND and Control](#6.6)
- * 6.7 [The relationship between PCTLOAN and Control](#6.7)
- * 6.8 [cost and median debts with earning](#6.8)
- * 6.9 [default with earning](#6.9)
- * 6.10 [degree type vs. earning](#6.10)
- * 6.11 [predominant degree, control  and earning after 6 year of entry](#6.11)
- * 6.12 [predominant degree + control (count)](#6.12)
- * 6.13 [control + earning](#6.13)
-* 7. [Findings and Conclusion](#7)
+ * 6.1 NPT4_COMBINE VS State
+ * 6.2 Debt VS Cost
+ * 6.3 Percentage of Pell Grant VS college type(CONTROL)
+ * 6.4 Percentage first-generation students VS Statu
+ * 6.5 family income with Pell Grant
+ * 6.6 cost and median debts with earning
+ * 6.7 default with earning
+ * 6.8 degree type vs. earning
+ * 6.9 predominant degree, control  and earning after 6 year of entry
+ * 6.10 predominant degree + control (count)
+ * 6.11 control + earning
+* 7. Comparison between variables
+* 8. Findings and Conclusion
 
@@
 
-#2. Goal<a id="Goal"></a>
-#3. Dataset Preparation<a id="dataprep"></a>
+#2. Goal
+#3. Dataset Preparation
 Our analysis focuses on the data of year 2012 -2013 of 23 variables.
@@
 
-Considering there is some special characters in the ID coloumn, we first rename the first column's name to be selectable. 
-```{R}
-colnames(data)[1] = 'UNITID'
-```
-
 The full dataset contains 1743 variables. 36 variables were intially selected to analyze.
@@
 
-variable = select(data, UNITID, INSTNM, STABBR, PREDDEG, LOCALE2, MENONLY, WOMENONLY, CONTROL, ADM_RATE, SATMTMID, SATVRMID, SATWRMID, SAT_AVG,
+variable = select(data, STABBR, PREDDEG, LOCALE2, MENONLY, WOMENONLY, CONTROL, ADM_RATE, SATMTMID, SATVRMID, SATWRMID, SAT_AVG,
                   NPT4_PUB,
@@
 
-#need to be commented out 
-class(as.data.frame(apply(variable[,-c(2,3,4,8)],2,as.numeric)))
+class(as.data.frame(apply(variable[,-c(1,2,6)],2,as.numeric)))
 
-#get rid of factor variables: INSTNM, STABBR, PREDDEG, CONTROL and change the data types to numeric 
-variable[,-c(2,3,4,8)] %>%
+#numeric
+variable[,-c(1,2,6)] %>%
     apply(2, as.numeric) %>%
@@
 
-#change the categorical level of the categorical variables:PREDDEG, CONTROL
+#categorical
 variable$preF=factor(variable$PREDDEG,levels=c(0,1,2,3,4),labels = c("NotClassified","Certificate-degree","Associate's-degree","Bachelor's-degree","Graduate-degree"))
-
 variable$ctrF=factor(variable$CONTROL,levels = c(1,2,3),labels = c("Public","Private Nonprofit","Private For-profit"))
@@
 #combine the three datafram to generate a new dataset. 
-data_new <- cbind(variable_num, variable$preF, variable$ctrF,variable$INSTNM, variable$STABBR)
+data_new <- cbind(variable_num, variable$preF, variable$ctrF,variable$STABBR)
+dim(data_new)
 
-#rename the categorical varaibles to approriate readable names
-colnames(data_new)[35] = 'PREDDEG_factor'
-colnames(data_new)[36] = 'Control_factor'
-colnames(data_new)[37] = 'INSTNM_factor'
-colnames(data_new)[38] = 'STABBR_factor'
-
-#need to be commented out
 mean(is.na(data_new$NPT4_PUB))
@@
 mean(is.na(data_new$NPT4_PRIV))
-data_new = data_new[,-10] 
-colnames(data_new)[10] <- "NPT4_COMBINE"
-
-mean(is.na(data_new$NPT4_COMBINE))
+data_new = data_new[,-9] 
+colnames(data_new)[9] <- "NPT4_COMBINE"
 ```
@@
 In order to choose the variables that have enough valid data related to the topic, we filtered and considered variables with percent missing value less than 25%.
-The new data set was named as ???data_clean???
-for further analysis.
+The new data set was named as “data_clean” for further analysis.
 ```{r, message=FALSE,warning=FALSE}
@@
 data_clean = data_new[,colMeans(is.na(data_new)) < 0.25]
-
-#rearrange the coloumn names
-data_clean = data_clean[c(1,17,18,15,16,2:14)]
-
 write.csv(data_clean,"data_clean.csv")
 ```
-The data_clean will be the cleaned dataset we used in the further variable summaries and comparisons
 
@@
 Our data set now contains 120 rows and 15 columns.
-120 rows consist of 20 years???
-of data for each of the chosen six countries.
+120 rows consist of 20 years’ of data for each of the chosen six countries.
 The 15 columns consist of country name, country code, year, and the twelve indicators.
 
-| Variable Name    | Data Type |  Unit  |  Description                                                          |    Category in Documentation       |
-| --------------   | ----------|--------|-----------------------------------------------------------------------| --------------------------------|
-| NPT4_COMBINE    | Numeric   | $         |average net price for TITLE IV institutions of both public and private.|  Cost                       |
-| DEP_STAT_PCT_IND | Numeric  | $         |Percentage of students who are financially independent                |     Cost                        |
-| PAR_ED_PCT_1STGEN| Numeric  | %         |Percentage first-generation students                                  |      Sudent Body                |
-| MD_FAMINC        | Numeric  | $         |Median family income in real 2015 dollars                             |       Earning                   |
-| PCTPELL          | Numeric  | %         |Percentage of undergraduates who receive a Pell Grant                 |        Financial Aid            |
-| PCTFLOAN         | Numeric  | %         |Percent of all federal undergraduate students receiving a federal student loan      |  Financial Aid    |
-| GRAD_DEBT_MDN    | Numeric  | $         |The median debt for students who have completed                       |   Financial Aid                 |
-| WDRAW_DEBT_MDN   | Numeric  | $         |The median debt for sutdents who have not completed                   |Financial Aid |
-| MD_EARN_WNE_P6   | Numeric  | $         |Median income six years after entry                                   |     earning   |
-| MD_EARN_WNE_P10  | Numeric  | $         |Median income ten years after entry                                   |     earning  |
-| RPY_3YR_RT_SUPP  | Numeric  | %         |Repayment rate 3 years after entering repayment                       |Repayment   |
-| CDR3             | Numeric  | %         |Loan default rate 3 years afer entering repayment                     |Repayment        |
-| PREDDEG_factor    | Factor  | N/A          |Predominant degree offered                                            |School       |
-| Control_factor    | Factor  | N/A          |Control of institution                                                |School     |
-| STABBR_factor  | Factor     | N/A       |State                                                                 |School     |
-| APPL_SCH_PCT_GE2 | Numeric  | %         |Percentage of students submitting 2 or more than 2 applications       | Student        |
-| UNITID | Numeric | N/A      | Unique ID for the institutions | unknown |
-| INSTNM | Factor |  N/A      |name for the institutions       | unknown |
+| Variable Name    | Data Type         | Description                                                          |    Category in Documentation       |
+| --------------  | ------------------ | ---------------------------------------------------------------------| --------------------------------|
+| NPT4_COMBINE    | Numeric            |average net price for TITLE IV institutions of both public and private.|  Cost                       |
+| DEP_STAT_PCT_IND | Numeric           |Percentage of students who are financially independent                |     Cost                        |
+| PAR_ED_PCT_1STGEN| Numeric           |Percentage first-generation students                                  |      Sudent Body                |
+| MD_FAMINC        | Numeric           |Median family income in real 2015 dollars                             |       Earning                   |
+| PCTPELL          | Numeric           |Percentage of undergraduates who receive a Pell Grant                 |        Financial Aid            |
+| PCTFLOAN         | Numeric           |Percent of all federal undergraduate students receiving a federal student loan      |  Financial Aid    |
+| GRAD_DEBT_MDN    | Numeric           |The median debt for students who have completed                       |   Financial Aid                 |
+| WDRAW_DEBT_MDN   | Numeric           |The median debt for sutdents who have not completed                   |Financial Aid |
+| MD_EARN_WNE_P6   | Numeric           |Median income six years after entry                                   |     earning   |
+| MD_EARN_WNE_P10  | Numeric           |Median income ten years after entry                                   |     earning  |
+| RPY_3YR_RT_SUPP  | Numeric           |Repayment rate 3 years after entering repayment                       |Repayment   |
+| CDR3             | Numeric           |Loan default rate 3 years afer entering repayment                     |Repayment        |
+| variable$qreF    | Factor            |Predominant degree offered                                            |School       |
+| variable$ctrF    | Factor            |Control of institution                                                |School     |
+| variable$STABBR  | Factor            |State                                                                 |School     |
+| APPL_SCH_PCT_GE2 | Numeric           |Percentage of students submitting 2 or more than 2 applications       | Student        |
 
-we will summarize the data set using dplyr and ggplot funtions in the variables summary section.
+we will summarize the data set using dplyr functions in the variables summary section.
 
-#4. Objective<a id="objective"></a>
-Over years, Many students see college as an investment to help them earn more and live better lives after graduation.
-Amount of findints have been conducted to analyze the influence of college quality on students' post graudation earning based on *colleges characterestis, student's demographic information, as well as family charateristics as well as federal finanical support to low-income students.
-It has become more and more important to understand the fac- tors that contribute to post-graduation earnings and the ability to repay student loans.
-
-Our objective in this stage is to conduct an exploration analysis on the CollegeScore dataset, investigating the individual variable's summary statistics and the relationship between these variabls, so as to better prepare for further analaysis on the factors affecting students' post college earnings with some advanced data mining techniques.
-
-#5. Summary of variables<a id="variablesummary"></a>
+#4. Objective
+#5. Summary of variables
 ##5.1 Individual variable summary
@@
 ####5.1.1.1. State
-State variable has 59 levels, which include 55 states in the United States, as well as.........
-From the following frequncy table, we see that California has the most University-equivalent institutions across the nation, followed by Texas and Ner York.
-It has about 800 universities!
-A bar chart is plotted below to show the top states with the most universities.
-Then summary of frequency table also shows that the state having the least university only has one, and 50% of state has more than 132 universities in the state.
+State variable has 59 levels, which include 55 states in the United States and 4 districts outside mainland US.
+To better understand how many universities each state has, we create a frequency table as shown below, and also a bar chart to visuzalize the top 15 states having the most universities.
+Some basic summary statistics are also displayed.
 ```{r, message=FALSE}
@@
 state_freq = count(data_clean,"variable$STABBR")[order(-count(data_clean,'variable$STABBR')$freq),]
-state_freq[1:15,]
+state_freq
 summary(state_freq)
@@
 ```
+The frequency table counts the number of records belonging to each state, since each records represents a institution, each cell also equals to the number of university in that state.
+
+From frequncy table, we see that California has the most  institutions across the nation, followed by Texas and Ner York.
+It has about 800 universities!
+A bar chart is plotted below to show the top 15 states in the frequency table.
+Then summary of frequency table also shows that the state having the least university only has one, and 50% of state has more than 132 universities in the state.
 
@@
 There are four types of degree taken into accout: Certificate-degree, Associate's degree, Bachelor's degree, and Graduate Degree.
-About 5% of the population are classified in terms of the predomiant degree they offer.
-43.2% offer Certificate degree as their predominant degree, 27.3% predominantly offer Bachelor degree and only 4.23% are classified as Graduate degree predominant.
+A frequency table in a format of percentage and a histogram are used to help understand the distributiton of the Prodominant Degree.
 
@@
 ```
+From above frequency table, we see that about 5% of the population are unclassified in terms of the predomiant degree they offer.
+43.2% offer Certificate degree as their predominant degree, 27.3% predominantly offer Bachelor degree and only 4.23% are classified as Graduate degree predominant.
 
 ####5.1.1.3. Control of Institution
+
 In addition to State and predominant degree, control of institution is another categorical variable in the dataset.
-The label of Control is to indicate whether or not a institution is institution is privately owned and whether or not it is a for-profit organization.
+The label of Control is to indicate whether or not a institution is privately owned and whether or not it is a for-profit organization.
 There are three levels of this variable: Public, Private Nonprofit, and Private-profit.
-Public school is assumed to be non-profit so it was not further divided into two subgroups.
+Public school is assumed to be non-profit so it is not further divided into two subgroups.
 Below we show the percentage summary of each category and a bar char is plotted to facilitate understanding.
-About 50% out of all institution are private for-profit university, which is the *most dominant type*.
-Public and Private Nonprofit are pretty tie in terms of number of observation in each category.
 
@@
 
+About 50% out of all institution are private for-profit university, which is the *most dominant type*.
+Public and Private Nonprofit are pretty tiex in terms of the number of observation in each category.
+
 ###5.1.2 Student Body
-####5.1.2.1 Percentage first-generation students<a id="5.1.2.1"></a>
+####5.1.2.1 Percentage first-generation students
 This element identifies the share of federally aided students at the institution who report that they are first generaion on the FAFSA (Free Application for Federal Student Aid).
@@
 
-With the followin codes, we extracted the institions with the highest and lowest percentage of first-generation students.
-Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
+We are using the code below to summarize this variable.
+
 ```{R, message=FALSE, warinig = FALSE}
-
-#the institions with the highest and lowest percentage of first-generation students
-data_clean %>%
- slice(c(which.min(PAR_ED_PCT_1STGEN),which.max(PAR_ED_PCT_1STGEN))) %>%
-  select (c(1,2,3,8))
 
@@
 ```
-From the output, we can tell that the highest value existed in  La Belle Beauty Academy with the value of 95.25%.
-The lowest value is 6.87%, which existed in Wheaton College.
+From the output, we can tell that the highest value is 95.25% while the lowest value is 6.87%.
 From the mean value of 48.08%, we can tell that about for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college.
@@
 
-####5.1.2.2 Number of applications<a id="5.1.2.2"></a>
-The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who list two and more than two institutions when he apply for financial aid when apply for universities.
-This information is only available to students who submit their applications for federal aid, and not cover the whole student body.
-Nevertheless, this variable still can provides some useful insights about how many universities a student usually apply to.
+####5.1.2.2 Number of applications
 
-The summry statistics show that median percentage of student submitting two or more than two application is 41.67 across all univerisities.
-The institution having the least percentage only has 4% of student applying to more than one university, while the school with the highest percent has as high as 90% of student applying to more than one university.
-The boxplot compare the the percent of student applying to only one university and that of student applying to more than 1.
-We can clearly see that a larger portion of students only list one instituiton when submitting their application for financial aid.
+The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who list two and more than two institutions when he applied for financial aid when submitting application for university.
+This information is only available to students who submitted their applications for federal aid, and does not cover the whole student body.
+Nevertheless, this variable still can provides some useful insights about how many universities a student usually apply for.
 
@@
 
-###5.1.3 Cost<a id="5.1.3"></a>
-####5.1.3.1 Average net price for Title IV institutions(both public and private institutions)<a id="5.1.3.1"></a>
+The summry statistics show that median percentage of student submitting two or more than two application is 41.67% across all univerisities.
+The institution having the least percentage only has 4% of student applying to more than one university, while the school with the highest percent has as high as 90% of student applying to more than one university.
+The boxplot compares the the percent of student applying to only one university with that of student applying for more than one.
+We can clearly see that a larger portion of students only listed one instituiton when submitting their application for financial aid.
 
-With the followin codes, we extracted the IV institutions with the highest and lowest Average net price.
-Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
+###5.1.3 Cost
 
+####5.1.3.1 Average net price for Title IV institutions(both public and private institutions)
+
+We are using the code below to summarize the variable. 
 ```{R, message=FALSE, warinig = FALSE}
-
-data_clean %>%
- slice(c(which.min(NPT4_COMBINE),which.max(NPT4_COMBINE))) %>%
-  select (c(1,2,3,6))
 
@@
 ```
-From the summary results, we learned that the maximum value of the variable is $87,414, which is for Aerosim Flight Academy located in FL.
-while the minimum value is about $-19,360 - the value for Berea College located in KY.
+From the summary results, we learned that the maximum value of the variable is \$87,414 while the minimum value is about \$-19,360.
 Considering the elements are collected from Title IV institutions, we reasoned that the negative values may be due to the fact that some intituions's finainal aid supoort can cover the students' all the costs, meanwhile overbalancing the cost amount.
-There are not much differerences between mean value and median value, which is around $15,000.
+There are not much differerences between mean value and median value, which is around \$15,000.
 We can conclude that the average net price costs for most institutuins are falling into the this range.
@@
 
-####5.1.3.2 Percentage of students who are financially independent<a id="5.1.3.2"></a>
+####5.1.3.2 Percentage of students who are financially independent
 
-With the followin codes, we extracted the institions which have the highest and lowest percentage of students who are financially independent.
-Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
+Here is the code to generate the summary statistics: 
 ```{R}
-
-data_clean %>%
- slice(c(which.min(DEP_STAT_PCT_IND),which.max(DEP_STAT_PCT_IND))) %>%
-  select (c(1,2,3,7))
 
@@
 ```
-From the summary results, we learned that the maximum value of the variable is 98.56%(Capella University) while the minimum value is about 0.73%(University of Notre Dame).
+From the summary results, we learned that the maximum value of the variable is 98.56% while the minimum value is about 0，73%.
 From the mean value, we can tell that about in most colleges, around 56% of the students are financially independent.
@@
 Below is the code to generate a visual representation in the histgram.
-
 ```{R}
@@
 
+Then we grouped by the data into the control type(Private For-profit, Public, Private Nonprofit) to take a further analysis.
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(variable$ctrF) %>%
+    dplyr::summarise(minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND ,na.rm=TRUE),    
+    maxDEP_STAT_PCT_IND=max(DEP_STAT_PCT_IND,na.rm=TRUE),medianDEP_STAT_PCT_IND=median(DEP_STAT_PCT_IND,na.rm=TRUE))  %>%
+       arrange(desc(medianDEP_STAT_PCT_IND))   %>%
+{.}  ->DEP_STAT_PCT_IND_CONTROL
+
+DEP_STAT_PCT_IND_CONTROL
+
+```
+From the result, we can tell the highest median value is in Privately For-profit colleges, which means that in most private for-profit colleges, the students are more likely to be financial independent.
+
 ###5.1.4 Financial Aid 
-####5.1.4.1 Percentage of undergraduates who receive a Pell Grant<a id="5.1.4.1"></a>
+####5.1.4.1 Percentage of undergraduates who receive a Pell Grant
 
 This elemnt, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received Pell Grants in a given year, which is an important measure of the access an insiitution provides to low-income students.
-
-We are using the code below to summarize this variable and extraced the instition with the highest and lowest value.
+We are using the code below to summarize this variable.
 
 ```{R, message=FALSE, warinig = FALSE}
-
-data_clean %>%
- slice(c(which.min(PCTPELL),which.max(PCTPELL))) %>%
-  select (c(1,2,3,10))
 
@@
 ```
-From the output, we can tell that the range is between 0 to 100%.	No students in Thunderbird School of Global Management has received the Pell grant while all of the students in American Conservatory Theater have receiced the Pell Grant.
-Among the colleges in the sample, half colleges have this rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs.
+From the output, we can tell that the range is between 0 to 100%.	 Among the colleges in the sample, half colleges have this rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs.
 The standard deviation value is 22.57%,  indicating a relative high difference of this element among different colleges, which can also be noticed from the largest range (100%).
@@
 
-####5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan<a id="5.1.4.2"></a>
+####5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan
 
@@
 It can provide important context to figures related to debt, repayment and non-repayment.
-We are using the code below to summarize this variable and extracted the institutions with highest and lowest value.
+We are using the code below to summarize this variable.
 
 ```{R, message=FALSE, warinig = FALSE}
-
-data_clean %>%
- slice(c(which.min(PCTFLOAN),which.max(PCTFLOAN))) %>%
-  select (c(1,2,3,11))
 
@@
 ```
-From the output, we can tell that the range is between 0 to 100%.	As we can recall from the previous variable(PCTPELL), American Conservatory Theater has both the PCTLOAN and PCTPELL the value of 100%.
-Gadsden State Community College has no students reveiced federal loans in the given year.
-Among the colleges in the sample, half colleges have this rate as high as 60.08% , which means around 60% of undergraduare students are able to receive the fedaral student loans.
+From the output, we can tell that the range is between 0 to 100%.	 Among the colleges in the sample, half colleges have this rate as high as 60.08% , which means around 60% of undergraduare students are able to receive the fedaral student loans.
 The standard deviation value is 29.26%,  indicating a relative high difference of this element among different colleges, which can also be noticed from the largest range (100%).
@@
 
-####5.1.4.3 The median debt for students who have completed<a id="5.1.4.3"></a>
-
-This element is the median loan debt acculated at the institution by all student borrowers of federal lonans who have habe completed the education.
-We are using the code below to summarize this variable.
-```{R, message=FALSE, warinig = FALSE}
+Then we grouped by the data into the college type level to take a further analysis.
+```{R message=FALSE, warinig = FALSE}
 
 data_clean %>%
- slice(c(which.min(GRAD_DEBT_MDN),which.max(GRAD_DEBT_MDN))) %>%
-  select (c(1,2,3,12))
+  group_by(variable$STABBR ) %>%
+    dplyr::summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
+    maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
+       arrange(medianPCTFLOAN)   %>%
+{.}  ->PCTFLOAN_STATE
+
+PCTFLOAN_STATE
+
+data_clean %>%
+  group_by(variable$ctrF ) %>%
+    dplyr::summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
+    maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
+       arrange(medianPCTFLOAN)   %>%
+{.}  ->PCTFLOAN_CONTROL
+
+PCTFLOAN_CONTROL
+
+#stacked histogram
+data_clean %>%
+  ggplot(aes(x=PCTFLOAN)) +
+  geom_bar(binwidth=0.1,aes(fill=variable$ctrF ))
+
+```
+
+From the result, the Private institutions have a much larger value than the public institutions.
+With a much higher costs burden, it makes sence that the low-income sutdents in private institutions are much more likely to seek financial aid than the public institution students.
+
+####5.1.4.3 The median debt for students who have completed
+
+This element is the median loan debt acculated at the institution by all student borrowers of federal lonans who have habe completed the education.We are using the code below to summarize this variable.
+```{R, message=FALSE, warinig = FALSE}
 
@@
 ```
-From the output, we can tell that the highest value is 48,250 - Southwest University of Visual Arts-Tucson, while the lowest value is 1448 - South Florida Institute of Technology.
+From the output, we can tell that the highest value is 48,250	 while the lowest value is 1448.
 From the median value of 14000, we can tell that among the colleges in the sample, the students who have completed the deducaiton from half of them had the median loan debt of 14000.
@@
 
-####5.1.4.4 The median debt for students who have not completed<a id="5.1.4.4"></a>
-WDRAW_DEBT_MDN describe the median number of debt fot students who ahve not completed their study.
-In other words, they are still attending the university.
-The summary statistics show that student in 50% of institution have a median debt of 5500.The median debt for attending students can be as much as 27620 for some school and as low as 600 for other.
-So we also calculate standard deviation to measure the dispenser of the distribution.
-The standard deviation is 2415.623 about 95% of institution have a median debt ranging from 1167.754  to 10830.246.
-The histogram that the distribution of the median roughly follows a standard distribution with a sharp peak and slightly heavy tail.
+####5.1.4.4 The median debt for students who have not completed
+
+WDRAW_DEBT_MDN describes the median amount of debt fot students who ahve not completed their study. In other words, they are still attending the university.
+
 ```{r, message=FALSE}
@@
 ```
+We use summary statistics, standard deviation and histogram to understand the distribution of this variable.
+
+The summary statistics show that students in 50% of institution have a median debt of more than 5500, which is not very unaffordable.
+The median debt for attending students can be as much as 27620 for some school and as low as 600 for others, which indicates it may have a large variance.
+So we also calculate standard deviation to measure the dispersion of the distribution.
+The standard deviation is 2415.623 about 95% of institution have a median debt ranging from 1167.754  to 10830.246.
+The histogram shows that the distribution of the median roughly follows a standard distribution with a sharp peak and slightly heavy tail.
 
 ###5.1.5 Earning
-####5.1.5.1 Median family income in real 2015 dollars<a id="5.1.5.1"></a>
+####5.1.5.1 Median family income in real 2015 dollars
 This elemnt is recording the median family income of students, one of the undergraduate student demographics for earnings cohorts.
@@
 ```{R, message=FALSE, warinig = FALSE}
-
-data_clean %>%
- slice(c(which.min(MD_FAMINC),which.max(MD_FAMINC))) %>%
-  select (c(1,2,3,9))
 
@@
 ```
-From the output, we can tell that the highest value is 155,606.6 -Jewish Theological Seminary of America, while the lowest value is 0 -Associated Technical College-Los Angeles.
+From the output, we can tell that the highest value is 155,606.6	 while the lowest value is 0.
 From the median value of 19,262.26, we can tell that among the colleges in the sample, the students in half of them had the family income below 19,262.26	.
@@
 ```
+
 From the histogram, we are able to conclude that in most colleges, students' family income are falling into the range of 12,000 - 18,000.
@@
 
-####5.1.5.2 Earning after graduation<a id="5.1.5.2"></a>
+####5.1.5.2 Earning after graduation
+
 We selected two variables MD_EARN_WNE_P6, MD_EARN_WNE_P10 in the earning group to study students' post-graduation earning.
-MD_EARN_WNE_P6 is the median earning for students six years after their entry, which corresponds to two years after gradutaion.
-Likewise, MD_EARN_WNE_P10 is the median earning ten years after their entry, which corresponds to four years after graduation.
-These two metrics give us an idea of how much money students are making after their graduation, in both short and medium terms,
+MD_EARN_WNE_P6 is the median earning for students after six years of their entry, which corresponds to two years after gradutaion for a four-year program.
+Likewise, MD_EARN_WNE_P10 is the median earning ten years after their entry, which corresponds to four years after graduation usually.
+These two metrics give us an idea of how much money students are making after their graduation, in both short and medium terms.
 
-In addition to these two variables, a derived variable is also calculated as the difference between the salary 2 years post-graduation and another two years later.
-The median earning of 50% institutions is below the 25300 below, while the maximum median earining is 120200.
-After two years, the median earning increase by about $5500.
+In addition to these two variables, a derived variable is also calculated as the difference between  MD_EARN_WNE_P6 and MD_EARN_WNE_P10.
+The six-year-after-entry median earning of 50% institutions is below the 25300 below, while the maximum median earining is 120200, while ten-year-after-entry median earning is higher by about $5500.
 The histogram gives a more clear idea of the earning distribution.
-Most of institution with students having a median earining between 10000 to 50000 two years after graduation.
-The distribution of earining 4 years after graduation falls at the right of the distribution of 2 years post-graduation, which means that the earning increases as students' experience increase.
-Beside, the fout-year post-graduation earning has a wider spread than that of two years.
-That indicate that the there is a greater gap between who are earning more and who are earning less, which can be confirmed by the increase in standard deviation.
+Most of institution with students having a median earining between 10000 to 50000 six years after entry.
+The distribution of ten-year-after-entry falls at the right side of the six-year distribution , which means that the earning increases as students' experience increase.
+Beside, the ten-year post-graduation earning has a wider spread than that of six years.
+That indicate that the there is a greater gap between who are earning more and who are earning less, which can be confirmed by an increase in standard deviation.
 
@@
 ###5.1.6 Repayment 
-####5.1.6.1 3-year Repayment rate<a id="5.1.6.1"></a>
-The variable RPY_3YR_RT_SUPP measure the percentage of student who entered repayemtn from October 1, 2010 to September 30, 2011, and made progress in the following three years.
-Repayment rates are generally considered most sensitive than default rates(we study later), which measure only the worse-case scenario for repayment outcomes.
-From the summary statistics, we see that 50% institutions have 40% of students who are making progress in their repayment.
-A histogram is also plotted to show the distribution, clearly with a heavy tail.
+####5.1.6.1 3-year Repayment rate
+The variable RPY_3YR_RT_SUPP measures the percentage of students who entered repayemtn from October 1, 2010 to September 30, 2011, and made progress in the following three years.
+Repayment rates are generally considered most sensitive than default rates(we study later), which measures only the worse-case scenario for repayment outcomes.
 
@@
 
-####5.1.6.2 3-year Default rate<a id="5.1.6.2"></a>
-The three-year cohort default rate describe the repayment in the opposite of repayment rate.
-In this dataset, it was calculated using the cohort of borrowers who entered repayment on their federal student loans between Octorber 1, 2010 to September 30, 2011, and who defaulted before September 2012.
-The median default rate aross all universities is 13%, which means that 50% of universities have 13 out of 100 students defaulted their loan!
+From the summary statistics, we see that 50% institutions have 40% of students who were making progress in their repayment.
+A histogram is also plotted to show the distribution, clearly with a heavy tail.
+
+####5.1.6.2 3-year Default rate
+
+The three-year cohort default rate describes the repayment in the opposite of repayment rate.
+In this dataset, it was calculated using the cohort of borrowers who entered repayment on their federal student loans between Octorber 1, 2010 to September 30, 2011, and  defaulted before September 2012.
 
@@
 
-#6. Relationship between variables<a id="6"></a>
+The median default rate aross all universities is 13%, which means that more than 50% of universities have 13 out of 100 students defaulted their loan!
 
-##6.1. The relationship between NPT4_COMBINE and State<a id="6.1"></a>
+#6. Relationship between variables
+##6.1. NPT4_COMBINE VS State
 
 To analyze the relationship between average net cost for Title IV institutions and State, we grouped the data into the state level to take a further look.
-We would like to investigate the state distribution of the Title IV instituions based on the average net costs.
 ```{R message=FALSE, warinig = FALSE}
@@
 
-##6.2. The relationship between Debt and Cost<a id="6.2"></a>
+##6.2. Debt VS Cost
 We assume there would be a strong linear relationship between the average net price for Title IV institutions and the median debt for students who have completed.
@@
 
-##6.3. The relationship between Percentage of Pell Grant and college type(CONTROL)<a id="6.3"></a>
+##6.3. Percentage of Pell Grant VS college type(CONTROL)
 
-To compare the percentages of Pell Grant in different control type, we grouped the Percentage of Pell Grant data into the college type level to take a further analysis.
-We would like to investigate the percentage of Pell Grant based on the  college type.
+To compare the percentages of Pell Grant in different control type(), we grouped the Percentage of Pell Grant data into the college type level to take a further analysis.
 
@@
 {.}  ->PCTPELL_CONTROL
+
 PCTPELL_CONTROL
@@
 
-##6.4. The relationship between Percentage first-generation students and States<a id="6.4"></a>
-To compare the Percentage first-generation students with the different states, we grouped data into the state level to take a further analysis.We would like to investigate the percentage of first-generation students based on the status distribution.
+##6.4  Percentage first-generation students VS Status
+To compare the Percentage first-generation students with the different states, we grouped data into the state level to take a further analysis.
 
@@
 
-##6.5. The relationship between family income and percentage of Pell Grant<a id="6.5"></a>
+##6.5 family income with Pell Grant
 We assume there would be a strong linear relationship between family income and the percentage of students who can receive the Pell Grant.
@@
 From the scatter plot, it is obvious that the MD_FAMINC is in a negative linear relationship with PCTPELL: the data show a downhill pattern from left to right; As the MD_FAMINC increase, the value of PCTPELL decreases.
-It makes sense in the real world applcation: for institutions with low percentage of students who receive the Pell Grant, the financial aids like Pell Grant may not be as necessary as they are in the instituions with high value of PCTPELL, thus these collges students will tend to have a high median family income after they graduate.
 
-##6.6. The relationship between DEP_STAT_PCT_IND and Control<a id="6.6"></a>
+##6.6 cost and median debts with earning
 
-To compare the percentage of financially-independent students in institutions with different college type, we grouped by the data into the control type(Private For-profit, Public, Private Nonprofit) to take a further analysis.
-```{R message=FALSE, warinig = FALSE}
-
-data_clean %>%
-  group_by(variable$ctrF) %>%
-    dplyr::summarise(minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND ,na.rm=TRUE),    
-    maxDEP_STAT_PCT_IND=max(DEP_STAT_PCT_IND,na.rm=TRUE),medianDEP_STAT_PCT_IND=median(DEP_STAT_PCT_IND,na.rm=TRUE))  %>%
-       arrange(desc(medianDEP_STAT_PCT_IND))   %>%
-{.}  ->DEP_STAT_PCT_IND_CONTROL
-
-DEP_STAT_PCT_IND_CONTROL
-
-```
-From the result, we can tell the highest median value is in Privately For-profit colleges, which means that in most private for-profit colleges, the students are more likely to be financial independent.
-
-##6.7. The relationship between PCTLOAN and Control.<a id="6.7"></a>
-
-To compare the percentage of students who receive the federal loans in the given year based on the schools control type,we grouped by the data into the control type(Private For-profit, Public, Private Nonprofit) to take a further analysis.
-A stacked histogram is also requested to generate a more vidual presentation.
-
-```{R message=FALSE, warinig = FALSE}
-
-data_clean %>%
-  group_by(variable$ctrF ) %>%
-    dplyr::summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
-    maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
-       arrange(medianPCTFLOAN)   %>%
-{.}  ->PCTFLOAN_CONTROL
-
-PCTFLOAN_CONTROL
-
-#stacked histogram
-data_clean %>%
-  ggplot(aes(x=PCTFLOAN)) +
-  geom_bar(binwidth=0.1,aes(fill=Control_factor ))
-
-```
-
-From the result, the Private institutions have a much larger median value than the public institutions.
-With a much higher costs burden, it makes sence that the low-income sutdents in private institutions are much more likely to seek financial aid than the public institution students.
-From the stacked histogram, we can tell that for all the three college type institutions, the most frequent ranges appear between 0.6 and 0.8, which is consistent with the findings we observe in the individual variable summary section.
-
-##6.8 cost and median debts with earning
 Next, we analyze the relationship between the amount the debt and post-graduation earning.
-There are variable in the dataset to summarize the median amount for student body, one for students who already graduated and one for students who have not completed their study.
-We decide debts for students who have not completed to study in this problem, since it may can be used as a predictor in the future modeling.
-First, we use scatter plot to visualize the relationship between earning and the amount of debts, the plot clearly shows a positive relationship.
-Then we further calculate pearson correlation to confirm our hypothesis.
-The resulting coefficient 0.44 indicatesa moderate linear relationship between earning and debts.
+There are two variables in the dataset to summarize the median amount of debt for student body, one for students who already graduated and one for students who have not completed their study.
+We decide debts for students who have not completed to study in this problem, since it makes more sense for the purpose of prediction.
 
@@
 ```
+First, we use scatter plot to visualize the relationship between earning and the amount of debts, the plot clearly shows a positive relationship.
+Then we further calculate pearson correlation to confirm our hypothesis.
+The resulting coefficient 0.44 indicatesa moderate linear relationship between earning and debts.
 
-##6.9 default with earning
+##6.7 default with earning
 
 Understanding the relationship between the default rate of load repayment and post-graduation earning on one hand can help federal financial commitee to improve their decision about load granting.
-If the the default rate increases as the earning increase, the commitee may prefer to grant more financial aid to institution whose students have higher earning.
+If  the default rate decreases as the earning increase, the commitee may prefer to grant more financial aid to institution whose students have higher earning.
 One the other hand, the default rate of repayment for a univeristy can also help predict the post-graduation earning for its students.
-A scatter plot is created to visualized the relationship between these two variables.
-From the chart, we see the earning seem to show a decreasing trend as the default rate increases, which shows a negative relationship between earning and default rate.
-Then we create the correlation coefficient of these two variables.
-The resulting r = -0.33 shows that correlation between earning and default rate is negative but the relationship is not very strong, this may be due the influences of some outliers.
-To conclude, the negative relationship between earning and default rate of loan repayment needs larger population or detail data processing to confirm.
 
@@
 ```
+A scatter plot is created to visualize the relationship between these two variables.
+From the chart, we see the earning seems to show a decreasing trend as the default rate increases, which shows a negative relationship between earning and default rate.
+Then we calculate the correlation coefficient between these two variables.
+The resulting r = -0.33 shows that correlation between earning and default rate is negative but the relationship is not very strong, which may be due to the influences of some outliers.
+To conclude, the negative relationship between earning and default rate of loan repayment needs extra data to confirm.
 
-##6.10 degree type vs. earning
+##6.8 degree type vs. earning
 
@@
 In this question, we would like to study if it's worthwhile to invest in higher degree.
-From the boxplot below, we see that the higher the education, the higher the median earning 2 years after graduation.
-Especially, a graduate-degree give a hugh rise to one's earning.
 
@@
 
-##6.11 predominant degree, control  and earning after 6 year of entry
+From the boxplot below, we see that the higher the education, the higher the median earning 6 years after entry Especially, a graduate-degree gives a hugh rise to one's earning.
+
+##6.9 predominant degree, control  and earning after 6 year of entry
 The following plot shows two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry.
@@
 For the range of average earning, the darkest color shows the lowest earning, while the lightest color shows the highest value.
-According to the plot, we realize that ???Private For-profit college???
-whose predominant degree award is ???Graduate degree???
-has the lowest average earning with the minimum average value at \$3100, while ???Public college???
-whose predominant degree award is ???Graduate degree???
-has the highest average earning with the maximum average value at $45850.
+According to the plot, we realize that ‘Private For-profit college’ whose predominant degree award is ‘Graduate degree’ has the lowest average earning with the minimum average value at \$3100, while ‘Public college’ whose predominant degree award is ‘Graduate degree’ has the highest average earning with the maximum average value at $45850.
 
-##6.12 predominant degree + control (count)
+##6.10 predominant degree + control (count)
 The following plot shows the relationship between two factor variables (predominant degree award variable and control variable).
@@
 ```
-According to the histogram comparison of predominant degree award among three types of colleges, we realize that for public college, Associate???s-degree is most frequent awarded in the public colleges.
-Bachelor???s-degree is most frequent awarded in the private nonprofit colleges.
+According to the histogram comparison of predominant degree award among three types of colleges, we realize that for public college, Associate’s-degree is most frequent awarded in the public colleges.
+Bachelor’s-degree is most frequent awarded in the private nonprofit colleges.
 Certificate-degree is most frequent awarded in private for-profit colleges.
 
-Although the ???Graduate-degree???
-awarded by public colleges shows the least frequency, it gains the highest average earnings of students working and not enrolled 6 years after entry compared with other degrees awarded by different colleges.
+Although the ‘Graduate-degree’ awarded by public colleges shows the least frequency, it gains the highest average earnings of students working and not enrolled 6 years after entry compared with other degrees awarded by different colleges.
 
@@
 We use the
-geom_violin() using the default scale = ???area???
-argument and geom_boxplot() function to explore the relationship.
+geom_violin() using the default scale = “area” argument and geom_boxplot() function to explore the relationship.
 The plot superimposed with boxplots shows side-by-side violin plots of median earnings of students working and not enrolled 6 years after entry according to different college types.
@@
 Obvious differences contain in these three boxplots.
-The boxplot of ???Public college???
-is comparatively short, while the boxplot of ???Private Nonprofit???
-is comparatively tall.
-The ???Private For-profit college???
-has the lowest median earning.
+The boxplot of ‘Public college’ is comparatively short, while the boxplot of ‘Private Nonprofit’ is comparatively tall.
+The ‘Private For-profit college’ has the lowest median earning.
 
-#7. Findings and Conclusion
+#7. Comparison between variables
+ #7.1 PCTPELL VS PCTFLOAN
+
+```{r}
+
+library(ggplot2)
+ggplot(data = data_clean) +
+  geom_histogram(breaks=seq(0, 1, by =0.05),aes(x = PCTPELL, y=(..count..)), 
+                       alpha=0.9, fill ="red",position="dodge") +
+  geom_histogram(breaks=seq(0, 1, by =0.05),aes(x = PCTFLOAN, y=(..count..)), 
+                       alpha=0.9,fill ="blue",position="dodge") +
+  labs(title="Overlaying Histogram of PCTPELL and PCTFLOAN ") +
+    theme(plot.title = element_text(hjust = 0.5)) + 
+  scale_colour_manual("", 
+                      breaks = c("PCTPELL", "PCTFLOAN"),
+                      values = c("PCTPELL"="red", "PCTFLOAN"="blue"))
+
+```
+
+aid_compare -> data_clean[c(5:6)]
+class(aid_compare)
+
+#8. Findings and Conclusion

### 2017-02-20 Draft-xuezhou-sjb.Rmd

--- 2017-02-19 Draft-xuezhou-sjb.Rmd
@@
 * 1.[Introduction](#Introduction)
+
 * 2.[Goal](#Goal)
+
+College is an investment.
+It's well-known that a major motivation for students to attend a university is higher post-graduation earning.
+So will your school make sure you get a return?
+With collge costs and student-debt levels reaching an all-time high, it's important to know which schools produce the highest earning alumni and what factors can help us predict post-graduate earnings for students in a university.
+
+The goal of our study is to understand what factors are predictive of students' post-graduate earning through individual variable analysis and relationship detection.
+To this end, we first explore a set of variables that are carefully selected to represent a school's profile, getting to know the distribution of each variable.
+Then we further analyze how these variables are related, particually the relationship between earning and other variables.
+
 * 3.[Dataset Preparation](#dataprep)
@@
 #change the categorical level of the categorical variables:PREDDEG, CONTROL
-variable$preF=factor(variable$PREDDEG,levels=c(0,1,2,3,4),labels = c("NotClassified","Certificate-degree","Associate's-degree","Bachelor's-degree","Graduate-degree"))
+PREDDEG_factor=factor(variable$PREDDEG,levels=c(0,1,2,3,4),labels = c("NotClassified","Certificate-degree","Associate's-degree","Bachelor's-degree","Graduate-degree"))
 
-variable$ctrF=factor(variable$CONTROL,levels = c(1,2,3),labels = c("Public","Private Nonprofit","Private For-profit"))
+Control_factor=factor(variable$CONTROL,levels = c(1,2,3),labels = c("Public","Private Nonprofit","Private For-profit"))
 
 #combine the three datafram to generate a new dataset. 
-data_new <- cbind(variable_num, variable$preF, variable$ctrF,variable$INSTNM, variable$STABBR)
+data_new <- cbind(variable_num, PREDDEG_factor, Control_factor,variable$INSTNM, variable$STABBR)
 
@@
 ####5.1.1.1. State<a id="5.1.1.1"></a>
-State variable has 59 levels, which include 55 states in the United States, as well as.........
-From the following frequncy table, we see that California has the most University-equivalent institutions across the nation, followed by Texas and Ner York.
-It has about 800 universities!
-A bar chart is plotted below to show the top states with the most universities.
-Then summary of frequency table also shows that the state having the least university only has one, and 50% of state has more than 132 universities in the state.
+
+State variable has 59 levels, which include 55 states in the United States and 4 districts outside mainland US.
+To better understand how many universities each state has, we create a frequency table as shown below, and also a bar chart to visuzalize the top 15 states having the most universities.
+Some basic summary statistics are also displayed.
+
 ```{r, message=FALSE}
 
-state_freq = count(data_clean,"variable$STABBR")[order(-count(data_clean,'variable$STABBR')$freq),]
-state_freq[1:15,]
+state_freq = count(data_clean,"STABBR_factor")[order(-count(data_clean,'STABBR_factor')$freq),]
+state_freq
 summary(state_freq)
-state_freq[1:15,] %>% ggplot(aes(x=factor(variable.STABBR), y=freq)) +
+state_freq[1:15,] %>% ggplot(aes(x=STABBR_factor, y=freq)) +
   geom_bar(stat = 'identity')
@@
 There are four types of degree taken into accout: Certificate-degree, Associate's degree, Bachelor's degree, and Graduate Degree.
-About 5% of the population are classified in terms of the predomiant degree they offer.
+A frequency table in a format of percentage and a histogram are used to help understand the distributiton of the Prodominant Degree.
+
+```{r, message=FALSE}
+prop.table(table(data_clean$`PREDDEG_factor`)) * 100
+data_clean %>% ggplot(aes(x=PREDDEG_factor, fill = 'blue')) + 
+  geom_histogram(stat = "count") +
+  xlab("Degree") 
+```
+From above frequency table, we see that about 5% of the population are unclassified in terms of the predomiant degree they offer.
 43.2% offer Certificate degree as their predominant degree, 27.3% predominantly offer Bachelor degree and only 4.23% are classified as Graduate degree predominant.
 
+####5.1.1.3. Control of Institution<a id="5.1.1.3"></a> 
+In addition to State and predominant degree, control of institution is another categorical variable in the dataset.
+The label of Control is to indicate whether or not a institution is privately owned and whether or not it is a for-profit organization.
+There are three levels of this variable: Public, Private Nonprofit, and Private-profit.
+Public school is assumed to be non-profit so it is not further divided into two subgroups.
+Below we show the percentage summary of each category and a bar char is plotted to facilitate understanding.
+
 ```{r, message=FALSE}
-prop.table(table(data_clean$`variable$preF`)) * 100
-data_clean %>% ggplot(aes(x=variable$`preF`, fill = 'blue')) + 
+prop.table(table(data_clean$`Control_factor`)) * 100
+data_clean %>% ggplot(aes(x=Control_factor, fill = 'blue')) + 
   geom_histogram(stat = "count") +
@@
 
-####5.1.1.3. Control of Institution<a id="5.1.1.3"></a> 
-In addition to State and predominant degree, control of institution is another categorical variable in the dataset.
-The label of Control is to indicate whether or not a institution is institution is privately owned and whether or not it is a for-profit organization.
-There are three levels of this variable: Public, Private Nonprofit, and Private-profit.
-Public school is assumed to be non-profit so it was not further divided into two subgroups.
-Below we show the percentage summary of each category and a bar char is plotted to facilitate understanding.
 About 50% out of all institution are private for-profit university, which is the *most dominant type*.
-Public and Private Nonprofit are pretty tie in terms of number of observation in each category.
-
-```{r, message=FALSE}
-prop.table(table(data_clean$`variable$ctrF`)) * 100
-data_clean %>% ggplot(aes(x=variable$`ctrF`, fill = 'blue')) + 
-  geom_histogram(stat = "count") +
-  xlab("Degree")
-```
+Public and Private Nonprofit are pretty tiex in terms of the number of observation in each category.
 
@@
 ####5.1.2.2 Number of applications<a id="5.1.2.2"></a>
-The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who list two and more than two institutions when he apply for financial aid when apply for universities.
-This information is only available to students who submit their applications for federal aid, and not cover the whole student body.
-Nevertheless, this variable still can provides some useful insights about how many universities a student usually apply to.
 
-The summry statistics show that median percentage of student submitting two or more than two application is 41.67 across all univerisities.
-The institution having the least percentage only has 4% of student applying to more than one university, while the school with the highest percent has as high as 90% of student applying to more than one university.
-The boxplot compare the the percent of student applying to only one university and that of student applying to more than 1.
-We can clearly see that a larger portion of students only list one instituiton when submitting their application for financial aid.
+The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who list two and more than two institutions when he applied for financial aid when submitting application for university.
+This information is only available to students who submitted their applications for federal aid, and does not cover the whole student body.
+Nevertheless, this variable still can provides some useful insights about how many universities a student usually apply for.
 
@@
 ```
+
+The summry statistics show that median percentage of student submitting two or more than two application is 41.67% across all univerisities.
+The institution having the least percentage only has 4% of student applying to more than one university, while the school with the highest percent has as high as 90% of student applying to more than one university.
+The boxplot compares the the percent of student applying to only one university with that of student applying for more than one.
+We can clearly see that a larger portion of students only listed one instituiton when submitting their application for financial aid.
 
@@
 ####5.1.4.4 The median debt for students who have not completed<a id="5.1.4.4"></a>
-WDRAW_DEBT_MDN describe the median number of debt fot students who ahve not completed their study.
-In other words, they are still attending the university.
-The summary statistics show that student in 50% of institution have a median debt of 5500.The median debt for attending students can be as much as 27620 for some school and as low as 600 for other.
-So we also calculate standard deviation to measure the dispenser of the distribution.
-The standard deviation is 2415.623 about 95% of institution have a median debt ranging from 1167.754  to 10830.246.
-The histogram that the distribution of the median roughly follows a standard distribution with a sharp peak and slightly heavy tail.
+WDRAW_DEBT_MDN describes the median amount of debt fot students who ahve not completed their study. In other words, they are still attending the university.
+
 ```{r, message=FALSE}
@@
 ```
+We use summary statistics, standard deviation and histogram to understand the distribution of this variable.
+The summary statistics show that students in 50% of institution have a median debt of more than 5500, which is not very unaffordable.
+The median debt for attending students can be as much as 27620 for some school and as low as 600 for others, which indicates it may have a large variance.
+So we also calculate standard deviation to measure the dispersion of the distribution.
+The standard deviation is 2415.623 about 95% of institution have a median debt ranging from 1167.754  to 10830.246.
+The histogram shows that the distribution of the median roughly follows a standard distribution with a sharp peak and slightly heavy tail.
 
@@
 ####5.1.5.2 Earning after graduation<a id="5.1.5.2"></a>
+
 We selected two variables MD_EARN_WNE_P6, MD_EARN_WNE_P10 in the earning group to study students' post-graduation earning.
-MD_EARN_WNE_P6 is the median earning for students six years after their entry, which corresponds to two years after gradutaion.
-Likewise, MD_EARN_WNE_P10 is the median earning ten years after their entry, which corresponds to four years after graduation.
-These two metrics give us an idea of how much money students are making after their graduation, in both short and medium terms,
+MD_EARN_WNE_P6 is the median earning for students after six years of their entry, which corresponds to two years after gradutaion for a four-year program.
+Likewise, MD_EARN_WNE_P10 is the median earning ten years after their entry, which corresponds to four years after graduation usually.
+These two metrics give us an idea of how much money students are making after their graduation, in both short and medium terms.
 
-In addition to these two variables, a derived variable is also calculated as the difference between the salary 2 years post-graduation and another two years later.
-The median earning of 50% institutions is below the 25300 below, while the maximum median earining is 120200.
-After two years, the median earning increase by about $5500.
-The histogram gives a more clear idea of the earning distribution.
-Most of institution with students having a median earining between 10000 to 50000 two years after graduation.
-The distribution of earining 4 years after graduation falls at the right of the distribution of 2 years post-graduation, which means that the earning increases as students' experience increase.
-Beside, the fout-year post-graduation earning has a wider spread than that of two years.
-That indicate that the there is a greater gap between who are earning more and who are earning less, which can be confirmed by the increase in standard deviation.
+In addition to these two variables, a derived variable is also calculated as the difference between  MD_EARN_WNE_P6 and MD_EARN_WNE_P10.
 
@@
 
+The six-year-after-entry median earning of 50% institutions is below the 25300 below, while the maximum median earining is 120200, while ten-year-after-entry median earning is higher by about $5500.
+The histogram gives a more clear idea of the earning distribution.
+Most of institution with students having a median earining between 10000 to 50000 six years after entry.
+The distribution of ten-year-after-entry falls at the right side of the six-year distribution , which means that the earning increases as students' experience increase.
+Beside, the ten-year post-graduation earning has a wider spread than that of six years.
+That indicate that the there is a greater gap between who are earning more and who are earning less, which can be confirmed by an increase in standard deviation.
+
 ###5.1.6 Repayment
+
 ####5.1.6.1 3-year Repayment rate<a id="5.1.6.1"></a>
-The variable RPY_3YR_RT_SUPP measure the percentage of student who entered repayemtn from October 1, 2010 to September 30, 2011, and made progress in the following three years.
-Repayment rates are generally considered most sensitive than default rates(we study later), which measure only the worse-case scenario for repayment outcomes.
-From the summary statistics, we see that 50% institutions have 40% of students who are making progress in their repayment.
-A histogram is also plotted to show the distribution, clearly with a heavy tail.
+The variable RPY_3YR_RT_SUPP measures the percentage of students who entered repayemtn from October 1, 2010 to September 30, 2011, and made progress in the following three years.
+Repayment rates are generally considered most sensitive than default rates(we study later), which measures only the worse-case scenario for repayment outcomes.
 
@@
 
+From the summary statistics, we see that 50% institutions have 40% of students who were making progress in their repayment.
+A histogram is also plotted to show the distribution, clearly with a heavy tail.
+
 ####5.1.6.2 3-year Default rate<a id="5.1.6.2"></a>
-The three-year cohort default rate describe the repayment in the opposite of repayment rate.
-In this dataset, it was calculated using the cohort of borrowers who entered repayment on their federal student loans between Octorber 1, 2010 to September 30, 2011, and who defaulted before September 2012.
-The median default rate aross all universities is 13%, which means that 50% of universities have 13 out of 100 students defaulted their loan!
+The three-year cohort default rate describes the repayment in the opposite of repayment rate.
+In this dataset, it was calculated using the cohort of borrowers who entered repayment on their federal student loans between Octorber 1, 2010 to September 30, 2011, and  defaulted before September 2012.
 
@@
 ```
+
+The median default rate aross all universities is 13%, which means that more than 50% of universities have 13 out of 100 students defaulted their loan!
 
@@
 data_clean %>%
-  group_by(variable$STABBR) %>%
+  group_by(STABBR_factor) %>%
     dplyr::summarise(minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),    
@@
 data_clean %>%
-  group_by(variable$ctrF ) %>%
+  group_by(Control_factor ) %>%
     dplyr::summarise(minPCTPELL=min(PCTPELL ,na.rm=TRUE),    
@@
 
-ggplot(data_clean, aes(x=PCTPELL, fill=variable$ctrF)) + geom_histogram(alpha=0.9, position="identity") +
+ggplot(data_clean, aes(x=PCTPELL, fill=Control_factor)) + geom_histogram(alpha=0.9, position="identity") +
   labs(title="Overlaying histogram between PCTPELL and Control ") +
@@
 data_clean %>%
-  group_by(variable$STABBR ) %>%
+  group_by(STABBR_factor ) %>%
     dplyr::summarise(minPAR_ED_PCT_1STGEN=min(PAR_ED_PCT_1STGEN ,na.rm=TRUE),    
@@
 data_clean %>%
-  group_by(variable$ctrF) %>%
+  group_by(Control_factor) %>%
     dplyr::summarise(minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND ,na.rm=TRUE),    
@@
 data_clean %>%
-  group_by(variable$ctrF ) %>%
+  group_by(Control_factor ) %>%
     dplyr::summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
@@
 Next, we analyze the relationship between the amount the debt and post-graduation earning.
-There are variable in the dataset to summarize the median amount for student body, one for students who already graduated and one for students who have not completed their study.
-We decide debts for students who have not completed to study in this problem, since it may can be used as a predictor in the future modeling.
-First, we use scatter plot to visualize the relationship between earning and the amount of debts, the plot clearly shows a positive relationship.
-Then we further calculate pearson correlation to confirm our hypothesis.
-The resulting coefficient 0.44 indicatesa moderate linear relationship between earning and debts.
+There are two variables in the dataset to summarize the median amount of debt for student body, one for students who already graduated and one for students who have not completed their study.
+We decide debts for students who have not completed to study in this problem, since it makes more sense for the purpose of prediction.
 
@@
 ```
+First, we use scatter plot to visualize the relationship between earning and the amount of debts, the plot clearly shows a positive relationship.
+Then we further calculate pearson correlation to confirm our hypothesis.
+The resulting coefficient 0.44 indicatesa moderate linear relationship between earning and debts.
 
@@
 Understanding the relationship between the default rate of load repayment and post-graduation earning on one hand can help federal financial commitee to improve their decision about load granting.
-If the the default rate increases as the earning increase, the commitee may prefer to grant more financial aid to institution whose students have higher earning.
+If  the default rate decreases as the earning increase, the commitee may prefer to grant more financial aid to institution whose students have higher earning.
 One the other hand, the default rate of repayment for a univeristy can also help predict the post-graduation earning for its students.
-A scatter plot is created to visualized the relationship between these two variables.
-From the chart, we see the earning seem to show a decreasing trend as the default rate increases, which shows a negative relationship between earning and default rate.
-Then we create the correlation coefficient of these two variables.
-The resulting r = -0.33 shows that correlation between earning and default rate is negative but the relationship is not very strong, this may be due the influences of some outliers.
-To conclude, the negative relationship between earning and default rate of loan repayment needs larger population or detail data processing to confirm.
 
@@
 ```
+A scatter plot is created to visualize the relationship between these two variables.
+From the chart, we see the earning seems to show a decreasing trend as the default rate increases, which shows a negative relationship between earning and default rate.
+Then we calculate the correlation coefficient between these two variables.
+The resulting r = -0.33 shows that correlation between earning and default rate is negative but the relationship is not very strong, which may be due to the influences of some outliers.
+To conclude, the negative relationship between earning and default rate of loan repayment needs extra data to confirm.
 
@@
 data_clean %>%
-  ggplot(aes(x=`variable$preF`,
+  ggplot(aes(x=`PREDDEG_factor`,
              y=MD_EARN_WNE_P6)) +
-  geom_boxplot(aes(fill = `variable$preF`))
+  geom_boxplot(aes(fill = `PREDDEG_factor`))
 
 ```
+
+From the boxplot below, we see that the higher the education, the higher the median earning 6 years after entry Especially, a graduate-degree gives a hugh rise to one's earning.
 

### 2017-02-20 Draft-xuezhou-sjb.Rmd
delete pngs and fix goals

@@
 * 1.[Introduction](#Introduction)
-
 * 2.[Goal](#Goal)
-
-College is an investment.
-It's well-known that a major motivation for students to attend a university is higher post-graduation earning.
-So will your school make sure you get a return?
-With collge costs and student-debt levels reaching an all-time high, it's important to know which schools produce the highest earning alumni and what factors can help us predict post-graduate earnings for students in a university.
-
-The goal of our study is to understand what factors are predictive of students' post-graduate earning through individual variable analysis and relationship detection.
-To this end, we first explore a set of variables that are carefully selected to represent a school's profile, getting to know the distribution of each variable.
-Then we further analyze how these variables are related, particually the relationship between earning and other variables.
-
 * 3.[Dataset Preparation](#dataprep)
@@
 #2. Goal<a id="Goal"></a>
+College is an investment.
+It's well-known that a major motivation for students to attend a university is higher post-graduation earning.
+So will your school make sure you get a return?
+With collge costs and student-debt levels reaching an all-time high, it's important to know which schools produce the highest earning alumni and what factors can help us predict post-graduate earnings for students in a university.
+
+The goal of our study is to understand what factors are predictive of students' post-graduate earning through individual variable analysis and relationship detection.
+To this end, we first explore a set of variables that are carefully selected to represent a school's profile, getting to know the distribution of each variable.
+Then we further analyze how these variables are related, particually the relationship between earning and other variables.
+
 #3. Dataset Preparation<a id="dataprep"></a>

### 2017-02-21 Draft-xuezhou-sjb.Rmd
Add the second paragraph of intro part..

@@
 
-These data are collected by United States Department of Education with the goal of helping students and families make more educated decisions about college and compare vital information easily.
+The data is collected by United States Department of Education, which corresponds to all undergraduate degree-granting institutions of higher education from 1996 through 2016.
+The dataset is pretty rich and covers a wide range of topics.
+It has 1743 associated columns.
+Each corresponds to one named data variable.
+All data variables are divided into 10 categories ranging from ‘Root’ category to ‘Earnings’ category.
+The data type contains integer, string and float.
+Data elements with an ‘IPEDS’ value in the ‘Source’ field are derived directly from the Integrated Postsecondary Education Data System (IPEDS).
+For further detailed description of the dataset, you can find it in the CollegeScorecardDataDictionary.xlsx excel file.
 

### 2017-02-21 Draft-xuezhou-sjb.Rmd
added findings

@@
 Each corresponds to one named data variable.
-All data variables are divided into 10 categories ranging from ‘Root’ category to ‘Earnings’ category.
+All data variables are divided into 10 categories ranging from Root category to  Earnings category.
 The data type contains integer, string and float.
-Data elements with an ‘IPEDS’ value in the ‘Source’ field are derived directly from the Integrated Postsecondary Education Data System (IPEDS).
+Data elements with an IPEDS value in the Source field are derived directly from the Integrated Postsecondary Education Data System (IPEDS).
 For further detailed description of the dataset, you can find it in the CollegeScorecardDataDictionary.xlsx excel file.
@@
 
-#need to be commented out
-mean(is.na(data_new$NPT4_PUB))
 ```
@@
 }
+
 mean(is.na(data_new$NPT4_PRIV))
@@
 In order to choose the variables that have enough valid data related to the topic, we filtered and considered variables with percent missing value less than 25%.
-The new data set was named as ???data_clean???
-for further analysis.
+The new data set was named as data_clean for further analysis.
 ```{r, message=FALSE,warning=FALSE}
@@
 We also use visualizations such as hisgroms and boxplots to understand the distribution of the values.
-For categorical variables, we use frequency table to understand the frequcies of each categories.Box chart is used as a visualization tool.
+For categorical variables, we use frequency table to understand the frequcies of each categories.
+Box chart is used as a visualization tool.
 
-The table below lists the variables that will be used for the further analysis: 12 of them are continuous variables while 3 of them are categorical variables.
-Our data set now contains 120 rows and 15 columns.
-120 rows consist of 20 years???
-of data for each of the chosen six countries.
-The 15 columns consist of country name, country code, year, and the twelve indicators.
+The table below lists the variables that will be used for the further analysis: 13 of them are continuous variables while 5 of them are categorical variables or Unique ID.
+After initial data cleaning process, the new data set now contain 18 variables and 7793 observations.
 
@@
 | NPT4_COMBINE    | Numeric   | $         |average net price for TITLE IV institutions of both public and private.|  Cost                       |
-| DEP_STAT_PCT_IND | Numeric  | $         |Percentage of students who are financially independent                |     Cost                        |
+| DEP_STAT_PCT_IND | Numeric  | %         |Percentage of students who are financially independent                |     Cost                        |
 | PAR_ED_PCT_1STGEN| Numeric  | %         |Percentage first-generation students                                  |      Sudent Body                |
@@
 | APPL_SCH_PCT_GE2 | Numeric  | %         |Percentage of students submitting 2 or more than 2 applications       | Student        |
-| UNITID | Numeric | N/A      | Unique ID for the institutions | unknown |
-| INSTNM | Factor |  N/A      |name for the institutions       | unknown |
+| UNITID | Numeric | N/A      | Unique ID for the institutions | Root |
+| INSTNM | String |  N/A      |name for the institutions       | Institution |
 
@@
 Our objective in this stage is to conduct an exploration analysis on the CollegeScore dataset, investigating the individual variable's summary statistics and the relationship between these variabls, so as to better prepare for further analaysis on the factors affecting students' post college earnings with some advanced data mining techniques.
+The summary statistics, dplyr function, ggplot functions as well as some visualization techniques will be mainly used in this project.
 
@@
 
+#summarize the variable of PAR_ED_PCT_1STGEN
 summarize(data_clean,
@@
 The lowest value is 6.87%, which existed in Wheaton College.
-From the mean value of 48.08%, we can tell that about for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college.
+From the mean value of 48.08%, we can tell that for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college.
 The standard deviation value of 12.2% indicated a relative high difference of this element among different colleges.
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for PAR_ED_PCT_1STGEN") +
+    labs(title="Histogram for Percentage first-generation students") +
     labs(x="PAR_ED_PCT_1STGEN", y="Count") +
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for NPT_COM") +
-    labs(x="NPT_COM", y="Count") +
+    labs(title="Histogram for Average Net Price for Title IV institutions") +
+    labs(x="Average Net Price", y="Count") +
     theme(plot.title = element_text(hjust = 0.5))
 ```
-From the histogram, we notice that the average net cost for IV institutions typically is 16000 -18000.
-The values between 16000 - 20000 are quite frequent.
+From the histogram, we notice that the average net cost for IV institutions typically is $16000 - $18000.
+The values between $16000 - $20000 are quite frequent.
 There were an unsually lagrge number of costs.
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for DEP_STAT_PCT_IND") +
+    labs(title="Histogram for Percentage of Students Who are Financially Independent") +
     labs(x="DEP_STAT_PCT_IND", y="Count") +
@@
 This elemnt, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received Pell Grants in a given year, which is an important measure of the access an insiitution provides to low-income students.
-
 We are using the code below to summarize this variable and extraced the instition with the highest and lowest value.
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for PCTPELL") +
+    labs(title="Histogram for Percentage of Undergraduates who Receive a Pell Grant") +
     labs(x="PCTPELL", y="Count") +
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for PCTFLOAN") +
+    labs(title="Histogram for Percent of All Federal Undergraduate Students Receiving a Federal Student Loan") +
     labs(x="PCTFLOAN", y="Count") +
@@
 This element is the median loan debt acculated at the institution by all student borrowers of federal lonans who have habe completed the education.
-We are using the code below to summarize this variable.
+We are using the code below to summarize this variable???
+extracting the largest and lowest value for this variable.
 ```{R, message=FALSE, warinig = FALSE}
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for GRAD_DEBT_MDN") +
+    labs(title="Histogram for median debt for students who have completed") +
     labs(x="GRAD_DEBT_MDN", y="Count") +
@@
 ####5.1.5.1 Median family income in real 2015 dollars<a id="5.1.5.1"></a>
-This elemnt is recording the median family income of students, one of the undergraduate student demographics for earnings cohorts.
-
-We are using the code below to summarize this variable.
+This elemnt is recording the median family income of students, one of the undergraduate student demographics for earnings cohorts. We are using the code below to summarize this variable.
 
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for MD_FAMINC") +
+    labs(title="Histogram for Median family income in real 2015 dollars") +
     labs(x="MD_FAMINC", y="Count") +
@@
 
-##6.3. The relationship between Percentage of Pell Grant and college type(CONTROL)<a id="6.3"></a>
+##6.3. The relationship between Percentage of Pell Grant and College Type(CONTROL)<a id="6.3"></a>
 
-To compare the percentages of Pell Grant in different control type, we grouped the Percentage of Pell Grant data into the college type level to take a further analysis.
-We would like to investigate the percentage of Pell Grant based on the  college type.
+To compare the percentages of Pell Grant in different control type, we grouped the data into the college type level to take a further analysis.
 
@@
 
+Below is the stacked histogram between these two variables. 
 ```{r}
@@
 ggplot(data_clean, aes(x=PCTPELL, fill=Control_factor)) + geom_histogram(alpha=0.9, position="identity") +
-  labs(title="Overlaying histogram between PCTPELL and Control ") +
+  labs(title="Stacked histogram between PCTPELL and Control ") +
     theme(plot.title = element_text(hjust = 0.5))
@@
 From the result, the Private institutions have a much larger median value than the public institutions.
-With a much higher costs burden, it makes sence that the low-income sutdents in private institutions are much more likely to seek financial aid than the public institution students.
-From the stacked histogram, we can tell that for all the three college type institutions, the most frequent ranges appear between 0.6 and 0.8, which is consistent with the findings we observe in the individual variable summary section.
+With a much higher costs burden, it makes sense that the low-income sutdents in private institutions are much more likely to seek financial aid than the public institution students.
+
+From the stacked histogram, we can tell that for all the three college type institutions, the most frequent ranges appear between 0.6 and 0.8, which is consistent with the findings we observe in the previous individual variable summary section.
 
@@
 #7. Findings and Conclusion<a id="7"></a>
+
+Through the process of summarizing these variables and analyzing the relationship between these variables, we found that for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college.The average net price costs for most institutuins are falling into the range between $16000 - $20000.
+Among the 7703 institution samples, more than 50% of the students are able to receive Pell Grant or Federal Loans as means of finanical support???
+and on average, 60% of students are financial independent.
+In most colleges, the students are able to reach the family income between $12000 and $18000.
+
+Our findings also suggest that the Title IV institions located in DC, VT, MA, NH, RI have a relative high amount of average net costs than those in other states.
+The students in private non-profit institutions and public institutions are less likely to seek Pell Grant as financial assistance than the students in the private for-profit colleges.In most private for-profit colleges, the students are more likely to be financial independet.
+
+It is also indicated that there exists a negative correlation between the percentage of students who receive Pell Grant and the students median family income after they graduate: for institutions with low percentage of students who receive the Pell Grant, the financial aids like Pell Grant may not be as necessary as they are in the instituions with high value of PCTPELL, thus these collges students will tend to have a high median family income after they graduate.
+
+As a result of all the statistical analysis, we conclude that XXXXXXXXXXXX. Attributes such as                will affec the students' post graudate earnings.

### 2017-02-21 Draft-xuezhou-sjb.Rmd
findings

@@
 
-Through the process of summarizing these variables and analyzing the relationship between these variables, we found that for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college.The average net price costs for most institutuins are falling into the range between $16000 - $20000.
-Among the 7703 institution samples, more than 50% of the students are able to receive Pell Grant or Federal Loans as means of finanical support???
-and on average, 60% of students are financial independent.
+Through the process of summarizing these variables and analyzing the relationship between these variables, we found that for federally aided students in most colleges, around half of the students' parents did not complete the college.The average net price costs for most institutuins are falling into the range between $16000 - $20000.
+Among the 7703 institution samples, more than 50% of the students are able to receive Pell Grant or Federal Loans as means of finanical support, and on average, 60% of students are financial independent.
 In most colleges, the students are able to reach the family income between $12000 and $18000.

### 2017-02-21 Draft-xuezhou-sjb.Rmd
Xiang

--- 2017-02-21 Draft-xuezhou-sjb.Rmd
@@
 * 1.[Introduction](#Introduction)
+
 * 2.[Goal](#Goal)
+
 * 3.[Dataset Preparation](#dataprep)
@@
 
-The data is collected by United States Department of Education, which corresponds to all undergraduate degree-granting institutions of higher education from 1996 through 2016.
-The dataset is pretty rich and covers a wide range of topics.
-It has 1743 associated columns.
-Each corresponds to one named data variable.
-All data variables are divided into 10 categories ranging from ‘Root’ category to ‘Earnings’ category.
-The data type contains integer, string and float.
-Data elements with an ‘IPEDS’ value in the ‘Source’ field are derived directly from the Integrated Postsecondary Education Data System (IPEDS).
-For further detailed description of the dataset, you can find it in the CollegeScorecardDataDictionary.xlsx excel file.
+These data are collected by United States Department of Education with the goal of helping students and families make more educated decisions about college and compare vital information easily.
 
 #2. Goal<a id="Goal"></a>
+
 College is an investment.
@@
 state_freq[1:15,] %>% ggplot(aes(x=STABBR_factor, y=freq)) +
-  geom_bar(stat = 'identity')
+  geom_bar(stat = 'identity') +
+  xlab('State') +
+  ylab('Number of Institution') +
+  ggtitle('Number of Institution in Each State') +
+  theme(plot.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=20, hjust=0.5))
 
@@
   geom_histogram(stat = "count") +
-  xlab("Degree") 
+  xlab("Degree")  +
+  ylab('Count') +
+  ggtitle('Number of Institution in Each Predominant Degree type') +
+  theme(plot.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=20, hjust=0.5)) 
 ```
@@
   geom_histogram(stat = "count") +
-  xlab("Degree")
+  xlab("Type") +
+  ylab('Count') +
+  ggtitle('Number of Institution in Each School type') +
+  theme(plot.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=20, hjust=0.5)) 
 ```
@@
 
+college_app$num_app = as.factor(college_app$num_app)
+levels(college_app$num_app)[levels(college_app$num_app)=="college_app_1"] <- "one"
+
 college_app %>%
     ggplot(aes(factor(num_app), perc_student)) +
-    geom_boxplot() 
+    geom_boxplot() +
+    xlab("Number of application submitted") +
+    ylab("Percentage of Student") +
+    labs(title="Percentage of Students Submitting One/Two or Greater Applications") +
+    theme(plot.title = element_text(hjust = 0.5))
+
 ```
@@
     geom_density(alpha = 0.5) + 
-    xlim(0, 150000)
+    xlim(0, 150000) +
+    xlab("Median Earning") +
+    scale_fill_discrete(name="Years from Entry",
+                         breaks=c("MD_EARN_WNE_P10", "MD_EARN_WNE_P6"),
+                         labels=c("10 years", "6 years"))
 
@@
 summary(repayment_clean)
-qplot(repayment_clean, geom = "histogram")
+qplot(repayment_clean, geom = "histogram") +
+xlab("Repayment Rate") +
+ylab("Count") +
+labs(title="Histogram for Repayment Rate") +
+theme(plot.title = element_text(hjust = 0.5))
+
 ```
@@
 summary(default_clean)
-qplot(default_clean, geom = "histogram")
+qplot(default_clean, geom = "histogram") +
+xlab("Default Rate") +
+ylab("Count") +
+labs(title="Histogram for Default Rate") +
+theme(plot.title = element_text(hjust = 0.5))
 
@@
 
-##6.8 cost and median debts with earning<a id="6.8"></a> 
+##6.8 Median debts with earning<a id="6.8"></a> 
 Next, we analyze the relationship between the amount the debt and post-graduation earning.
@@
                 geom_jitter() +
-                geom_smooth(method = "lm", se = FALSE, col = 'red')
+                geom_smooth(method = "lm", se = FALSE, col = 'red') +
+                xlab("Median Amount Debt for Students who have not completed") +
+                ylab("Earning  6 Years from Entry")
 
@@
                 geom_jitter() +
-                geom_smooth(method = "lm", se = FALSE, col = 'red')
+                geom_smooth(method = "lm", se = FALSE, col = 'red')+
+                xlab("Three-year Cohort Default Rate") +
+                ylab("Earning  6 Years from Entry")
 
@@
              y=MD_EARN_WNE_P6)) +
-  geom_boxplot(aes(fill = `PREDDEG_factor`))
+  geom_boxplot(aes(fill = `PREDDEG_factor`)) +
+  xlab("Degree Type") +
+  ylab("Earning  6 Years from Entry") +
+  scale_fill_discrete(name="Degree Type")
 
+S
 ```

### 2017-02-21 Draft-xuezhou-sjb.Rmd
update the R code from 6.11 to 6.13
update the R code from 6.11 to 6.13  and the 1st and 2nd paragraph of 6.11 and 6.12

@@
 * 1.[Introduction](#Introduction)
-
 * 2.[Goal](#Goal)
-
 * 3.[Dataset Preparation](#dataprep)
@@
 
-These data are collected by United States Department of Education with the goal of helping students and families make more educated decisions about college and compare vital information easily.
+The data is collected by United States Department of Education, which corresponds to all undergraduate degree-granting institutions of higher education from 1996 through 2016.
+The dataset is pretty rich and covers a wide range of topics.
+It has 1743 associated columns.
+Each corresponds to one named data variable.
+All data variables are divided into 10 categories ranging from ‘Root’ category to ‘Earnings’ category.
+The data type contains integer, string and float.
+Data elements with an ‘IPEDS’ value in the ‘Source’ field are derived directly from the Integrated Postsecondary Education Data System (IPEDS).
+For further detailed description of the dataset, you can find it in the CollegeScorecardDataDictionary.xlsx excel file.
 
 #2. Goal<a id="Goal"></a>
-
 College is an investment.
@@
 state_freq[1:15,] %>% ggplot(aes(x=STABBR_factor, y=freq)) +
-  geom_bar(stat = 'identity') +
-  xlab('State') +
-  ylab('Number of Institution') +
-  ggtitle('Number of Institution in Each State') +
-  theme(plot.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=20, hjust=0.5))
+  geom_bar(stat = 'identity')
 
@@
   geom_histogram(stat = "count") +
-  xlab("Degree")  +
-  ylab('Count') +
-  ggtitle('Number of Institution in Each Predominant Degree type') +
-  theme(plot.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=20, hjust=0.5)) 
+  xlab("Degree") 
 ```
@@
   geom_histogram(stat = "count") +
-  xlab("Type") +
-  ylab('Count') +
-  ggtitle('Number of Institution in Each School type') +
-  theme(plot.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=20, hjust=0.5)) 
+  xlab("Degree")
 ```
@@
 
-college_app$num_app = as.factor(college_app$num_app)
-levels(college_app$num_app)[levels(college_app$num_app)=="college_app_1"] <- "one"
-
 college_app %>%
     ggplot(aes(factor(num_app), perc_student)) +
-    geom_boxplot() +
-    xlab("Number of application submitted") +
-    ylab("Percentage of Student") +
-    labs(title="Percentage of Students Submitting One/Two or Greater Applications") +
-    theme(plot.title = element_text(hjust = 0.5))
-
+    geom_boxplot() 
 ```
@@
     geom_density(alpha = 0.5) + 
-    xlim(0, 150000) +
-    xlab("Median Earning") +
-    scale_fill_discrete(name="Years from Entry",
-                         breaks=c("MD_EARN_WNE_P10", "MD_EARN_WNE_P6"),
-                         labels=c("10 years", "6 years"))
+    xlim(0, 150000)
 
@@
 summary(repayment_clean)
-qplot(repayment_clean, geom = "histogram") +
-xlab("Repayment Rate") +
-ylab("Count") +
-labs(title="Histogram for Repayment Rate") +
-theme(plot.title = element_text(hjust = 0.5))
-
+qplot(repayment_clean, geom = "histogram")
 ```
@@
 summary(default_clean)
-qplot(default_clean, geom = "histogram") +
-xlab("Default Rate") +
-ylab("Count") +
-labs(title="Histogram for Default Rate") +
-theme(plot.title = element_text(hjust = 0.5))
+qplot(default_clean, geom = "histogram")
 
@@
 
-##6.8 Median debts with earning<a id="6.8"></a> 
+##6.8 cost and median debts with earning<a id="6.8"></a> 
 Next, we analyze the relationship between the amount the debt and post-graduation earning.
@@
                 geom_jitter() +
-                geom_smooth(method = "lm", se = FALSE, col = 'red') +
-                xlab("Median Amount Debt for Students who have not completed") +
-                ylab("Earning  6 Years from Entry")
+                geom_smooth(method = "lm", se = FALSE, col = 'red')
 
@@
                 geom_jitter() +
-                geom_smooth(method = "lm", se = FALSE, col = 'red')+
-                xlab("Three-year Cohort Default Rate") +
-                ylab("Earning  6 Years from Entry")
+                geom_smooth(method = "lm", se = FALSE, col = 'red')
 
@@
              y=MD_EARN_WNE_P6)) +
-  geom_boxplot(aes(fill = `PREDDEG_factor`)) +
-  xlab("Degree Type") +
-  ylab("Earning  6 Years from Entry") +
-  scale_fill_discrete(name="Degree Type")
+  geom_boxplot(aes(fill = `PREDDEG_factor`))
 
-S
 ```
@@
 ##6.11 predominant degree, control  and earning after 6 year of entry<a id="6.11"></a> 
-The following plot shows two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry.
-Our team uses the function geom_tile()[ggplot2 package] to visualize the correlation matrix.
+The following plot shows two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry).
+
+For the R code, we first group the dataset by PREDDEG factor and Control factor.
+Then, we summarize the average earnings.
+After that, we set the aesthetic mapping between Control_factor and PREDDEG_factor as the horizontal axis and the vertical axis.
+Then, we use the function geom_tile()[ggplot2 package] to create a heat map to visualize the correlation between variables.
+The default color scheme maps a very dark blue color for the lowest abundance values, up to a very light blue for the highest abundance values.
 
@@
 data_clean %>%
- group_by(preF, ctrF) %>%
+group_by(PREDDEG_factor, Control_factor) %>%
 summarize(avg.earn_p6=mean(MD_EARN_WNE_P6)) %>%
- ggplot(aes(x=ctrF, y=preF)) +
-  geom_tile(aes(fill=avg.earn_p6))
+ggplot(aes(x=Control_factor, y=PREDDEG_factor)) +
+geom_tile(aes(fill=avg.earn_p6)) +
+ggtitle("Heat map between Average Earnings and PREDDEG&Control Factors")
 
@@
 The following plot shows the relationship between two factor variables (predominant degree award variable and control variable).
-Our team uses the function geom_bar() to map the counting values to y, ctrF variable to x and the preF variable to the fill of the bar.
+
+For the R code, our team first set the aesthetic mapping on Control_factor as the horizontal axis.
+Then, we use the function geom_bar() to fill different control types using the counting values of PREDDEG factor.
+After that, we further use the function xtabs() to calculate the exact number for each predominant degree awarded by different types of colleges.
 
@@
 data_clean %>%
-+ ggplot(aes(x=ctrF))+
-+ geom_bar(aes(fill=preF), position = 'dodge')
+ggplot(aes(x=Control_factor))+
+geom_bar(aes(fill=PREDDEG_factor), position = 'dodge')+
+ggtitle("Bar plot between PREDDEG and Control Factor")
+
 ```
@@
 data_clean %>%
- xtabs(formula = ~preF+ctrF)
+xtabs(formula = ~PREDDEG_factor+Control_factor)
+
 ```
@@
   data_clean %>%
-+ ggplot(aes(x=ctrF, y=MD_EARN_WNE_P6)) +
-+ geom_violin(aes(fill=ctrF), scale = 'area') +
-+ geom_boxplot(aes(fill=ctrF), width = 0.25, outlier.size = 1.25)
++ ggplot(aes(x=Control_factor, y=MD_EARN_WNE_P6)) +
++ geom_violin(aes(fill=Control_factor), scale = 'area')+ 
++ geom_boxplot(aes(fill=Control_factor), width = 0.25, outlier.size = 1.25)+
++ ggtitle("Violin plot between Control Factor and Median Earnings")
+
+```
+```{r}
+summary(data_clean$MD_EARN_WNE_P6[data_clean$Control_factor=="Public"])
+summary(data_clean$MD_EARN_WNE_P6[data_clean$Control_factor=="Private Nonprofit"])
+summary(data_clean$MD_EARN_WNE_P6[data_clean$Control_factor=="Private For-profit"])
 ```

### 2017-02-21 Draft-xuezhou-sjb.Rmd

--- 2017-02-21 Draft-xuezhou-sjb.Rmd
@@
 #2. Goal<a id="Goal"></a>
+
 College is an investment.
@@
              y=MD_EARN_WNE_P6)) +
-  geom_boxplot(aes(fill = `PREDDEG_factor`))
-
+  geom_boxplot(aes(fill = `PREDDEG_factor`)) +
+  xlab("Degree Type") +
+  ylab("Earning  6 Years from Entry") +
+  scale_fill_discrete(name="Degree Type")
+
+S
 ```
@@
 
+We also found that California has the most universities across the nation, with 797 institutions.
+In terms of predominant degree offered, 43.2% are classified as Certificate-degree predominant institution, which is the most predominant type.
+
 Our findings also suggest that the Title IV institions located in DC, VT, MA, NH, RI have a relative high amount of average net costs than those in other states.
@@
 
-As a result of all the statistical analysis, we conclude that XXXXXXXXXXXX. Attributes such as                will affec the students' post graudate earnings.
+As to the factors that influence post-graduate earning, we explore the relationship between median earning six years from entry and amount of debt, default rate for student loan and predominant degree type.
+The analysis suggests that there is a positive relationship between the median debt amount for students who have not completed and post-graduation earnings.
+This may be due to the fact that top universities usually have tuitions but may also indicates that students who are willing to take more debts for their college education are likely to make more money after graduation.
+In addition, default rate of student loan is negatively related to post-graduate earning.
+This finding is reasonable since students making more money are more likely to pay off their debt.
+The variable can be used a school-level predictor to predict overall post-graduate earning for each university.
+The degree type, unsurprisingly, has an impact on the earnings students are making after graduation.
+As the level of graduation increases, the post-graduate earning also increase significantly, which suggests that investing in education is worthwhile.
+
+As a result of all the statistical analysis, we conclude that XXXXXXXXXXXX. Attributes such as               the students' post graudate earnings.

### 2017-02-21 Draft-xuezhou-sjb.Rmd

--- 2017-02-21 Draft-xuezhou-sjb.Rmd
@@
 #7. Findings and Conclusion<a id="7"></a>
+
+Through the process of summarizing these variables and analyzing the relationship between these variables, we found that for federally aided students in most colleges, around half of the students' parents did not complete the college.The average net price costs for most institutuins are falling into the range between $16000 - $20000.
+Among the 7703 institution samples, more than 50% of the students are able to receive Pell Grant or Federal Loans as means of finanical support, and on average, 60% of students are financial independent.
+In most colleges, the students are able to reach the family income between $12000 and $18000.
+
+We also found that California has the most universities across the nation, with 797 institutions.
+In terms of predominant degree offered, 43.2% are classified as Certificate-degree predominant institution, which is the most predominant type.
+
+Our findings also suggest that the Title IV institions located in DC, VT, MA, NH, RI have a relative high amount of average net costs than those in other states.
+The students in private non-profit institutions and public institutions are less likely to seek Pell Grant as financial assistance than the students in the private for-profit colleges.In most private for-profit colleges, the students are more likely to be financial independet.
+
+It is also indicated that there exists a negative correlation between the percentage of students who receive Pell Grant and the students median family income after they graduate: for institutions with low percentage of students who receive the Pell Grant, the financial aids like Pell Grant may not be as necessary as they are in the instituions with high value of PCTPELL, thus these collges students will tend to have a high median family income after they graduate.
+
+As to the factors that influence post-graduate earning, we explore the relationship between median earning six years from entry and amount of debt, default rate for student loan and predominant degree type.
+The analysis suggests that there is a positive relationship between the median debt amount for students who have not completed and post-graduation earnings.
+This may be due to the fact that top universities usually have tuitions but may also indicates that students who are willing to take more debts for their college education are likely to make more money after graduation.
+In addition, default rate of student loan is negatively related to post-graduate earning.
+This finding is reasonable since students making more money are more likely to pay off their debt.
+The variable can be used a school-level predictor to predict overall post-graduate earning for each university.
+The degree type, unsurprisingly, has an impact on the earnings students are making after graduation.
+As the level of graduation increases, the post-graduate earning also increase significantly, which suggests that investing in education is worthwhile.
+
+As a result of all the statistical analysis, we conclude that XXXXXXXXXXXX. Attributes such as               the students' post graudate earnings.

### 2017-02-21 Draft-xuezhou-sjb.Rmd
update 02211209

@@
 Each corresponds to one named data variable.
-All data variables are divided into 10 categories ranging from ‘Root’ category to ‘Earnings’ category.
+All data variables are divided into 10 categories ranging from ???Root???
+category to ???Earnings???
+category.
 The data type contains integer, string and float.
-Data elements with an ‘IPEDS’ value in the ‘Source’ field are derived directly from the Integrated Postsecondary Education Data System (IPEDS).
+Data elements with an ???IPEDS???
+value in the ???Source???
+field are derived directly from the Integrated Postsecondary Education Data System (IPEDS).
 For further detailed description of the dataset, you can find it in the CollegeScorecardDataDictionary.xlsx excel file.
@@
 }
-mean(is.na(data_new$NPT4_PRIV))
+
 data_new = data_new[,-10] 
@@
 
-mean(is.na(data_new$NPT4_COMBINE))
 ```
@@
 We also use visualizations such as hisgroms and boxplots to understand the distribution of the values.
-For categorical variables, we use frequency table to understand the frequcies of each categories.Box chart is used as a visualization tool.
-
-The table below lists the variables that will be used for the further analysis: 12 of them are continuous variables while 3 of them are categorical variables.
-Our data set now contains 120 rows and 15 columns.
-120 rows consist of 20 years???
-of data for each of the chosen six countries.
-The 15 columns consist of country name, country code, year, and the twelve indicators.
+For categorical variables, we use frequency table to understand the frequcies of each categories.
+Box chart is used as a visualization tool.
+
+The table below lists the variables that will be used for the further analysis: 13 of them are continuous variables while 5 of them are categorical variables or Unique ID.
+After initial data cleaning process, the new data set now contain 18 variables and 7793 observations.
 
@@
 | APPL_SCH_PCT_GE2 | Numeric  | %         |Percentage of students submitting 2 or more than 2 applications       | Student        |
-| UNITID | Numeric | N/A      | Unique ID for the institutions | unknown |
-| INSTNM | Factor |  N/A      |name for the institutions       | unknown |
+| UNITID | Numeric | N/A      | Unique ID for the institutions | Root  |
+| INSTNM | Factor |  N/A      | name for the institutions       | Institution |
 
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for NPT_COM") +
+    labs(title="Histogram for Percentage first-generation students") +
     labs(x="NPT_COM", y="Count") +
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for DEP_STAT_PCT_IND") +
+    labs(title="Histogram for Percentage of students who are financially independent") +
     labs(x="DEP_STAT_PCT_IND", y="Count") +
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for PCTPELL") +
+    labs(title="Histogram for Percentage of undergraduates who receive a Pell Grant") +
     labs(x="PCTPELL", y="Count") +
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for PCTFLOAN") +
+    labs(title="Histogram for Percent of all federal undergraduate students receiving a federal student loan") +
     labs(x="PCTFLOAN", y="Count") +
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for GRAD_DEBT_MDN") +
+    labs(title="Histogram for The median debt for students who have completed") +
     labs(x="GRAD_DEBT_MDN", y="Count") +
@@
 
-##6.1. The relationship between NPT4_COMBINE and State<a id="6.1"></a>
+##6.1. The relationship between average net cost for Title IV institutions and State<a id="6.1"></a>
 
@@
 
-##6.2. The relationship between Debt and Cost<a id="6.2"></a>
+##6.2. The relationship between median debt and average net price for Title IV institutions<a id="6.2"></a>
 We assume there would be a strong linear relationship between the average net price for Title IV institutions and the median debt for students who have completed.
@@
 ggplot(data_clean, aes(x =NPT4_COMBINE, y = GRAD_DEBT_MDN)) + geom_point(color="red")+  geom_smooth(method = "lm", se = TRUE)+
-  labs(title="Scatterplot between Cost and Debt ") +
+  labs(title="Scatterplot between median debt and average net price for Title IV institutions ") +
     theme(plot.title = element_text(hjust = 0.5))
@@
 To compare the percentages of Pell Grant in different control type, we grouped the Percentage of Pell Grant data into the college type level to take a further analysis.
-We would like to investigate the percentage of Pell Grant based on the  college type.
 
@@
   ggplot(aes(x=PCTFLOAN)) +
-  geom_bar(binwidth=0.1,aes(fill=Control_factor ))
+  labs(title="Stacked histgram between Control and PCTLOAN ")+
+  geom_bar(binwidth=0.1,aes(fill=Control_factor )) + theme(plot.title = element_text(hjust = 0.5))
 
@@
 
-S
 ```
@@
 ```{r}
+
 data_clean %>%
   group_by(PREDDEG_factor, Control_factor) %>%
-summarize(avg.earn_p6=mean(MD_EARN_WNE_P6)) %>%
+    dplyr::summarize(avg.earn_p6=mean(MD_EARN_WNE_P6,na.rm=TRUE))%>%
       ggplot(aes(x=Control_factor, y=PREDDEG_factor)) +
@@
   data_clean %>%
-+ ggplot(aes(x=Control_factor, y=MD_EARN_WNE_P6)) +
-+ geom_violin(aes(fill=Control_factor), scale = 'area')+ 
-+ geom_boxplot(aes(fill=Control_factor), width = 0.25, outlier.size = 1.25)+
-+ ggtitle("Violin plot between Control Factor and Median Earnings")
+  ggplot(aes(x=Control_factor, y=MD_EARN_WNE_P6)) +
+  geom_violin(aes(fill=Control_factor), scale = 'area')+ 
+  geom_boxplot(aes(fill=Control_factor), width = 0.25, outlier.size = 1.25)+
+  ggtitle("Violin plot between Control Factor and Median Earnings")
 
@@
 
-As a result of all the statistical analysis, we conclude that XXXXXXXXXXXX. Attributes such as               the students' post graudate earnings.
+As a result of all the statistical analysis, we conclude that the attributes such as percentage of students who receive Pell Grant, median earning six years from entry, amount of debt, default rate for student loan and predominant degree type will affect the students' post graudate earnings.

### 2017-02-21 Draft-xuezhou-sjb.Rmd
fix formats and depress the messages from output

@@
 ---
-title: "Draft-assinment1"
+title: "Exploratory Data Analysis for College Scorecard Dataset"
 author: "Kun Shen, Xiang Li, Xue Zhou"
-date: "February 10, 2017"
+date: "February 20, 2017"
 output: html_document
@@
 ```{r setup, include=FALSE,warning=FALSE, message = FALSE}
-knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
+knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message = FALSE)
 ```
@@
 
-#need to be commented out 
-class(as.data.frame(apply(variable[,-c(2,3,4,8)],2,as.numeric)))
-
 #get rid of factor variables: INSTNM, STABBR, PREDDEG, CONTROL and change the data types to numeric 
@@
 ```
-The data_clean will be the cleaned dataset we used in the further variable summaries and comparisons
-
-For numeric variabls, we calculate the basic summary statistics such as median, min, max, mode and quantile.
+
+The data_clean will be the cleaned dataset we used in the further variable summaries and comparisons.For numeric variabls, we calculate the basic summary statistics such as median, min, max, mode and quantile.
 We also use visualizations such as hisgroms and boxplots to understand the distribution of the values.
@@
 ```
+
 The frequency table counts the number of records belonging to each state, since each records represents a institution, each cell also equals to the number of university in that state.
@@
 ```
+
 From above frequency table, we see that about 5% of the population are unclassified in terms of the predomiant degree they offer.
@@
 ```
+
 From the histogram, we can tell the share of first generarion students is most frequently falling into the range between 0.5 and 0.6.
@@
 ```
+
 From the summary results, we learned that the maximum value of the variable is $87,414, which is for Aerosim Flight Academy located in FL.
@@
 ```
+
 From the histogram, we notice that the average net cost for IV institutions typically is 16000 -18000.
@@
 ```
+
 From the summary results, we learned that the maximum value of the variable is 98.56%(Capella University) while the minimum value is about 0.73%(University of Notre Dame).
@@
 ```
+
 From the histogram, we notice that the value between 0.7 to 0.8 are quite frequent. There was no obvious extrme value in the sample, thus the histogram is normally distributed.
@@
 ```
+
 From the output, we can tell that the range is between 0 to 100%.	No students in Thunderbird School of Global Management has received the Pell grant while all of the students in American Conservatory Theater have receiced the Pell Grant.
@@
 ```
+
 From the histogram, we are able to conclude that in most colleges, 40% to 50% of the undergraduate students are able to receive the Pell Grant.
@@
 ```
+
 From the output, we can tell that the range is between 0 to 100%.	As we can recall from the previous variable(PCTPELL), American Conservatory Theater has both the PCTLOAN and PCTPELL the value of 100%.
@@
 ```
+
 From the histogram, it's obvious that there are two peaks in the graph. The rates between 0- 10% and the rates between 60% to 80% are similarly frequent compated to the other ranges.
@@
 ```
+
 From the output, we can tell that the highest value is 48,250 - Southwest University of Visual Arts-Tucson, while the lowest value is 1448 - South Florida Institute of Technology.
@@
 ```
+
 From the histogram, it is obvious that most students in the sampled colleges still have 8000- 10000 debt amount after they completed the education.
@@
 ```
+
 We use summary statistics, standard deviation and histogram to understand the distribution of this variable.
@@
 ```
+
 From the output, we can tell that the highest value is 155,606.6 -Jewish Theological Seminary of America, while the lowest value is 0 -Associated Technical College-Los Angeles.
@@
 ```
+
 Baesd on the output, we can see the highest median value for avreage net price exists in DC while the lowest median value is in AS(American Samoa).Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are all states conceiving amount of well-known educational institutions.
@@
 ```
+
 From the result, the Private Nonprofit institutions and Public institutions have a similar low median values while the provate For- profit institutions have a hgihest value.
@@
 ggplot(data_clean, aes(x=PCTPELL, fill=Control_factor)) + geom_histogram(alpha=0.9, position="identity") +
-  labs(title="Overlaying histogram between PCTPELL and Control ") +
-    theme(plot.title = element_text(hjust = 0.5))
-
-```
+  labs(title="Stacked Histogram between PCTPELL and Control ") +
+    theme(plot.title = element_text(hjust = 0.5))
+
+```
+
 From the overlaying histograms, the percentage of students who receive Pell Grant are quite frequent for Private For-profit institutions between 0.7 and 0.75 ; the percentage of students who receive Pell Grant are quite frequent around 0.4 for Public institutions and Private Nonprofit institutions.
@@
 ##6.4. The relationship between Percentage first-generation students and States<a id="6.4"></a>
-To compare the Percentage first-generation students with the different states, we grouped data into the state level to take a further analysis.We would like to investigate the percentage of first-generation students based on the status distribution.
+To compare the Percentage first-generation students with the different states, we grouped data into the state level to take a further analysis.
+We would like to investigate the percentage of first-generation students based on the status distribution.
 
@@
 ```
+
 From the scatter plot, it is obvious that the MD_FAMINC is in a negative linear relationship with PCTPELL: the data show a downhill pattern from left to right; As the MD_FAMINC increase, the value of PCTPELL decreases.
@@
 ```
+
 From the result, we can tell the highest median value is in Privately For-profit colleges, which means that in most private for-profit colleges, the students are more likely to be financial independent.
@@
 ```
+
 First, we use scatter plot to visualize the relationship between earning and the amount of debts, the plot clearly shows a positive relationship.
@@
 ```
+
 A scatter plot is created to visualize the relationship between these two variables.
@@
 ```
+
 Mean value of the earning is more representative for the two categories, preF and ctrF.
@@
 ```
+
 According to the histogram comparison of predominant degree award among three types of colleges, we realize that for public college, Associate???s-degree is most frequent awarded in the public colleges.
@@
 ```
+
 Obvious differences contain in these three boxplots.

### 2017-02-21 Draft-xuezhou-sjb.Rmd
update

@@
 #1. Introduction<a id="Introduction"></a>
-The College Scorecard project is designed to increase the transparency of the college data, working as a navigator for students to make informed decisions about their choice for higher education.
+The College Scorecard project was initially designed to increase the transparency of the college data, working as a navigator for students to make informed decisions about their choice for higher education.
 The information gathered in the dataset also gives students and families more power to differentiate postsecondary institutions and get an idea about how well colleges are preparing their students to be successful.
-The dataset is created to address the lack of information about college costs, earnings, federal student aid, admission rate etc.
+The dataset was created to address the lack of information about college costs, earnings, federal student aid, admission rate, etc.
 The data provides insights into the performance of institutions that receive federal aid dollars, and the outcomes of the students of those institutions.
-It also gives extra support for students and families to weight the tradeoffs on choosing the right college according to their own needs and educational goals.
-
-The data is collected by United States Department of Education, which corresponds to all undergraduate degree-granting institutions of higher education from 1996 through 2016.
+It also gives extra support for students and families to weigh the tradeoffs on choosing the right college according to their own needs and educational goals.
+
+The data was collected by the United States Department of Education, which corresponds to all undergraduate degree-granting institutions of higher education from 1996 through 2016.
 The dataset is pretty rich and covers a wide range of topics.
@@
 Each corresponds to one named data variable.
-All data variables are divided into 10 categories ranging from ???Root???
-category to ???Earnings???
-category.
+All data variables are divided into 10 categories ranging from Root category to Earnings category.
 The data type contains integer, string and float.
-Data elements with an ???IPEDS???
-value in the ???Source???
-field are derived directly from the Integrated Postsecondary Education Data System (IPEDS).
-For further detailed description of the dataset, you can find it in the CollegeScorecardDataDictionary.xlsx excel file.
+Data elements with an IPEDS value in the Source field are derived directly from the Integrated Postsecondary Education Data System (IPEDS).
+Further detailed description of the dataset can be found in the CollegeScorecardDataDictionary.xlsx excel file.
 
@@
 College is an investment.
-It's well-known that a major motivation for students to attend a university is higher post-graduation earning.
+It's well-known that a major motivation for students to attend a university is higher post-graduation earnings.
 So will your school make sure you get a return?
-With collge costs and student-debt levels reaching an all-time high, it's important to know which schools produce the highest earning alumni and what factors can help us predict post-graduate earnings for students in a university.
-
-The goal of our study is to understand what factors are predictive of students' post-graduate earning through individual variable analysis and relationship detection.
-To this end, we first explore a set of variables that are carefully selected to represent a school's profile, getting to know the distribution of each variable.
-Then we further analyze how these variables are related, particually the relationship between earning and other variables.
+With college costs and student-debt levels reaching an all-time high, it's important to know which schools produce the highest earning alumni and what factors can help us predict post-graduation earnings for students in a university.
+
+The goal of our study is to understand what factors are predictive of students' post-graduation earnings through individual variable analysis and relationship detection.
+To this end, we first explored a set of variables that were carefully selected to represent a school's profile, getting to know the distribution of each variable.
+Then we further analyzed how these variables are related, particually the relationship between earnings and other variables.
 
@@
 Our analysis focuses on the data of year 2012 -2013 of 23 variables.
-We chose the median income 6 years after graduation as our predictor variable, and eliminated many fields with bunch of missing values as well as some fields not pertainig enougn to post-graduation income.
-We then fosued primarily on a set of 23 variables, including gender, age, ethnic, acedemic costs, family income, completetion rates and replayment rate.
+We chose the median income 6 years after entry as our predictor variable, and eliminated many fields with a bunch of missing values as well as some fields not pertaining enougn to post-graduation income.
+We then fosued primarily on a set of 23 variables, including age, acedemic costs, family income, completetion rates and repayment rate.
 
@@
 
-We loaded ScoreCard from the US Dapartment of Education.
-The main dataset spans nearly 20 years of data and covering multiple sources including IPEDS, NSLDS, and Department of Treasury.
-To ensure the maximum data accurary and availablity, we choose the subset of year 2012 - 2013 as our analysis sample.
+The dataset spans nearly 20 years of data and covers multiple sources including IPEDS, NSLDS, and Department of Treasury.
+Since our goal was to study post-graduation for students in each university, we used 2012-2013 dataset which contains tha most recent earning information.
+
 ```{r, message=FALSE,warning=FALSE}
@@
 
-Considering there is some special characters in the ID coloumn, we first rename the first column's name to be selectable. 
+Considering there is some special characters in the ID column, we first renamed the first column to make it easier to understand.
 ```{R}
@@
 
-The full dataset contains 1743 variables. 36 variables were intially selected to analyze.
+The full dataset contains 1743 variables. 36 variables were intially selected to be analyze.
 ```{r, message=FALSE,warning=FALSE}
@@
 
-After loading the dataset, we change variables to the correct data types. 
+After loading the dataset, we converted variables to the correct data types.
+
 ```{r, message=FALSE,warning=FALSE}
@@
 
-Considering the variable NPT4_PUB and NPT4_PRIV are mutually exclusive, to maximize the data integrity, we integrated the information of the two coloumns into a single cololumn, named with "NPT4_COMBINE".
-The new coloumn records the average net price for TITLE IV institutions of both public and private.
+Considering the variable NPT4_PUB and NPT4_PRIV are mutually exclusive both indicate net price.
+we integrated the information of the two coloumns into a single column, named as NPT4_COMBINE.
+The new column records the average net price for both private and public TITLE IV institutions.
 ```{r, message=FALSE,warning=FALSE}
@@
 
-In order to choose the variables that have enough valid data related to the topic, we filtered and considered variables with percent missing value less than 25%.
-The new data set was named as ???data_clean???
-for further analysis.
+In order to choose the variables that have enough valid data related to the topic, we only considered variables with percent missing value less than 25%. The new data set was named as data_clean. 
 ```{r, message=FALSE,warning=FALSE}
@@
 
-The data_clean will be the cleaned dataset we used in the further variable summaries and comparisons.For numeric variabls, we calculate the basic summary statistics such as median, min, max, mode and quantile.
-We also use visualizations such as hisgroms and boxplots to understand the distribution of the values.
-For categorical variables, we use frequency table to understand the frequcies of each categories.
-Box chart is used as a visualization tool.
+The data_clean will be the clean dataset we used in the further variable summary and comparison.For numerical variables, we calculate the basic summary statistics such as median, min, max, mode and quantile.
+We also use visualizations such as historgams and boxplots to understand the distribution of the values.
+For categorical variables, we use frequency table to understand the frequencies of each category.
 
 The table below lists the variables that will be used for the further analysis: 13 of them are continuous variables while 5 of them are categorical variables or Unique ID.
-After initial data cleaning process, the new data set now contain 18 variables and 7793 observations.
+After initial data cleaning process, the new data set now contains 18 variables and 7793 observations.
 
@@
 
-we will summarize the data set using dplyr and ggplot funtions in the variables summary section.
+we will summarize the data set using functions provided in dplyr and ggplot funtions in the variables summary section.
 
 #4. Objective<a id="objective"></a>
-Over years, Many students see college as an investment to help them earn more and live better lives after graduation.
-Amount of findints have been conducted to analyze the influence of college quality on students' post graudation earning based on *colleges characterestis, student's demographic information, as well as family charateristics as well as federal finanical support to low-income students.
-It has become more and more important to understand the fac- tors that contribute to post-graduation earnings and the ability to repay student loans.
-
-Our objective in this stage is to conduct an exploration analysis on the CollegeScore dataset, investigating the individual variable's summary statistics and the relationship between these variabls, so as to better prepare for further analaysis on the factors affecting students' post college earnings with some advanced data mining techniques.
+Over the year, many students see college as an investment to help them earn more and live better lives after graduation.
+A large number of research has been conducted to analyze the influence of college quality on students' post-graudation earnings based on colleges' characteristics, students' demographic information, family charateristics as well as federal finanical support to low-income students.
+It has become more and more important to understand the factors that contribute to post-graduation earnings and the ability to repay student loans.
+
+Our objective at this stage is to conduct an exploration analysis on the CollegeScore dataset, investigating the individual variable's summary statistics and the relationship between these variabls, so as to better prepare for further analaysis on the factors affecting students' post college earnings with some advanced data mining techniques.
 The summary statistics, dplyr function, ggplot functions as well as some visualization techniques will be mainly used in this project.
@@
 State variable has 59 levels, which include 55 states in the United States and 4 districts outside mainland US.
-To better understand how many universities each state has, we create a frequency table as shown below, and also a bar chart to visuzalize the top 15 states having the most universities.
+To better understand how many universities each state has, we created a frequency table as shown below, and also a bar chart to visuzalize the top 15 states having the most universities.
 Some basic summary statistics are also displayed.
@@
 
-The frequency table counts the number of records belonging to each state, since each records represents a institution, each cell also equals to the number of university in that state.
-
-From frequncy table, we see that California has the most  institutions across the nation, followed by Texas and Ner York.
+The frequency table counts the number of records belonging to each state, since each record represents a institution, each cell also equals to the number of university in that state.
+
+From the frequncy table, we see that California has the most institutions across the nation, followed by Texas and Ner York.
 It has about 800 universities!
 A bar chart is plotted below to show the top 15 states in the frequency table.
-Then summary of frequency table also shows that the state having the least university only has one, and 50% of state has more than 132 universities in the state.
+Then the summary of frequency table also shows that the state having the least university only has one, and 50% of states have more than 132 universities in the state.
 
@@
 Then we analyze the type of degree universities mainly offer.
-There are four types of degree taken into accout: Certificate-degree, Associate's degree, Bachelor's degree, and Graduate Degree.
-A frequency table in a format of percentage and a histogram are used to help understand the distributiton of the Prodominant Degree.
+There are four types of degree taken into account: certificate degree, associate degree, bachelor degree, and graduate degree.
+A frequency table in a format of percentage and a histogram are used to help understand the distribution of the predominant degree.
 
@@
 
-From above frequency table, we see that about 5% of the population are unclassified in terms of the predomiant degree they offer.
-43.2% offer Certificate degree as their predominant degree, 27.3% predominantly offer Bachelor degree and only 4.23% are classified as Graduate degree predominant.
+From above frequency table, we see that about 5% of the population are unclassified in terms of the predominant degree they offer.
+43.2% offer certificate degree as their predominant degree, 27.3% predominantly offer Bachelor degree and only 4.23% offer graduate degree as their predominant degree.
 
 ####5.1.1.3. Control of Institution<a id="5.1.1.3"></a> 
-In addition to State and predominant degree, control of institution is another categorical variable in the dataset.
-The label of Control is to indicate whether or not a institution is privately owned and whether or not it is a for-profit organization.
+In addition to state and predominant degree, control of institution is another categorical variable in the dataset.
+The label of Control is to indicate whether or not an institution is privately owned and whether or not it is a for-profit organization.
 There are three levels of this variable: Public, Private Nonprofit, and Private-profit.
 Public school is assumed to be non-profit so it is not further divided into two subgroups.
-Below we show the percentage summary of each category and a bar char is plotted to facilitate understanding.
+Below we show the percentage summary of each category and a bar chart is plotted to facilitate understanding.
 
@@
 
-About 50% out of all institution are private for-profit university, which is the *most dominant type*.
-Public and Private Nonprofit are pretty tiex in terms of the number of observation in each category.
+About 50% of all institutions are private for-profit university, which is the most dominant type. Public and private nonprofit are pretty tied in terms of the number of observation in each category.
 
@@
 ####5.1.2.1 Percentage first-generation students<a id="5.1.2.1"></a>
-This element identifies the share of federally aided students at the institution who report that they are first generaion on the FAFSA (Free Application for Federal Student Aid).
-First generation students are defined as sudents in which neither parents completed college.
-
-With the followin codes, we extracted the institions with the highest and lowest percentage of first-generation students.
-Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
+This element identifies the share of federally aided students at the institution who report that they are the first generaion on the FAFSA (Free Application for Federal Student Aid).
+First generation students are defined as students whose parents didn't attend a college.
+
+With the following codes, we extracted the institutions with the highest and lowest percentage of first-generation students.
+Then we observed this varible's basic summary statistics: min, max, mean, median and standard deviation.
 ```{R, message=FALSE, warinig = FALSE}
@@
 From the mean value of 48.08%, we can tell that about for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college.
-The standard deviation value of 12.2% indicated a relative high difference of this element among different colleges.
-
-Below is the code to generate a visual representation in the histgram. 
+The standard deviation value of 12.2% indicated a relatively high difference of this element among different colleges.
+
+Below is the code to generate a visual representation in the histogram. 
 ```{R}
@@
 
-From the histogram, we can tell the share of first generarion students is most frequently falling into the range between 0.5 and 0.6.
-The value with the value more than 0.8 are rare.
-The colleges with the share of first generarion students value less than 0.5 are much more than the ones with the value larger than 0.5.
-Thus the histogram is seriously left skewed.
+From the histogram, we can tell the share of the first-generarion students is most frequently falling into the range between 0.5 and 0.6. The values with the value more than 0.8 are rare.
 
@@
 The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who list two and more than two institutions when he applied for financial aid when submitting application for university.
-This information is only available to students who submitted their applications for federal aid, and does not cover the whole student body.
-Nevertheless, this variable still can provides some useful insights about how many universities a student usually apply for.
+This information is only available to students who submitted their applications for federal aid and does not cover the whole student body.
+Nevertheless, this variable still can provide some useful insights about how many universities a student usually apply for.
 
@@
 
-The summry statistics show that median percentage of student submitting two or more than two application is 41.67% across all univerisities.
-The institution having the least percentage only has 4% of student applying to more than one university, while the school with the highest percent has as high as 90% of student applying to more than one university.
-The boxplot compares the the percent of student applying to only one university with that of student applying for more than one.
-We can clearly see that a larger portion of students only listed one instituiton when submitting their application for financial aid.
+The summary statistics show that the median percentage of those students submitting two or more than two application is 41.67% across all universities The institution having the least percentage only has 4% of students applying for more than one university, while the school with the highest percent has as high as 90% of students applying for more than one university.
+The boxplot compares the percent of student applying for only one university with the percent of students applying for more than one.
+We can clearly see that a larger portion of students only listed one institution when submitting their application for financial aid.
 
@@
 while the minimum value is about $-19,360 - the value for Berea College located in KY.
-Considering the elements are collected from Title IV institutions, we reasoned that the negative values may be due to the fact that some intituions's finainal aid supoort can cover the students' all the costs, meanwhile overbalancing the cost amount.
-There are not much differerences between mean value and median value, which is around $15,000.
-We can conclude that the average net price costs for most institutuins are falling into the this range.
+Considering the elements are collected from Title IV institutions, we reasoned that the negative values may be due to the fact that some institution's financial supoort can cover the students' all the costs, even overbalancing the cost amount.
+There are not much differerence between the mean value and the median value, which is around $15,000.
+We can conclude that the average net price costs for most institutions are falling into  this range.
 The standard deviation value is 7604.884 which is relatively large, indicating a great spread in the data.
 
-Below is the histogram for this variable which is able to provide us with a visual distrubution presentation. 
+Below is the histogram for this variable which is able to provide us with a visual distribution presentation. 
 ```{R}
@@
 
-From the histogram, we notice that the average net cost for IV institutions typically is 16000 -18000.
-The values between 16000 - 20000 are quite frequent.
-There were an unsually lagrge number of costs.
-The institutions with cost of more than 50000 were atypical.
-Due to the atypical large value, the histogram is abviously right skewed.
-Without those extreme values, the histogram would be reasonably symmetric.
+From the histogram, we notice that the average net cost for IV institutions typically is $16000-$18000. The values between $16000 -$20000 are quite frequent.
 
@@
 
-With the followin codes, we extracted the institions which have the highest and lowest percentage of students who are financially independent.
-Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
+With the following codes, we extracted the institutions which have the highest and lowest percentage of students who are financially independent.
+Then we observed this varible's basic summary statistics: min, max, mean, median and standard deviation.
 ```{R}
@@
 From the summary results, we learned that the maximum value of the variable is 98.56%(Capella University) while the minimum value is about 0.73%(University of Notre Dame).
-From the mean value, we can tell that about in most colleges, around 56% of the students are financially independent.
-The standatd deviation value of 23.3% indicates a relative high different among different colleges, which can also be concluded from the difference between maximum value and minimum value.
-Below is the code to generate a visual representation in the histgram.
+From the mean value, we can tell that  in most colleges, around 56% of the students are financially independent.
+The standard deviation value 23.3% indicates a relatively high difference among different colleges, which can also be concluded from the difference between the maximum value and the minimum value.
+Below is the code to generate a visual representation in the histogram
 
@@
 
-From the histogram, we notice that the value between 0.7 to 0.8 are quite frequent. There was no obvious extrme value in the sample, thus the histogram is normally distributed.
+From the histogram, we notice that the values between 0.7 to 0.8 are quite frequent. There was no obvious extreme value in the sample, thus the histogram is normally distributed.
 
@@
 
-This elemnt, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received Pell Grants in a given year, which is an important measure of the access an insiitution provides to low-income students.
-We are using the code below to summarize this variable and extraced the instition with the highest and lowest value.
+This element, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received Pell Grants in a given year, which is an important measure of the financial aid an institution provides to low-income students.
+We are using the code below to summarize this variable and extract the institutions with the highest and lowest value.
 
@@
 
-From the output, we can tell that the range is between 0 to 100%.	No students in Thunderbird School of Global Management has received the Pell grant while all of the students in American Conservatory Theater have receiced the Pell Grant.
-Among the colleges in the sample, half colleges have this rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs.
-The standard deviation value is 22.57%,  indicating a relative high difference of this element among different colleges, which can also be noticed from the largest range (100%).
-Below is the code to generate a visual representation in the histgram.
+From the output, we can tell that the range is between 0 to 100%.	No students in Thunderbird School of Global Management has received the Pell grant while all of the students in American Conservatory Theater received the Pell Grant.
+Among the colleges in the sample, half of the colleges colleges have the rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs.
+The standard deviation value is 22.57%,  indicating a relatively high difference of this element among different colleges, which can also be noticed from the largest range (100%).
+Below is the code to generate a visual representation in the histogram.
 
@@
 From the histogram, we are able to conclude that in most colleges, 40% to 50% of the undergraduate students are able to receive the Pell Grant.
-With little skewness, the histogram is resonalbe symmetric.
+With little skewness, the histogram is reasonably symmetric.
 
@@
 
-This elemnt, as reported in the IPEDs Student Finaicial Aid component, show the share of undergraduate students who received federal loans in a given year.
+This element, as reported in the IPEDs Student Finaicial Aid component, shows the share of undergraduate students who received federal loans in a given year.
 It can provide important context to figures related to debt, repayment and non-repayment.
-We are using the code below to summarize this variable and extracted the institutions with highest and lowest value.
+We are using the code below to summarize this variable and extract the institutions with highest and lowest value.
 
@@
 
-From the output, we can tell that the range is between 0 to 100%.	As we can recall from the previous variable(PCTPELL), American Conservatory Theater has both the PCTLOAN and PCTPELL the value of 100%.
-Gadsden State Community College has no students reveiced federal loans in the given year.
-Among the colleges in the sample, half colleges have this rate as high as 60.08% , which means around 60% of undergraduare students are able to receive the fedaral student loans.
-The standard deviation value is 29.26%,  indicating a relative high difference of this element among different colleges, which can also be noticed from the largest range (100%).
-Below is the code to generate a visual representation in the histgram.
+From the output, we can tell that the range is between 0 to 100%.	As we can recall from the previous variable(PCTPELL), Gadsden State Community College has no students receiving federal loans in the given year.
+Among the colleges in the sample, half of colleges have this rate as high as 60.08% , which means around 60% of undergraduate students receive the fedaral student loans.
+The standard deviation value is 29.26%,  indicating a relatively high difference of this element among different colleges, which can also be noticed from the largest range (100%).
+Below is the code to generate a visual representation in the histogram.
 
@@
 
-From the histogram, it's obvious that there are two peaks in the graph. The rates between 0- 10% and the rates between 60% to 80% are similarly frequent compated to the other ranges.
+From the histogram, it's obvious that there are two peaks in the graph. The rate between 0-10% and the rates between 60% to 80% have similarly frequency compared to the other ranges.
 
@@
 
-This element is the median loan debt acculated at the institution by all student borrowers of federal lonans who have habe completed the education.
-We are using the code below to summarize this variable.
+This variable recorded debt for students who have completed the education. We are using the code below to summarize this variable. 
 ```{R, message=FALSE, warinig = FALSE}
@@
 
-From the output, we can tell that the highest value is 48,250 - Southwest University of Visual Arts-Tucson, while the lowest value is 1448 - South Florida Institute of Technology.
-From the median value of 14000, we can tell that among the colleges in the sample, the students who have completed the deducaiton from half of them had the median loan debt of 14000.
-The standard deviation value is 8020.958,  indicating a relative high difference of this element among different colleges.
-Below is the code to generate a visual representation in the histgram.
+From the output, we can tell that the highest value is $48,250, belonging to Southwest University of Visual Arts-Tucson, while the lowest value is 1448, belonging to South Florida Institute of Technology.
+From the median value of 14000, we can tell that among the colleges in the sample, for students who have completed the education, half of them have  median loan debt of $14000.
+The standard deviation value is $8020.958,  indicating a relatively high difference of this element among different colleges.
+Below is the code to generate a visual representation in the histogram.
 
@@
 
-From the histogram, it is obvious that most students in the sampled colleges still have 8000- 10000 debt amount after they completed the education.
+From the histogram, it is obvious that most students in the sample colleges still have $8000 - $10000 debt amount after they completed the education.
 This amount is related to the federal loan amount provided by each institution.
-Additionally, it is asscociated with students' ability to repay the loan during the acdemic period.
-For example, for the students who were capabiblity of working through the acedemic years or securing jobs before graduation may have a less value for this variable.
+Additionally, it is associated with students' ability to repay the loan during the academic period.
+For example, for the students who were capable of working through the acedemic years or securing jobs before graduation may have a less value for this variable.
 
 ####5.1.4.4 The median debt for students who have not completed<a id="5.1.4.4"></a>
-WDRAW_DEBT_MDN describes the median amount of debt fot students who ahve not completed their study. In other words, they are still attending the university.
+
+WDRAW_DEBT_MDN describes the median amount of debt for students who have not completed their study. In other words, they are still attending the university.
 
@@
 We use summary statistics, standard deviation and histogram to understand the distribution of this variable.
-The summary statistics show that students in 50% of institution have a median debt of more than 5500, which is not very unaffordable.
-The median debt for attending students can be as much as 27620 for some school and as low as 600 for others, which indicates it may have a large variance.
+The summary statistics show that students in more than 50% of institution have a median debt of more than $5500, which is not very unaffordable.
+The median debt for attending students can be as much as $27620 for some school and as low as $600 for others, which indicates that it may have a large variance.
 So we also calculate standard deviation to measure the dispersion of the distribution.
-The standard deviation is 2415.623 about 95% of institution have a median debt ranging from 1167.754  to 10830.246.
+The standard deviation is $2415.623, which means about 95% of institution have  median debt ranging from $1167.754  to $10830.246.
 The histogram shows that the distribution of the median roughly follows a standard distribution with a sharp peak and slightly heavy tail.
@@
 ####5.1.5.1 Median family income in real 2015 dollars<a id="5.1.5.1"></a>
-This elemnt is recording the median family income of students, one of the undergraduate student demographics for earnings cohorts. We are using the code below to summarize this variable.
+This element is recorded the median family income of students, one of the undergraduate student demographics for earnings cohorts. We are using the code below to summarize this variable.
 
@@
 
-From the output, we can tell that the highest value is 155,606.6 -Jewish Theological Seminary of America, while the lowest value is 0 -Associated Technical College-Los Angeles.
-From the median value of 19,262.26, we can tell that among the colleges in the sample, the students in half of them had the family income below 19,262.26	.
-The standard deviation value is 18,824.31,  indicating a relative high difference of this element among different colleges.
-Below is the code to generate a visual representation in the histgram.
+From the output, we can tell that the highest value is $155,606.6 for Jewish Theological Seminary of America, while the lowest value is $0 for Associated Technical College-Los Angeles.
+From the median value of $19,262.26, we can tell that among the colleges in the sample, the students in half of them had the family income below $19,262.26	.
+The standard deviation value is $18,824.31,  indicating a relatively high difference of this element among different colleges.
+Below is the code to generate a visual representation in the histogram
 
@@
 
-From the histogram, we are able to conclude that in most colleges, students' family income are falling into the range of 12,000 - 18,000.
-Due to the atypical large values, the histogram is obviouslu skewed to right.
-The value larger than 18,000 are tend to be extreme and the cases larger than 75000 are rare and atypical, which means rare colleges have the students with family income more than 75000.
+From the histogram, we are able to conclude that in most colleges, students' family income fall into the range of $12,000 - $18,000.
+Due to the some influential observations, the histogram is obviously skewed to right.
+The values larger than $18,000 are tend to be extreme and the cases larger than $75000 are rare, which means that rare colleges have the students with family income more than $75000.
 
@@
 
-We selected two variables MD_EARN_WNE_P6, MD_EARN_WNE_P10 in the earning group to study students' post-graduation earning.
-MD_EARN_WNE_P6 is the median earning for students after six years of their entry, which corresponds to two years after gradutaion for a four-year program.
-Likewise, MD_EARN_WNE_P10 is the median earning ten years after their entry, which corresponds to four years after graduation usually.
-These two metrics give us an idea of how much money students are making after their graduation, in both short and medium terms.
+We selected two variables MD_EARN_WNE_P6, MD_EARN_WNE_P10 in the earning group to study students' post-graduation earnings.
+MD_EARN_WNE_P6 is the median earning for students after six years of their entry, which corresponds to two years after graduation for a four-year program.
+Likewise, MD_EARN_WNE_P10 is the median earning ten years after their entry, which usually corresponds to four years after graduation These two metrics give us an idea of how much money students are making after their graduation, in both short and medium terms.
 
@@
 
-The six-year-after-entry median earning of 50% institutions is below the 25300 below, while the maximum median earining is 120200, while ten-year-after-entry median earning is higher by about $5500.
-The histogram gives a more clear idea of the earning distribution.
-Most of institution with students having a median earining between 10000 to 50000 six years after entry.
-The distribution of ten-year-after-entry falls at the right side of the six-year distribution , which means that the earning increases as students' experience increase.
-Beside, the ten-year post-graduation earning has a wider spread than that of six years.
-That indicate that the there is a greater gap between who are earning more and who are earning less, which can be confirmed by an increase in standard deviation.
+The six-year after-entry median earnings of 50% institutions is below the $25,300, while ten-year-after-entry median earning is higher by about $5,500.
+The histogram gives us a more clear idea of the earning distribution.
+Most institutions with students having a median earnings between $10,000 to $50,000 six years after entry.
+The distribution of ten-year-after-entry falls at the right side of the six-year distribution, which means that the earning increases as students' experience increases.
+Besides, the ten-year post-graduation earning has a wider spread than that of the six-years.
+This indicates that there is a greater gap between who are earning more and who are earning less, which can be confirmed by an increase in the standard deviation.
 
@@
 ####5.1.6.1 3-year Repayment rate<a id="5.1.6.1"></a>
-The variable RPY_3YR_RT_SUPP measures the percentage of students who entered repayemtn from October 1, 2010 to September 30, 2011, and made progress in the following three years.
-Repayment rates are generally considered most sensitive than default rates(we study later), which measures only the worse-case scenario for repayment outcomes.
+The variable RPY_3YR_RT_SUPP measures the percentage of students who entered repayment from October 1, 2010 to September 30, 2011, and made progress in the following three years.
+Repayment rates are generally considered more sensitive than default rates(we will study later), which measures only the worst-case scenario for repayment outcomes.
 
@@
 
-From the summary statistics, we see that 50% institutions have 40% of students who were making progress in their repayment.
+From the summary statistics, we see that 50% institutions have more than 40% of students who were making progress in their repayment.
 A histogram is also plotted to show the distribution, clearly with a heavy tail.
@@
 ####5.1.6.2 3-year Default rate<a id="5.1.6.2"></a>
+
 The three-year cohort default rate describes the repayment in the opposite of repayment rate.
@@
 
-The median default rate aross all universities is 13%, which means that more than 50% of universities have 13 out of 100 students defaulted their loan!
+The median default rate arosses all universities is 13%, which means that more than 50% of the universities have more than 13 out of 100 students defaulted their loan!
 
@@
 
-To analyze the relationship between average net cost for Title IV institutions and State, we grouped the data into the state level to take a further look.
-We would like to investigate the state distribution of the Title IV instituions based on the average net costs.
+To analyze the relationship between the average net cost for Title IV institutions and State, we grouped the data into the state level to take a further look.
+We would like to investigate the distribution of average net price in each state.
+
 ```{R message=FALSE, warinig = FALSE}
@@
 
-Baesd on the output, we can see the highest median value for avreage net price exists in DC while the lowest median value is in AS(American Samoa).Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are all states conceiving amount of well-known educational institutions.
+Based on the output, we can see that the highest median value for the avreage net price exists in DC while the lowest median value is in AS(American Samoa).
+Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are all the states have  well-known educational institutions.
 
 ##6.2. The relationship between median debt and average net price for Title IV institutions<a id="6.2"></a>
-We assume there would be a strong linear relationship between the average net price for Title IV institutions and the median debt for students who have completed.
-We hypothese that the students who completed education in colleges with high costs will have a highe value of median debt.
-We use the following code to generate the scatterplot between these two variables to verify our assumptions.
+
+ We hypothesised that the students who completed education in colleges with high costs will have a higher value of median debt.
+We use the following code to generate the scatterplot between these two variables to verify our hypothesis.
 
@@
 
-From the scatter plot, it is obvious that the NPT4_COMBINE is in positive linear relationship with GRAD_DEBT_MDN: the data show an uphill pattern from left to right; As the average net price for Title IV institutions increase, the value of median debt for students who have completed increases as well.
+From the scatter plot, it is obvious that the NPT4_COMBINE is in a positive linear relationship with GRAD_DEBT_MDN: the data shows an uphill pattern from left to right; As the average net price for Title IV institutions increases, the value of median debt for students who have completed the college increases as well.
 
@@
 
-To compare the percentages of Pell Grant in different control type, we grouped the Percentage of Pell Grant data into the college type level to take a further analysis.
+To compare the percentage of Pell Grant in different control types, we grouped the Percentage of Pell Grant data usubf college type to take a further analysis.
 
@@
 
-From the result, the Private Nonprofit institutions and Public institutions have a similar low median values while the provate For- profit institutions have a hgihest value.
-This is resonable because generally speanking, the private for-profit colleges may have much higher costs than the private nonprofit and public institutions.
-The low-income students were more likely to seek a Pell Grant for assistant.
-Additionally, in some states, state financial aid may be sufficient to cover costs at community colleges; thus the students in thoes schools may not need to seek for addtional financial aid like Pell Grant.
+From the result, the private nonprofit institutions and the public institutions both have a  low median values while the private for-profit institutions have a higher value.
+This is reasonable because the private for-profit colleges may have much higher costs than the private nonprofit and public institutions.
+The low-income students were more likely to seek a Pell Grant as assistance.
+Additionally, in some states, state financial aid may be sufficient to cover costs at community colleges; thus the students in those schools may not need to seek for additional financial aid like Pell Grant.
 
@@
 
-From the overlaying histograms, the percentage of students who receive Pell Grant are quite frequent for Private For-profit institutions between 0.7 and 0.75 ; the percentage of students who receive Pell Grant are quite frequent around 0.4 for Public institutions and Private Nonprofit institutions.
-The histogram for private for profit is slightly left skewed due to atypical small value near 0.
-The histogram for public and private nonprofit institions are reasonably symmetric
+From the overlaying histograms, the percentages of students who received the Pell Grant for private for-profit institutions are quite frequent between 0.7 and 0.75 ; the percentages of students who receive the Pell Grant for public institutions and private nonprofit institutions are quite frequent around 0.4.
+The histogram for private for-profit is slightly left skewed due to small values near 0.
+The histogram for public and private non-profit institutions are reasonably symmetric.
 
 ##6.4. The relationship between Percentage first-generation students and States<a id="6.4"></a>
-To compare the Percentage first-generation students with the different states, we grouped data into the state level to take a further analysis.
-We would like to investigate the percentage of first-generation students based on the status distribution.
+
+To compare the percentage of first-generation students in each states, we grouped data into the state level to take a further analysis.
 
@@
 
-The results showed that the universities located in ND, SD, VT, MN and MT had the lowest value of Percentage first-generation students, which indicated that the students in thoses five states would have a better family education background.
-The five states with the highest share of parents first-generation are LA, NJ, MP, MV, MH.
+The results show that the universities located in ND, SD, VT, MN and MT have the lowest percentage first-generation students, which indicates that the students in these five states would have a better family education background.
+The five states with the highest share of first-generation students are LA, NJ, MP, MV, MH.
 
 ##6.5. The relationship between family income and percentage of Pell Grant<a id="6.5"></a>
- We assume there would be a strong linear relationship between family income and the percentage of students who can receive the Pell Grant.
-We hypothese that colleges with small value of family income would tend to have high percentages of Pell Grant.
-The following codes are used to generate the scatter pkot between these two varaibles.
+
+We hypothesed that colleges with small value of family income would tend to have high percentages of Pell Grant. The following codes are used to generate the scatter plots between these two variables.
 
@@
 
-From the scatter plot, it is obvious that the MD_FAMINC is in a negative linear relationship with PCTPELL: the data show a downhill pattern from left to right; As the MD_FAMINC increase, the value of PCTPELL decreases.
-It makes sense in the real world applcation: for institutions with low percentage of students who receive the Pell Grant, the financial aids like Pell Grant may not be as necessary as they are in the instituions with high value of PCTPELL, thus these collges students will tend to have a high median family income after they graduate.
+From the scatter plot, it is obvious that the MD_FAMINC is in a negative linear relationship with PCTPELL: the data shows a downhill pattern from left to right; As the MD_FAMINC increases, the value of PCTPELL decreases.
+It makes sense in the real world; for institutions with low percentage of students who receive the Pell Grant, the financial aid like Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, since these collges students tend to have a higher median family income.
 
@@
 
-To compare the percentage of financially-independent students in institutions with different college type, we grouped by the data into the control type(Private For-profit, Public, Private Nonprofit) to take a further analysis.
+To compare the percentage of financially-independent students in the institutions with different college type, we grouped the data by the control type(private for-profit, public, private nonprofit) to take a further analysis.
+
 ```{R message=FALSE, warinig = FALSE}
@@
 
-From the result, we can tell the highest median value is in Privately For-profit colleges, which means that in most private for-profit colleges, the students are more likely to be financial independent.
+From the result, we can tell the highest median value is in privately for-profit colleges, which means that in most private for-profit colleges, the students are more likely to be financially independent.
 
@@
 
-To compare the percentage of students who receive the federal loans in the given year based on the schools control type,we grouped by the data into the control type(Private For-profit, Public, Private Nonprofit) to take a further analysis.
-A stacked histogram is also requested to generate a more vidual presentation.
+To compare the percentage of students who receive the federal loans in the given year in each schools control type, we grouped the data by the control type(private for-profit, public, private non-profit) to take a further analysis.
+A stacked histogram is also used to generate a more visual presentation.
 
@@
 
-From the result, the Private institutions have a much larger median value than the public institutions.
-With a much higher costs burden, it makes sence that the low-income sutdents in private institutions are much more likely to seek financial aid than the public institution students.
-From the stacked histogram, we can tell that for all the three college type institutions, the most frequent ranges appear between 0.6 and 0.8, which is consistent with the findings we observe in the individual variable summary section.
+From the result,  private institutions have a much larger median value than the public institutions.
+With a much higher costs burden, it makes sense that the low-income students in private institutions are much more likely to seek financial aid than the public institution students.
+From the stacked histogram, we can tell that for all  three college types, the most frequent range appears between 0.6 and 0.8, which is consistent with the findings we observe in the individual variable summary section.
 
 ##6.8 Median debts with earning<a id="6.8"></a> 
-Next, we analyze the relationship between the amount the debt and post-graduation earning.
+Next, we analyze the relationship between the amount the debt and post-graduation earnings.
 There are two variables in the dataset to summarize the median amount of debt for student body, one for students who already graduated and one for students who have not completed their study.
-We decide debts for students who have not completed to study in this problem, since it makes more sense for the purpose of prediction.
+We decided to use debt for students who have not completed their study in this problem, since it makes more sense for the purpose of prediction.
 
@@
 
-First, we use scatter plot to visualize the relationship between earning and the amount of debts, the plot clearly shows a positive relationship.
-Then we further calculate pearson correlation to confirm our hypothesis.
-The resulting coefficient 0.44 indicatesa moderate linear relationship between earning and debts.
+First, we used a scatter plot to visualize the relationship between earnings and the amount of debt, the plot clearly shows a positive relationship.
+Then we further calculated pearson correlation to confirm our hypothesis.
+The resulting coefficient 0.44 indicates a moderate linear relationship between earnings and debt.
 
@@
 
-Understanding the relationship between the default rate of load repayment and post-graduation earning on one hand can help federal financial commitee to improve their decision about load granting.
-If  the default rate decreases as the earning increase, the commitee may prefer to grant more financial aid to institution whose students have higher earning.
-One the other hand, the default rate of repayment for a univeristy can also help predict the post-graduation earning for its students.
+For one thing, understanding the relationship between the default rate of loan repayment and post-graduation earnings can help federal financial committee to improve their decision about load granting.
+If the default rate decreases as the earnings increase, the committee may prefer to grant more financial aid to institutions whose students have higher earnings.
+For another, the default rate of repayment for a university can also help predict the post-graduation earnings for its students.
 
@@
 A scatter plot is created to visualize the relationship between these two variables.
-From the chart, we see the earning seems to show a decreasing trend as the default rate increases, which shows a negative relationship between earning and default rate.
+From the chart, we see that earnings seems to show a decreasing trend as the default rate increases, which shows a negative relationship between earnings and default rate.
 Then we calculate the correlation coefficient between these two variables.
-The resulting r = -0.33 shows that correlation between earning and default rate is negative but the relationship is not very strong, which may be due to the influences of some outliers.
-To conclude, the negative relationship between earning and default rate of loan repayment needs extra data to confirm.
+The resulting r = -0.33 shows that the correlation between earnings and default rate is negative but the relationship is not very strong, which may be due to the influences of some outliers.
+To conclude, the negative relationship between earnings and default rate of loan repayment needs extra data to confirm.
 
@@
 
-From the boxplot below, we see that the higher the education, the higher the median earning 6 years after entry Especially, a graduate-degree gives a hugh rise to one's earning.
+From the boxplot below, we see that the higher the education, the higher the median earning 6 years after entry. Especially, a graduate-degree gives a hugh rise to one's earning.
 
 ##6.11 predominant degree, control  and earning after 6 year of entry<a id="6.11"></a>
+
 The following plot shows two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry).
 
-For the R code, we first group the dataset by PREDDEG factor and Control factor.
-Then, we summarize the average earnings.
+For the R code, we first grouped the dataset by PREDDEG factor and Control factor.
+Then, we summarized the average earnings.
 After that, we set the aesthetic mapping between Control_factor and PREDDEG_factor as the horizontal axis and the vertical axis.
@@
 
-Mean value of the earning is more representative for the two categories, preF and ctrF.
-For the range of average earning, the darkest color shows the lowest earning, while the lightest color shows the highest value.
-According to the plot, we realize that ???Private For-profit college???
-whose predominant degree award is ???Graduate degree???
-has the lowest average earning with the minimum average value at \$3100, while ???Public college???
-whose predominant degree award is ???Graduate degree???
-has the highest average earning with the maximum average value at $45850.
+Mean value of the earnings is more representative for the two categories, predominant degree and control According to the plot, we realized that private for-profit colleges whose predominant degree awarded is graduate degree has the lowest average earning with the minimum average value at $3100, while Public college whose predominant degree awarded is graduate degree has the highest average earnings with the maximum average value at $45,850.
 
 ##6.12 predominant degree + control (count)<a id="6.12"></a>
+
 The following plot shows the relationship between two factor variables (predominant degree award variable and control variable).
@@
 For the R code, our team first set the aesthetic mapping on Control_factor as the horizontal axis.
-Then, we use the function geom_bar() to fill different control types using the counting values of PREDDEG factor.
+Then, we use the function geom_bar() to fill different control types using the counting values of PREDDEG, which creates bars based on a grouped frequency distribution.
 After that, we further use the function xtabs() to calculate the exact number for each predominant degree awarded by different types of colleges.
@@
 
-According to the histogram comparison of predominant degree award among three types of colleges, we realize that for public college, Associate???s-degree is most frequent awarded in the public colleges.
-Bachelor???s-degree is most frequent awarded in the private nonprofit colleges.
-Certificate-degree is most frequent awarded in private for-profit colleges.
-
-Although the ???Graduate-degree???
-awarded by public colleges shows the least frequency, it gains the highest average earnings of students working and not enrolled 6 years after entry compared with other degrees awarded by different colleges.
-
-Our team further uses the function xtabs() to calculate the exact number for each predominant degree awarded by different types of colleges.
 ```{R}
@@
 
+From the graph we can see that for the public type of school, associate degree is mostly awarded and it is in the middle of the distribution, with fewer NotClassified degree and fewer graduate degree in the extremes.
+But the distribution is symmetric compared with the other two types.
+For the Private Nonprofit type of school, mostly awarded degrees (Bachelor degree and Graduate degree) are located at the right of the distribution, with fewer NotClassified degree on the left.
+The distribution shows a negative skew: the awarded degrees extend to the left farther than they do to the right.
+For the private for-profit type of school, mostly awarded degree (Certificate degree) is on the left of the distribution, with fewer graduate degree on the right.
+The distribution shows a positive skew: the awarded degrees extend to the right farther than they do to the left.
+
 #6.13 control + earning<a id="6.13"></a>
+
 The following plot shows one factor variable (control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry).
-We use the
-geom_violin() using the default scale = ???area???
-argument and geom_boxplot() function to explore the relationship.
+
+We use the ggplot() function to initialize the ggplot object and declare the input data frame for a graphic with x equals Control_factor and y equals Median Earnings.
+Then, we use the geom_violin() function which shows a compact display of a continuous distribution.
+The geom_violin() function is a blend of geom_boxplot() and geom_density(), which shows a mirrored density plot.
+For better visualization, we use geom_boxplot() to create a box and whiskers plot, which shows five summary statistics (the median, two hinges and two whiskers), and all “outlying” points individually.
+And the summary() function further displays the exact values.
 The plot superimposed with boxplots shows side-by-side violin plots of median earnings of students working and not enrolled 6 years after entry according to different college types.
@@
 
+All the three types of school have far out values above, which can be explored further.
 Obvious differences contain in these three boxplots.
-The boxplot of ???Public college???
-is comparatively short, while the boxplot of ???Private Nonprofit???
-is comparatively tall.
-The ???Private For-profit college???
-has the lowest median earning.
+The boxplot of public college is comparatively short, while the boxplot of private nonprofit is comparatively tall.
+The private for-profit college has the lowest median earning.
+
+For Public type of school, the earnings has a high density around the level of $25,000.
+From the box plot summary, we realize that half of the public school student earnings are between  $21,880 and $30,320.
+Also, a larger mean value than median value indicates a positive skew distribution.
+
+For the Private Nonprofit type of school, the earnings has a high density around the level of $31,250.
+From the box plot summary, we realize that half of the private nonprofit school student earnings are between $13,000 and $35,800.
+Also, a lower mean value than median value indicates a negative skew distribution.
+
+For the Private For-profit type of school, the earnings has a high density around the level of $18,750.
+The distribution is not symmetric.
+From the box plot summary, we realize that half of the private for-profit school student earnings are between $12,080 and $24,700.
+Also, a lower mean value than median value indicates a negative skew distribution.
 
@@
 
-Through the process of summarizing these variables and analyzing the relationship between these variables, we found that for federally aided students in most colleges, around half of the students' parents did not complete the college.The average net price costs for most institutuins are falling into the range between $16000 - $20000.
-Among the 7703 institution samples, more than 50% of the students are able to receive Pell Grant or Federal Loans as means of finanical support, and on average, 60% of students are financial independent.
-In most colleges, the students are able to reach the family income between $12000 and $18000.
+Through the process of summarizing these variables and analyzing the relationship between these variables, we found that for federally-aided students in most colleges, around half of the students' parents did not complete college.The average net price costs for most institutions are falling into the range between $16000 - $20000.
+Among the 7703 institution samples, more than 50% of the students receive a Pell Grant or Federal Loans as means of finanical support, and on average, 60% of students are financially independent.
+In most colleges, the students have the family income between $12000 and $18000.
 
 We also found that California has the most universities across the nation, with 797 institutions.
-In terms of predominant degree offered, 43.2% are classified as Certificate-degree predominant institution, which is the most predominant type.
-
-Our findings also suggest that the Title IV institions located in DC, VT, MA, NH, RI have a relative high amount of average net costs than those in other states.
-The students in private non-profit institutions and public institutions are less likely to seek Pell Grant as financial assistance than the students in the private for-profit colleges.In most private for-profit colleges, the students are more likely to be financial independet.
-
-It is also indicated that there exists a negative correlation between the percentage of students who receive Pell Grant and the students median family income after they graduate: for institutions with low percentage of students who receive the Pell Grant, the financial aids like Pell Grant may not be as necessary as they are in the instituions with high value of PCTPELL, thus these collges students will tend to have a high median family income after they graduate.
-
-As to the factors that influence post-graduate earning, we explore the relationship between median earning six years from entry and amount of debt, default rate for student loan and predominant degree type.
-The analysis suggests that there is a positive relationship between the median debt amount for students who have not completed and post-graduation earnings.
-This may be due to the fact that top universities usually have tuitions but may also indicates that students who are willing to take more debts for their college education are likely to make more money after graduation.
-In addition, default rate of student loan is negatively related to post-graduate earning.
+In terms of predominant degrees offered, 43.2% are classified as a certificate degree predominant institution, which is the most prevalent type.
+
+Our findings also suggest that the Title IV institutions located in DC, VT, MA, NH, RI have a relatively high amount of average net costs than those in other states.
+The students in private non-profit institutions and public institutions are less likely to seek a Pell Grant as financial assistance than the students in the private for-profit colleges.In most private for-profit colleges, the students are more likely to be financially independet.
+
+It indicates that there exists a negative correlation between the percentage of students who receive a Pell Grant and the students' median family income; for institutions with low percentage of students who receive the Pell Grant, the financial aid like the Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, since these college students will tend to have a high median family income.
+
+As to the factors that influence post-graduation earnings, we explored the relationship between median earnings six years from entry and other six variables.
+The analysis suggests that there is a positive relationship between the median debt amount for students who have not completed college and post-graduation earnings.
+This may be due to the fact that top universities usually have higher tuition but may also indicate that students who are willing to take more debts for their college education are likely to make more money after graduation.
+In addition, the default rate of student loans is negatively related to post-graduation earnings.
 This finding is reasonable since students making more money are more likely to pay off their debt.
-The variable can be used a school-level predictor to predict overall post-graduate earning for each university.
+The variable can be used as a school-level predictor to predict overall post-graduation earnings for each university.
 The degree type, unsurprisingly, has an impact on the earnings students are making after graduation.
-As the level of graduation increases, the post-graduate earning also increase significantly, which suggests that investing in education is worthwhile.
-
-As a result of all the statistical analysis, we conclude that the attributes such as percentage of students who receive Pell Grant, median earning six years from entry, amount of debt, default rate for student loan and predominant degree type will affect the students' post graudate earnings.
+As the level of education increases, the post-graduate earnings also increase significantly, which suggests that investing in education is worthwhile.
+
+When it comes to relationship between control of institution and earning, private for-profit institutions whose predominant degree awarded is graduate degree has the lowest average earnings with the minimum average value at $3100.
+Public institutions whose predominant degree awarded is graduate degree has the highest average earning with the maximum average value at $45850.
+Although the graduate degree awarded by public colleges shows the least frequency, it contains the most value because those students (working and not enrolled 6 years after entry) graduating from public colleges holding the graduate degree show the highest average earnings compared with students holding other degrees awarded by different institutions.Also, students graduate from Private Nonprofit institution generally has a higher earning than those students graduate from other types of institutions.
+
+According to all analysis, we conclude that the attributes such as percentage of students who receive a Pell Grant, median earning six years from entry, amount of debt, default rate for student loans and predominant degree type can be used to predict the students' post-graudation earnings.

### 2017-02-22 Draft-xuezhou-sjb.Rmd
update

@@
 
-From the histogram, we notice that the average net cost for IV institutions typically is $16000-$18000. The values between $16000 -$20000 are quite frequent.
+From the histogram, we notice that the average net cost for IV institutions typically is $16000-$18000.
+The values between $16000 -$20000 are quite frequent.
+There were an unsually lagrge number of costs.
+The institutions with cost of more than 50000 were atypical.
+Due to the atypical large value, the histogram is abviously right skewed.
+Without those extreme values, the histogram would be reasonably symmetric.
 
@@
 With the following codes, we extracted the institutions which have the highest and lowest percentage of students who are financially independent.
-Then we observed this varible's basic summary statistics: min, max, mean, median and standard deviation.
+Then we observed this variable's basic summary statistics: min, max, mean, median and standard deviation.
 ```{R}
@@
 From the histogram, we are able to conclude that in most colleges, students' family income fall into the range of $12,000 - $18,000.
-Due to the some influential observations, the histogram is obviously skewed to right.
+Due to some atypical large values, the histogram is obviously skewed to right.
 The values larger than $18,000 are tend to be extreme and the cases larger than $75000 are rare, which means that rare colleges have the students with family income more than $75000.
@@
 
-The median default rate arosses all universities is 13%, which means that more than 50% of the universities have more than 13 out of 100 students defaulted their loan!
+The median default rate across all universities is 13%, which means that more than 50% of the universities have more than 13 out of 100 students defaulted their loan!
 
@@
 
-To compare the percentage of Pell Grant in different control types, we grouped the Percentage of Pell Grant data usubf college type to take a further analysis.
+To compare the percentage of Pell Grant in different control types, we grouped the Percentage of Pell Grant data using college type to take a further analysis.
 
@@
 The geom_violin() function is a blend of geom_boxplot() and geom_density(), which shows a mirrored density plot.
-For better visualization, we use geom_boxplot() to create a box and whiskers plot, which shows five summary statistics (the median, two hinges and two whiskers), and all “outlying” points individually.
+For better visualization, we use geom_boxplot() to create a box and whiskers plot, which shows five summary statistics (the median, two hinges and two whiskers), and all ???outlying???
+points individually.
 And the summary() function further displays the exact values.

### 2017-02-22 Draft-xuezhou-sjb.Rmd
fix index and grammers

@@
 * 4.[Objective](#objective)
-* 5.[Summary of variables](#variablesummary)
-* 5.1 [Individual variable summary](#5.1)
+* 5.[Summary of Variables](#variablesummary)
+* 5.1 [Individual Variable Summary](#5.1)
   * 5.1.1 [School](#5.1.1)
@@
   * 5.1.2 Student Body
-    * 5.1.2.1 [Percentage first-generation students](#5.1.2.1)
-    * 5.1.2.2 [Number of applications](#5.1.2.2)
+    * 5.1.2.1 [Percentage First-generation Students](#5.1.2.1)
+    * 5.1.2.2 [Number of Applications](#5.1.2.2)
   * 5.1.3 Cost 
-    * 5.1.3.1 [Average net price for Title IV institutions(both public and private institutions)](#5.1.3.1)
-    * 5.1.3.2 [Percentage of students who are financially independent](#5.1.3.2)
+    * 5.1.3.1 [Average Net Price for Title IV Institutions (Both Public and Private Institutions)](#5.1.3.1)
+    * 5.1.3.2 [Percentage of Students Who are Financially Independent](#5.1.3.2)
   * 5.1.4 Financial Aid
-    * 5.1.4.1 [Percentage of undergraduates who receive a Pell Grant](#5.1.4.1)
-    * 5.1.4.2 [Percent of all federal undergraduate students receiving a federal student loan](#5.1.4.2l)
-    * 5.1.4.3 [The median debt for students who have completed](#5.1.4.3)
-    * 5.1.4.4 [The median debt for students who have not completed](#5.1.4.4)
+    * 5.1.4.1 [Percentage of Undergraduates Who Receive a Pell Grant](#5.1.4.1)
+    * 5.1.4.2 [Percent of all Federal Undergraduate Students Receiving a Federal Student Loan](#5.1.4.2l)
+    * 5.1.4.3 [The Median Debt for Students Who Have Completed](#5.1.4.3)
+    * 5.1.4.4 [The Median Debt for Students Who Have not Completed](#5.1.4.4)
   * 5.1.5 Earning
-    * 5.1.5.1 [Median family income in real 2015 dollars](#5.1.5.1)
+    * 5.1.5.1 [Median Family Income in real 2015 Dollars](#5.1.5.1)
     * 5.1.5.2 [Earning after graduation](#5.1.5.2)
@@
 * 6.Relationship between variables
- * 6.1 [The relationship between NPT4_COMBINE and State](#6.1)
- * 6.2 [The relationship between Debt and Cost](#6.2)
- * 6.3 [The relationship between Percentage of Pell Grant and college type(CONTROL)](#6.3)
- * 6.4 [The relationship between Percentage first-generation students and States](#6.4)
- * 6.5 [The relationship between family income and percentage of Pell Grant](#6.5)
- * 6.6 [The relationship between DEP_STAT_PCT_IND and Control](#6.6)
- * 6.7 [The relationship between PCTLOAN and Control](#6.7)
- * 6.8 [cost and median debts with earning](#6.8)
- * 6.9 [default with earning](#6.9)
- * 6.10 [degree type vs. earning](#6.10)
- * 6.11 [predominant degree, control  and earning after 6 year of entry](#6.11)
- * 6.12 [predominant degree + control (count)](#6.12)
- * 6.13 [control + earning](#6.13)
+ * 6.1 [The Relationship between NPT4_COMBINE and State](#6.1)
+ * 6.2 [The Relationship between Debt and Cost](#6.2)
+ * 6.3 [The Relationship between Percentage of Pell Grant and College Type (CONTROL)](#6.3)
+ * 6.4 [The Relationship between Percentage First-generation Students and States](#6.4)
+ * 6.5 [The Relationship between Median Family Income and Percentage of Pell Grant](#6.5)
+ * 6.6 [The Relationship between the Percentage of Financially-independent Students and College Type (Control)](#6.6)
+ * 6.7 [The Relationship between Percentage of Students Who Receive the Federal Loans and College Type (Control).](#6.7)
+ * 6.8 [The Relationsip between Median Debts and Post-Graduation Earnings](#6.8)
+ * 6.9 [The Relashionship between the Default Rate of Loan Repayment and Post-graduation Earnings ](#6.9)
+ * 6.10 [The Relashionship between Earning 6 Years from Entry and Degree Type](#6.10)
+ * 6.11 [The Relashionship among Predominant Degree, College Type(Control) and Earning after 6 year of Entry](#6.11)
+ * 6.12 [The Relashionship between Predominant Degree and College Type (Control)](#6.12)
+ * 6.13 [The Relashionship between College Type (Control) and Median earnings of Students Working and Not Enrolled 6 Years after Entry](#6.13)
 * 7. [Findings and Conclusion](#7)
@@
 The data provides insights into the performance of institutions that receive federal aid dollars, and the outcomes of the students of those institutions.
-It also gives extra support for students and families to weigh the tradeoffs on choosing the right college according to their own needs and educational goals.
+It also gives extra support for students and families to weigh the tradeoffs on choosing the right colleges according to their own needs and educational goals.
 
 The data was collected by the United States Department of Education, which corresponds to all undergraduate degree-granting institutions of higher education from 1996 through 2016.
-The dataset is pretty rich and covers a wide range of topics.
-It has 1743 associated columns.
-Each corresponds to one named data variable.
+The dataset is pretty rich and covering a wide range of topics.
+It has 1743 associated columns, each corresponds to one named data variable.
 All data variables are divided into 10 categories ranging from Root category to Earnings category.
@@
 The goal of our study is to understand what factors are predictive of students' post-graduation earnings through individual variable analysis and relationship detection.
-To this end, we first explored a set of variables that were carefully selected to represent a school's profile, getting to know the distribution of each variable.
+To this end, we first explored a set of variables that were carefully selected to represent a school's profile and students demographic information, getting to know the distribution of each variable.
 Then we further analyzed how these variables are related, particually the relationship between earnings and other variables.
@@
 #3. Dataset Preparation<a id="dataprep"></a>
-Our analysis focuses on the data of year 2012 -2013 of 23 variables.
+Our analysis focuses on the data of year 2012 -2013 of 38 variables.
 We chose the median income 6 years after entry as our predictor variable, and eliminated many fields with a bunch of missing values as well as some fields not pertaining enougn to post-graduation income.
@@
 
-The full dataset contains 1743 variables. 36 variables were intially selected to be analyze.
+The full dataset contains 1743 variables, as we metioned before, 38 variables were intially selected to be analyze.
 ```{r, message=FALSE,warning=FALSE}
@@
 
-Considering the variable NPT4_PUB and NPT4_PRIV are mutually exclusive both indicate net price.
-we integrated the information of the two coloumns into a single column, named as NPT4_COMBINE.
+Considering the variable NPT4_PUB and NPT4_PRIV are mutually exclusive, which both represent average net price, we integrated the information of the two coloumns into a single column, named as NPT4_COMBINE.
 The new column records the average net price for both private and public TITLE IV institutions.
@@
 
-In order to choose the variables that have enough valid data related to the topic, we only considered variables with percent missing value less than 25%. The new data set was named as data_clean. 
+In order to choose the variables that have enough valid observations related to the topic, we only considered variables with the percent missing value less than 25%.
+The new data was named as data_clean.
 ```{r, message=FALSE,warning=FALSE}
@@
 
-The data_clean will be the clean dataset we used in the further variable summary and comparison.For numerical variables, we calculate the basic summary statistics such as median, min, max, mode and quantile.
-We also use visualizations such as historgams and boxplots to understand the distribution of the values.
-For categorical variables, we use frequency table to understand the frequencies of each category.
-
-The table below lists the variables that will be used for the further analysis: 13 of them are continuous variables while 5 of them are categorical variables or Unique ID.
+For numerical variables, we calculated the basic summary statistics such as median, min, max, standard deviation and mean.
+We also used visualizations such as historgams and boxplots to understand the distribution of the values.
+For categorical variables, we used frequency table to understand the frequencies of each category.
+
+The table below listed the variables that will be used for the further analysis: 13 of them are continuous variables while 5 of them are categorical variables or Unique ID.
 After initial data cleaning process, the new data set now contains 18 variables and 7793 observations.
@@
 
-we will summarize the data set using functions provided in dplyr and ggplot funtions in the variables summary section.
+we summarized the data set using functions provided in dplyr and ggplot in the variables summary section.
 
@@
 
-Our objective at this stage is to conduct an exploration analysis on the CollegeScore dataset, investigating the individual variable's summary statistics and the relationship between these variabls, so as to better prepare for further analaysis on the factors affecting students' post college earnings with some advanced data mining techniques.
-The summary statistics, dplyr function, ggplot functions as well as some visualization techniques will be mainly used in this project.
+Our objective at this stage is to conduct an exploratory data analysis on the CollegeScore dataset, investigating the individual variables' summary statistics and the relationship between these variabls, so as to better prepare for further analaysis on the factors affecting students' post college earnings with some advanced data mining techniques.
+The summary statistics, dplyr function, ggplot functions as well as some visualization techniques were mainly used in this project.
 
@@
 ##5.1 Individual variable summary<a id="5.1"></a> 
-
 ###5.1.1 School<a id="5.1.1"></a> 
@@
 
-State variable has 59 levels, which include 55 states in the United States and 4 districts outside mainland US.
+The State variable has 59 levels, which include 55 states in the United States and 4 districts outside mainland US.
 To better understand how many universities each state has, we created a frequency table as shown below, and also a bar chart to visuzalize the top 15 states having the most universities.
@@
 
-####5.1.1.3. Control of Institution<a id="5.1.1.3"></a> 
+####5.1.1.3. Control of Institutions <a id="5.1.1.3"></a> 
 In addition to state and predominant degree, control of institution is another categorical variable in the dataset.
@@
 The lowest value is 6.87%, which existed in Wheaton College.
-From the mean value of 48.08%, we can tell that about for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college.
+From the mean value of 48.08%, we can tell that for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college.
 The standard deviation value of 12.2% indicated a relatively high difference of this element among different colleges.
@@
 The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who list two and more than two institutions when he applied for financial aid when submitting application for university.
-This information is only available to students who submitted their applications for federal aid and does not cover the whole student body.
+This information is only available to the students who submitted their applications for federal aid and does not cover the whole student body.
 Nevertheless, this variable still can provide some useful insights about how many universities a student usually apply for.
@@
 
-The summary statistics show that the median percentage of those students submitting two or more than two application is 41.67% across all universities The institution having the least percentage only has 4% of students applying for more than one university, while the school with the highest percent has as high as 90% of students applying for more than one university.
-The boxplot compares the percent of student applying for only one university with the percent of students applying for more than one.
+The summary statistics show that the median percentage of those students submitting two or more than two application is 41.67% across all universities.
+The institution having the least percentage only has 4% of students applying for more than one university, while the school with the highest percent has as high as 90% of students applying for more than one university.
+The boxplot compares the percent of students applying for only one university with the percent of students applying for more than one.
 We can clearly see that a larger portion of students only listed one institution when submitting their application for financial aid.
@@
 ###5.1.3 Cost<a id="5.1.3"></a>
-####5.1.3.1 Average net price for Title IV institutions(both public and private institutions)<a id="5.1.3.1"></a>
-
-With the followin codes, we extracted the IV institutions with the highest and lowest Average net price.
+####5.1.3.1 Average Net Price for Title IV Institutions(both public and private institutions)<a id="5.1.3.1"></a>
+
+With the following codes, we extracted the IV institutions with the highest and lowest average net price.
 Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
@@
 Considering the elements are collected from Title IV institutions, we reasoned that the negative values may be due to the fact that some institution's financial supoort can cover the students' all the costs, even overbalancing the cost amount.
-There are not much differerence between the mean value and the median value, which is around $15,000.
+There are not much differerences between the mean value and the median value, which is around $15,000.
 We can conclude that the average net price costs for most institutions are falling into  this range.
-The standard deviation value is 7604.884 which is relatively large, indicating a great spread in the data.
+The standard deviation value is $7604.884 which is relatively large, indicating a great spread in the data.
 
@@
 There were an unsually lagrge number of costs.
-The institutions with cost of more than 50000 were atypical.
+The institutions with cost of more than $50000 were atypical.
 Due to the atypical large value, the histogram is abviously right skewed.
@@
 
-From the summary results, we learned that the maximum value of the variable is 98.56%(Capella University) while the minimum value is about 0.73%(University of Notre Dame).
+From the summary results, we learned that the maximum value of the variable is 98.56%(Capella University) while the minimum value is about 0.73% (University of Notre Dame).
 From the mean value, we can tell that in most colleges, around 56% of the students are financially independent.
@@
 
-From the histogram, we notice that the values between 0.7 to 0.8 are quite frequent. There was no obvious extreme value in the sample, thus the histogram is normally distributed.
+From the histogram, we notice that the values between 0.7 to 0.8 are quite frequent. There is no obvious extreme value in the sample, thus the histogram is normally distributed.
 
@@
 From the output, we can tell that the range is between 0 to 100%.	No students in Thunderbird School of Global Management has received the Pell grant while all of the students in American Conservatory Theater received the Pell Grant.
-Among the colleges in the sample, half of the colleges colleges have the rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs.
+Among the colleges in the sample, half of the colleges have the rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs.
 The standard deviation value is 22.57%,  indicating a relatively high difference of this element among different colleges, which can also be noticed from the largest range (100%).
@@
 This element, as reported in the IPEDs Student Finaicial Aid component, shows the share of undergraduate students who received federal loans in a given year.
-It can provide important context to figures related to debt, repayment and non-repayment.
-We are using the code below to summarize this variable and extract the institutions with highest and lowest value.
+It can provide important context to figures related to debt, repayment and non-repayment.We are using the code below to summarize this variable and extract the institutions with highest and lowest value.
 
@@
 
-From the output, we can tell that the range is between 0 to 100%.	As we can recall from the previous variable(PCTPELL), Gadsden State Community College has no students receiving federal loans in the given year.
-Among the colleges in the sample, half of colleges have this rate as high as 60.08% , which means around 60% of undergraduate students receive the fedaral student loans.
+From the output, we can tell that the range is between 0 to 100%.	As we can recall from the previous variable (PCTPELL), Gadsden State Community College has no students receiving federal loans in the given year.
+Among the colleges in the sample, half of colleges have this rate as high as 60.08% , which means around 60% of undergraduate students received the fedaral student loans in the given year.
 The standard deviation value is 29.26%,  indicating a relatively high difference of this element among different colleges, which can also be noticed from the largest range (100%).
@@
 
-From the histogram, it's obvious that there are two peaks in the graph. The rate between 0-10% and the rates between 60% to 80% have similarly frequency compared to the other ranges.
+From the histogram, it's obvious that there are two peaks in the graph. The rate between 0-10% and the rates between 60% to 80% have similar frequency compared to the other ranges.
 
@@
 
-This variable recorded debt for students who have completed the education. We are using the code below to summarize this variable. 
+This variable recorded debt for students who have completed the education. We are using the code below to summarize this variable and extracted the institutions with largest and smallest value.
+
 ```{R, message=FALSE, warinig = FALSE}
@@
 We use summary statistics, standard deviation and histogram to understand the distribution of this variable.
-The summary statistics show that students in more than 50% of institution have a median debt of more than $5500, which is not very unaffordable.
-The median debt for attending students can be as much as $27620 for some school and as low as $600 for others, which indicates that it may have a large variance.
+The summary statistics show that students in more than 50% of institutions have a median debt of more than $5500, which is not very unaffordable.
+The median debt for attending students could be as much as $27620 for some schools and as low as $600 for others, indicating that it may have a large variance.
 So we also calculate standard deviation to measure the dispersion of the distribution.
@@
 From the output, we can tell that the highest value is $155,606.6 for Jewish Theological Seminary of America, while the lowest value is $0 for Associated Technical College-Los Angeles.
-From the median value of $19,262.26, we can tell that among the colleges in the sample, the students in half of them had the family income below $19,262.26	.
+From the median value of $19,262.26, we can tell that among the colleges in the sample, the students in half of them had the family income below $19,262.26.
 The standard deviation value is $18,824.31, indicating a relatively high difference of this element among different colleges.
@@
 Due to some atypical large values, the histogram is obviously skewed to right.
-The values larger than $18,000 are tend to be extreme and the cases larger than $75000 are rare, which means that rare colleges have the students with family income more than $75000.
+The values larger than $18,000 are tending to be extreme and the cases larger than $75000 are rare, which means that rare colleges have the students with family income more than $75000.
 
@@
 
-We selected two variables MD_EARN_WNE_P6, MD_EARN_WNE_P10 in the earning group to study students' post-graduation earnings.
+MD_EARN_WNE_P6 and MD_EARN_WNE_P10 are two of the earning varibles we selected to study students' post-graduation earnings.
 MD_EARN_WNE_P6 is the median earning for students after six years of their entry, which corresponds to two years after graduation for a four-year program.
-Likewise, MD_EARN_WNE_P10 is the median earning ten years after their entry, which usually corresponds to four years after graduation These two metrics give us an idea of how much money students are making after their graduation, in both short and medium terms.
+Likewise, MD_EARN_WNE_P10 is the median earning ten years after their entry, which usually corresponds to four years after graduation.
+These two metrics give us an idea of how much money students are making after their graduation, in both short and medium terms.
 
@@
 Besides, the ten-year post-graduation earning has a wider spread than that of the six-years.
-This indicates that there is a greater gap between who are earning more and who are earning less, which can be confirmed by an increase in the standard deviation.
+This indicates that there is a greater gap of the earning amount among the different institutions, which can be confirmed by an increase in the standard deviation.
 
@@
 ####5.1.6.1 3-year Repayment rate<a id="5.1.6.1"></a>
-The variable RPY_3YR_RT_SUPP measures the percentage of students who entered repayment from October 1, 2010 to September 30, 2011, and made progress in the following three years.
+The variable RPY_3YR_RT_SUPP measures the percentage of students who entered repayment from October 1, 2010 to September 30, 2011 and made progress in the following three years.
 Repayment rates are generally considered more sensitive than default rates(we will study later), which measures only the worst-case scenario for repayment outcomes.
@@
 #6. Relationship between variables<a id="6"></a>
-
 ##6.1. The relationship between average net cost for Title IV institutions and State<a id="6.1"></a>
@@
 To analyze the relationship between the average net cost for Title IV institutions and State, we grouped the data into the state level to take a further look.
-We would like to investigate the distribution of average net price in each state.
+We would like to investigate the distribution of institutions ' average net price in each state.
 
@@
 Based on the output, we can see that the highest median value for the avreage net price exists in DC while the lowest median value is in AS(American Samoa).
-Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are all the states have  well-known educational institutions.
+Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are the states with certain well-known educational institutions.
 
@@
 The results show that the universities located in ND, SD, VT, MN and MT have the lowest percentage first-generation students, which indicates that the students in these five states would have a better family education background.
-The five states with the highest share of first-generation students are LA, NJ, MP, MV, MH.
+The five states with the highest share of first-generation students are LA, NJ, MP, MV, and MH.
 
@@
 
-##6.6. The relationship between DEP_STAT_PCT_IND and Control<a id="6.6"></a>
+##6.6. The Relationship between the Percentage of Financially-independent Students and Control<a id="6.6"></a>
 
@@
 
-##6.7. The relationship between PCTLOAN and Control.<a id="6.7"></a>
+##6.7. The Relationship between Percentage of Students Who Receive the Federal Loans and Control.<a id="6.7"></a>
 
@@
 From the result,  private institutions have a much larger median value than the public institutions.
-With a much higher costs burden, it makes sense that the low-income students in private institutions are much more likely to seek financial aid than the public institution students.
+With a much higher costs burden, it makes sense that the low-income students in private institutions are much more likely to seek financial aids than the students in public institutions.
+
 From the stacked histogram, we can tell that for all  three college types, the most frequent range appears between 0.6 and 0.8, which is consistent with the findings we observe in the individual variable summary section.
 
-##6.8 Median debts with earning<a id="6.8"></a> 
-Next, we analyze the relationship between the amount the debt and post-graduation earnings.
+##6.8 The Relationsip between Median Debts and Post-Graduation Earnings<a id="6.8"></a> 
+Next, we analyze the relationship between the amount of the debt and post-graduation earnings.
 There are two variables in the dataset to summarize the median amount of debt for student body, one for students who already graduated and one for students who have not completed their study.
@@
 
-##6.9 default with earning<a id="6.9"></a>
-
-For one thing, understanding the relationship between the default rate of loan repayment and post-graduation earnings can help federal financial committee to improve their decision about load granting.
-If the default rate decreases as the earnings increase, the committee may prefer to grant more financial aid to institutions whose students have higher earnings.
+##6.9 The Relashionship between the Default Rate of Loan Repayment and Post-graduation Earnings<a id="6.9"></a> 
+For one thing, understanding the relationship between the default rate of loan repayment and post-graduation earnings can help federal financial committee to improve their decisions about load granting.
+If the default rate decreases as the earnings increases, the committee may prefer to grant more financial aid to institutions whose students have higher earnings.
 For another, the default rate of repayment for a university can also help predict the post-graduation earnings for its students.
@@
 The resulting r = -0.33 shows that the correlation between earnings and default rate is negative but the relationship is not very strong, which may be due to the influences of some outliers.
-To conclude, the negative relationship between earnings and default rate of loan repayment needs extra data to confirm.
-
-##6.10 degree type vs. earning<a id="6.10"></a>
+To conclude, the negative relationship between earnings and default rate of loan repayment needs extra data to validate.
+
+##6.10 The Relashionship between Earning 6 Years from Entry and Degree Type<a id="6.10"></a>
 
 The higher education one receives, the higher the earning is after graduation.
-This is an assumption we are making when investing higher education.
-But is it really the case.
+This is a common assumption we are making when investing higher education, and this is it really the case.
 In this question, we would like to study if it's worthwhile to invest in higher degree.
@@
 
-From the boxplot below, we see that the higher the education, the higher the median earning 6 years after entry. Especially, a graduate-degree gives a hugh rise to one's earning.
-
-##6.11 predominant degree, control  and earning after 6 year of entry<a id="6.11"></a>
-
-The following plot shows two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry).
+From the boxplot above, we see that the higher the education, the higher the median earning 6 years after entry. Especially, a graduate-degree gives a hugh rise to one's earning.
+
+##6.11 The Relashionship among Predominant Degree, Control and Earning after 6 year of Entry<a id="6.11"></a>
+
+The following plot shows the relationship among two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry).
 
@@
 
-Mean value of the earnings is more representative for the two categories, predominant degree and control According to the plot, we realized that private for-profit colleges whose predominant degree awarded is graduate degree has the lowest average earning with the minimum average value at $3100, while Public college whose predominant degree awarded is graduate degree has the highest average earnings with the maximum average value at $45,850.
-
-##6.12 predominant degree + control (count)<a id="6.12"></a>
+Mean value of the earnings is more representative for the two categories, predominant degree and control.
+According to the plot, we realized that private for-profit colleges whose predominant degree awarded is graduate degree have the lowest average earning with the minimum average value at $3100, while Public colleges whose predominant degree awarded are graduate degree has the highest average earnings with the maximum average value at $45,850.
+
+##6.12 The Relashionship between Predominant Degree and Control<a id="6.12"></a>
 
@@
 For the R code, our team first set the aesthetic mapping on Control_factor as the horizontal axis.
-Then, we use the function geom_bar() to fill different control types using the counting values of PREDDEG, which creates bars based on a grouped frequency distribution.
-After that, we further use the function xtabs() to calculate the exact number for each predominant degree awarded by different types of colleges.
+Then, we used the function geom_bar() to fill different control types using the counting values of PREDDEG, which created bars based on a grouped frequency distribution.
+After that, we used the function xtabs() to calculate the exact number for each predominant degree awarded by different types of colleges.
 
@@
 
-#6.13 control + earning<a id="6.13"></a>
+#6.13 The Relashionship between Control and Median earnings of Students Working and Not Enrolled 6 Years after Entry<a id="6.13"></a>
 
@@
 
-We use the ggplot() function to initialize the ggplot object and declare the input data frame for a graphic with x equals Control_factor and y equals Median Earnings.
-Then, we use the geom_violin() function which shows a compact display of a continuous distribution.
+We used the ggplot() function to initialize the ggplot object and declared the input data frame for a graphic with x equals Control_factor and y equals Median Earnings.
+Then, we used the geom_violin() function which showed a compact display of a continuous distribution.
 The geom_violin() function is a blend of geom_boxplot() and geom_density(), which shows a mirrored density plot.
-For better visualization, we use geom_boxplot() to create a box and whiskers plot, which shows five summary statistics (the median, two hinges and two whiskers), and all ???outlying???
-points individually.
-And the summary() function further displays the exact values.
-The plot superimposed with boxplots shows side-by-side violin plots of median earnings of students working and not enrolled 6 years after entry according to different college types.
+For better visualization, we used geom_boxplot() to create a box and whiskers plot, which showed five summary statistics (the median, two hinges and two whiskers), and all outlying points individually.
+And the summary() function further displayed the exact values.
+The plot superimposed with boxplots showed side-by-side violin plots of median earnings of students working and not enrolled 6 years after entry according to different college types.
 
@@
 
-All the three types of school have far out values above, which can be explored further.
-Obvious differences contain in these three boxplots.
+All the three types of schools have far out values above, which can be explored further.
+Obvious differences were contained in these three boxplots.
 The boxplot of public college is comparatively short, while the boxplot of private nonprofit is comparatively tall.
@@
 
-Through the process of summarizing these variables and analyzing the relationship between these variables, we found that for federally-aided students in most colleges, around half of the students' parents did not complete college.The average net price costs for most institutions are falling into the range between $16000 - $20000.
-Among the 7703 institution samples, more than 50% of the students receive a Pell Grant or Federal Loans as means of finanical support, and on average, 60% of students are financially independent.
-In most colleges, the students have the family income between $12000 and $18000.
+Through the process of summarizing these variables and analyzing the relationship between these variables, we found that for federally-aided students in most colleges, around half of the students' parents did not complete colleges.The average net price costs for most institutions are falling into the range between $16000 - $20000.
+Among the 7703 institution samples, more than 50% of the students received a Pell Grant or Federal Loan as means of finanical support, and on average, 60% of students are financially independent.
+In most colleges, the students are able to reach the family income between $12000 and $18000 after they graduate.
 
@@
 
-It indicates that there exists a negative correlation between the percentage of students who receive a Pell Grant and the students' median family income; for institutions with low percentage of students who receive the Pell Grant, the financial aid like the Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, since these college students will tend to have a high median family income.
+It indicates that there exists a negative correlation between the percentage of students who receive a Pell Grant and the students' median family income; for institutions with low percentage of students who receive the Pell Grant, the financial aid like the Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, thus these college students will tend to have a high median family income after they graduate.
 

### 2017-02-22 Draft-xuezhou-sjb.Rmd
further fix

@@
 ###5.1.2 Student Body<a id="5.1.2"></a> 
-####5.1.2.1 Percentage first-generation students<a id="5.1.2.1"></a>
+####5.1.2.1 Percentage of First-generation Students<a id="5.1.2.1"></a>
 This element identifies the share of federally aided students at the institution who report that they are the first generaion on the FAFSA (Free Application for Federal Student Aid).
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for Percentage first-generation students") +
+    labs(title="Histogram for Percentage First-Generation Students") +
     labs(x="PAR_ED_PCT_1STGEN", y="Count") +
@@
 
-With the following codes, we extracted the IV institutions with the highest and lowest average net price.
+With the following codes, we extracted the IV institutions' basic demographic information with the highest and lowest average net price.
 Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
@@
 
-From the summary results, we learned that the maximum value of the variable is 98.56%(Capella University) while the minimum value is about 0.73% (University of Notre Dame).
+From the summary results, we learned that the maximum value of the variable is 98.56% (Capella University) while the minimum value is about 0.73% (University of Notre Dame).
 From the mean value, we can tell that in most colleges, around 56% of the students are financially independent.
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for Percentage of students who are financially independent") +
+    labs(title="Histogram for Percentage of Students Who Are Financially Independent") +
     labs(x="DEP_STAT_PCT_IND", y="Count") +
@@
 This element, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received Pell Grants in a given year, which is an important measure of the financial aid an institution provides to low-income students.
-We are using the code below to summarize this variable and extract the institutions with the highest and lowest value.
+We are using the code below to summarize this variable(min, max, mean, median and standard deviation) and extract the institutions with the highest and lowest value.
 
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for Percentage of undergraduates who receive a Pell Grant") +
+    labs(title="Histogram for Percentage of Undergraduates Who Receive a Pell Grant") +
     labs(x="PCTPELL", y="Count") +
@@
 
-####5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan<a id="5.1.4.2"></a>
+####5.1.4.2 Percent of All Federal Undergraduate Students Receiving a Federal Student Loan<a id="5.1.4.2"></a>
 
 This element, as reported in the IPEDs Student Finaicial Aid component, shows the share of undergraduate students who received federal loans in a given year.
-It can provide important context to figures related to debt, repayment and non-repayment.We are using the code below to summarize this variable and extract the institutions with highest and lowest value.
+It can provide important context to figures related to debt, repayment and non-repayment.
+We were using the code below to summarize this variable(min, max, mean, median and standard deviation) and extracted the institutions with highest and lowest value.
 
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for Percent of all federal undergraduate students receiving a federal student loan") +
+    labs(title="Histogram for Percent of all Federal Undergraduate Students Receiving a Federal Student Loan") +
     labs(x="PCTFLOAN", y="Count") +
@@
 
-This variable recorded debt for students who have completed the education. We are using the code below to summarize this variable and extracted the institutions with largest and smallest value.
+This variable recorded debt for students who have completed the education.
+We were using the code below to summarize this variable(min, max, mean, median and standard deviation) and extracted the institutions with largest and smallest value.
 
@@
 #6. Relationship between variables<a id="6"></a>
-##6.1. The relationship between average net cost for Title IV institutions and State<a id="6.1"></a>
-
-To analyze the relationship between the average net cost for Title IV institutions and State, we grouped the data into the state level to take a further look.
-We would like to investigate the distribution of institutions ' average net price in each state.
+##6.1. The Relationship between Average Net Cost for Title IV institutions and State<a id="6.1"></a>
+
+To analyze the relationship between the average net cost for Title IV institutions and State, we grouped the data into state level to take a further analysis on each state's min, max and median value of the Average Net Cost - We would like to investigate the distribution of institutions ' average net price in each state.
 
@@
 
-To compare the percentage of Pell Grant in different control types, we grouped the Percentage of Pell Grant data using college type to take a further analysis.
+To compare the percentage of Pell Grant in different control types, we grouped the Percentage of Pell Grant data using college type to take a further analysis on each college type's min, max and median value of Percentage of Pell Grant.
 
@@
 
-To compare the percentage of first-generation students in each states, we grouped data into the state level to take a further analysis.
+To compare the percentage of first-generation students in each states, we grouped data into the state level to take a further analysis on the min, max and median value of Percentage first-generation students on the state base.
+The top 5 states as well as the lowest 5 states are extracted.
 
@@
 
-##6.5. The relationship between family income and percentage of Pell Grant<a id="6.5"></a>
+##6.5. The relationship between Family Income and Percentage of Pell Grant<a id="6.5"></a>
 
@@
 
-To compare the percentage of financially-independent students in the institutions with different college type, we grouped the data by the control type(private for-profit, public, private nonprofit) to take a further analysis.
+To compare the percentage of financially-independent students in the institutions with different college type, we grouped the data by the control type(private for-profit, public, private nonprofit) to take a further analysis on the min, max and median value for each type.
 
@@
 
-To compare the percentage of students who receive the federal loans in the given year in each schools control type, we grouped the data by the control type(private for-profit, public, private non-profit) to take a further analysis.
+To compare the percentage of students who receive the federal loans in the given year in each schools control type, we grouped the data by the control type(private for-profit, public, private non-profit) to take a further analysis on the min, max and median value of the variable seperately.
 A stacked histogram is also used to generate a more visual presentation.
@@
 
-#6.13 The Relashionship between Control and Median earnings of Students Working and Not Enrolled 6 Years after Entry<a id="6.13"></a>
+##6.13 The Relashionship between Control and Median earnings of Students Working and Not Enrolled 6 Years after Entry<a id="6.13"></a>
 

### 2017-02-22 Draft-xuezhou-sjb.Rmd
fix

@@
 ####5.1.5.1 Median family income in real 2015 dollars<a id="5.1.5.1"></a>
-This element is recorded the median family income of students, one of the undergraduate student demographics for earnings cohorts. We are using the code below to summarize this variable.
+This element is recorded the median family income of students, one of the undergraduate student demographics for earnings cohorts.
+We are using the code below to summarize this variable(min, max, mean, median and standard deviation).
 
@@
 
-##6.9 The Relashionship between the Default Rate of Loan Repayment and Post-graduation Earnings<a id="6.9"></a> 
+##6.9 The Relationship between the Default Rate of Loan Repayment and Post-graduation Earnings<a id="6.9"></a> 
 For one thing, understanding the relationship between the default rate of loan repayment and post-graduation earnings can help federal financial committee to improve their decisions about load granting.
@@
 
-##6.10 The Relashionship between Earning 6 Years from Entry and Degree Type<a id="6.10"></a>
+##6.10 The Relationship between Earning 6 Years from Entry and Degree Type<a id="6.10"></a>
 
@@
 
-##6.11 The Relashionship among Predominant Degree, Control and Earning after 6 year of Entry<a id="6.11"></a>
+##6.11 The Relationship among Predominant Degree, Control and Earning after 6 year of Entry<a id="6.11"></a>
 
@@
 
-##6.12 The Relashionship between Predominant Degree and Control<a id="6.12"></a>
+##6.12 The Relationship between Predominant Degree and Control<a id="6.12"></a>
 
@@
 
-##6.13 The Relashionship between Control and Median earnings of Students Working and Not Enrolled 6 Years after Entry<a id="6.13"></a>
+##6.13 The Relationship between Control and Median earnings of Students Working and Not Enrolled 6 Years after Entry<a id="6.13"></a>
 

### 2017-02-22 Draft-xuezhou-sjb.Rmd
fix
fix typo

@@
  * 6.8 [The Relationsip between Median Debts and Post-Graduation Earnings](#6.8)
- * 6.9 [The Relashionship between the Default Rate of Loan Repayment and Post-graduation Earnings ](#6.9)
- * 6.10 [The Relashionship between Earning 6 Years from Entry and Degree Type](#6.10)
- * 6.11 [The Relashionship among Predominant Degree, College Type(Control) and Earning after 6 year of Entry](#6.11)
- * 6.12 [The Relashionship between Predominant Degree and College Type (Control)](#6.12)
- * 6.13 [The Relashionship between College Type (Control) and Median earnings of Students Working and Not Enrolled 6 Years after Entry](#6.13)
+ * 6.9 [The Relationship between the Default Rate of Loan Repayment and Post-graduation Earnings ](#6.9)
+ * 6.10 [The Relationship between Earning 6 Years from Entry and Degree Type](#6.10)
+ * 6.11 [The Relationship among Predominant Degree, College Type(Control) and Earning after 6 year of Entry](#6.11)
+ * 6.12 [The Relationship between Predominant Degree and College Type (Control)](#6.12)
+ * 6.13 [The Relationship between College Type (Control) and Median earnings of Students Working and Not Enrolled 6 Years after Entry](#6.13)
 * 7. [Findings and Conclusion](#7)
@@
 data_clean %>% ggplot(aes(x = WDRAW_DEBT_MDN)) +
-                geom_histogram(binwidth = 450) #warning?
+                geom_histogram(binwidth = 450)
 

### 2017-02-22 Draft-xuezhou-sjb.Rmd
fix

@@
 * 6.Relationship between variables
- * 6.1 [The Relationship between NPT4_COMBINE and State](#6.1)
- * 6.2 [The Relationship between Debt and Cost](#6.2)
+ * 6.1 [The Relationship between Average Net Cost for Title IV institutions and State](#6.1)
+ * 6.2 [The Relationship between Median Debt and Average Net Price for Title IV institutions](#6.2)
  * 6.3 [The Relationship between Percentage of Pell Grant and College Type (CONTROL)](#6.3)

### 2017-02-22 Draft-xuezhou-sjb.Rmd
update

@@
 #1. Introduction<a id="Introduction"></a>
+
 The College Scorecard project was initially designed to increase the transparency of the college data, working as a navigator for students to make informed decisions about their choice for higher education.
@@
 The data was collected by the United States Department of Education, which corresponds to all undergraduate degree-granting institutions of higher education from 1996 through 2016.
-The dataset is pretty rich and covering a wide range of topics.
-It has 1743 associated columns, each corresponds to one named data variable.
-All data variables are divided into 10 categories ranging from Root category to Earnings category.
+The dataset is pretty rich and it covers a wide range of topics.
+It has 1743 associated columns, each corresponding to one named data variable.
+All data variables are divided into 10 categories including basic information of institution, earnings, financial aid, student body, etc.
 The data type contains integer, string and float.
 Data elements with an IPEDS value in the Source field are derived directly from the Integrated Postsecondary Education Data System (IPEDS).
-Further detailed description of the dataset can be found in the CollegeScorecardDataDictionary.xlsx excel file.
+Further detailed description of the dataset can be found in the full data documentation.
 
@@
 #3. Dataset Preparation<a id="dataprep"></a>
-Our analysis focuses on the data of year 2012 -2013 of 38 variables.
-We chose the median income 6 years after entry as our predictor variable, and eliminated many fields with a bunch of missing values as well as some fields not pertaining enougn to post-graduation income.
-We then fosued primarily on a set of 23 variables, including age, acedemic costs, family income, completetion rates and repayment rate.
+Our analysis focuses on the data of year 2012 -2013.
+We chose the median income 6 years after entry as our response variable, and eliminated many fields with a bunch of missing values as well as some fields not pertaining to post-graduation income.
+We fosued primarily on a set of 38 variables, including age, academic costs, family income, completetion rates, repayment rate, etc.
 
@@
 
-The dataset spans nearly 20 years of data and covers multiple sources including IPEDS, NSLDS, and Department of Treasury.
-Since our goal was to study post-graduation for students in each university, we used 2012-2013 dataset which contains tha most recent earning information.
+The dataset spans nearly 20 years and covers multiple sources including IPEDS, NSLDS, and Department of Treasury.
+Since our goal was to study post-graduation earnings for students in each university, we used 2012-2013 dataset which contains tha most recent earning information.
 
@@
 
-Considering there is some special characters in the ID column, we first renamed the first column to make it easier to understand.
+Considering there are some special characters in the ID column, we first renamed the first column to make it easier to understand.
 ```{R}
@@
 
-The full dataset contains 1743 variables, as we metioned before, 38 variables were intially selected to be analyze.
+The full dataset contains 1743 variables, as we metioned before, 38 variables were intially selected to be analyzed.
 ```{r, message=FALSE,warning=FALSE}
@@
 
-In order to choose the variables that have enough valid observations related to the topic, we only considered variables with the percent missing value less than 25%.
+In order to choose the variables that have enough valid observations related to the topic, we only considered variables with the percent of missing value less than 25%.
 The new data was named as data_clean.
@@
 
-The table below listed the variables that will be used for the further analysis: 13 of them are continuous variables while 5 of them are categorical variables or Unique ID.
-After initial data cleaning process, the new data set now contains 18 variables and 7793 observations.
+The table below lists the variables that will be used for the further analysis: 13 of them are continuous variables while 5 of them are categorical variables or Unique ID.
+After initial data cleaning process, the new data set contained 18 variables and 7,793 observations.
 
@@
 ####5.1.1.2. Predominant Degree<a id="5.1.1.2"></a> 
-Then we analyze the type of degree universities mainly offer.
+Then we analyzed the type of degree universities mainly offer.
 There are four types of degree taken into account: certificate degree, associate degree, bachelor degree, and graduate degree.
-A frequency table in a format of percentage and a histogram are used to help understand the distribution of the predominant degree.
+A frequency table in a format of percentage and a histogram were used to help understand the distribution of the predominant degree.
 
@@
 In addition to state and predominant degree, control of institution is another categorical variable in the dataset.
-The label of Control is to indicate whether or not an institution is privately owned and whether or not it is a for-profit organization.
-There are three levels of this variable: Public, Private Nonprofit, and Private-profit.
-Public school is assumed to be non-profit so it is not further divided into two subgroups.
+The label of control is to indicate whether or not an institution is privately owned and whether or not it is a for-profit organization.
+There are three levels of this variable: public, private non-profit, and private-profit.
+Public schools are assumed to be non-profit so they are not further divided into two subgroups.
 Below we show the percentage summary of each category and a bar chart is plotted to facilitate understanding.
@@
 
-About 50% of all institutions are private for-profit university, which is the most dominant type. Public and private nonprofit are pretty tied in terms of the number of observation in each category.
+About 50% of all institutions are private for-profit university, which is the most dominant type. Public and private non-profit are pretty tied in terms of the number of observation in each category.
 
@@
 ####5.1.2.1 Percentage of First-generation Students<a id="5.1.2.1"></a>
-This element identifies the share of federally aided students at the institution who report that they are the first generaion on the FAFSA (Free Application for Federal Student Aid).
+This element identifies the share of federally aided students at the institutions who reported that they were the first generaion on the FAFSA (Free Application for Federal Student Aid).
 First generation students are defined as students whose parents didn't attend a college.
@@
 With the following codes, we extracted the institutions with the highest and lowest percentage of first-generation students.
-Then we observed this varible's basic summary statistics: min, max, mean, median and standard deviation.
+Then we observed this variable's basic summary statistics: min, max, mean, median and standard deviation.
 ```{R, message=FALSE, warinig = FALSE}
@@
 
-Below is the code to generate a visual representation in the histogram. 
+Below are the codes to generate a visual representation in the histogram. 
 ```{R}
@@
 
-From the histogram, we can tell the share of the first-generarion students is most frequently falling into the range between 0.5 and 0.6. The values with the value more than 0.8 are rare.
+From the histogram, we can tell the share of the first-generarion students is most frequently falling into the range between 0.5 and 0.6. The observations with the value more than 0.8 are rare.
 
@@
 
-The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who list two and more than two institutions when he applied for financial aid when submitting application for university.
+The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who listed two and more than two institutions when he applied for financial aid when submitting applications for university.
 This information is only available to the students who submitted their applications for federal aid and does not cover the whole student body.
@@
 
-The summary statistics show that the median percentage of those students submitting two or more than two application is 41.67% across all universities.
+The summary statistics show that the median percentage of those students submitting two or more than two applications is 41.67% across all universities.
 The institution having the least percentage only has 4% of students applying for more than one university, while the school with the highest percent has as high as 90% of students applying for more than one university.
@@
 
-With the following codes, we extracted the IV institutions' basic demographic information with the highest and lowest average net price.
+With the following codes, we extracted the Title IV institutions' basic demographic information with the highest and lowest average net price.
 Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
@@
 From the summary results, we learned that the maximum value of the variable is $87,414, which is for Aerosim Flight Academy located in FL.
-while the minimum value is about $-19,360 - the value for Berea College located in KY.
+while the minimum value is about $-19,360, for Berea College located in KY.
 Considering the elements are collected from Title IV institutions, we reasoned that the negative values may be due to the fact that some institution's financial supoort can cover the students' all the costs, even overbalancing the cost amount.
-There are not much differerences between the mean value and the median value, which is around $15,000.
+There is not much differerence between the mean value and the median value, which is around $15,000.
 We can conclude that the average net price costs for most institutions are falling into this range.
@@
 
-Below is the histogram for this variable which is able to provide us with a visual distribution presentation. 
+Below shows the histogram for this variable, which is able to provide us with a visual distribution presentation. 
 ```{R}
@@
 
-From the histogram, we notice that the average net cost for IV institutions typically is $16000-$18000.
-The values between $16000 -$20000 are quite frequent.
-There were an unsually lagrge number of costs.
-The institutions with cost of more than $50000 were atypical.
-Due to the atypical large value, the histogram is abviously right skewed.
+From the histogram, we notice that the average net cost for most IV institutions falls between $16,000-$20,000.
+There were some unusually large observations.
+The institutions with cost of more than $50,000 were atypical.
+Due to the these large values, the histogram is abviously right-skewed.
 Without those extreme values, the histogram would be reasonably symmetric.

### 2017-02-22 Draft-xuezhou-sjb.Rmd
update family income

@@
     * 5.1.2.2 [Number of Applications](#5.1.2.2)
+    * 5.1.2.3 [Median Family Income in Real 2015 Dollars](#5.1.2.3)
   * 5.1.3 Cost 
@@
   * 5.1.5 Earning
-    * 5.1.5.1 [Median Family Income in real 2015 Dollars](#5.1.5.1)
-    * 5.1.5.2 [Earning after graduation](#5.1.5.2)
+    * 5.1.5.1 [Earning after graduation](#5.1.5.1)
   * 5.1.6 Repayment
@@
  * 6.12 [The Relationship between Predominant Degree and College Type (Control)](#6.12)
- * 6.13 [The Relationship between College Type (Control) and Median earnings of Students Working and Not Enrolled 6 Years after Entry](#6.13)
+ * 6.13 [The Relationship between College Type (Control) and Median Earnings of Students Working and Not Enrolled 6 Years after Entry](#6.13)
 * 7. [Findings and Conclusion](#7)
@@
 | PAR_ED_PCT_1STGEN| Numeric  | %         |Percentage first-generation students                                  |      Sudent Body                |
-| MD_FAMINC        | Numeric  | $         |Median family income in real 2015 dollars                             |       Earning                   |
+| MD_FAMINC        | Numeric  | $         |Median family income in real 2015 dollars                             |    Student Body                   |
 | PCTPELL          | Numeric  | %         |Percentage of undergraduates who receive a Pell Grant                 |        Financial Aid            |
@@
 
+####5.1.2.3 Median family income in real 2015 dollars<a id="5.1.2.3"></a>
+This element is recorded the median family income of students, one of the undergraduate student demographics for earnings cohorts.
+We are using the code below to summarize this variable(min, max, mean, median and standard deviation).
+
+```{R, message=FALSE, warinig = FALSE}
+
+data_clean %>%
+ slice(c(which.min(MD_FAMINC),which.max(MD_FAMINC))) %>%
+  select (c(1,2,3,9))
+
+summarize(data_clean,
+          minMD_FAMINC=min(MD_FAMINC ,na.rm=TRUE),
+          maxMD_FAMINC=max(MD_FAMINC,na.rm=TRUE),
+          meanMD_FAMINC=mean(MD_FAMINC,na.rm=TRUE),
+          medianMD_FAMINC=median(MD_FAMINC,na.rm=TRUE),
+          sdMD_FAMINC = sd(MD_FAMINC,na.rm=TRUE))
+
+```
+
+From the output, we can tell that the highest value is $155,606.6 for Jewish Theological Seminary of America, while the lowest value is $0 for Associated Technical College-Los Angeles.
+From the median value of $19,262.26, we can tell that among the colleges in the sample, the students in half of them had the family income below $19,262.26.
+The standard deviation value is $18,824.31, indicating a relatively high difference of this element among different colleges.
+Below is the code to generate a visual representation in the histogram
+
+```{R}
+
+  ggplot(data=data_clean , aes(MD_FAMINC)) + 
+    geom_histogram(breaks=seq(0, 156000, by =3000), 
+                   col="red", 
+                   aes(fill=..count..),na.rm=TRUE)+
+    labs(title="Histogram for Median family income in real 2015 dollars") +
+    labs(x="MD_FAMINC", y="Count") +
+    theme(plot.title = element_text(hjust = 0.5))
+```
+
+From the histogram, we are able to conclude that in most colleges, students' family income fall into the range of $12,000 - $18,000.
+Due to some atypical large values, the histogram is obviously skewed to right.
+The values larger than $18,000 are tending to be extreme and the cases larger than $75000 are rare, which means that rare colleges have the students with family income more than $75000.
+
 ###5.1.3 Cost<a id="5.1.3"></a>
@@
 ###5.1.5 Earning<a id="5.1.5"></a>
-####5.1.5.1 Median family income in real 2015 dollars<a id="5.1.5.1"></a>
-This element is recorded the median family income of students, one of the undergraduate student demographics for earnings cohorts.
-We are using the code below to summarize this variable(min, max, mean, median and standard deviation).
-
-```{R, message=FALSE, warinig = FALSE}
-
-data_clean %>%
- slice(c(which.min(MD_FAMINC),which.max(MD_FAMINC))) %>%
-  select (c(1,2,3,9))
-
-summarize(data_clean,
-          minMD_FAMINC=min(MD_FAMINC ,na.rm=TRUE),
-          maxMD_FAMINC=max(MD_FAMINC,na.rm=TRUE),
-          meanMD_FAMINC=mean(MD_FAMINC,na.rm=TRUE),
-          medianMD_FAMINC=median(MD_FAMINC,na.rm=TRUE),
-          sdMD_FAMINC = sd(MD_FAMINC,na.rm=TRUE))
-
-```
-
-From the output, we can tell that the highest value is $155,606.6 for Jewish Theological Seminary of America, while the lowest value is $0 for Associated Technical College-Los Angeles.
-From the median value of $19,262.26, we can tell that among the colleges in the sample, the students in half of them had the family income below $19,262.26.
-The standard deviation value is $18,824.31, indicating a relatively high difference of this element among different colleges.
-Below is the code to generate a visual representation in the histogram
-
-```{R}
-
-  ggplot(data=data_clean , aes(MD_FAMINC)) + 
-    geom_histogram(breaks=seq(0, 156000, by =3000), 
-                   col="red", 
-                   aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for Median family income in real 2015 dollars") +
-    labs(x="MD_FAMINC", y="Count") +
-    theme(plot.title = element_text(hjust = 0.5))
-```
-
-From the histogram, we are able to conclude that in most colleges, students' family income fall into the range of $12,000 - $18,000.
-Due to some atypical large values, the histogram is obviously skewed to right.
-The values larger than $18,000 are tending to be extreme and the cases larger than $75000 are rare, which means that rare colleges have the students with family income more than $75000.
-
-####5.1.5.2 Earning after graduation<a id="5.1.5.2"></a>
+
+####5.1.5.1 Earning after graduation<a id="5.1.5.1"></a>
 
@@
 From the scatter plot, it is obvious that the MD_FAMINC is in a negative linear relationship with PCTPELL: the data shows a downhill pattern from left to right; As the MD_FAMINC increases, the value of PCTPELL decreases.
-It makes sense in the real world; for institutions with low percentage of students who receive the Pell Grant, the financial aid like Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, since these collges students tend to have a higher median family income.
+It makes sense in the real world; for institutions with low percentage of students who receive the Pell Grant, the financial aid like Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, as these college students may have enough median family income to cover their costs.
 
@@
 
-##6.8 The Relationsip between Median Debts and Post-Graduation Earnings<a id="6.8"></a> 
+##6.8 The Relationship between Median Debts and Post-Graduation Earnings<a id="6.8"></a> 
 Next, we analyze the relationship between the amount of the debt and post-graduation earnings.
@@
 
-It indicates that there exists a negative correlation between the percentage of students who receive a Pell Grant and the students' median family income; for institutions with low percentage of students who receive the Pell Grant, the financial aid like the Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, thus these college students will tend to have a high median family income after they graduate.
+It indicates that there exists a negative correlation between the percentage of students who receive a Pell Grant and the students' median family income; for institutions with low percentage of students who receive the Pell Grant, the financial aid like the Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, since these college students may have sufficient family income to support their education.
 

### 2017-02-23 Draft-xuezhou-sjb.Rmd
final version

@@
     * 5.1.4.4 [The Median Debt for Students Who Have not Completed](#5.1.4.4)
-  * 5.1.5 Earning
-    * 5.1.5.1 [Earning after graduation](#5.1.5.1)
-  * 5.1.6 Repayment
-    * 5.1.6.1 [3-year Repayment rate](#5.1.6.1)
-    * 5.1.6.2 [3-year Default rate](#5.1.6.2)
+  * 5.1.5 Repayment
+    * 5.1.5.1 [3-year Repayment rate](#5.1.5.1)
+    * 5.1.5.2 [3-year Default rate](#5.1.5.2)
 * 6.Relationship between variables
@@
  * 6.10 [The Relationship between Earning 6 Years from Entry and Degree Type](#6.10)
- * 6.11 [The Relationship among Predominant Degree, College Type(Control) and Earning after 6 year of Entry](#6.11)
- * 6.12 [The Relationship between Predominant Degree and College Type (Control)](#6.12)
- * 6.13 [The Relationship between College Type (Control) and Median Earnings of Students Working and Not Enrolled 6 Years after Entry](#6.13)
+ * 6.11 [The Relationship between Earning 6 Years from Entry and 10 Years from Entry](#6.11)
+ * 6.12 [The Relationship among Predominant Degree, College Type(Control) and Earning after 6 year of Entry](#6.12)
+ * 6.13 [The Relationship between Predominant Degree and College Type (Control)](#6.13)
+ * 6.14 [The Relationship between College Type (Control) and Median Earnings of Students Working and Not Enrolled 6 Years after Entry](#6.14)
 * 7. [Findings and Conclusion](#7)
@@
 The standard deviation value 23.3% indicates a relatively high difference among different colleges, which can also be concluded from the difference between the maximum value and the minimum value.
-Below is the code to generate a visual representation in the histogram
+Below is the code to generate a visual representation in the histogram.
 
@@
 
-From the histogram, we notice that the values between 0.7 to 0.8 are quite frequent. There is no obvious extreme value in the sample, thus the histogram is normally distributed.
+From the histogram, we notice that the values between 0.7 to 0.8 are quite frequent. There is no obvious extreme value in the sample, and the histogram appears to be normally distributed.
 
@@
 
-This element, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received Pell Grants in a given year, which is an important measure of the financial aid an institution provides to low-income students.
-We are using the code below to summarize this variable(min, max, mean, median and standard deviation) and extract the institutions with the highest and lowest value.
+This element, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received the Pell Grant in a given year, which is an important measure of the financial aid an institution provides to low-income students.
+We were using the code below to summarize this variable(min, max, mean, median and standard deviation) and extract the institutions with the highest and lowest value.
 
@@
 
-From the output, we can tell that the range is between 0 to 100%.	No students in Thunderbird School of Global Management has received the Pell grant while all of the students in American Conservatory Theater received the Pell Grant.
+From the output, we can tell that the range is between 0 to 100%.	No student in Thunderbird School of Global Management has received the Pell grant while all of the students in American Conservatory Theater received the Pell Grant.
 Among the colleges in the sample, half of the colleges have the rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs.
@@
 This element, as reported in the IPEDs Student Finaicial Aid component, shows the share of undergraduate students who received federal loans in a given year.
-It can provide important context to figures related to debt, repayment and non-repayment.
-We were using the code below to summarize this variable(min, max, mean, median and standard deviation) and extracted the institutions with highest and lowest value.
+It can provide important context to data related to debt, repayment and non-repayment.
+We were using the code below to summarize this variable(min, max, mean, median and standard deviation) and extract the institutions with highest and lowest value.
 
@@
 
-From the output, we can tell that the range is between 0 to 100%.	As we can recall from the previous variable (PCTPELL), Gadsden State Community College has no students receiving federal loans in the given year.
+From the output, we can tell that the range is between 0 to 100%.	As we can recall from the previous variable (PCTPELL), Gadsden State Community College has no student receiving federal loans in the given year.
 Among the colleges in the sample, half of colleges have this rate as high as 60.08% , which means around 60% of undergraduate students received the fedaral student loans in the given year.
@@
 
-From the histogram, it's obvious that there are two peaks in the graph. The rate between 0-10% and the rates between 60% to 80% have similar frequency compared to the other ranges.
+From the histogram, it's obvious that there are two peaks in the graph. The rate between 0-10% and the rate between 60% to 80% have similar frequency compared to the other ranges.
 
@@
 This variable recorded debt for students who have completed the education.
-We were using the code below to summarize this variable(min, max, mean, median and standard deviation) and extracted the institutions with largest and smallest value.
+We were using the code below to summarize this variable(min, max, mean, median and standard deviation) and extract the institutions with largest and smallest value.
 
@@
 
-From the output, we can tell that the highest value is $48,250, belonging to Southwest University of Visual Arts-Tucson, while the lowest value is 1448, belonging to South Florida Institute of Technology.
-From the median value of 14000, we can tell that among the colleges in the sample, for students who have completed the education, half of them have  median loan debt of $14000.
-The standard deviation value is $8020.958,  indicating a relatively high difference of this element among different colleges.
+From the output, we can tell that the highest value is $48,250, belonging to Southwest University of Visual Arts-Tucson, while the lowest value is $1,448, belonging to South Florida Institute of Technology.
+From the median value of $14,000, we can tell that among the colleges in the sample, for students who have completed the education, half of them have median loan debt of $14,000.
+The standard deviation value is $8,020.958,  indicating a relatively high difference of this element among different colleges.
 Below is the code to generate a visual representation in the histogram.
@@
 
-From the histogram, it is obvious that most students in the sample colleges still have $8000 - $10000 debt amount after they completed the education.
+From the histogram, it is obvious that most students in the sample colleges still have $8,000 - $10,000 debt amount after they completed the education.
 This amount is related to the federal loan amount provided by each institution.
@@
 We use summary statistics, standard deviation and histogram to understand the distribution of this variable.
-The summary statistics show that students in more than 50% of institutions have a median debt of more than $5500, which is not very unaffordable.
-The median debt for attending students could be as much as $27620 for some schools and as low as $600 for others, indicating that it may have a large variance.
-So we also calculate standard deviation to measure the dispersion of the distribution.
-The standard deviation is $2415.623, which means about 95% of institution have  median debt ranging from $1167.754  to $10830.246.
-The histogram shows that the distribution of the median roughly follows a standard distribution with a sharp peak and slightly heavy tail.
-
-###5.1.5 Earning<a id="5.1.5"></a>
-
-####5.1.5.1 Earning after graduation<a id="5.1.5.1"></a>
-
-MD_EARN_WNE_P6 and MD_EARN_WNE_P10 are two of the earning varibles we selected to study students' post-graduation earnings.
+The summary statistics show that students in more than 50% of institutions have a median debt of $5,500, which is not very unaffordable.
+The median debt for attending students could be as much as $27,620 for some schools and as low as $600 for others, indicating that it may have a large variance.
+So we also calculated standard deviation to measure the dispersion of the distribution.
+The standard deviation is $2,415.623, which means about 95% of institutions' students have  median debt ranging from $1167.754  to $10,830.246.
+The histogram shows that the distribution of the median roughly follows a normal distribution with a sharp peak and slightly heavy tail.
+
+###5.1.5 Repayment
+
+####5.1.5.1 3-year Repayment rate<a id="5.1.6.1"></a>
+The variable RPY_3YR_RT_SUPP measures the percentage of students who entered repayment from October 1, 2010 to September 30, 2011 and made progress in the following three years.
+Repayment rates are generally considered more sensitive than default rates(we will study later), which measures only the worst-case scenario for repayment outcomes.
+
+```{r,message=FALSE}
+repayment_clean = na.omit(data_clean$RPY_3YR_RT_SUPP)
+summary(repayment_clean)
+qplot(repayment_clean, geom = "histogram") +
+xlab("Repayment Rate") +
+ylab("Count") +
+labs(title="Histogram for Repayment Rate") +
+theme(plot.title = element_text(hjust = 0.5))
+
+```
+
+From the summary statistics, we see that 50% institutions have more than 40% of students who were making progress in their repayment.
+A histogram is also plotted to show the distribution, clearly with a heavy tail.
+
+####5.1.5.2 3-year Default rate<a id="5.1.6.2"></a>
+
+The three-year cohort default rate describes the repayment in the opposite of repayment rate.
+In this dataset, it was calculated using the cohort of borrowers who entered repayment on their federal student loans between Octorber 1, 2010 to September 30, 2011, and defaulted before September 2012.
+
+```{r,message=FALSE}
+default_clean = na.omit(data_clean$CDR3)
+summary(default_clean)
+qplot(default_clean, geom = "histogram") +
+xlab("Default Rate") +
+ylab("Count") +
+labs(title="Histogram for Default Rate") +
+theme(plot.title = element_text(hjust = 0.5))
+
+summary(data_clean$RPY_3YR_RT_SUPP + data_clean$CDR3)
+```
+
+The median default rate across all universities is 13%, which means that more than 50% of the universities have more than 13% of the students defaulted their loan!
+
+#6. Relationship between variables<a id="6"></a>
+##6.1. The Relationship between Average Net Cost for Title IV institutions and State<a id="6.1"></a>
+
+To analyze the relationship between the average net cost for Title IV institutions and State, we grouped the data into state level to take a further analysis on each state's min, max and median value of the Average Net Cost; we would like to investigate the distribution of institutions' average net price in each state.
+
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(STABBR_factor) %>%
+    dplyr::summarise(minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),    
+    maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE))  %>%
+       arrange(desc(medianNPT4_COMBINE))   %>%
+{.}  ->NPT4_STATE
+
+NPT4_STATE
+
+```
+
+Based on the output, we can see that the highest median value for the avreage net price exists in DC while the lowest median value is in AS(American Samoa).
+Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are the states with certain well-known educational institutions.
+
+##6.2. The relationship between median debt and average net price for Title IV institutions<a id="6.2"></a>
+
+We hypothesized that the students who completed education in colleges with high costs will have a higher value of median debt.
+We use the following code to generate the scatterplot between these two variables to verify our hypothesis.
+
+```{R}
+
+ggplot(data_clean, aes(x =NPT4_COMBINE, y = GRAD_DEBT_MDN)) + geom_point(color="red")+  geom_smooth(method = "lm", se = TRUE)+
+  labs(title="Scatterplot between median debt and average net price for Title IV institutions ") +
+    theme(plot.title = element_text(hjust = 0.5))
+
+```
+
+From the scatter plot, it is obvious that the NPT4_COMBINE is in a positive linear relationship with GRAD_DEBT_MDN: the data shows an uphill pattern from left to right.
+As the average net price for Title IV institutions increases, the value of median debt for students who have completed the college increases as well.
+
+##6.3. The relationship between Percentage of Pell Grant and college type(CONTROL)<a id="6.3"></a>
+
+To compare the percentage of Pell Grant in different control types, we grouped the Percentage of Pell Grant data using college type to take a further analysis on each college type's min, max and median value of Percentage of Pell Grant.
+
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(Control_factor ) %>%
+    dplyr::summarise(minPCTPELL=min(PCTPELL ,na.rm=TRUE),    
+    maxPCTPELL=max(PCTPELL,na.rm=TRUE),medianPCTPELL=median(PCTPELL,na.rm=TRUE))  %>%
+       arrange(medianPCTPELL)   %>%
+{.}  ->PCTPELL_CONTROL
+PCTPELL_CONTROL
+
+```
+
+From the result, the private nonprofit institutions and the public institutions both have a low median values while the private for-profit institutions have a higher value.
+This is reasonable because the private for-profit colleges may have much higher costs than the private nonprofit and public institutions.
+The low-income students were more likely to seek a Pell Grant as financial assistance.
+Additionally, in some states, the state financial aid may be sufficient to cover costs at community colleges; thus the students in those schools may not need to seek for additional financial aid like Pell Grant.
+
+Below is the stacked histogram between these two variables. 
+```{r}
+
+ggplot(data_clean, aes(x=PCTPELL, fill=Control_factor)) + geom_histogram(alpha=0.9, position="identity") +
+  labs(title="Stacked Histogram between PCTPELL and Control ") +
+    theme(plot.title = element_text(hjust = 0.5))
+
+```
+
+From the overlaying histogram, the percentages of students who received the Pell Grant for private for-profit institutions are quite frequent between 0.7 and 0.75 ; the percentages of students who received the Pell Grant for public institutions and private nonprofit institutions are quite frequent around 0.4.
+The histogram for private for-profit is slightly left-skewed due to small values near 0.
+The histogram for public and private non-profit institutions are reasonably symmetric.
+
+##6.4. The relationship between Percentage first-generation students and States<a id="6.4"></a>
+
+To compare the percentage of first-generation students in each state, we grouped data into the state level to take a further analysis on the min, max and median value of Percentage first-generation students on the state base.
+The top 5 states as well as the lowest 5 states are extracted.
+
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(STABBR_factor ) %>%
+    dplyr::summarise(minPAR_ED_PCT_1STGEN=min(PAR_ED_PCT_1STGEN ,na.rm=TRUE),    
+    maxPAR_ED_PCT_1STGEN=max(PAR_ED_PCT_1STGEN,na.rm=TRUE),medianPAR_ED_PCT_1STGEN=median(PAR_ED_PCT_1STGEN,na.rm=TRUE))  %>%
+       arrange(medianPAR_ED_PCT_1STGEN)   %>%
+{.}  ->PAR_ED_PCT_1STGEN_STATE
+
+PAR_ED_PCT_1STGEN_STATE
+
+```
+
+The results show that the universities located in ND, SD, VT, MN and MT have the lowest percentage of first-generation students, which indicates that the students in these five states would have a better family education background.
+The five states with the highest share of first-generation students are LA, NJ, MP, MV, and MH.
+
+##6.5. The relationship between Family Income and Percentage of Pell Grant<a id="6.5"></a>
+
+We hypothesized that colleges with small value of family income would tend to have high percentages of Pell Grant.
+The following codes are used to generate the scatter plot between these two variables.
+
+```{r}
+
+ggplot(data_clean, aes(x =MD_FAMINC, y = PCTPELL)) + geom_point(color="red")+  geom_smooth(method = "lm", se = TRUE)+
+  labs(title="Scatterplot between MD_FAMINC and PCTPELL ") +
+    theme(plot.title = element_text(hjust = 0.5))
+
+```
+
+From the scatter plot, it is obvious that the MD_FAMINC is in a negative linear relationship with PCTPELL: the plot shows a downhill pattern from left to right; As the MD_FAMINC increases, the value of PCTPELL decreases.
+It makes sense in the real world; for institutions with low percentage of students who receive the Pell Grant, the financial aid like Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, as these college students may have enough median family income to cover their costs.
+
+##6.6. The Relationship between the Percentage of Financially-independent Students and Control<a id="6.6"></a>
+
+To compare the percentage of financially-independent students in the institutions of different college type, we grouped the data by the control type(private for-profit, public, private nonprofit) to take a further analysis on the min, max and median value for each type.
+
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(Control_factor) %>%
+    dplyr::summarise(minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND ,na.rm=TRUE),    
+    maxDEP_STAT_PCT_IND=max(DEP_STAT_PCT_IND,na.rm=TRUE),medianDEP_STAT_PCT_IND=median(DEP_STAT_PCT_IND,na.rm=TRUE))  %>%
+       arrange(desc(medianDEP_STAT_PCT_IND))   %>%
+{.}  ->DEP_STAT_PCT_IND_CONTROL
+
+DEP_STAT_PCT_IND_CONTROL
+
+```
+
+From the result, we can tell that the highest median value is in private for-profit colleges, which means that in most private for-profit colleges, the students are more likely to be financially independent.
+
+##6.7. The Relationship between Percentage of Students Who Receive the Federal Loans and Control.<a id="6.7"></a>
+
+To compare the percentage of students who received the federal loans in the given year in each control type, we grouped the data by the control type(private for-profit, public, private non-profit) to take a further analysis on the min, max and median value of the variable seperately.
+A stacked histogram is also used to generate a more visual presentation.
+
+```{R message=FALSE, warinig = FALSE}
+
+data_clean %>%
+  group_by(Control_factor ) %>%
+    dplyr::summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
+    maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
+       arrange(medianPCTFLOAN)   %>%
+{.}  ->PCTFLOAN_CONTROL
+
+PCTFLOAN_CONTROL
+
+#stacked histogram
+data_clean %>%
+  ggplot(aes(x=PCTFLOAN)) +
+  labs(title="Stacked histgram between Control and PCTLOAN ")+
+  geom_bar(binwidth=0.1,aes(fill=Control_factor )) + theme(plot.title = element_text(hjust = 0.5))
+
+```
+
+From the result,  private institutions have a much larger median value than the public institutions.
+With a much higher costs burden, it makes sense that the low-income students in private institutions are much more likely to seek financial aid than the students in public institutions.
+
+From the stacked histogram, we can tell that for all three college types, the most frequent range appears between 0.6 and 0.8, which is consistent with the findings we observe in the individual variable summary section.
+
+##6.8 The Relationship between Median Debts and Post-Graduation Earnings<a id="6.8"></a> 
+Next, we analyzed the relationship between the amount of the debt and post-graduation earnings.
+There are two variables in the dataset to summarize the median amount of debt for student body, one for students who already graduated and one for students who have not completed their study.
+We decided to use debt for students who have not completed their study in this problem, since it makes more sense for the purpose of prediction.
+
+```{r, message=FALSE}
+
+data_clean %>% ggplot(aes(x = WDRAW_DEBT_MDN, y = MD_EARN_WNE_P6)) + 
+                geom_jitter() +
+                geom_smooth(method = "lm", se = FALSE, col = 'red') +
+                xlab("Median Amount Debt for Students who have not completed") +
+                ylab("Earning  6 Years from Entry")
+
+data_clean %>% select(WDRAW_DEBT_MDN, MD_EARN_WNE_P6) %>%
+      na.omit() %>%
+       {.} -> earn_debt
+
+cor(earn_debt$WDRAW_DEBT_MDN,earn_debt$MD_EARN_WNE_P6)
+
+```
+
+First, we used a scatter plot to visualize the relationship between earnings and the amount of debt, the plot clearly shows a positive relationship.
+Then we further calculated pearson correlation to confirm our hypothesis.
+The resulting coefficient 0.44 indicates a moderate linear relationship between earnings and debt.
+
+##6.9 The Relationship between the Default Rate of Loan Repayment and Post-graduation Earnings<a id="6.9"></a> 
+For one thing, understanding the relationship between the default rate of loan repayment and post-graduation earnings can help federal financial committee to improve their decisions about loan granting.
+If the default rate decreases as the earnings increases, the committee may prefer to grant more financial aid to institutions whose students have higher earnings.
+For another, the default rate of repayment for a university can also help predict the post-graduation earnings for its students.
+
+```{r, message=FALSE}
+data_clean %>% ggplot(aes(x = CDR3, y = MD_EARN_WNE_P6)) + 
+                geom_jitter() +
+                geom_smooth(method = "lm", se = FALSE, col = 'red')+
+                xlab("Three-year Cohort Default Rate") +
+                ylab("Earning  6 Years from Entry")
+
+data_clean %>% select(CDR3, MD_EARN_WNE_P6) %>%
+      na.omit() %>%
+       {.} -> earn_default
+
+cor(earn_default$CDR3,earn_default$MD_EARN_WNE_P6)
+
+```
+
+A scatter plot is created to visualize the relationship between these two variables.
+From the chart, we see that earnings seem to show a decreasing trend as the default rate increases, which shows a negative relationship between earnings and default rate.
+Then we calculated the correlation coefficient between these two variables.
+The resulting r = -0.33 shows that the correlation between earnings and default rate is negative but the relationship is not very strong, which may be due to the influences of some outliers.
+To conclude, the negative relationship between earnings and default rate of loan repayment needs extra data to validate.
+
+##6.10 The Relationship between Earning 6 Years from Entry and Degree Type<a id="6.10"></a>
+
+The higher education one receives, the higher the earning will be after graduation.
+This is an assumption we are making when investing in higher education.
+But is it really the case.
+In this question, we would like to study if it's worthwhile to invest in higher degree.
+```{r,message = FALSE}
+data_clean %>%
+  ggplot(aes(x=`PREDDEG_factor`,
+             y=MD_EARN_WNE_P6)) +
+  geom_boxplot(aes(fill = `PREDDEG_factor`)) +
+  xlab("Degree Type") +
+  ylab("Earning  6 Years from Entry") +
+  scale_fill_discrete(name="Degree Type")
+
+```
+From the boxplot above, we see that the higher the education, the higher the median earning 6 years after entry. Especially, a graduate-degree gives a hugh rise to one's earning.
+
+##6.11 The Relationship between Earning 6 Years from Entry and 10 Years from Entry<a id="6.11"></a>
+MD_EARN_WNE_P6 and MD_EARN_WNE_P10 are two varibles we selected to study students' post-graduation earnings.
 MD_EARN_WNE_P6 is the median earning for students after six years of their entry, which corresponds to two years after graduation for a four-year program.
@@
 
-The six-year after-entry median earnings of 50% institutions is below the $25,300, while ten-year-after-entry median earning is higher by about $5,500.
+The six-year after-entry median earnings of 50% institutions are below $25,300, while ten-year after-entry median earnings are higher by about $5,500.
 The histogram gives us a more clear idea of the earning distribution.
@@
 
-###5.1.6 Repayment
-
-####5.1.6.1 3-year Repayment rate<a id="5.1.6.1"></a>
-The variable RPY_3YR_RT_SUPP measures the percentage of students who entered repayment from October 1, 2010 to September 30, 2011 and made progress in the following three years.
-Repayment rates are generally considered more sensitive than default rates(we will study later), which measures only the worst-case scenario for repayment outcomes.
-
-```{r,message=FALSE}
-repayment_clean = na.omit(data_clean$RPY_3YR_RT_SUPP)
-summary(repayment_clean)
-qplot(repayment_clean, geom = "histogram") +
-xlab("Repayment Rate") +
-ylab("Count") +
-labs(title="Histogram for Repayment Rate") +
-theme(plot.title = element_text(hjust = 0.5))
-
-```
-
-From the summary statistics, we see that 50% institutions have more than 40% of students who were making progress in their repayment.
-A histogram is also plotted to show the distribution, clearly with a heavy tail.
-
-####5.1.6.2 3-year Default rate<a id="5.1.6.2"></a>
-
-The three-year cohort default rate describes the repayment in the opposite of repayment rate.
-In this dataset, it was calculated using the cohort of borrowers who entered repayment on their federal student loans between Octorber 1, 2010 to September 30, 2011, and defaulted before September 2012.
-
-```{r,message=FALSE}
-default_clean = na.omit(data_clean$CDR3)
-summary(default_clean)
-qplot(default_clean, geom = "histogram") +
-xlab("Default Rate") +
-ylab("Count") +
-labs(title="Histogram for Default Rate") +
-theme(plot.title = element_text(hjust = 0.5))
-
-summary(data_clean$RPY_3YR_RT_SUPP + data_clean$CDR3)
-```
-
-The median default rate across all universities is 13%, which means that more than 50% of the universities have more than 13 out of 100 students defaulted their loan!
-
-#6. Relationship between variables<a id="6"></a>
-##6.1. The Relationship between Average Net Cost for Title IV institutions and State<a id="6.1"></a>
-
-To analyze the relationship between the average net cost for Title IV institutions and State, we grouped the data into state level to take a further analysis on each state's min, max and median value of the Average Net Cost - We would like to investigate the distribution of institutions ' average net price in each state.
-
-```{R message=FALSE, warinig = FALSE}
-
-data_clean %>%
-  group_by(STABBR_factor) %>%
-    dplyr::summarise(minNPT4_COMBINE=min(NPT4_COMBINE ,na.rm=TRUE),    
-    maxNPT4_COMBINE=max(NPT4_COMBINE,na.rm=TRUE),medianNPT4_COMBINE=median(NPT4_COMBINE,na.rm=TRUE))  %>%
-       arrange(desc(medianNPT4_COMBINE))   %>%
-{.}  ->NPT4_STATE
-
-NPT4_STATE
-
-```
-
-Based on the output, we can see that the highest median value for the avreage net price exists in DC while the lowest median value is in AS(American Samoa).
-Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are the states with certain well-known educational institutions.
-
-##6.2. The relationship between median debt and average net price for Title IV institutions<a id="6.2"></a>
-
-We hypothesised that the students who completed education in colleges with high costs will have a higher value of median debt.
-We use the following code to generate the scatterplot between these two variables to verify our hypothesis.
-
-```{R}
-
-ggplot(data_clean, aes(x =NPT4_COMBINE, y = GRAD_DEBT_MDN)) + geom_point(color="red")+  geom_smooth(method = "lm", se = TRUE)+
-  labs(title="Scatterplot between median debt and average net price for Title IV institutions ") +
-    theme(plot.title = element_text(hjust = 0.5))
-
-```
-
-From the scatter plot, it is obvious that the NPT4_COMBINE is in a positive linear relationship with GRAD_DEBT_MDN: the data shows an uphill pattern from left to right; As the average net price for Title IV institutions increases, the value of median debt for students who have completed the college increases as well.
-
-##6.3. The relationship between Percentage of Pell Grant and college type(CONTROL)<a id="6.3"></a>
-
-To compare the percentage of Pell Grant in different control types, we grouped the Percentage of Pell Grant data using college type to take a further analysis on each college type's min, max and median value of Percentage of Pell Grant.
-
-```{R message=FALSE, warinig = FALSE}
-
-data_clean %>%
-  group_by(Control_factor ) %>%
-    dplyr::summarise(minPCTPELL=min(PCTPELL ,na.rm=TRUE),    
-    maxPCTPELL=max(PCTPELL,na.rm=TRUE),medianPCTPELL=median(PCTPELL,na.rm=TRUE))  %>%
-       arrange(medianPCTPELL)   %>%
-{.}  ->PCTPELL_CONTROL
-PCTPELL_CONTROL
-
-```
-
-From the result, the private nonprofit institutions and the public institutions both have a low median values while the private for-profit institutions have a higher value.
-This is reasonable because the private for-profit colleges may have much higher costs than the private nonprofit and public institutions.
-The low-income students were more likely to seek a Pell Grant as assistance.
-Additionally, in some states, state financial aid may be sufficient to cover costs at community colleges; thus the students in those schools may not need to seek for additional financial aid like Pell Grant.
-
-Below is the stacked histogram between these two variables. 
-```{r}
-
-ggplot(data_clean, aes(x=PCTPELL, fill=Control_factor)) + geom_histogram(alpha=0.9, position="identity") +
-  labs(title="Stacked Histogram between PCTPELL and Control ") +
-    theme(plot.title = element_text(hjust = 0.5))
-
-```
-
-From the overlaying histograms, the percentages of students who received the Pell Grant for private for-profit institutions are quite frequent between 0.7 and 0.75 ; the percentages of students who receive the Pell Grant for public institutions and private nonprofit institutions are quite frequent around 0.4.
-The histogram for private for-profit is slightly left skewed due to small values near 0.
-The histogram for public and private non-profit institutions are reasonably symmetric.
-
-##6.4. The relationship between Percentage first-generation students and States<a id="6.4"></a>
-
-To compare the percentage of first-generation students in each states, we grouped data into the state level to take a further analysis on the min, max and median value of Percentage first-generation students on the state base.
-The top 5 states as well as the lowest 5 states are extracted.
-
-```{R message=FALSE, warinig = FALSE}
-
-data_clean %>%
-  group_by(STABBR_factor ) %>%
-    dplyr::summarise(minPAR_ED_PCT_1STGEN=min(PAR_ED_PCT_1STGEN ,na.rm=TRUE),    
-    maxPAR_ED_PCT_1STGEN=max(PAR_ED_PCT_1STGEN,na.rm=TRUE),medianPAR_ED_PCT_1STGEN=median(PAR_ED_PCT_1STGEN,na.rm=TRUE))  %>%
-       arrange(medianPAR_ED_PCT_1STGEN)   %>%
-{.}  ->PAR_ED_PCT_1STGEN_STATE
-
-PAR_ED_PCT_1STGEN_STATE
-
-```
-
-The results show that the universities located in ND, SD, VT, MN and MT have the lowest percentage first-generation students, which indicates that the students in these five states would have a better family education background.
-The five states with the highest share of first-generation students are LA, NJ, MP, MV, and MH.
-
-##6.5. The relationship between Family Income and Percentage of Pell Grant<a id="6.5"></a>
-
-We hypothesed that colleges with small value of family income would tend to have high percentages of Pell Grant. The following codes are used to generate the scatter plots between these two variables.
-
-```{r}
-
-ggplot(data_clean, aes(x =MD_FAMINC, y = PCTPELL)) + geom_point(color="red")+  geom_smooth(method = "lm", se = TRUE)+
-  labs(title="Scatterplot between MD_FAMINC and PCTPELL ") +
-    theme(plot.title = element_text(hjust = 0.5))
-
-```
-
-From the scatter plot, it is obvious that the MD_FAMINC is in a negative linear relationship with PCTPELL: the data shows a downhill pattern from left to right; As the MD_FAMINC increases, the value of PCTPELL decreases.
-It makes sense in the real world; for institutions with low percentage of students who receive the Pell Grant, the financial aid like Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, as these college students may have enough median family income to cover their costs.
-
-##6.6. The Relationship between the Percentage of Financially-independent Students and Control<a id="6.6"></a>
-
-To compare the percentage of financially-independent students in the institutions with different college type, we grouped the data by the control type(private for-profit, public, private nonprofit) to take a further analysis on the min, max and median value for each type.
-
-```{R message=FALSE, warinig = FALSE}
-
-data_clean %>%
-  group_by(Control_factor) %>%
-    dplyr::summarise(minDEP_STAT_PCT_IND=min(DEP_STAT_PCT_IND ,na.rm=TRUE),    
-    maxDEP_STAT_PCT_IND=max(DEP_STAT_PCT_IND,na.rm=TRUE),medianDEP_STAT_PCT_IND=median(DEP_STAT_PCT_IND,na.rm=TRUE))  %>%
-       arrange(desc(medianDEP_STAT_PCT_IND))   %>%
-{.}  ->DEP_STAT_PCT_IND_CONTROL
-
-DEP_STAT_PCT_IND_CONTROL
-
-```
-
-From the result, we can tell the highest median value is in privately for-profit colleges, which means that in most private for-profit colleges, the students are more likely to be financially independent.
-
-##6.7. The Relationship between Percentage of Students Who Receive the Federal Loans and Control.<a id="6.7"></a>
-
-To compare the percentage of students who receive the federal loans in the given year in each schools control type, we grouped the data by the control type(private for-profit, public, private non-profit) to take a further analysis on the min, max and median value of the variable seperately.
-A stacked histogram is also used to generate a more visual presentation.
-
-```{R message=FALSE, warinig = FALSE}
-
-data_clean %>%
-  group_by(Control_factor ) %>%
-    dplyr::summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
-    maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
-       arrange(medianPCTFLOAN)   %>%
-{.}  ->PCTFLOAN_CONTROL
-
-PCTFLOAN_CONTROL
-
-#stacked histogram
-data_clean %>%
-  ggplot(aes(x=PCTFLOAN)) +
-  labs(title="Stacked histgram between Control and PCTLOAN ")+
-  geom_bar(binwidth=0.1,aes(fill=Control_factor )) + theme(plot.title = element_text(hjust = 0.5))
-
-```
-
-From the result,  private institutions have a much larger median value than the public institutions.
-With a much higher costs burden, it makes sense that the low-income students in private institutions are much more likely to seek financial aids than the students in public institutions.
-
-From the stacked histogram, we can tell that for all three college types, the most frequent range appears between 0.6 and 0.8, which is consistent with the findings we observe in the individual variable summary section.
-
-##6.8 The Relationship between Median Debts and Post-Graduation Earnings<a id="6.8"></a> 
-Next, we analyze the relationship between the amount of the debt and post-graduation earnings.
-There are two variables in the dataset to summarize the median amount of debt for student body, one for students who already graduated and one for students who have not completed their study.
-We decided to use debt for students who have not completed their study in this problem, since it makes more sense for the purpose of prediction.
-
-```{r, message=FALSE}
-
-data_clean %>% ggplot(aes(x = WDRAW_DEBT_MDN, y = MD_EARN_WNE_P6)) + 
-                geom_jitter() +
-                geom_smooth(method = "lm", se = FALSE, col = 'red') +
-                xlab("Median Amount Debt for Students who have not completed") +
-                ylab("Earning  6 Years from Entry")
-
-data_clean %>% select(WDRAW_DEBT_MDN, MD_EARN_WNE_P6) %>%
-      na.omit() %>%
-       {.} -> earn_debt
-
-cor(earn_debt$WDRAW_DEBT_MDN,earn_debt$MD_EARN_WNE_P6)
-
-```
-
-First, we used a scatter plot to visualize the relationship between earnings and the amount of debt, the plot clearly shows a positive relationship.
-Then we further calculated pearson correlation to confirm our hypothesis.
-The resulting coefficient 0.44 indicates a moderate linear relationship between earnings and debt.
-
-##6.9 The Relationship between the Default Rate of Loan Repayment and Post-graduation Earnings<a id="6.9"></a> 
-For one thing, understanding the relationship between the default rate of loan repayment and post-graduation earnings can help federal financial committee to improve their decisions about load granting.
-If the default rate decreases as the earnings increases, the committee may prefer to grant more financial aid to institutions whose students have higher earnings.
-For another, the default rate of repayment for a university can also help predict the post-graduation earnings for its students.
-
-```{r, message=FALSE}
-data_clean %>% ggplot(aes(x = CDR3, y = MD_EARN_WNE_P6)) + 
-                geom_jitter() +
-                geom_smooth(method = "lm", se = FALSE, col = 'red')+
-                xlab("Three-year Cohort Default Rate") +
-                ylab("Earning  6 Years from Entry")
-
-data_clean %>% select(CDR3, MD_EARN_WNE_P6) %>%
-      na.omit() %>%
-       {.} -> earn_default
-
-cor(earn_default$CDR3,earn_default$MD_EARN_WNE_P6)
-
-```
-
-A scatter plot is created to visualize the relationship between these two variables.
-From the chart, we see that earnings seems to show a decreasing trend as the default rate increases, which shows a negative relationship between earnings and default rate.
-Then we calculate the correlation coefficient between these two variables.
-The resulting r = -0.33 shows that the correlation between earnings and default rate is negative but the relationship is not very strong, which may be due to the influences of some outliers.
-To conclude, the negative relationship between earnings and default rate of loan repayment needs extra data to validate.
-
-##6.10 The Relationship between Earning 6 Years from Entry and Degree Type<a id="6.10"></a>
-
-The higher education one receives, the higher the earning is after graduation.
-This is a common assumption we are making when investing higher education, and this is it really the case.
-In this question, we would like to study if it's worthwhile to invest in higher degree.
-
-```{r, message=FALSE}
-
-data_clean %>%
-  ggplot(aes(x=`PREDDEG_factor`,
-             y=MD_EARN_WNE_P6)) +
-  geom_boxplot(aes(fill = `PREDDEG_factor`)) +
-  xlab("Degree Type") +
-  ylab("Earning  6 Years from Entry") +
-  scale_fill_discrete(name="Degree Type")
-
-```
-
-From the boxplot above, we see that the higher the education, the higher the median earning 6 years after entry. Especially, a graduate-degree gives a hugh rise to one's earning.
-
-##6.11 The Relationship among Predominant Degree, Control and Earning after 6 year of Entry<a id="6.11"></a>
-
-The following plot shows the relationship among two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry).
-
-For the R code, we first grouped the dataset by PREDDEG factor and Control factor.
-Then, we summarized the average earnings.
+##6.12 The Relationship among Predominant Degree, Control and Earning after 6 year of Entry<a id="6.12"></a>
+
+The following plot shows two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry).
+
+For the R code, we first group the dataset by PREDDEG factor and Control factor.
+Then, we summarize the average earnings.
 After that, we set the aesthetic mapping between Control_factor and PREDDEG_factor as the horizontal axis and the vertical axis.
@@
 
-Mean value of the earnings is more representative for the two categories, predominant degree and control.
-According to the plot, we realized that private for-profit colleges whose predominant degree awarded is graduate degree have the lowest average earning with the minimum average value at $3100, while Public colleges whose predominant degree awarded are graduate degree has the highest average earnings with the maximum average value at $45,850.
-
-##6.12 The Relationship between Predominant Degree and Control<a id="6.12"></a>
-
-The following plot shows the relationship between two factor variables (predominant degree award variable and control variable).
+The mean value of the earning is more representative for the two categories: predominant degree factor and control factor.
+According to the plot’s color contrast, we realize that private for-profit colleges whose predominant degree awarded are certificate degree have the lowest average earning, while both the public colleges and private nonprofit colleges whose predominant degree awarded is graduate degree have the highest average earning.
+
+##6.13 The Relationship between Predominant Degree and Control<a id="6.13"></a>
+
+The following plot shows the relationship between two factor variables (predominant degree variable and control variable).
 
@@
 
-```
-
-```{R}
 data_clean %>%
@@
 
-From the graph we can see that for the public type of school, associate degree is mostly awarded and it is in the middle of the distribution, with fewer NotClassified degree and fewer graduate degree in the extremes.
+From the graph we can see that for the public schools, associate degree is mostly awarded, with fewer NotClassified degree and fewer graduate degree.
 But the distribution is symmetric compared with the other two types.
-For the Private Nonprofit type of school, mostly awarded degrees (Bachelor degree and Graduate degree) are located at the right of the distribution, with fewer NotClassified degree on the left.
-The distribution shows a negative skew: the awarded degrees extend to the left farther than they do to the right.
-For the private for-profit type of school, mostly awarded degree (Certificate degree) is on the left of the distribution, with fewer graduate degree on the right.
-The distribution shows a positive skew: the awarded degrees extend to the right farther than they do to the left.
-
-##6.13 The Relationship between Control and Median earnings of Students Working and Not Enrolled 6 Years after Entry<a id="6.13"></a>
+For the private nonprofit schools, mostly awarded degrees are bachelor degree and graduate degree.
+The distribution shows a negative skew.
+For the private for-profit type of school, mostly awarded degree is certificate degree.
+
+##6.14 The Relationship between Control and Median earnings of Students Working and Not Enrolled 6 Years after Entry<a id="6.14"></a>
 
@@
 
-```
-```{r}
 summary(data_clean$MD_EARN_WNE_P6[data_clean$Control_factor=="Public"])
@@
 
-For Public type of school, the earnings has a high density around the level of $25,000.
+For public type of school, the earnings has a high density around the level of $25,000.
 From the box plot summary, we realize that half of the public school student earnings are between $21,880 and $30,320.
-Also, a larger mean value than median value indicates a positive skew distribution.
-
-For the Private Nonprofit type of school, the earnings has a high density around the level of $31,250.
+Also, a larger mean value than median value indicates a positive-skewed distribution.
+
+For the private nonprofit type of school, the earnings has a high density around the level of $31,250.
 From the box plot summary, we realize that half of the private nonprofit school student earnings are between $13,000 and $35,800.
-Also, a lower mean value than median value indicates a negative skew distribution.
-
-For the Private For-profit type of school, the earnings has a high density around the level of $18,750.
+Also, a lower mean value than median value indicates a negative-skewed distribution.
+
+For the private for-profit type of school, the earnings has a high density around the level of $18,750.
 The distribution is not symmetric.
 From the box plot summary, we realize that half of the private for-profit school student earnings are between $12,080 and $24,700.
-Also, a lower mean value than median value indicates a negative skew distribution.
+Also, a lower mean value than median value indicates a negative-skewed distribution.
 
@@
 In terms of predominant degrees offered, 43.2% are classified as a certificate degree predominant institution, which is the most prevalent type.
-
-Our findings also suggest that the Title IV institutions located in DC, VT, MA, NH, RI have a relatively high amount of average net costs than those in other states.
+It is also suggested that the Title IV institutions located in DC, VT, MA, NH, RI have a relatively high amount of average net costs than those in other states.
 The students in private non-profit institutions and public institutions are less likely to seek a Pell Grant as financial assistance than the students in the private for-profit colleges.In most private for-profit colleges, the students are more likely to be financially independet.
@@
 
-As to the factors that influence post-graduation earnings, we explored the relationship between median earnings six years from entry and other six variables.
-The analysis suggests that there is a positive relationship between the median debt amount for students who have not completed college and post-graduation earnings.
+As to the factors that influence post-graduation earnings, the analysis suggests that there is a positive relationship between the median debt amount for students who have not completed college and post-graduation earnings.
 This may be due to the fact that top universities usually have higher tuition but may also indicate that students who are willing to take more debts for their college education are likely to make more money after graduation.
@@
 This finding is reasonable since students making more money are more likely to pay off their debt.
-The variable can be used as a school-level predictor to predict overall post-graduation earnings for each university.
-The degree type, unsurprisingly, has an impact on the earnings students are making after graduation.
-As the level of education increases, the post-graduate earnings also increase significantly, which suggests that investing in education is worthwhile.
-
-When it comes to relationship between control of institution and earning, private for-profit institutions whose predominant degree awarded is graduate degree has the lowest average earnings with the minimum average value at $3100.
-Public institutions whose predominant degree awarded is graduate degree has the highest average earning with the maximum average value at $45850.
+
+When it comes to relationship between control of institution and earning, private for-profit institutions whose predominant degree awarded is graduate degree has the lowest average earnings with the minimum average value at $3,100.
+Public institutions whose predominant degree awarded is graduate degree has the highest average earning with the maximum average value at $45,850.
 Although the graduate degree awarded by public colleges shows the least frequency, it contains the most value because those students (working and not enrolled 6 years after entry) graduating from public colleges holding the graduate degree show the highest average earnings compared with students holding other degrees awarded by different institutions.Also, students graduate from Private Nonprofit institution generally has a higher earning than those students graduate from other types of institutions.
 
-According to all analysis, we conclude that the attributes such as percentage of students who receive a Pell Grant, median earning six years from entry, amount of debt, default rate for student loans and predominant degree type can be used to predict the students' post-graudation earnings.
+According to all analysis, we conclude that the attributes such as percentage of students who receive a Pell Grant, median earning six years from entry, amount of debt, default rate for student loans and predominant degree type can be used to predict the students' post-graudation earnings in the further analysis.

### 2017-02-23 Draft-xuezhou-sjb.Rmd
final
update 6.11 6.12 6.13

--- 2017-02-22 Draft-xuezhou-sjb.Rmd
@@
 * 4.[Objective](#objective)
-* 5.[Summary of Variables](#variablesummary)
-* 5.1 [Individual Variable Summary](#5.1)
+* 5.[Summary of variables](#variablesummary)
+* 5.1 [Individual variable summary](#5.1)
   * 5.1.1 [School](#5.1.1)
@@
   * 5.1.2 Student Body
-    * 5.1.2.1 [Percentage First-generation Students](#5.1.2.1)
-    * 5.1.2.2 [Number of Applications](#5.1.2.2)
+    * 5.1.2.1 [Percentage first-generation students](#5.1.2.1)
+    * 5.1.2.2 [Number of applications](#5.1.2.2)
   * 5.1.3 Cost 
-    * 5.1.3.1 [Average Net Price for Title IV Institutions (Both Public and Private Institutions)](#5.1.3.1)
-    * 5.1.3.2 [Percentage of Students Who are Financially Independent](#5.1.3.2)
+    * 5.1.3.1 [Average net price for Title IV institutions(both public and private institutions)](#5.1.3.1)
+    * 5.1.3.2 [Percentage of students who are financially independent](#5.1.3.2)
   * 5.1.4 Financial Aid
-    * 5.1.4.1 [Percentage of Undergraduates Who Receive a Pell Grant](#5.1.4.1)
-    * 5.1.4.2 [Percent of all Federal Undergraduate Students Receiving a Federal Student Loan](#5.1.4.2l)
-    * 5.1.4.3 [The Median Debt for Students Who Have Completed](#5.1.4.3)
-    * 5.1.4.4 [The Median Debt for Students Who Have not Completed](#5.1.4.4)
+    * 5.1.4.1 [Percentage of undergraduates who receive a Pell Grant](#5.1.4.1)
+    * 5.1.4.2 [Percent of all federal undergraduate students receiving a federal student loan](#5.1.4.2l)
+    * 5.1.4.3 [The median debt for students who have completed](#5.1.4.3)
+    * 5.1.4.4 [The median debt for students who have not completed](#5.1.4.4)
   * 5.1.5 Earning
-    * 5.1.5.1 [Median Family Income in real 2015 Dollars](#5.1.5.1)
+    * 5.1.5.1 [Median family income in real 2015 dollars](#5.1.5.1)
     * 5.1.5.2 [Earning after graduation](#5.1.5.2)
@@
 * 6.Relationship between variables
- * 6.1 [The Relationship between Average Net Cost for Title IV institutions and State](#6.1)
- * 6.2 [The Relationship between Median Debt and Average Net Price for Title IV institutions](#6.2)
- * 6.3 [The Relationship between Percentage of Pell Grant and College Type (CONTROL)](#6.3)
- * 6.4 [The Relationship between Percentage First-generation Students and States](#6.4)
- * 6.5 [The Relationship between Median Family Income and Percentage of Pell Grant](#6.5)
- * 6.6 [The Relationship between the Percentage of Financially-independent Students and College Type (Control)](#6.6)
- * 6.7 [The Relationship between Percentage of Students Who Receive the Federal Loans and College Type (Control).](#6.7)
- * 6.8 [The Relationsip between Median Debts and Post-Graduation Earnings](#6.8)
- * 6.9 [The Relationship between the Default Rate of Loan Repayment and Post-graduation Earnings ](#6.9)
- * 6.10 [The Relationship between Earning 6 Years from Entry and Degree Type](#6.10)
- * 6.11 [The Relationship among Predominant Degree, College Type(Control) and Earning after 6 year of Entry](#6.11)
- * 6.12 [The Relationship between Predominant Degree and College Type (Control)](#6.12)
- * 6.13 [The Relationship between College Type (Control) and Median earnings of Students Working and Not Enrolled 6 Years after Entry](#6.13)
+ * 6.1 [The relationship between NPT4_COMBINE and State](#6.1)
+ * 6.2 [The relationship between Debt and Cost](#6.2)
+ * 6.3 [The relationship between Percentage of Pell Grant and college type(CONTROL)](#6.3)
+ * 6.4 [The relationship between Percentage first-generation students and States](#6.4)
+ * 6.5 [The relationship between family income and percentage of Pell Grant](#6.5)
+ * 6.6 [The relationship between DEP_STAT_PCT_IND and Control](#6.6)
+ * 6.7 [The relationship between PCTLOAN and Control](#6.7)
+ * 6.8 [cost and median debts with earning](#6.8)
+ * 6.9 [default with earning](#6.9)
+ * 6.10 [degree type vs. earning](#6.10)
+ * 6.11 [predominant degree, control  and earning after 6 year of entry](#6.11)
+ * 6.12 [predominant degree + control (count)](#6.12)
+ * 6.13 [control + earning](#6.13)
 * 7. [Findings and Conclusion](#7)
@@
 #1. Introduction<a id="Introduction"></a>
-
 The College Scorecard project was initially designed to increase the transparency of the college data, working as a navigator for students to make informed decisions about their choice for higher education.
@@
 The data provides insights into the performance of institutions that receive federal aid dollars, and the outcomes of the students of those institutions.
-It also gives extra support for students and families to weigh the tradeoffs on choosing the right colleges according to their own needs and educational goals.
+It also gives extra support for students and families to weigh the tradeoffs on choosing the right college according to their own needs and educational goals.
 
 The data was collected by the United States Department of Education, which corresponds to all undergraduate degree-granting institutions of higher education from 1996 through 2016.
-The dataset is pretty rich and it covers a wide range of topics.
-It has 1743 associated columns, each corresponding to one named data variable.
-All data variables are divided into 10 categories including basic information of institution, earnings, financial aid, student body, etc.
+The dataset is pretty rich and covers a wide range of topics.
+It has 1743 associated columns.
+Each corresponds to one named data variable.
+All data variables are divided into 10 categories ranging from Root category to Earnings category.
 The data type contains integer, string and float.
 Data elements with an IPEDS value in the Source field are derived directly from the Integrated Postsecondary Education Data System (IPEDS).
-Further detailed description of the dataset can be found in the full data documentation.
+Further detailed description of the dataset can be found in the CollegeScorecardDataDictionary.xlsx excel file.
 
@@
 The goal of our study is to understand what factors are predictive of students' post-graduation earnings through individual variable analysis and relationship detection.
-To this end, we first explored a set of variables that were carefully selected to represent a school's profile and students demographic information, getting to know the distribution of each variable.
+To this end, we first explored a set of variables that were carefully selected to represent a school's profile, getting to know the distribution of each variable.
 Then we further analyzed how these variables are related, particually the relationship between earnings and other variables.
@@
 #3. Dataset Preparation<a id="dataprep"></a>
-Our analysis focuses on the data of year 2012 -2013.
-We chose the median income 6 years after entry as our response variable, and eliminated many fields with a bunch of missing values as well as some fields not pertaining to post-graduation income.
-We fosued primarily on a set of 38 variables, including age, academic costs, family income, completetion rates, repayment rate, etc.
+Our analysis focuses on the data of year 2012 -2013 of 23 variables.
+We chose the median income 6 years after entry as our predictor variable, and eliminated many fields with a bunch of missing values as well as some fields not pertaining enougn to post-graduation income.
+We then fosued primarily on a set of 23 variables, including age, acedemic costs, family income, completetion rates and repayment rate.
 
@@
 
-The dataset spans nearly 20 years and covers multiple sources including IPEDS, NSLDS, and Department of Treasury.
-Since our goal was to study post-graduation earnings for students in each university, we used 2012-2013 dataset which contains tha most recent earning information.
+The dataset spans nearly 20 years of data and covers multiple sources including IPEDS, NSLDS, and Department of Treasury.
+Since our goal was to study post-graduation for students in each university, we used 2012-2013 dataset which contains tha most recent earning information.
 
@@
 
-Considering there are some special characters in the ID column, we first renamed the first column to make it easier to understand.
+Considering there is some special characters in the ID column, we first renamed the first column to make it easier to understand.
 ```{R}
@@
 
-The full dataset contains 1743 variables, as we metioned before, 38 variables were intially selected to be analyzed.
+The full dataset contains 1743 variables. 36 variables were intially selected to be analyze.
 ```{r, message=FALSE,warning=FALSE}
@@
 
-Considering the variable NPT4_PUB and NPT4_PRIV are mutually exclusive, which both represent average net price, we integrated the information of the two coloumns into a single column, named as NPT4_COMBINE.
+Considering the variable NPT4_PUB and NPT4_PRIV are mutually exclusive both indicate net price.
+we integrated the information of the two coloumns into a single column, named as NPT4_COMBINE.
 The new column records the average net price for both private and public TITLE IV institutions.
@@
 
-In order to choose the variables that have enough valid observations related to the topic, we only considered variables with the percent of missing value less than 25%.
-The new data was named as data_clean.
+In order to choose the variables that have enough valid data related to the topic, we only considered variables with percent missing value less than 25%. The new data set was named as data_clean. 
 ```{r, message=FALSE,warning=FALSE}
@@
 
-For numerical variables, we calculated the basic summary statistics such as median, min, max, standard deviation and mean.
-We also used visualizations such as historgams and boxplots to understand the distribution of the values.
-For categorical variables, we used frequency table to understand the frequencies of each category.
+The data_clean will be the clean dataset we used in the further variable summary and comparison.For numerical variables, we calculate the basic summary statistics such as median, min, max, mode and quantile.
+We also use visualizations such as historgams and boxplots to understand the distribution of the values.
+For categorical variables, we use frequency table to understand the frequencies of each category.
 
 The table below lists the variables that will be used for the further analysis: 13 of them are continuous variables while 5 of them are categorical variables or Unique ID.
-After initial data cleaning process, the new data set contained 18 variables and 7,793 observations.
+After initial data cleaning process, the new data set now contains 18 variables and 7793 observations.
 
@@
 
-we summarized the data set using functions provided in dplyr and ggplot in the variables summary section.
+we will summarize the data set using functions provided in dplyr and ggplot funtions in the variables summary section.
 
@@
 
-Our objective at this stage is to conduct an exploratory data analysis on the CollegeScore dataset, investigating the individual variables' summary statistics and the relationship between these variabls, so as to better prepare for further analaysis on the factors affecting students' post college earnings with some advanced data mining techniques.
-The summary statistics, dplyr function, ggplot functions as well as some visualization techniques were mainly used in this project.
+Our objective at this stage is to conduct an exploration analysis on the CollegeScore dataset, investigating the individual variable's summary statistics and the relationship between these variabls, so as to better prepare for further analaysis on the factors affecting students' post college earnings with some advanced data mining techniques.
+The summary statistics, dplyr function, ggplot functions as well as some visualization techniques will be mainly used in this project.
 
@@
 ##5.1 Individual variable summary<a id="5.1"></a>
+
 ###5.1.1 School<a id="5.1.1"></a> 
@@
 
-The State variable has 59 levels, which include 55 states in the United States and 4 districts outside mainland US.
+State variable has 59 levels, which include 55 states in the United States and 4 districts outside mainland US.
 To better understand how many universities each state has, we created a frequency table as shown below, and also a bar chart to visuzalize the top 15 states having the most universities.
@@
 ####5.1.1.2. Predominant Degree<a id="5.1.1.2"></a> 
-Then we analyzed the type of degree universities mainly offer.
+Then we analyze the type of degree universities mainly offer.
 There are four types of degree taken into account: certificate degree, associate degree, bachelor degree, and graduate degree.
-A frequency table in a format of percentage and a histogram were used to help understand the distribution of the predominant degree.
+A frequency table in a format of percentage and a histogram are used to help understand the distribution of the predominant degree.
 
@@
 
-####5.1.1.3. Control of Institutions <a id="5.1.1.3"></a> 
+####5.1.1.3. Control of Institution<a id="5.1.1.3"></a> 
 In addition to state and predominant degree, control of institution is another categorical variable in the dataset.
-The label of control is to indicate whether or not an institution is privately owned and whether or not it is a for-profit organization.
-There are three levels of this variable: public, private non-profit, and private-profit.
-Public schools are assumed to be non-profit so they are not further divided into two subgroups.
+The label of Control is to indicate whether or not an institution is privately owned and whether or not it is a for-profit organization.
+There are three levels of this variable: Public, Private Nonprofit, and Private-profit.
+Public school is assumed to be non-profit so it is not further divided into two subgroups.
 Below we show the percentage summary of each category and a bar chart is plotted to facilitate understanding.
@@
 
-About 50% of all institutions are private for-profit university, which is the most dominant type. Public and private non-profit are pretty tied in terms of the number of observation in each category.
+About 50% of all institutions are private for-profit university, which is the most dominant type. Public and private nonprofit are pretty tied in terms of the number of observation in each category.
 
 ###5.1.2 Student Body<a id="5.1.2"></a> 
-####5.1.2.1 Percentage of First-generation Students<a id="5.1.2.1"></a>
-This element identifies the share of federally aided students at the institutions who reported that they were the first generaion on the FAFSA (Free Application for Federal Student Aid).
+####5.1.2.1 Percentage first-generation students<a id="5.1.2.1"></a>
+This element identifies the share of federally aided students at the institution who report that they are the first generaion on the FAFSA (Free Application for Federal Student Aid).
 First generation students are defined as students whose parents didn't attend a college.
@@
 With the following codes, we extracted the institutions with the highest and lowest percentage of first-generation students.
-Then we observed this variable's basic summary statistics: min, max, mean, median and standard deviation.
+Then we observed this varible's basic summary statistics: min, max, mean, median and standard deviation.
 ```{R, message=FALSE, warinig = FALSE}
@@
 The lowest value is 6.87%, which existed in Wheaton College.
-From the mean value of 48.08%, we can tell that for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college.
+From the mean value of 48.08%, we can tell that about for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college.
 The standard deviation value of 12.2% indicated a relatively high difference of this element among different colleges.
 
-Below are the codes to generate a visual representation in the histogram. 
+Below is the code to generate a visual representation in the histogram. 
 ```{R}
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for Percentage First-Generation Students") +
+    labs(title="Histogram for Percentage first-generation students") +
     labs(x="PAR_ED_PCT_1STGEN", y="Count") +
@@
 
-From the histogram, we can tell the share of the first-generarion students is most frequently falling into the range between 0.5 and 0.6. The observations with the value more than 0.8 are rare.
+From the histogram, we can tell the share of the first-generarion students is most frequently falling into the range between 0.5 and 0.6. The values with the value more than 0.8 are rare.
 
@@
 
-The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who listed two and more than two institutions when he applied for financial aid when submitting applications for university.
-This information is only available to the students who submitted their applications for federal aid and does not cover the whole student body.
+The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who list two and more than two institutions when he applied for financial aid when submitting application for university.
+This information is only available to students who submitted their applications for federal aid and does not cover the whole student body.
 Nevertheless, this variable still can provide some useful insights about how many universities a student usually apply for.
@@
 
-The summary statistics show that the median percentage of those students submitting two or more than two applications is 41.67% across all universities.
-The institution having the least percentage only has 4% of students applying for more than one university, while the school with the highest percent has as high as 90% of students applying for more than one university.
-The boxplot compares the percent of students applying for only one university with the percent of students applying for more than one.
+The summary statistics show that the median percentage of those students submitting two or more than two application is 41.67% across all universities The institution having the least percentage only has 4% of students applying for more than one university, while the school with the highest percent has as high as 90% of students applying for more than one university.
+The boxplot compares the percent of student applying for only one university with the percent of students applying for more than one.
 We can clearly see that a larger portion of students only listed one institution when submitting their application for financial aid.
@@
 ###5.1.3 Cost<a id="5.1.3"></a>
-####5.1.3.1 Average Net Price for Title IV Institutions(both public and private institutions)<a id="5.1.3.1"></a>
-
-With the following codes, we extracted the Title IV institutions' basic demographic information with the highest and lowest average net price.
+####5.1.3.1 Average net price for Title IV institutions(both public and private institutions)<a id="5.1.3.1"></a>
+
+With the followin codes, we extracted the IV institutions with the highest and lowest Average net price.
 Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
@@
 From the summary results, we learned that the maximum value of the variable is $87,414, which is for Aerosim Flight Academy located in FL.
-while the minimum value is about $-19,360, for Berea College located in KY.
+while the minimum value is about $-19,360 - the value for Berea College located in KY.
 Considering the elements are collected from Title IV institutions, we reasoned that the negative values may be due to the fact that some institution's financial supoort can cover the students' all the costs, even overbalancing the cost amount.
-There is not much differerence between the mean value and the median value, which is around $15,000.
+There are not much differerence between the mean value and the median value, which is around $15,000.
 We can conclude that the average net price costs for most institutions are falling into  this range.
-The standard deviation value is $7604.884 which is relatively large, indicating a great spread in the data.
-
-Below shows the histogram for this variable, which is able to provide us with a visual distribution presentation. 
+The standard deviation value is 7604.884 which is relatively large, indicating a great spread in the data.
+
+Below is the histogram for this variable which is able to provide us with a visual distribution presentation. 
 ```{R}
@@
 
-From the histogram, we notice that the average net cost for most IV institutions falls between $16,000-$20,000.
-There were some unusually large observations.
-The institutions with cost of more than $50,000 were atypical.
-Due to the these large values, the histogram is abviously right-skewed.
+From the histogram, we notice that the average net cost for IV institutions typically is $16000-$18000.
+The values between $16000 -$20000 are quite frequent.
+There were an unsually lagrge number of costs.
+The institutions with cost of more than 50000 were atypical.
+Due to the atypical large value, the histogram is abviously right skewed.
 Without those extreme values, the histogram would be reasonably symmetric.
@@
 
-From the summary results, we learned that the maximum value of the variable is 98.56% (Capella University) while the minimum value is about 0.73% (University of Notre Dame).
+From the summary results, we learned that the maximum value of the variable is 98.56%(Capella University) while the minimum value is about 0.73%(University of Notre Dame).
 From the mean value, we can tell that  in most colleges, around 56% of the students are financially independent.
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for Percentage of Students Who Are Financially Independent") +
+    labs(title="Histogram for Percentage of students who are financially independent") +
     labs(x="DEP_STAT_PCT_IND", y="Count") +
@@
 
-From the histogram, we notice that the values between 0.7 to 0.8 are quite frequent. There is no obvious extreme value in the sample, thus the histogram is normally distributed.
+From the histogram, we notice that the values between 0.7 to 0.8 are quite frequent. There was no obvious extreme value in the sample, thus the histogram is normally distributed.
 
@@
 This element, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received Pell Grants in a given year, which is an important measure of the financial aid an institution provides to low-income students.
-We are using the code below to summarize this variable(min, max, mean, median and standard deviation) and extract the institutions with the highest and lowest value.
+We are using the code below to summarize this variable and extract the institutions with the highest and lowest value.
 
@@
 From the output, we can tell that the range is between 0 to 100%.	No students in Thunderbird School of Global Management has received the Pell grant while all of the students in American Conservatory Theater received the Pell Grant.
-Among the colleges in the sample, half of the colleges have the rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs.
+Among the colleges in the sample, half of the colleges colleges have the rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs.
 The standard deviation value is 22.57%,  indicating a relatively high difference of this element among different colleges, which can also be noticed from the largest range (100%).
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for Percentage of Undergraduates Who Receive a Pell Grant") +
+    labs(title="Histogram for Percentage of undergraduates who receive a Pell Grant") +
     labs(x="PCTPELL", y="Count") +
@@
 
-####5.1.4.2 Percent of All Federal Undergraduate Students Receiving a Federal Student Loan<a id="5.1.4.2"></a>
+####5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan<a id="5.1.4.2"></a>
 
@@
 It can provide important context to figures related to debt, repayment and non-repayment.
-We were using the code below to summarize this variable(min, max, mean, median and standard deviation) and extracted the institutions with highest and lowest value.
+We are using the code below to summarize this variable and extract the institutions with highest and lowest value.
 
@@
 
-From the output, we can tell that the range is between 0 to 100%.	As we can recall from the previous variable (PCTPELL), Gadsden State Community College has no students receiving federal loans in the given year.
-Among the colleges in the sample, half of colleges have this rate as high as 60.08% , which means around 60% of undergraduate students received the fedaral student loans in the given year.
+From the output, we can tell that the range is between 0 to 100%.	As we can recall from the previous variable(PCTPELL), Gadsden State Community College has no students receiving federal loans in the given year.
+Among the colleges in the sample, half of colleges have this rate as high as 60.08% , which means around 60% of undergraduate students receive the fedaral student loans.
 The standard deviation value is 29.26%,  indicating a relatively high difference of this element among different colleges, which can also be noticed from the largest range (100%).
@@
                    aes(fill=..count..),na.rm=TRUE)+
-    labs(title="Histogram for Percent of all Federal Undergraduate Students Receiving a Federal Student Loan") +
+    labs(title="Histogram for Percent of all federal undergraduate students receiving a federal student loan") +
     labs(x="PCTFLOAN", y="Count") +
@@
 
-From the histogram, it's obvious that there are two peaks in the graph. The rate between 0-10% and the rates between 60% to 80% have similar frequency compared to the other ranges.
+From the histogram, it's obvious that there are two peaks in the graph. The rate between 0-10% and the rates between 60% to 80% have similarly frequency compared to the other ranges.
 
@@
 
-This variable recorded debt for students who have completed the education.
-We were using the code below to summarize this variable(min, max, mean, median and standard deviation) and extracted the institutions with largest and smallest value.
-
+This variable recorded debt for students who have completed the education. We are using the code below to summarize this variable. 
 ```{R, message=FALSE, warinig = FALSE}
@@
 data_clean %>% ggplot(aes(x = WDRAW_DEBT_MDN)) +
-                geom_histogram(binwidth = 450)
+                geom_histogram(binwidth = 450) #warning?
 
@@
 We use summary statistics, standard deviation and histogram to understand the distribution of this variable.
-The summary statistics show that students in more than 50% of institutions have a median debt of more than $5500, which is not very unaffordable.
-The median debt for attending students could be as much as $27620 for some schools and as low as $600 for others, indicating that it may have a large variance.
+The summary statistics show that students in more than 50% of institution have a median debt of more than $5500, which is not very unaffordable.
+The median debt for attending students can be as much as $27620 for some school and as low as $600 for others, which indicates that it may have a large variance.
 So we also calculate standard deviation to measure the dispersion of the distribution.
@@
 ####5.1.5.1 Median family income in real 2015 dollars<a id="5.1.5.1"></a>
-This element is recorded the median family income of students, one of the undergraduate student demographics for earnings cohorts.
-We are using the code below to summarize this variable(min, max, mean, median and standard deviation).
+This element is recorded the median family income of students, one of the undergraduate student demographics for earnings cohorts. We are using the code below to summarize this variable.
 
@@
 From the output, we can tell that the highest value is $155,606.6 for Jewish Theological Seminary of America, while the lowest value is $0 for Associated Technical College-Los Angeles.
-From the median value of $19,262.26, we can tell that among the colleges in the sample, the students in half of them had the family income below $19,262.26.
+From the median value of $19,262.26, we can tell that among the colleges in the sample, the students in half of them had the family income below $19,262.26	.
 The standard deviation value is $18,824.31,  indicating a relatively high difference of this element among different colleges.
@@
 Due to some atypical large values, the histogram is obviously skewed to right.
-The values larger than $18,000 are tending to be extreme and the cases larger than $75000 are rare, which means that rare colleges have the students with family income more than $75000.
+The values larger than $18,000 are tend to be extreme and the cases larger than $75000 are rare, which means that rare colleges have the students with family income more than $75000.
 
@@
 
-MD_EARN_WNE_P6 and MD_EARN_WNE_P10 are two of the earning varibles we selected to study students' post-graduation earnings.
+We selected two variables MD_EARN_WNE_P6, MD_EARN_WNE_P10 in the earning group to study students' post-graduation earnings.
 MD_EARN_WNE_P6 is the median earning for students after six years of their entry, which corresponds to two years after graduation for a four-year program.
-Likewise, MD_EARN_WNE_P10 is the median earning ten years after their entry, which usually corresponds to four years after graduation.
-These two metrics give us an idea of how much money students are making after their graduation, in both short and medium terms.
+Likewise, MD_EARN_WNE_P10 is the median earning ten years after their entry, which usually corresponds to four years after graduation These two metrics give us an idea of how much money students are making after their graduation, in both short and medium terms.
 
@@
 Besides, the ten-year post-graduation earning has a wider spread than that of the six-years.
-This indicates that there is a greater gap of the earning amount among the different institutions, which can be confirmed by an increase in the standard deviation.
+This indicates that there is a greater gap between who are earning more and who are earning less, which can be confirmed by an increase in the standard deviation.
 
@@
 ####5.1.6.1 3-year Repayment rate<a id="5.1.6.1"></a>
-The variable RPY_3YR_RT_SUPP measures the percentage of students who entered repayment from October 1, 2010 to September 30, 2011 and made progress in the following three years.
+The variable RPY_3YR_RT_SUPP measures the percentage of students who entered repayment from October 1, 2010 to September 30, 2011, and made progress in the following three years.
 Repayment rates are generally considered more sensitive than default rates(we will study later), which measures only the worst-case scenario for repayment outcomes.
@@
 #6. Relationship between variables<a id="6"></a>
-##6.1. The Relationship between Average Net Cost for Title IV institutions and State<a id="6.1"></a>
-
-To analyze the relationship between the average net cost for Title IV institutions and State, we grouped the data into state level to take a further analysis on each state's min, max and median value of the Average Net Cost - We would like to investigate the distribution of institutions ' average net price in each state.
+
+##6.1. The relationship between average net cost for Title IV institutions and State<a id="6.1"></a>
+
+To analyze the relationship between the average net cost for Title IV institutions and State, we grouped the data into the state level to take a further look.
+We would like to investigate the distribution of average net price in each state.
 
@@
 Based on the output, we can see that the highest median value for the avreage net price exists in DC while the lowest median value is in AS(American Samoa).
-Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are the states with certain well-known educational institutions.
+Moreover, we found that the top 5 states are DC, VT, MA, NH, RI, which are all the states have  well-known educational institutions.
 
@@
 
-To compare the percentage of Pell Grant in different control types, we grouped the Percentage of Pell Grant data using college type to take a further analysis on each college type's min, max and median value of Percentage of Pell Grant.
+To compare the percentage of Pell Grant in different control types, we grouped the Percentage of Pell Grant data using college type to take a further analysis.
 
@@
 
-To compare the percentage of first-generation students in each states, we grouped data into the state level to take a further analysis on the min, max and median value of Percentage first-generation students on the state base.
-The top 5 states as well as the lowest 5 states are extracted.
+To compare the percentage of first-generation students in each states, we grouped data into the state level to take a further analysis.
 
@@
 The results show that the universities located in ND, SD, VT, MN and MT have the lowest percentage first-generation students, which indicates that the students in these five states would have a better family education background.
-The five states with the highest share of first-generation students are LA, NJ, MP, MV, and MH.
-
-##6.5. The relationship between Family Income and Percentage of Pell Grant<a id="6.5"></a>
+The five states with the highest share of first-generation students are LA, NJ, MP, MV, MH.
+
+##6.5. The relationship between family income and percentage of Pell Grant<a id="6.5"></a>
 
@@
 
-##6.6. The Relationship between the Percentage of Financially-independent Students and Control<a id="6.6"></a>
-
-To compare the percentage of financially-independent students in the institutions with different college type, we grouped the data by the control type(private for-profit, public, private nonprofit) to take a further analysis on the min, max and median value for each type.
+##6.6. The relationship between DEP_STAT_PCT_IND and Control<a id="6.6"></a>
+
+To compare the percentage of financially-independent students in the institutions with different college type, we grouped the data by the control type(private for-profit, public, private nonprofit) to take a further analysis.
 
@@
 
-##6.7. The Relationship between Percentage of Students Who Receive the Federal Loans and Control.<a id="6.7"></a>
-
-To compare the percentage of students who receive the federal loans in the given year in each schools control type, we grouped the data by the control type(private for-profit, public, private non-profit) to take a further analysis on the min, max and median value of the variable seperately.
+##6.7. The relationship between PCTLOAN and Control.<a id="6.7"></a>
+
+To compare the percentage of students who receive the federal loans in the given year in each schools control type, we grouped the data by the control type(private for-profit, public, private non-profit) to take a further analysis.
 A stacked histogram is also used to generate a more visual presentation.
@@
 From the result,  private institutions have a much larger median value than the public institutions.
-With a much higher costs burden, it makes sense that the low-income students in private institutions are much more likely to seek financial aids than the students in public institutions.
-
+With a much higher costs burden, it makes sense that the low-income students in private institutions are much more likely to seek financial aid than the public institution students.
 From the stacked histogram, we can tell that for all  three college types, the most frequent range appears between 0.6 and 0.8, which is consistent with the findings we observe in the individual variable summary section.
 
-##6.8 The Relationsip between Median Debts and Post-Graduation Earnings<a id="6.8"></a> 
-Next, we analyze the relationship between the amount of the debt and post-graduation earnings.
+##6.8 Median debts with earning<a id="6.8"></a> 
+Next, we analyze the relationship between the amount the debt and post-graduation earnings.
 There are two variables in the dataset to summarize the median amount of debt for student body, one for students who already graduated and one for students who have not completed their study.
@@
 
-##6.9 The Relationship between the Default Rate of Loan Repayment and Post-graduation Earnings<a id="6.9"></a> 
-For one thing, understanding the relationship between the default rate of loan repayment and post-graduation earnings can help federal financial committee to improve their decisions about load granting.
-If the default rate decreases as the earnings increases, the committee may prefer to grant more financial aid to institutions whose students have higher earnings.
+##6.9 default with earning<a id="6.9"></a>
+
+For one thing, understanding the relationship between the default rate of loan repayment and post-graduation earnings can help federal financial committee to improve their decision about load granting.
+If the default rate decreases as the earnings increase, the committee may prefer to grant more financial aid to institutions whose students have higher earnings.
 For another, the default rate of repayment for a university can also help predict the post-graduation earnings for its students.
@@
 The resulting r = -0.33 shows that the correlation between earnings and default rate is negative but the relationship is not very strong, which may be due to the influences of some outliers.
-To conclude, the negative relationship between earnings and default rate of loan repayment needs extra data to validate.
-
-##6.10 The Relationship between Earning 6 Years from Entry and Degree Type<a id="6.10"></a>
+To conclude, the negative relationship between earnings and default rate of loan repayment needs extra data to confirm.
+
+##6.10 degree type vs. earning<a id="6.10"></a>
 
 The higher education one receives, the higher the earning is after graduation.
-This is a common assumption we are making when investing higher education, and this is it really the case.
+This is an assumption we are making when investing higher education.
+But is it really the case.
 In this question, we would like to study if it's worthwhile to invest in higher degree.
@@
 
-From the boxplot above, we see that the higher the education, the higher the median earning 6 years after entry. Especially, a graduate-degree gives a hugh rise to one's earning.
-
-##6.11 The Relationship among Predominant Degree, Control and Earning after 6 year of Entry<a id="6.11"></a>
-
-The following plot shows the relationship among two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry).
+From the boxplot below, we see that the higher the education, the higher the median earning 6 years after entry. Especially, a graduate-degree gives a hugh rise to one's earning.
+
+##6.11 predominant degree, control  and earning after 6 year of entry<a id="6.11"></a>
+
+The following plot shows two factor variables (predominant degree award variable and control variable) associated with one numeric variable (MD_EARN_WNE_P6: Median earnings of students working and not enrolled 6 years after entry).
 
@@
 
-Mean value of the earnings is more representative for the two categories, predominant degree and control.
-According to the plot, we realized that private for-profit colleges whose predominant degree awarded is graduate degree have the lowest average earning with the minimum average value at $3100, while Public colleges whose predominant degree awarded are graduate degree has the highest average earnings with the maximum average value at $45,850.
-
-##6.12 The Relationship between Predominant Degree and Control<a id="6.12"></a>
+Mean value of the earning is more representative for the two categories, predominant degree factor and control factor.
+According to the plot’s color contrast, we realize that private for-profit college whose predominant degree award is certificate degree has the lowest average earning, while both the public college and private nonprofit college whose predominant degree award is graduate degree has the highest average earning.
+
+##6.12 predominant degree + control (count)<a id="6.12"></a>
 
@@
 For the R code, our team first set the aesthetic mapping on Control_factor as the horizontal axis.
-Then, we used the function geom_bar() to fill different control types using the counting values of PREDDEG, which created bars based on a grouped frequency distribution.
-After that, we used the function xtabs() to calculate the exact number for each predominant degree awarded by different types of colleges.
+Then, we use the function geom_bar() to fill different control types using the counting values of PREDDEG, which creates bars based on a grouped frequency distribution.
+After that, we further use the function xtabs() to calculate the exact number for each predominant degree awarded by different types of colleges.
 
@@
 
-From the graph we can see that for the public type of school, associate degree is mostly awarded and it is in the middle of the distribution, with fewer NotClassified degree and fewer graduate degree in the extremes.
+From the graph we can see that, for the public type of school, Associate degree is mostly awarded and it is in the middle of the distribution, with fewer NotClassified degree and fewer Graduate degree in the extremes.
 But the distribution is symmetric compared with the other two types.
@@
 The distribution shows a negative skew: the awarded degrees extend to the left farther than they do to the right.
-For the private for-profit type of school, mostly awarded degree (Certificate degree) is on the left of the distribution, with fewer graduate degree on the right.
+For the Private For-profit type of school, mostly awarded degree (Certificate degree) is on the left of the distribution, with fewer Graduate degree on the right.
 The distribution shows a positive skew: the awarded degrees extend to the right farther than they do to the left.
 
-##6.13 The Relationship between Control and Median earnings of Students Working and Not Enrolled 6 Years after Entry<a id="6.13"></a>
+#6.13 control + earning<a id="6.13"></a>
 
@@
 
-We used the ggplot() function to initialize the ggplot object and declared the input data frame for a graphic with x equals Control_factor and y equals Median Earnings.
-Then, we used the geom_violin() function which showed a compact display of a continuous distribution.
+We use the ggplot() function to initialize the ggplot object and declare the input data frame for a graphic with x equals Control_factor and y equals Median Earnings.
+Then, we use the geom_violin() function which shows a compact display of a continuous distribution.
 The geom_violin() function is a blend of geom_boxplot() and geom_density(), which shows a mirrored density plot.
-For better visualization, we used geom_boxplot() to create a box and whiskers plot, which showed five summary statistics (the median, two hinges and two whiskers), and all outlying points individually.
-And the summary() function further displayed the exact values.
-The plot superimposed with boxplots showed side-by-side violin plots of median earnings of students working and not enrolled 6 years after entry according to different college types.
+For better visualization, we use geom_boxplot() to create a box and whiskers plot, which shows five summary statistics (the median, two hinges and two whiskers), and all ???outlying???
+points individually.
+And the summary() function further displays the exact values.
+The plot superimposed with boxplots shows side-by-side violin plots of median earnings of students working and not enrolled 6 years after entry according to different college types.
 
@@
 
-All the three types of schools have far out values above, which can be explored further.
-Obvious differences were contained in these three boxplots.
-The boxplot of public college is comparatively short, while the boxplot of private nonprofit is comparatively tall.
-The private for-profit college has the lowest median earning.
-
-For Public type of school, the earnings has a high density around the level of $25,000.
-From the box plot summary, we realize that half of the public school student earnings are between  $21,880 and $30,320.
+All the three types of school have far out values above, which can be explored further.
+Obvious differences contain in these three boxplots.
+The boxplot of Public college is comparatively short, while the boxplot of Private Nonprofit is comparatively tall.
+The Private For-profit college has the lowest median earning.
+
+For public type of school, the earnings have a high density around the level of $25,000.
+From the boxplot summary, we realize that half of the public school student earnings are between $22,700 and $30,900.
 Also, a larger mean value than median value indicates a positive skew distribution.
 
-For the Private Nonprofit type of school, the earnings has a high density around the level of $31,250.
-From the box plot summary, we realize that half of the private nonprofit school student earnings are between $13,000 and $35,800.
-Also, a lower mean value than median value indicates a negative skew distribution.
-
-For the Private For-profit type of school, the earnings has a high density around the level of $18,750.
+For the Private Nonprofit type of school, the earnings have a high density around the level of $31,250.
+From the boxplot summary, we realize that half of the private nonprofit school student earnings are between $25,400 and $38,100.
+Also, it shows a positive skew distribution with a larger mean value at $33,350 than the median value at $31,900.
+
+For the Private For-profit type of school, the earnings have a high density around the level of $18,750.
 The distribution is not symmetric.
-From the box plot summary, we realize that half of the private for-profit school student earnings are between $12,080 and $24,700.
-Also, a lower mean value than median value indicates a negative skew distribution.
+From the boxplot summary, we realize that half of the private for-profit school student earnings are between $17,000 and $27,300.
+Also, the mean value at $22,610 higher than median value indicates a positive skew distribution.
 
@@
 
-Through the process of summarizing these variables and analyzing the relationship between these variables, we found that for federally-aided students in most colleges, around half of the students' parents did not complete colleges.The average net price costs for most institutions are falling into the range between $16000 - $20000.
-Among the 7703 institution samples, more than 50% of the students received a Pell Grant or Federal Loan as means of finanical support, and on average, 60% of students are financially independent.
-In most colleges, the students are able to reach the family income between $12000 and $18000 after they graduate.
+Through the process of summarizing these variables and analyzing the relationship between these variables, we found that for federally-aided students in most colleges, around half of the students' parents did not complete college.The average net price costs for most institutions are falling into the range between $16000 - $20000.
+Among the 7703 institution samples, more than 50% of the students receive a Pell Grant or Federal Loans as means of finanical support, and on average, 60% of students are financially independent.
+In most colleges, the students have the family income between $12000 and $18000.
 
@@
 
-It indicates that there exists a negative correlation between the percentage of students who receive a Pell Grant and the students' median family income; for institutions with low percentage of students who receive the Pell Grant, the financial aid like the Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, thus these college students will tend to have a high median family income after they graduate.
+It indicates that there exists a negative correlation between the percentage of students who receive a Pell Grant and the students' median family income; for institutions with low percentage of students who receive the Pell Grant, the financial aid like the Pell Grant may not be as necessary as they are in the institutions with high value of PCTPELL, since these college students will tend to have a high median family income.
 

### 2017-02-23 Draft-xuezhou-sjb.Rmd
update conclusion

@@
 
-When it comes to relationship between control of institution and earning, private for-profit institutions whose predominant degree awarded is graduate degree has the lowest average earnings with the minimum average value at $3100.
-Public institutions whose predominant degree awarded is graduate degree has the highest average earning with the maximum average value at $45850.
-Although the graduate degree awarded by public colleges shows the least frequency, it contains the most value because those students (working and not enrolled 6 years after entry) graduating from public colleges holding the graduate degree show the highest average earnings compared with students holding other degrees awarded by different institutions.Also, students graduate from Private Nonprofit institution generally has a higher earning than those students graduate from other types of institutions.
+When it comes to relationship between control of institution and earning, although the ‘Graduate-degree’ awarded by public colleges and private nonprofit colleges shows a lower frequency, they contain the most value because those students (working and not enrolled 6 years after entry) graduating from public colleges and private nonprofit colleges holding the graduate degree show the highest average earnings compared with students holding other degrees awarded by different institutions.
+Also, students graduate from Private Nonprofit institution generally has a higher earning than those students graduate from other types of institutions.
 

### 2017-03-15 Draft-xuezhou-sjb.Rmd
added redundant

--- 2017-02-23 Draft-xuezhou-sjb.Rmd
@@
 
-#5. Summary of variables<a id="5"></a>
+#5. Summary of variables<a id="variablesummary"></a>
 ##5.1 Individual variable summary<a id="5.1"></a> 
@@
 The mean value of the earning is more representative for the two categories: predominant degree factor and control factor.
-According to the plot’s color contrast, we realize that private for-profit colleges whose predominant degree awarded are certificate degree have the lowest average earning, while both the public colleges and private nonprofit colleges whose predominant degree awarded is graduate degree have the highest average earning.
+According to the plot???s color contrast, we realize that private for-profit colleges whose predominant degree awarded are certificate degree have the lowest average earning, while both the public colleges and private nonprofit colleges whose predominant degree awarded is graduate degree have the highest average earning.
 
